### 2019-01-15 main.tex
init commit - with some latex compile errors

%% For double-blind review submission, w/o CCS and ACM Reference (max submission space)
%\documentclass[acmsmall,review,anonymous]{acmart}\settopmatter{printfolios=true,printccs=false,printacmref=false}
%% For double-blind review submission, w/ CCS and ACM Reference
%\documentclass[acmsmall,review,anonymous]{acmart}\settopmatter{printfolios=true}
%% For single-blind review submission, w/o CCS and ACM Reference (max submission space)
%\documentclass[acmsmall,review]{acmart}\settopmatter{printfolios=true,printccs=false,printacmref=false}
%% For single-blind review submission, w/ CCS and ACM Reference
%\documentclass[acmsmall,review]{acmart}\settopmatter{printfolios=true}
%% For final camera-ready submission, w/ required CCS and ACM Reference
\documentclass[acmsmall]{acmart}\settopmatter{}

%% Journal information
%% Supplied to authors by publisher for camera-ready submission;
%% use defaults for review submission.
\acmJournal{PACMPL}
\acmVolume{1}
\acmNumber{ICFP} % CONF = POPL or ICFP or OOPSLA
\acmArticle{1}
\acmYear{2018}
\acmMonth{1}
\acmDOI{} % \acmDOI{10.1145/nnnnnnn.nnnnnnn}
\startPage{1}

%% Copyright information
%% Supplied to authors (based on authors' rights management selection;
%% see authors.acm.org) by publisher for camera-ready submission;
%% use 'none' for review submission.
\setcopyright{none}
%\setcopyright{acmcopyright}
%\setcopyright{acmlicensed}
%\setcopyright{rightsretained}
%\copyrightyear{2018}           %% If different from \acmYear

%% Bibliography style
\bibliographystyle{ACM-Reference-Format}
%% Citation style
%% Note: author/year citations are required for papers published as an
%% issue of PACMPL.
\citestyle{acmauthoryear}   %% For author/year citations

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% Note: Authors migrating a paper from PACMPL format to traditional
%% SIGPLAN proceedings format must update the '\documentclass' and
%% topmatter commands above; see 'acmart-sigplanproc-template.tex'.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%% Some recommended packages.
\usepackage{booktabs}   %% For formal tables:
                        %% http://ctan.org/pkg/booktabs
\usepackage{subcaption} %% For complex figures with subfigures/subcaptions
                        %% http://ctan.org/pkg/subcaption

\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{amstext}
\usepackage{enumerate}

\usepackage{multirow}
\usepackage{color}

\usepackage{xparse}
\usepackage{listings}
\usepackage{lstabella}

\usepackage{mathpartir}
\usepackage{comment}

\usepackage{mathtools}
\usepackage{graphicx}

\input{sections/macros.tex}

\begin{document}

%% Title information
\title[Short Title]{Full Title}         %% [Short Title] is optional;
                                        %% when present, will be used in
                                        %% header instead of Full Title.
%\titlenote{with title note}             %% \titlenote is optional;
                                        %% can be repeated if necessary;
                                        %% contents suppressed with 'anonymous'
%\subtitle{Subtitle}                     %% \subtitle is optional
%\subtitlenote{with subtitle note}       %% \subtitlenote is optional;
                                        %% can be repeated if necessary;
                                        %% contents suppressed with 'anonymous'

%% Author information
%% Contents and number of authors suppressed with 'anonymous'.
%% Each author should be introduced by \author, followed by
%% \authornote (optional), \orcid (optional), \affiliation, and
%% \email.
%% An author may have multiple affiliations and/or emails; repeat the
%% appropriate command.
%% Many elements are not rendered, but should be provided for metadata
%% extraction tools.

%% Author with single affiliation.
\author{Jinxu Zhao}
%\authornote{with author1 note}          %% \authornote is optional;
                                        %% can be repeated if necessary
%\orcid{nnnn-nnnn-nnnn-nnnn}             %% \orcid is optional
\affiliation{
  %\position{Position1}
  \department{Department of Computer Science}              %% \department is recommended
  \institution{The University of Hong Kong}            %% \institution is required
  %\streetaddress{Pokfulam}
  \city{Hong Kong}
  %\state{State1}
  %\postcode{Post-Code1}
  \country{China}                    %% \country is recommended
}
\email{jxzhao@cs.hku.hk}          %% \email is recommended

%% Author with two affiliations and emails.
\author{Bruno C. d. S. Oliveira}
%\authornote{with author1 note}          %% \authornote is optional;
%% can be repeated if necessary
%\orcid{nnnn-nnnn-nnnn-nnnn}             %% \orcid is optional
\affiliation{
	%\position{Position1}
	\department{Department of Computer Science}              %% \department is recommended
	\institution{The University of Hong Kong}            %% \institution is required
	%\streetaddress{Pokfulam}
	\city{Hong Kong}
	%\state{State1}
	%\postcode{Post-Code1}
	\country{China}                    %% \country is recommended
}

%% Author with two affiliations and emails.
\author{Tom Schrijvers}
%\authornote{with author2 note}          %% \authornote is optional;
%% can be repeated if necessary
%\orcid{nnnn-nnnn-nnnn-nnnn}             %% \orcid is optional
\affiliation{
	%\position{Position2a}
	\department{Department of Computer Science}             %% \department is recommended
	\institution{KU Leuven}           %% \institution is required
	%\streetaddress{Street2a Address2a}
	\city{Leuven}
	%\state{State2a}
	%\postcode{Post-Code2a}
	\country{Belgium}                   %% \country is recommended
}
\email{tom.schrijvers@cs.kuleuven.be}

%% Abstract
%% Note: \begin{abstract}...\end{abstract} environment must come
%% before \maketitle command
\begin{abstract}
Text of abstract \ldots.
\end{abstract}

%% 2012 ACM Computing Classification System (CSS) concepts
%% Generate at 'http://dl.acm.org/ccs/ccs.cfm'.
\begin{CCSXML}
<ccs2012>
<concept>
<concept_id>10011007.10011006.10011008</concept_id>
<concept_desc>Software and its engineering~General programming languages</concept_desc>
<concept_significance>500</concept_significance>
</concept>
<concept>
<concept_id>10003456.10003457.10003521.10003525</concept_id>
<concept_desc>Social and professional topics~History of programming languages</concept_desc>
<concept_significance>300</concept_significance>
</concept>
</ccs2012>
\end{CCSXML}

\ccsdesc[500]{Software and its engineering~General programming languages}
\ccsdesc[300]{Social and professional topics~History of programming languages}
%% End of generated code

%% Keywords
%% comma separated list
\keywords{keyword1, keyword2, keyword3}  %% \keywords are mandatory in final camera-ready submission

%% \maketitle
%% Note: \maketitle command must come after title commands, author
%% commands, abstract environment, Computing Classification System
%% environment and commands, and keywords command.
\maketitle

\input{sections/introduction.tex}

\input{sections/algorithm.tex}

%% Acknowledgments
\begin{acks}                            %% acks environment is optional
                                        %% contents suppressed with 'anonymous'
  %% Commands \grantsponsor{<sponsorID>}{<name>}{<url>} and
  %% \grantnum[<url>]{<sponsorID>}{<number>} should be used to
  %% acknowledge financial support and will be used by metadata
  %% extraction tools.
  This material is based upon work supported by the
  \grantsponsor{GS100000001}{National Science
    Foundation}{http://dx.doi.org/10.13039/100000001} under Grant
  No.~\grantnum{GS100000001}{nnnnnnn} and Grant
  No.~\grantnum{GS100000001}{mmmmmmm}.  Any opinions, findings, and
  conclusions or recommendations expressed in this material are those
  of the author and do not necessarily reflect the views of the
  National Science Foundation.
\end{acks}

%% Bibliography
\bibliography{main}

%% Appendix
\appendix
\section{Appendix}

Text of appendix \ldots

\end{document}

### 2019-01-15 sections/algorithm.tex
init commit - with some latex compile errors

\section{Algorithm}

$$\begin{aligned}
\text{Type variables}\qquad&a, b\\
\text{Existential variables}\qquad&\a, \b\\[3mm]
\text{Types}\qquad&A', B', C' &::=&\quad 1 \mid a \mid \forall x. A' \mid A'\to B'\\
\text{Mono-types}\qquad&\tau &::=&\quad 1 \mid a \mid \tau_1\to \tau_2\\
\text{Context}\qquad&\P &::=&\quad \cdot \mid \P, a \mid \P, x:A' \color{red} \\[3mm]
% \text{Declarative Judgements}\qquad&\jo &::=&\quad \cdot \mid \P\vdash A' \le B' \jc \jo\\
% &&&\quad \mid \P\vdash e\Leftarrow A' \jc \jo \mid \P\vdash e\To A' \jc \jo \mid \P\vdash A' \bullet e\TTo C' \jc \jo\\
% \noalign{\jimmy{\text{What do the markers do: $\G \mark,a \mark,b\vdash \j_1;\j_2;\j_3$ means $\G,a,b\vdash \j_1 \land \G,a\vdash \j_1,\j_2 \land \G\vdash \j_3$}}}\\[3mm]
\text{Algorithmic types}\qquad&A, B, C &::=&\quad 1 \mid a \mid \a \mid \forall a. A \mid A\to B\\
\text{Algorithmic judgement chain}\qquad&\j &::=&\quad A \le B \mid e\Lto A \mid e\To_{a} \j \mid \appInfAlg{A}{e}{a}{\j}\\[3mm]
\text{Algorithmic worklist}\qquad&\G &::=&\quad \cdot \mid \G, a \mid \G, \a \mid \G, x: A \mid \G \mark \j\\[3mm]
\text{Terms}\qquad&e&::=&\quad x \mid () \mid \lam{x}{e} \mid e_1~e_2 \mid (e:A)
\end{aligned}$$

### 2019-01-15 sections/conclusion.tex
init commit - with some latex compile errors

\section{Conclusion and Future Work}

In this paper we have shown how to mechanise an algorithmic subtyping relation
for higher-order polymorphism, together with its proofs of soundness,
completeness and decidability, in the Abella proof assistant. 
In ongoing work we are extending our mechanisation with a bidirectional type
inference algorithm. The main difficulty there is communicating the
instantiations of existential variables from the subtyping algorithm to the
type inference. To make this possible we are exploring a continuation passing
style formulation, which generalises the worklist approach.
Another possible extension is to have the algorithm return an explicit witness
for the subtyping as part of type-directed elaboration into System F.

### 2019-01-15 sections/introduction.tex
init commit - with some latex compile errors

\section{Introduction}

Most statically typed functional languages support a form of
\emph{(implicit) parametric polymorphism}~\cite{reynolds1983types}.

### 2019-01-17 sections/Metatheory.tex
sections; add algo. rules; few text on algo

\section{Metatheory}

\subsection{Transfer to the Declarative System}
\jimmy{The transfer rule outputs declarative judgment chains, which are later processed by another declarative relation, handling all the guessing.}

\begin{figure}[t]
\framebox{$\Gm \sto \Om$} $\Gm$ transfers to $\Om$. \jimmy{Always substitute from the very first existential variable. Types in contexts ($\Gm,x:A,\ldots$) should be taken care.}
\begin{gather*}
\inferrule*[right=$\mathtt{{\sto}}\Om$]
{~}
{\Om \sto \Om}
\quad
\inferrule*[right=$\mathtt{{\sto}\al}$]
{\Om\vdash\tau \\ \Om,[\tau/\al]\Gm \sto \Om}
{\Om,\al,\Gm \sto \Om}
\end{gather*}
\caption{Transfer Rules}
\label{fig:trans}
\end{figure}

\subsection{Soundness}
\subsection{Completeness}
\subsection{Decidability}

\subsection{Abella}

### 2019-01-17 sections/algorithm.tex
Fix compile error
sections; add algo. rules; few text on algo

@@
 \section{Algorithm}
+
+\subsection{Syntax and Well-Formedness}
 
@@
 \text{Type variables}\qquad&a, b\\
-\text{Existential variables}\qquad&\a, \b\\[3mm]
-\text{Types}\qquad&A', B', C' &::=&\quad 1 \mid a \mid \forall x. A' \mid A'\to B'\\
+\text{Existential variables}\qquad&\al, \bt\\[3mm]
+%\text{Types}\qquad&A', B', C' &::=&\quad 1 \mid a \mid \forall x. A' \mid A'\to B'\\
 \text{Mono-types}\qquad&\tau &::=&\quad 1 \mid a \mid \tau_1\to \tau_2\\
-\text{Context}\qquad&\P &::=&\quad \cdot \mid \P, a \mid \P, x:A' \color{red} \\[3mm]
+%\text{Context}\qquad&\Psi &::=&\quad \cdot \mid \Psi, a \mid \Psi, x:A' \color{red} \\[3mm]
 % \text{Declarative Judgements}\qquad&\jo &::=&\quad \cdot \mid \P\vdash A' \le B' \jc \jo\\
 % &&&\quad \mid \P\vdash e\Leftarrow A' \jc \jo \mid \P\vdash e\To A' \jc \jo \mid \P\vdash A' \bullet e\TTo C' \jc \jo\\
-% \noalign{\jimmy{\text{What do the markers do: $\G \mark,a \mark,b\vdash \j_1;\j_2;\j_3$ means $\G,a,b\vdash \j_1 \land \G,a\vdash \j_1,\j_2 \land \G\vdash \j_3$}}}\\[3mm]
-\text{Algorithmic types}\qquad&A, B, C &::=&\quad 1 \mid a \mid \a \mid \forall a. A \mid A\to B\\
-\text{Algorithmic judgement chain}\qquad&\j &::=&\quad A \le B \mid e\Lto A \mid e\To_{a} \j \mid \appInfAlg{A}{e}{a}{\j}\\[3mm]
-\text{Algorithmic worklist}\qquad&\G &::=&\quad \cdot \mid \G, a \mid \G, \a \mid \G, x: A \mid \G \mark \j\\[3mm]
-\text{Terms}\qquad&e&::=&\quad x \mid () \mid \lam{x}{e} \mid e_1~e_2 \mid (e:A)
+% \noalign{\jimmy{\text{What do the markers do: $\G \vdash,a \vdash,b\vdash \j_1;\j_2;\j_3$ means $\G,a,b\vdash \j_1 \land \G,a\vdash \j_1,\j_2 \land \G\vdash \j_3$}}}\\[3mm]
+\text{Algorithmic types}\qquad&A, B, C &::=&\quad 1 \mid a \mid \al \mid \forall a. A \mid A\to B\\
+\text{Algorithmic judgement chain}\qquad&\jg &::=&\quad A \le B \mid e\Lto A \mid e\To_{a} \jg \mid \appInfAlg{A}{e}\\[3mm]
+\text{Algorithmic worklist}\qquad&\Gm &::=&\quad \cdot \mid \Gm, a \mid \Gm, \al \mid \Gm, x: A \mid \Gm \vdash \jg\\[3mm]
+\text{Terms}\qquad&e&::=&\quad x \mid () \mid \lam{e} \mid e_1~e_2 \mid (e:A)
 \end{aligned}$$
+
+\paragraph{Existential Variables} ...
+
+\paragraph{Judgment and judgment chain} ...
+\paragraph{Worklist Judgments} Our algorithmic judgment has a non-standard form.
+We combine the traditional contexts and judgment(s) into a single worklist.
+The worklist collects variable bindings and judgments in order, which represents their scopings:
+each variable is visible after it is declared.
+For example, $a,x:a,x \Lto a$ is a valid worklist, but $\textcolor{red}{x} \Lto a, x:\textcolor{red}{a},a$ is not.
+
+\paragraph{Hole Notation}...
+
+\begin{figure}[t]
+\framebox{$\Gm\vdash\jg$} Well-formed algorithmic judgment.
+\begin{gather*}
+\inferrule*[right=$\mathtt{wf{\le}}$]
+{\Gm\vdash A \\ \Gm\vdash B}
+{\Gm\vdash A\le B}
+\qquad
+\inferrule*[right=$\mathtt{wf{\Lto}}$]
+{\Gm\vdash e \\ \Gm\vdash A}
+{\Gm\vdash e \Lto A}
+\\
+\inferrule*[right=$\mathtt{wf{\To}}$]
+{\Gm\vdash e \\ \Gm, a\vdash \jg}
+{\Gm\vdash e \To_a \jg}
+\qquad
+\inferrule*[right=$\mathtt{wf{\TTo}}$]
+{\Gm\vdash A \\ \Gm, a\vdash \jg \\ \Gm\vdash e}
+{\Gm\vdash \appInfAlg{A}{e}}
+\end{gather*}
+
+\framebox{$\text{wf }\Gm$} Well-formed algorithmic worklist.
+\begin{gather*}
+\inferrule*[right=$\mathtt{wf\cdot}$]
+{~}{\text{wf }\cdot}
+\qquad
+\inferrule*[right=$\mathtt{wf_a}$]
+{\text{wf }\Gm}
+{\text{wf }\Gm, a}
+\qquad
+\inferrule*[right=$\mathtt{wf_{\al}}$]
+{\text{wf }\Gm}
+{\text{wf }\Gm, \al}
+\qquad
+\inferrule*[right=$\mathtt{wf_{of}}$]
+{\text{wf }\Gm \\ \Gm\vdash A}
+{\text{wf }\Gm, x:A}
+\qquad
+\inferrule*[right=$\mathtt{wf_{\jg}}$]
+{\text{wf }\Gm \\ \Gm\vdash\jg}
+{\text{wf }\Gm \vdash\jg}
+\end{gather*}
+\caption{Algorithmic Well-Formedness}\label{fig:alg}
+\end{figure}
+
+\subsection{Algorithmic System}
+
+% TODO rules too long
+\begin{figure}[t]
+\framebox{$\Gm\rto \Gm'$} $\Gm$ reduces to $\Gm'$.
+%\jimmy{Whenever there is a change on judgment count, marker should be properly added/removed. Decidability: all valid judgment chain should be reduced to nil.}
+\begin{gather*}
+\begin{aligned}
+\Gm, a &\rto \Gm \qquad\qquad
+\Gm, \al \rto \Gm \qquad\qquad
+\Gm, x:A \rto \Gm
+\\[2mm]
+\Gm \vdash 1\le 1 &\rto \Gm\\
+\Gm \vdash a\le a &\rto \Gm\\
+\Gm \vdash \al\le \al &\rto \Gm
+\\[2mm]
+\Gm \vdash A_1\to A_2 \le B_1\to B_2 &\rto \Gm \vdash A_2 \le B_2 \vdash B_1\le A_1\\
+\Gm \vdash \forall a. A\le B &\rto \Gm,\al \vdash [\al/a]A\le B \quad\text{\jimmy{No more existential variable "leaks"!}}\\
+\Gm \vdash A\le \forall b. B &\rto \Gm,b \vdash A\le B \quad\text{\jimmy{And no more type variable "leaks"!}}
+\\[2mm]
+%\\
+%\text{Let } \color{red}\Gm[\al \toto B // G_M] &:= \Gm_L, G_M, [B/\al]\Gm_R, \text{ when } \Gm[\al] = \Gm_L,\al,\Gm_R \ (G_M\text{ defaults to }\cdot)\\
+\Gm[\al] \vdash \al \le A\to B &\rto [\al[1]\to\al[2]/\al] (\Gm[\al[1], \al[2]] \vdash \al[1]\to \al[2] \le A \to B)\\
+ &\qquad \text{when }\al\notin FV(A)\cup FV(B)\\
+\Gm[\al] \vdash A\to B \le \al &\rto [\al[1]\to \al[2]/\al] (\Gm[\al[1], \al[2]] \vdash A \to B \le \al[1]\to \al[2])\\
+ &\qquad \text{when }\al\notin FV(A)\cup FV(B)
+ \\[2mm]
+\Gm[\al][\bt] \vdash \al \le \bt &\rto [\al/\bt](\Gm[\al][])\\
+\Gm[\al][\bt] \vdash \bt \le \al &\rto [\al/\bt](\Gm[\al][])\\
+\Gm[a][\bt] \vdash a \le \bt &\rto [a/\bt](\Gm[a][])\\
+\Gm[a][\bt] \vdash \bt \le a &\rto [a/\bt](\Gm[a][])\\
+\Gm[\bt] \vdash 1 \le \bt &\rto [1/\bt](\Gm[])\\
+\Gm[\bt] \vdash \bt \le 1 &\rto [1/\bt](\Gm[])
+\\[2mm]
+\Gm \vdash e \Lto B &\rto \Gm \vdash e\To_a a\le B\\
+\Gm \vdash () \Lto 1 &\rto \Gm\\
+\Gm \vdash e\Lto \forall a. A &\rto \Gm,a \vdash e\Lto A\\
+\Gm \vdash \lambda x. e \Lto A\to B &\rto \Gm, x:A  \vdash e \Lto B\\
+\Gm[\al] \vdash \lambda x. e \Lto \al &\rto [\al[1]\to \al[2] / \al](\Gm[\al[1],\al[2]], x:\al[1] \vdash e \Lto \al[2]) \quad\text{\jimmy{Additional case}}
+\\[2mm]
+\Gm \vdash x\To_a \jg &\rto \Gm \vdash [A/a] \jg \qquad (x:A)\in \Gm\\
+\Gm \vdash (e:A)\To_a \jg &\rto \Gm \vdash [A/a]\jg \vdash e \Lto A\\
+\Gm \vdash ()\To_a \jg &\rto \Gm \vdash [1/a]\jg \qquad\jimmy{\downarrow\text{Precise control on exvars: to the $2^{nd}$ chain}}\\
+\Gm \vdash \lambda x. e \To_a \jg &\rto
+	\Gm,\al,\bt \vdash [\al\to\bt/a]\jg, x:\al \vdash e\Lto \bt \quad\jimmy{\text{differs from Dunfield's}}\\
+\Gm \vdash e_1\ e_2 \To_a \jg &\rto \Gm \vdash e_1\To_{a_1} (\appInfAlg{a_1}{e_2})
+\\[2mm]
+\Gm \vdash \appInfAlg{\forall a. A}{e} &\rto \Gm,\al \vdash \appInfAlg{[\al/a]A}{e}\\
+\Gm \vdash \appInfAlg{A\to C}{e} &\rto \Gm \vdash [C/a]\jg \vdash e \Lto A\\
+\Gm[\al] \vdash \appInfAlg{\al}{e} &\rto
+	[\al[1]\to\al[2]/\al](\Gm[\al[1], \al[2]] \vdash [\al[2]/a]\jg \vdash e\Lto \al[1])\\
+ &\color{magenta} \makebox[0pt]{\qquad or}\phantom{{}\rto{}}
+ 	[\al[1]\to\al[2]/\al](\Gm[\al[1], \al[2]] \vdash \appInfAlg{\al[1]\to\al[2]}{e}) \quad\text{(1 step equiv.)}
+\end{aligned}
+\end{gather*}
+\caption{Algorithmic Typing}\label{fig:alg}
+\end{figure}

### 2019-01-18 sections/algorithm.tex
some alg. texts

@@
 
-$$\begin{aligned}
+\begin{figure}
+\begin{gather*}
+\begin{aligned}
 \text{Type variables}\qquad&a, b\\
-\text{Existential variables}\qquad&\al, \bt\\[3mm]
+\text{Existential variables}\qquad&\al, \bt
+\\[3mm]
 %\text{Types}\qquad&A', B', C' &::=&\quad 1 \mid a \mid \forall x. A' \mid A'\to B'\\
@@
 % &&&\quad \mid \P\vdash e\Leftarrow A' \jc \jo \mid \P\vdash e\To A' \jc \jo \mid \P\vdash A' \bullet e\TTo C' \jc \jo\\
-% \noalign{\jimmy{\text{What do the markers do: $\G \vdash,a \vdash,b\vdash \j_1;\j_2;\j_3$ means $\G,a,b\vdash \j_1 \land \G,a\vdash \j_1,\j_2 \land \G\vdash \j_3$}}}\\[3mm]
-\text{Algorithmic types}\qquad&A, B, C &::=&\quad 1 \mid a \mid \al \mid \forall a. A \mid A\to B\\
-\text{Algorithmic judgement chain}\qquad&\jg &::=&\quad A \le B \mid e\Lto A \mid e\To_{a} \jg \mid \appInfAlg{A}{e}\\[3mm]
-\text{Algorithmic worklist}\qquad&\Gm &::=&\quad \cdot \mid \Gm, a \mid \Gm, \al \mid \Gm, x: A \mid \Gm \vdash \jg\\[3mm]
+% \noalign{\jimmy{\text{What do the markers do: $\G \vdash,a \vdash,b\vdash \j_1;\j_2;\j_3$ means $\G,a,b\vdash \j_1 \land \G,a\vdash \j_1,\j_2 \land \G\vdash \j_3$}}} \\[3mm]
+\text{Algorithmic types}\qquad&A, B, C &::=&\quad 1 \mid a \mid \al \mid \forall a. A \mid A\to B
+\\[3mm]
 \text{Terms}\qquad&e&::=&\quad x \mid () \mid \lam{e} \mid e_1~e_2 \mid (e:A)
-\end{aligned}$$
+\\[3mm]
+\text{Algorithmic judgement chain}\qquad&\jg &::=&\quad A \le B \mid e\Lto A \mid e\To_{a} \jg \mid \appInfAlg{A}{e}
+\\[3mm]
+\text{Algorithmic worklist}\qquad&\Gm &::=&\quad \cdot \mid \Gm, a \mid \Gm, \al \mid \Gm, x: A \mid \Gm \vdash \jg\\
+\text{Declarative worklist}\qquad&\Om &::=&\quad \cdot \mid \Gm, a \mid \Gm, x: A \mid \Gm \vdash \jg
+\end{aligned}
+\end{gather*}
+\caption{Algorithmic Syntax}\label{fig:alg_syntax}
+\end{figure}
 
@@
 
-\paragraph{Judgment and judgment chain} ...
-\paragraph{Worklist Judgments} Our algorithmic judgment has a non-standard form.
+\paragraph{Judgments and judgment chains} Judgment is a core component during algorithmic type-checking.
+There are four kinds of judgments in our system: subtyping, checking, inference and application inference.
+Subtyping and checking are relatively simple, since the result could be either success or fail, and no values will be returned.
+For inference and application inference, they would return some type that leads to further computation, so we use continuation-passing-style encodings for them.
+For example, a judgment chain $e \To_a (a \le B)$ contains two judgments: first we want to infer the type of the expression $e$, and then check if that type is a subtype of $B$.
+In the rest of the paper, we usually refer to ``judgment chains'' by ``judgments''.
+
+\paragraph{Worklist Judgments} Our algorithm has a non-standard form.
 We combine the traditional contexts and judgment(s) into a single worklist.
-The worklist collects variable bindings and judgments in order, which represents their scopings:
-each variable is visible after it is declared.
-For example, $a,x:a,x \Lto a$ is a valid worklist, but $\textcolor{red}{x} \Lto a, x:\textcolor{red}{a},a$ is not.
+The worklist collects variable bindings and judgments in \emph{order}, which represents their scopings:
+each variable is visible only after it is declared.
+For example, $[a,x:a,x \Lto a]$ is a valid worklist, but $[\textcolor{red}{x} \Lto a, x:\textcolor{red}{a},a]$ is not (red-colored symbols are out-of-scope).
 
@@
 \begin{figure}[t]
-\framebox{$\Gm\vdash\jg$} Well-formed algorithmic judgment.
+\framebox{$\Gm\vdash\jg$} Well-formed algorithmic judgment
 \begin{gather*}
@@
 
-\framebox{$\text{wf }\Gm$} Well-formed algorithmic worklist.
+\framebox{$\text{wf }\Gm$} Well-formed algorithmic worklist
 \begin{gather*}
@@
 \end{gather*}
-\caption{Algorithmic Well-Formedness}\label{fig:alg}
+\caption{Algorithmic Well-Formedness}\label{fig:alg_wf}
 \end{figure}
@@
 \Gm \vdash A_1\to A_2 \le B_1\to B_2 &\rto \Gm \vdash A_2 \le B_2 \vdash B_1\le A_1\\
-\Gm \vdash \forall a. A\le B &\rto \Gm,\al \vdash [\al/a]A\le B \quad\text{\jimmy{No more existential variable "leaks"!}}\\
-\Gm \vdash A\le \forall b. B &\rto \Gm,b \vdash A\le B \quad\text{\jimmy{And no more type variable "leaks"!}}
+\Gm \vdash \forall a. A\le B &\rto \Gm,\al \vdash [\al/a]A\le B \quad\text{\jimmy{No more exvar leaks}}\\
+\Gm \vdash A\le \forall b. B &\rto \Gm,b \vdash A\le B \quad\text{\jimmy{No more typevar leaks}}
 \\[2mm]
@@
 \Gm \vdash \lambda x. e \Lto A\to B &\rto \Gm, x:A  \vdash e \Lto B\\
-\Gm[\al] \vdash \lambda x. e \Lto \al &\rto [\al[1]\to \al[2] / \al](\Gm[\al[1],\al[2]], x:\al[1] \vdash e \Lto \al[2]) \quad\text{\jimmy{Additional case}}
+\Gm[\al] \vdash \lambda x. e \Lto \al &\rto [\al[1]\to \al[2] / \al](\Gm[\al[1],\al[2]], x:\al[1] \vdash e \Lto \al[2]) \quad\text{\jimmy{Additional}}
 \\[2mm]
@@
 \Gm \vdash (e:A)\To_a \jg &\rto \Gm \vdash [A/a]\jg \vdash e \Lto A\\
-\Gm \vdash ()\To_a \jg &\rto \Gm \vdash [1/a]\jg \qquad\jimmy{\downarrow\text{Precise control on exvars: to the $2^{nd}$ chain}}\\
+\Gm \vdash ()\To_a \jg &\rto \Gm \vdash [1/a]\jg \qquad\jimmy{\downarrow\text{Precise control on exvars}}\\
 \Gm \vdash \lambda x. e \To_a \jg &\rto
-	\Gm,\al,\bt \vdash [\al\to\bt/a]\jg, x:\al \vdash e\Lto \bt \quad\jimmy{\text{differs from Dunfield's}}\\
+	\Gm,\al,\bt \vdash [\al\to\bt/a]\jg, x:\al \vdash e\Lto \bt \quad\jimmy{\text{differs}}\\
 \Gm \vdash e_1\ e_2 \To_a \jg &\rto \Gm \vdash e_1\To_{a_1} (\appInfAlg{a_1}{e_2})
@@
  &\color{magenta} \makebox[0pt]{\qquad or}\phantom{{}\rto{}}
- 	[\al[1]\to\al[2]/\al](\Gm[\al[1], \al[2]] \vdash \appInfAlg{\al[1]\to\al[2]}{e}) \quad\text{(1 step equiv.)}
+ 	[\al[1]\to\al[2]/\al](\Gm[\al[1], \al[2]] \vdash \appInfAlg{\al[1]\to\al[2]}{e})
 \end{aligned}

### 2019-01-18 sections/algorithm.tex
Add rule number for algrithmic system. More texts on alg.

@@
 \subsection{Syntax and Well-Formedness}
+
+Figure~\ref{fig:alg:syntax} shows the
+syntax and the well-formedness judgement.
 
@@
 \text{Existential variables}\qquad&\al, \bt
-\\[3mm]
+\\[2mm]
 %\text{Types}\qquad&A', B', C' &::=&\quad 1 \mid a \mid \forall x. A' \mid A'\to B'\\
 \text{Mono-types}\qquad&\tau &::=&\quad 1 \mid a \mid \tau_1\to \tau_2\\
-%\text{Context}\qquad&\Psi &::=&\quad \cdot \mid \Psi, a \mid \Psi, x:A' \color{red} \\[3mm]
+%\text{Context}\qquad&\Psi &::=&\quad \cdot \mid \Psi, a \mid \Psi, x:A' \color{red} \\[2mm]
 % \text{Declarative Judgements}\qquad&\jo &::=&\quad \cdot \mid \P\vdash A' \le B' \jc \jo\\
 % &&&\quad \mid \P\vdash e\Leftarrow A' \jc \jo \mid \P\vdash e\To A' \jc \jo \mid \P\vdash A' \bullet e\TTo C' \jc \jo\\
-% \noalign{\jimmy{\text{What do the markers do: $\G \vdash,a \vdash,b\vdash \j_1;\j_2;\j_3$ means $\G,a,b\vdash \j_1 \land \G,a\vdash \j_1,\j_2 \land \G\vdash \j_3$}}} \\[3mm]
+% \noalign{\jimmy{\text{What do the markers do: $\G \vdash,a \vdash,b\vdash \j_1;\j_2;\j_3$ means $\G,a,b\vdash \j_1 \land \G,a\vdash \j_1,\j_2 \land \G\vdash \j_3$}}} \\[2mm]
 \text{Algorithmic types}\qquad&A, B, C &::=&\quad 1 \mid a \mid \al \mid \forall a. A \mid A\to B
-\\[3mm]
+\\[2mm]
 \text{Terms}\qquad&e&::=&\quad x \mid () \mid \lam{e} \mid e_1~e_2 \mid (e:A)
-\\[3mm]
+\\[2mm]
 \text{Algorithmic judgement chain}\qquad&\jg &::=&\quad A \le B \mid e\Lto A \mid e\To_{a} \jg \mid \appInfAlg{A}{e}
-\\[3mm]
+\\[2mm]
 \text{Algorithmic worklist}\qquad&\Gm &::=&\quad \cdot \mid \Gm, a \mid \Gm, \al \mid \Gm, x: A \mid \Gm \vdash \jg\\
@@
 \end{gather*}
-\caption{Algorithmic Syntax}\label{fig:alg_syntax}
-\end{figure}
 
-\paragraph{Existential Variables} ...
-
-\paragraph{Judgments and judgment chains} Judgment is a core component during algorithmic type-checking.
-There are four kinds of judgments in our system: subtyping, checking, inference and application inference.
-Subtyping and checking are relatively simple, since the result could be either success or fail, and no values will be returned.
-For inference and application inference, they would return some type that leads to further computation, so we use continuation-passing-style encodings for them.
-For example, a judgment chain $e \To_a (a \le B)$ contains two judgments: first we want to infer the type of the expression $e$, and then check if that type is a subtype of $B$.
-In the rest of the paper, we usually refer to ``judgment chains'' by ``judgments''.
-
-\paragraph{Worklist Judgments} Our algorithm has a non-standard form.
-We combine the traditional contexts and judgment(s) into a single worklist.
-The worklist collects variable bindings and judgments in \emph{order}, which represents their scopings:
-each variable is visible only after it is declared.
-For example, $[a,x:a,x \Lto a]$ is a valid worklist, but $[\textcolor{red}{x} \Lto a, x:\textcolor{red}{a},a]$ is not (red-colored symbols are out-of-scope).
-
-\paragraph{Hole Notation}...
-
-\begin{figure}[t]
 \framebox{$\Gm\vdash\jg$} Well-formed algorithmic judgment
@@
 \end{gather*}
-\caption{Algorithmic Well-Formedness}\label{fig:alg_wf}
+\caption{Syntax and Well-Formedness for the Algorithmic System}\label{fig:alg:syntax}
 \end{figure}
 
+\paragraph{Existential Variables} ...
+
+\paragraph{Judgments and judgment chains} Judgment is a core component of algorithmic type-checking.
+There are four kinds of judgments in our system: subtyping, checking, inference and application inference.
+Subtyping and checking are relatively simple, since the result could be either success or fail, and no values will be returned.
+For inference and application inference, they would return some type that leads to further computation, so we use continuation-passing-style encodings for them.
+For example, a judgment chain $e \To_a (a \le B)$ contains two judgments: first we want to infer the type of the expression $e$, and then check if that type is a subtype of $B$.
+In the rest of the paper, we usually refer to ``judgment chains'' by ``judgments''.
+
+\paragraph{Worklist Judgments} Our algorithm has a non-standard form.
+We combine the traditional contexts and judgment(s) into a single worklist.
+The worklist collects variable bindings and judgments in \emph{order}, which represents their scopings:
+each variable is visible only after it is declared.
+For example, $[a,x:a,x \Lto a]$ is a valid worklist, but $[\textcolor{red}{x} \Lto a, x:\textcolor{red}{a},a]$ is not (red-colored symbols are out-of-scope).
+
+\paragraph{Hole Notation}...
+
 \subsection{Algorithmic System}
+
+\newcounter{algRuleCounter}
+\newcommand \algrule {\stepcounter{algRuleCounter}\rrule{\arabic{algRuleCounter}}}
 
@@
 \begin{aligned}
-\Gm, a &\rto \Gm \qquad\qquad
-\Gm, \al \rto \Gm \qquad\qquad
-\Gm, x:A \rto \Gm
+\Gm, a &\algrule \Gm \qquad
+\Gm, \al \algrule \Gm \qquad
+\Gm, x:A \algrule \Gm
 \\[2mm]
-\Gm \vdash 1\le 1 &\rto \Gm\\
-\Gm \vdash a\le a &\rto \Gm\\
-\Gm \vdash \al\le \al &\rto \Gm
-\\[2mm]
-\Gm \vdash A_1\to A_2 \le B_1\to B_2 &\rto \Gm \vdash A_2 \le B_2 \vdash B_1\le A_1\\
-\Gm \vdash \forall a. A\le B &\rto \Gm,\al \vdash [\al/a]A\le B \quad\text{\jimmy{No more exvar leaks}}\\
-\Gm \vdash A\le \forall b. B &\rto \Gm,b \vdash A\le B \quad\text{\jimmy{No more typevar leaks}}
+\Gm \vdash 1\le 1 &\algrule \Gm\\
+\Gm \vdash a\le a &\algrule \Gm\\
+\Gm \vdash \al\le \al &\algrule \Gm\\
+\Gm \vdash A_1\to A_2 \le B_1\to B_2 &\algrule \Gm \vdash A_2 \le B_2 \vdash B_1\le A_1\\
+\Gm \vdash \forall a. A\le B &\algrule \Gm,\al \vdash [\al/a]A\le B \quad\text{\jimmy{No more exvar leaks}}\\
+\Gm \vdash A\le \forall b. B &\algrule \Gm,b \vdash A\le B \quad\text{\jimmy{No more typevar leaks}}
 \\[2mm]
@@
 %\text{Let } \color{red}\Gm[\al \toto B // G_M] &:= \Gm_L, G_M, [B/\al]\Gm_R, \text{ when } \Gm[\al] = \Gm_L,\al,\Gm_R \ (G_M\text{ defaults to }\cdot)\\
-\Gm[\al] \vdash \al \le A\to B &\rto [\al[1]\to\al[2]/\al] (\Gm[\al[1], \al[2]] \vdash \al[1]\to \al[2] \le A \to B)\\
+\Gm[\al] \vdash \al \le A\to B &\algrule [\al[1]\to\al[2]/\al] (\Gm[\al[1], \al[2]] \vdash \al[1]\to \al[2] \le A \to B)\\
  &\qquad \text{when }\al\notin FV(A)\cup FV(B)\\
-\Gm[\al] \vdash A\to B \le \al &\rto [\al[1]\to \al[2]/\al] (\Gm[\al[1], \al[2]] \vdash A \to B \le \al[1]\to \al[2])\\
+\Gm[\al] \vdash A\to B \le \al &\algrule [\al[1]\to \al[2]/\al] (\Gm[\al[1], \al[2]] \vdash A \to B \le \al[1]\to \al[2])\\
  &\qquad \text{when }\al\notin FV(A)\cup FV(B)
  \\[2mm]
-\Gm[\al][\bt] \vdash \al \le \bt &\rto [\al/\bt](\Gm[\al][])\\
-\Gm[\al][\bt] \vdash \bt \le \al &\rto [\al/\bt](\Gm[\al][])\\
-\Gm[a][\bt] \vdash a \le \bt &\rto [a/\bt](\Gm[a][])\\
-\Gm[a][\bt] \vdash \bt \le a &\rto [a/\bt](\Gm[a][])\\
-\Gm[\bt] \vdash 1 \le \bt &\rto [1/\bt](\Gm[])\\
-\Gm[\bt] \vdash \bt \le 1 &\rto [1/\bt](\Gm[])
+\Gm[\al][\bt] \vdash \al \le \bt &\algrule [\al/\bt](\Gm[\al][])\\
+\Gm[\al][\bt] \vdash \bt \le \al &\algrule [\al/\bt](\Gm[\al][])\\
+\Gm[a][\bt] \vdash a \le \bt &\algrule [a/\bt](\Gm[a][])\\
+\Gm[a][\bt] \vdash \bt \le a &\algrule [a/\bt](\Gm[a][])\\
+\Gm[\bt] \vdash 1 \le \bt &\algrule [1/\bt](\Gm[])\\
+\Gm[\bt] \vdash \bt \le 1 &\algrule [1/\bt](\Gm[])
 \\[2mm]
-\Gm \vdash e \Lto B &\rto \Gm \vdash e\To_a a\le B\\
-\Gm \vdash () \Lto 1 &\rto \Gm\\
-\Gm \vdash e\Lto \forall a. A &\rto \Gm,a \vdash e\Lto A\\
-\Gm \vdash \lambda x. e \Lto A\to B &\rto \Gm, x:A  \vdash e \Lto B\\
-\Gm[\al] \vdash \lambda x. e \Lto \al &\rto [\al[1]\to \al[2] / \al](\Gm[\al[1],\al[2]], x:\al[1] \vdash e \Lto \al[2]) \quad\text{\jimmy{Additional}}
+\Gm \vdash e \Lto B &\algrule \Gm \vdash e\To_a a\le B\\
+\Gm \vdash () \Lto 1 &\algrule \Gm\\
+\Gm \vdash e\Lto \forall a. A &\algrule \Gm,a \vdash e\Lto A\\
+\Gm \vdash \lambda x. e \Lto A\to B &\algrule \Gm, x:A  \vdash e \Lto B\\
+\Gm[\al] \vdash \lambda x. e \Lto \al &\algrule [\al[1]\to \al[2] / \al](\Gm[\al[1],\al[2]], x:\al[1] \vdash e \Lto \al[2]) \quad\text{\jimmy{Additional}}
 \\[2mm]
-\Gm \vdash x\To_a \jg &\rto \Gm \vdash [A/a] \jg \qquad (x:A)\in \Gm\\
-\Gm \vdash (e:A)\To_a \jg &\rto \Gm \vdash [A/a]\jg \vdash e \Lto A\\
-\Gm \vdash ()\To_a \jg &\rto \Gm \vdash [1/a]\jg \qquad\jimmy{\downarrow\text{Precise control on exvars}}\\
-\Gm \vdash \lambda x. e \To_a \jg &\rto
+\Gm \vdash x\To_a \jg &\algrule \Gm \vdash [A/a] \jg \qquad (x:A)\in \Gm\\
+\Gm \vdash (e:A)\To_a \jg &\algrule \Gm \vdash [A/a]\jg \vdash e \Lto A\\
+\Gm \vdash ()\To_a \jg &\algrule \Gm \vdash [1/a]\jg \qquad\jimmy{\downarrow\text{Precise control on exvars}}\\
+\Gm \vdash \lambda x. e \To_a \jg &\algrule
 	\Gm,\al,\bt \vdash [\al\to\bt/a]\jg, x:\al \vdash e\Lto \bt \quad\jimmy{\text{differs}}\\
-\Gm \vdash e_1\ e_2 \To_a \jg &\rto \Gm \vdash e_1\To_{a_1} (\appInfAlg{a_1}{e_2})
+\Gm \vdash e_1\ e_2 \To_a \jg &\algrule \Gm \vdash e_1\To_{a_1} (\appInfAlg{a_1}{e_2})
 \\[2mm]
-\Gm \vdash \appInfAlg{\forall a. A}{e} &\rto \Gm,\al \vdash \appInfAlg{[\al/a]A}{e}\\
-\Gm \vdash \appInfAlg{A\to C}{e} &\rto \Gm \vdash [C/a]\jg \vdash e \Lto A\\
-\Gm[\al] \vdash \appInfAlg{\al}{e} &\rto
+\Gm \vdash \appInfAlg{\forall a. A}{e} &\algrule \Gm,\al \vdash \appInfAlg{[\al/a]A}{e}\\
+\Gm \vdash \appInfAlg{A\to C}{e} &\algrule \Gm \vdash [C/a]\jg \vdash e \Lto A\\
+\Gm[\al] \vdash \appInfAlg{\al}{e} &\algrule
 	[\al[1]\to\al[2]/\al](\Gm[\al[1], \al[2]] \vdash [\al[2]/a]\jg \vdash e\Lto \al[1])\\
- &\color{magenta} \makebox[0pt]{\qquad or}\phantom{{}\rto{}}
+ &\color{magenta} \makebox[0pt]{\qquad or} \phantom{{}\rrule{}{}}
  	[\al[1]\to\al[2]/\al](\Gm[\al[1], \al[2]] \vdash \appInfAlg{\al[1]\to\al[2]}{e})
@@
 \end{figure}
+
+The algorithmic typing reduction rules, defined in Figure~\ref{fig:alg}, has the form $\Gm \rto \Gm'$,
+where the worklists $\Gm, \Gm'$ contains variable declarations and judgments.
+In every step it pops the first judgment from the worklist for processing,
+possibly pushes new judgments to the worklist, and repeats this process until the list is empty.
+A worklist that could be finally reduced to empty is considered to be successfully type-checked.
+When the first element is a variable declaration, we simply continue the process (Rules 1-3).
+The other rules always deal with the first judgment in the worklist. Logically we can discern 6 group of rules.
+
+Firstly, we have six rules (Rules 4-9) that are similar to those of declarative subtyping.
+For instance, Rule 7 consumes one subtyping judgment and pushes two back to the worklist.
+Rule 8 differs from declarative Rule ${\le}{\forall}L$ by introducing an existential variable $\al$ instead of guessing the mono-type $\tau$ for instantiation.
+Each existential variable is later solved to a mono-type $\tau$ with the same context,
+so the final solution of $\al$, $\tau$, should be well-formed under context $\Gm$.
+
+\paragraph{Variable scopings within worklist}
+Rules 8 and 9 involve variable declaration, which demonstrates how our worklist treats variable scopings.
+Rule 8 introduces an existential variable $\al$ that is only visible to the judgment $[\al/a]A \le B$.
+Then reduction continues until all the subtyping judgments after $\al$ are satisfied.
+Finally we can safely remove $\al$ since no occurrence of $\al$ could leak to its left in the worklist.
+On the other hand, removing $\al$ according to the algorithm carbage-collects the variable at the perfect time: it leaves the environment immediately after unreferenced.
+
+The design of our judgment chain is closely related to the property of the judgments.
+Subtyping and checking does not return anything, so variables cannot leak anyway, as applied to Rules 8 and 9.
+However, inference and application inference may return a type that contains new variables.
+Take Rule 26 as an example, if it simply returns $\al \to \bt$, passes that to the next judgment and continues the type-checking process,
+variables $\al$ and $\bt$ in the next judgment are out of scope
+and will break the well-formedness of the worklist.
+By making use of the continuation-passing-style judgement chain,
+inner judgments always shares context with its parent judgment,
+therefore we preserve well-formedness during reduction.

### 2019-01-21 sections/introduction.tex
Sections and Skeleton for Overview
sections; add algo. rules; few text on algo
Intro sketch

@@
 
-Most statically typed functional languages support a form of
-\emph{(implicit) parametric polymorphism}~\cite{reynolds1983types}.
+Modern functional programming languages, such as Haskell or OCaml,
+use sophisticated forms of type inference.
+
+However, while being an
+important topic in the Programming Languages research, there is little
+work on the mechanization of the metatheory of type inference on
+theorem provers. In particular we are unaware of any complete
+formalization of the type inference algorithms that are the backbone of modern 
+functional languages.
+
+This paper presents the first full mechanical formalization of the metatheory for
+higher-ranked polymorphic type inference. The system that we formalize
+is the bidirectional type system by Dunfield and
+Krishnaswami (DK). The DK type system has two variants (a declarative
+and an algorithmic one) that have been been \emph{manually} proved to be
+\emph{sound}, \emph{complete} and \emph{decidable}.
+
+We present 
+a mechanical formalization in the Abella theorem prover 
+for DK's declarative type system using a different algorithm. We use a
+new algorithm for a few different reasons. Firstly, our new algorithm
+employs \emph{worklist judgements}, which allow precise capture of the
+scope of vairables and simplify the formalization of scoping in a
+theorem prover. Secondly, while DK's original formalization comes with very
+well-written manual proofs, there are several details missing,
+which complicate the task of writing a formla proof. 
+Perhaps more importantly, some proofs/lemmas are wrong and it is
+not clear how to replace these by other valid proofs/lemmas. 
+\bruno{others}
+
+Despite the use of a different algorithm we prove the
+same results by DK, although with significantly different proofs and
+proof techniques.
+
+The calculus and its metatheory
+have been fully formalized in the Abella theorem prover\cite{AbellaDesc}. % Add at least one ref to make compile work normally
+
+In summary, the contributions of this paper are:
+
+\begin{itemize}
+
+\item {\bf A full mechanical formalization of type inference with
+    higher-ranked types:}
+
+\item {\bf A new algorithm for DK:}
+
+\item {\bf Worlists with inference judgments:}
+
+\item {\bf Abella formalization:} The calculus and its metatheory
+have been fully formalized in the Abella theorem prover\cite{AbellaDesc}.
+
+\end{itemize}

### 2019-01-21 sections/Metatheory.tex
Metatheory: lemmas and theorems.

@@
 
+\begin{lemma}[Insert]
+If $\Gm_L\vdash \tau$ and $\Gm_L, [\tau/\al]\Gm_R \sto \Om$
+, then $\Gm_L, \al, \Gm_R \sto \Om$.
+\end{lemma}
+\begin{lemma}[Extract]
+If $\Gm_L, \al, \Gm_R \sto \Om$
+, then $\exists \tau$ s.t. $\Gm\vdash\tau$ and $\Gm_L, [\tau/\al]\Gm_R \sto \Om$.
+\end{lemma}
+
 \subsection{Soundness}
+
+\begin{theorem}[Soundness]
+If \emph{wf }$\Gm$ and $\Gm \redto \cdot$, then $\exists \Om$ s.t. $\Gm\sto\Om$ and $\Om\redto\cdot$.
+\end{theorem}
+
 \subsection{Completeness}
+
+\begin{theorem}[Completeness]
+If \emph{wf }$\Gm$ and $\exists \Om$ s.t. $\Gm\sto\Om$ and $\Om\redto\cdot$, then $\Gm \redto \cdot$.
+\end{theorem}
+
+\begin{lemma}[Prune Transfer for Instantiation]
+If $(\Gm \vdash \al \le A \to B) \sto (\Om \vdash C \le A_1 \to B_1)$ and $\Om \vdash_d C \le A_1 \to B_1$,
+then $\al\notin FV(A) \cup FV(B)$.
+\end{lemma}
+
 \subsection{Decidability}
+\begin{theorem}[Decidability]
+For all well-formed $\Gm$, it is decidable whether $\Gm\redto\cdot$ or not.
+\end{theorem}
+
+lexicographic group of induction measurements
+$\langle |\Gm|_e, |\Gm|_\Leftrightarrow, |\Gm|_\forall, |\Gm|_{\al}, |\Gm|_\to + |\Gm| \rangle$
+
+\begin{definition}[Worklist Measures]
+\[|\Gm|_e = \sum_{e\Lto A\in\Gm}|e|_e + \sum_{e\To_a \jg\in\Gm}|e|_e + \sum_{\appInfAlg{A}{e}\in\Gm}|e|_e\]
+\[|\Gm|_\Leftrightarrow = \#_{e\Lto A\in\Gm}
+    + 2\cdot\#_{e\To_a \jg\in\Gm} + 3\cdot\#_{\appInfAlg{A}{e}\in\Gm}\]
+\[|\Gm|_\forall = \sum_{e\Lto A\in\Gm}|A|_\forall + \sum_{\appInfAlg{A}{e}\in\Gm}|A|_\forall +
+    \sum_{A\le B\in\Gm} |A|_\forall + |B|_\forall\]
+\[|\Gm|_\to = \sum_{e\Lto A\in\Gm}|A|_\to + \sum_{\appInfAlg{A}{e}\in\Gm}|A|_\to +
+    \sum_{A\le B\in\Gm} |A|_\to + |B|_\to\]
+\[|\Gm|_{\al} = \#_{\al\in\Gm}\]
+\end{definition}
 

### 2019-01-21 main.tex
Sections and Skeleton for Overview
sections; add algo. rules; few text on algo

@@
 %% For double-blind review submission, w/o CCS and ACM Reference (max submission space)
-%\documentclass[acmsmall,review,anonymous]{acmart}\settopmatter{printfolios=true,printccs=false,printacmref=false}
+\documentclass[acmsmall,review,anonymous]{acmart}\settopmatter{printfolios=true,printccs=false,printacmref=false}
 %% For double-blind review submission, w/ CCS and ACM Reference
@@
 %% For final camera-ready submission, w/ required CCS and ACM Reference
-\documentclass[acmsmall]{acmart}\settopmatter{}
+%\documentclass[acmsmall]{acmart}\settopmatter{}
 
@@
 %% Title information
-\title[Short Title]{Full Title}         %% [Short Title] is optional;
+\title{A Mechanical Formalization of Higher-Ranked Polymorphic
+  Type Inference}         %% [Short Title] is optional;
                                         %% when present, will be used in
@@
 \begin{abstract}
-Text of abstract \ldots.
+Modern functional programming languages, such as Haskell or OCaml,
+use sophisticated forms of type inference. However, while being an
+important topic in the Programming Languages research, there is little
+work on the mechanization of the metatheory of type inference on
+theorem provers. In particular we are unaware of any complete
+formalization of the type inference algorithms that are the backbone of modern 
+functional languages.
+
+This paper presents the first full mechanical formalization of the metatheory for
+higher-ranked polymorphic type inference. The system that we formalize
+is the bidirectional type system by Dunfield and
+Krishnaswami (DK). The DK type system has two variants (a declarative
+and an algorithmic one) that have been been \emph{manually} proved to be
+\emph{sound}, \emph{complete} and \emph{decidable}. We present 
+a mechanical formalization in the Abella theorem prover 
+for DK's declarative type system using a different algorithm. We use a
+new algorithm for a few different reasons. Firstly, our new algorithm
+employs \emph{worklist judgements}, which allow precise capture of the
+scope of vairables and simplify the formalization of scoping in a
+theorem prover. Secondly, while DK's original formalization comes with very
+well-written manual proofs, there are several details missing,
+which complicate the task of writing a formla proof. 
+Perhaps more importantly, some proofs/lemmas are wrong and it is
+not clear how to replace these by other valid proofs/lemmas. 
+\bruno{others} Despite the use of a different algorithm we prove the
+same results by DK, although with significantly different proofs and
+proof techniques.  
 \end{abstract}
@@
 \input{sections/introduction.tex}
-
+\input{sections/Overview.tex}
+%\input{sections/Declarative.tex}
 \input{sections/algorithm.tex}
+\input{sections/Metatheory.tex}
+\input{sections/Discussion.tex}
+\input{sections/Related.tex}
+\input{sections/conclusion.tex}
 

### 2019-01-21 sections/Overview.tex

\section{Overview}

\bruno{I think some text/ideas-for-text can be borrowed from ITP.}

\subsection{DK's Declarative System}
Subtyping and Typing. \jimmy{Can you prepare the figures. Subtyping is already in ITP.}

\paragraph{Metatheory} Talk about the lemmas that are false, show counterexample.

\paragraph{Other Challenges} Scoping, Output Contexts?

\subsection{Small-Step Unification} Mention some algorithms for unification for 
dependent types that use a small-step approach. Credit them later for some ideas that 
we also employ.

Some issues to point out: Duplicated contexts (rather than shared contexts), which make 
the formalization harder since requires ``synchronizing'' things ...

\subsection{Worklist Algorithm for Subtyping}

Revise the approach in ITP.

\paragraph{Scoping Garbage} Scoping problems (extra garbage).

\paragraph{No Inference Judgments} 
Only subtyping (which is a \emph{checking} judgment). 
No approach to deal with \emph{inference} judgments.

\subsection{Algorithmic Type Inference for Higher-Ranked Types: Key Ideas}

Our algorithm borrows some ideas from previous work, while adding new ones. 
A small-step style processing worklists; \emph{Judgment Chains}; others.

### 2019-01-21 main.tex

@@
 Modern functional programming languages, such as Haskell or OCaml,
-use sophisticated forms of type inference. However, while being an
+use sophisticated forms of type inference. Yet, while it is an
 important topic in the Programming Languages research, there is little
-work on the mechanization of the metatheory of type inference on
+work on the mechanization of the metatheory of type inference in
 theorem provers. In particular we are unaware of any complete
@@
 Krishnaswami (DK). The DK type system has two variants (a declarative
-and an algorithmic one) that have been been \emph{manually} proved to be
+and an algorithmic one) that have been \emph{manually} proven
 \emph{sound}, \emph{complete} and \emph{decidable}. We present 
 a mechanical formalization in the Abella theorem prover 
-for DK's declarative type system using a different algorithm. We use a
-new algorithm for a few different reasons. Firstly, our new algorithm
+of DK's declarative type system with a novel algorithmic system. We have
+several reasons to use a new algorithm. Firstly, our new algorithm
 employs \emph{worklist judgements}, which allow precise capture of the
-scope of vairables and simplify the formalization of scoping in a
+scope of variables and simplify the formalization of scoping in a
 theorem prover. Secondly, while DK's original formalization comes with very
 well-written manual proofs, there are several details missing,
-which complicate the task of writing a formla proof. 
+which complicate the task of writing a formal proof. 
 Perhaps more importantly, some proofs/lemmas are wrong and it is
@@
 \bruno{others} Despite the use of a different algorithm we prove the
-same results by DK, although with significantly different proofs and
+same results as DK, although with significantly different proofs and
 proof techniques.  

### 2019-01-21 sections/Overview.tex
Skeleton for Section 2
discussion of our ITP paper in the Overview section

@@
 
+%-------------------------------------------------------------------------------
 \subsection{Worklist Algorithm for Subtyping}
 
-Revise the approach in ITP.
+In recent work~\cite{itp2018}, we have shown how to mechanise a variant of DK's
+subtyping algorithm and shown it correct with respect to DK's declarative
+subtyping judgment. This approach overcomes the problem's in DK's formulation
+by using a \emph{worklist}-based judgement of the form $\Gamma \vdash \Omega$
+where $\Omega$ is a worklist, i.e. a sequence, of subtyping problems of the
+form $A \leq B$.  The judgement is defined by case analysis on the first
+element of $\Omega$ and recursively processes the worklist until it is empty.
+Using both a single common environment $\Gamma$ and a worklist $\Omega$ greatly
+simplifies propagating the instantiation of type variables that happen in one
+subtyping problem to the remaining ones.
 
-\paragraph{Scoping Garbage} Scoping problems (extra garbage).
+Unfortunately, this work does not solve all problems. In particular, it has two
+major limitations that prevent it from scaling to the whole DK system.
+
+\paragraph{Scoping Garbage} Firstly, the single common type environment
+$\Gamma$ does not accurately reflect the type and unification variables
+currently in scope. Instead, it is an overapproximation that steadily accrues
+variables, and only drops those unification variables that are instantiated.
+In other words, $\Gamma$ contains ``garbage'' that is no longer in scope.
+This complicates establishing the relation with the declarative system.
 
 \paragraph{No Inference Judgments} 
-Only subtyping (which is a \emph{checking} judgment). 
-No approach to deal with \emph{inference} judgments.
+Secondly and more importantly, the approach only deals with a judgement for
+\emph{checking} whether one type is the subtype of another. While this may not
+so be difficult to scale to the \emph{checking} mode of term typing $\Gamma
+\vdash e \Leftarrow A$, it clearly lacks the functionality to support the
+\emph{inference} mode of term typing $\Gamma \vdash e \Rightarrow A$. Indeed,
+the latter requires not only the communication of unification variable
+instantiations from one typing problem to another, but also an inferred type.
 
+%-------------------------------------------------------------------------------
 \subsection{Algorithmic Type Inference for Higher-Ranked Types: Key Ideas}

### 2019-01-21 sections/Overview.tex
overview: new worklist + chains

@@
 element of $\Omega$ and recursively processes the worklist until it is empty.
-Using both a single common environment $\Gamma$ and a worklist $\Omega$ greatly
+Using both a single common ordered context $\Gamma$ and a worklist $\Omega$ greatly
 simplifies propagating the instantiation of type variables that happen in one
@@
 
-\paragraph{Scoping Garbage} Firstly, the single common type environment
+\paragraph{Scoping Garbage} Firstly, the single common type ordered context 
 $\Gamma$ does not accurately reflect the type and unification variables
@@
 
-Our algorithm borrows some ideas from previous work, while adding new ones. 
-A small-step style processing worklists; \emph{Judgment Chains}; others.
+Our new algorithmic type system addresses builds on the work above, but
+addresses the open problems.
+
+\paragraph{Scope Tracking}
+Firstly, we avoid scoping garbage by blending the ordered context and the
+worklist into a single syntactic sort $\Gamma$, our algorithmic worklist. This
+algorithmic worklist interleaves (type and term) variables with \emph{work}
+like checking or inferring types of expressions. The interleaving keeps track
+of the variable scopes in the usual, natural way: each variables is in scope of
+anything in front of it in the worklist. If there is nothing in front, the
+variable is no longer needed and can be popped from the worklist. This way, no
+garbage builds up.
+
+\paragraph{Inference Judgements}
+To express the DK inference judgements, we use a novel form of work entries in
+the worklist: our algorithmic judgement chains $\omega$. In its simplest form,
+such a judgement chain is just a check, like a subtyping check $A \leq B$ or a
+term typecheck $e \Leftarrow A$.  The non-trivial forms of chains capture an
+inference task together with the the work that depends on its outcome. For
+instance, a type inference task takes the form $e \Rightarrow_a \omega$. It
+denotes that we need to infer the type, say $A$, of expression $e$ and use it
+in the chain $\omega$ by subsituting it for the placeholder type variable $a$.
+
+% Our algorithm borrows some ideas from previous work, while adding new ones. 
+% A small-step style processing worklists; \emph{Judgment Chains}; others.
+
+% Importantly we now deal with inference judgement.

### 2019-01-22 sections/Metatheory.tex
Decl. subtyping figure; Vdash symbol; transfer rules

@@
 \jimmy{The transfer rule outputs declarative judgment chains, which are later processed by another declarative relation, handling all the guessing.}
+
+The transfer rules are shown in Figure~\ref{fig:trans}.
+The rules recursively replace the first existential variable to a well-scoped mono-type,
+so the algorithmic worklist will finally become a declarative one.
+In order to maintain well-formedness,
+the substitution should affect all the judgments and term variable bindings in the worklist.
 
@@
 
+\begin{figure}[t]
+\framebox{$\Om \rto \Om'$} $\Om$ reduces to $\Om'$.
+\begin{gather*}
+\begin{aligned}
+\Om,a &\rto \Om \qquad \Om,\al \rto \Om \qquad \Om,x:A \rto \Om\\
+\Om\Vdash A\le B &\rto \Om \text{ when } \Om\vdash A\le B\\
+\Om\Vdash e\Lto A &\rto \Om \text{ when } \Om\vdash e\Lto A\\
+\Om\Vdash e\To_a \jg &\rto \Om\Vdash[A/a]\jg \text{ when } \Om\vdash e\To A\\
+\Om\Vdash \appInfAlg{A}{e} &\rto \Om\Vdash[C/a]\jg \text{ when } \Om\vdash \appInf{A}{e}{C}\\
+\end{aligned}
+\end{gather*}
+\caption{Algorithmic Worklist Typing Rules}
+\label{fig:alg:typing}
+\end{figure}
+
 \begin{lemma}[Insert]
@@
 \begin{theorem}[Soundness]
-If \emph{wf }$\Gm$ and $\Gm \redto \cdot$, then $\exists \Om$ s.t. $\Gm\sto\Om$ and $\Om\redto\cdot$.
+If \emph{wf }$\Gm$ and $\Gm \redto \nil$, then $\exists \Om$ s.t. $\Gm\sto\Om$ and $\Om\redto\nil$.
 \end{theorem}
@@
 \begin{theorem}[Completeness]
-If \emph{wf }$\Gm$ and $\exists \Om$ s.t. $\Gm\sto\Om$ and $\Om\redto\cdot$, then $\Gm \redto \cdot$.
+If \emph{wf }$\Gm$ and $\exists \Om$ s.t. $\Gm\sto\Om$ and $\Om\redto\nil$, then $\Gm \redto \nil$.
 \end{theorem}
@@
 \begin{lemma}[Prune Transfer for Instantiation]
-If $(\Gm \vdash \al \le A \to B) \sto (\Om \vdash C \le A_1 \to B_1)$ and $\Om \vdash_d C \le A_1 \to B_1$,
+If $(\Gm \Vdash \al \le A \to B) \sto (\Om \Vdash C \le A_1 \to B_1)$ and $\Om \vdash C \le A_1 \to B_1$,
 then $\al\notin FV(A) \cup FV(B)$.
@@
 \begin{theorem}[Decidability]
-For all well-formed $\Gm$, it is decidable whether $\Gm\redto\cdot$ or not.
+For all well-formed $\Gm$, it is decidable whether $\Gm\redto\nil$ or not.
 \end{theorem}
@@
 \[|\Gm|_\Leftrightarrow = \#_{e\Lto A\in\Gm}
-    + 2\cdot\#_{e\To_a \jg\in\Gm} + 3\cdot\#_{\appInfAlg{A}{e}\in\Gm}\]
+    + 2\nil\#_{e\To_a \jg\in\Gm} + 3\nil\#_{\appInfAlg{A}{e}\in\Gm}\]
 \[|\Gm|_\forall = \sum_{e\Lto A\in\Gm}|A|_\forall + \sum_{\appInfAlg{A}{e}\in\Gm}|A|_\forall +

### 2019-01-22 sections/Overview.tex
Decl. subtyping figure; Vdash symbol; transfer rules

@@
 Subtyping and Typing. \jimmy{Can you prepare the figures. Subtyping is already in ITP.}
+
+\begin{figure}[t]
+\[
+\begin{array}{l@{\qquad}lcl}
+\text{Type variables}\qquad&a, b\\[3mm]
+\text{Types}\qquad&A, B, C &::=&\quad 1 \mid a \mid \forall a. A \mid A\to B\\
+\text{Monotypes}\qquad&\tau &::=&\quad 1 \mid a \mid \tau_1\to \tau_2\\
+\text{Contexts}\qquad&\Psi &::=&\quad \cdot \mid \Psi, a\\
+%\text{Judgments}\qquad&\exps &::=&\quad \cdot \mid A \le B : \exps
+\end{array}
+\]
+\caption{Syntax of Declarative System}\label{fig:decl:syntax}
+\end{figure}
+
+The syntax of this declarative system is shown in Figure~\ref{fig:decl:syntax}.
+
+\begin{figure}[t]
+\centering \framebox{$\Om \vdash A$}
+\begin{gather*}
+\inferrule*[right=$\mathtt{wf_d unit}$]
+    {~}{\Om\vdash 1}
+\qquad
+\inferrule*[right=$\mathtt{wf_d var}$]
+    {a\in\Om}{\Om\vdash a}
+\qquad
+\inferrule*[right=$\mathtt{wf_d{\to}}$]
+    {\Om\vdash A\quad \Om\vdash B}
+    {\Om\vdash A\to B}
+\qquad
+\inferrule*[right=$\mathtt{wf_d\forall}$]
+    {\Om, a\vdash A}
+    {\Om\vdash \forall a. A}
+\end{gather*}
+
+\centering \framebox{$\Om \vdash A \le B$}
+\begin{gather*}
+\inferrule*[right=$\mathtt{{\le}Var}$]
+    {a\in\Om}{\Om\vdash a\le a}
+\qquad
+\inferrule*[right=$\mathtt{{\le}Unit}$]
+    {~}{\Om \vdash 1 \le 1}
+\qquad
+\inferrule*[right=$\mathtt{{\le}{\to}}$]
+    {\Om \vdash B_1 \le A_1 \quad \Om \vdash A_2 \le B_2}
+    {\Om\vdash A_1\to A_2 \le B_1\to B_2}
+\\
+\inferrule*[right=$\mathtt{{\le}\forall L}$]
+    {\Om\vdash \tau \quad \Om\vdash [\tau/a] A \le B}
+    {\Om\vdash \forall a. A \le B}
+\qquad
+\inferrule*[right=$\mathtt{{\le}\forall R}$]
+    {\Om, a\vdash A\le B}
+    {\Om\vdash A \le \forall a. B}
+\end{gather*}
+\caption{Well-formedness of Declarative Types and Declarative Subtyping}\label{fig:decl:wf_sub}
+\end{figure}
+
+In Figure~\ref{fig:decl:wf_sub},
+we give the well-formedness and subtyping relation for the declarative system.
 

### 2019-01-22 sections/Overview.tex
overview: other problems with DK's algorithm

@@
 
-\paragraph{Other Challenges} Scoping, Output Contexts?
+%- - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - 
+\paragraph{Complex Scoping and Propagation}
 
+Besides the problematic lemmas in the DK metatheory, there are other reasons to
+look for an alternative algorithmic formulation of the type system that is more
+amenable to mechanisation. Indeed, two aspects that are particularly
+challenging to mechanise are the scoping of universal and existential type
+variables, and the propagation of the instantiation of existential type
+variables across judgments.
+
+DK is already quite disciplined on these accounts, in particular compared to
+traditional constraint-based type-inference algorithms like Algorithm W which
+features an implicit global scope for all type variables. Indeed, DK uses an
+explicit and ordered context that tracks the relative scope of universal and
+existential variables and it is careful to only instantiate existential
+variables in a well-scoped manner.
+
+Moreover, DK's algorithm carefully propagates instantiations by recording them
+into the context and threading this context through all judgments.  means of
+two This means that every judgments takes two contexts---an input and an output
+context---rather than the conventional single context. The output context
+records any new variable instantations; to propagate these instantiations to
+the remaining judgements, their predecessor's output context---which is their
+input context---is applied to them as a substitution.
+
+While it works, DK's approach is still fairly involved and thus hard to reason
+about in a mechanised setting. Indeed, the instantiations have to be recorded
+in the context and are applied incrementally to each remaining judgment in
+turn, rather than immediately to all remaining judgments at once. Also the
+stack discipline of the ordered contexts does not mesh well with the use of
+output contexts; explicit marker entries are needed in two places to
+demarcate the end of an existential variable's scope. These are compelling 
+reasons to look for a simpler algorithm that is easier to reason about in a 
+mechanised setting.
+
+%-------------------------------------------------------------------------------
 \subsection{Small-Step Unification} Mention some algorithms for unification for 

### 2019-01-24 sections/Overview.tex
Declarative typing rules

@@
 \text{Types}\qquad&A, B, C &::=&\quad 1 \mid a \mid \forall a. A \mid A\to B\\
-\text{Monotypes}\qquad&\tau &::=&\quad 1 \mid a \mid \tau_1\to \tau_2\\
+\text{Monotypes}\qquad&\tau,\sigma &::=&\quad 1 \mid a \mid \tau\to \sigma\\
 \text{Contexts}\qquad&\Psi &::=&\quad \cdot \mid \Psi, a\\
@@
 \begin{figure}[t]
-\centering \framebox{$\Om \vdash A$}
-\begin{gather*}
-\inferrule*[right=$\mathtt{wf_d unit}$]
-    {~}{\Om\vdash 1}
-\qquad
-\inferrule*[right=$\mathtt{wf_d var}$]
-    {a\in\Om}{\Om\vdash a}
-\qquad
-\inferrule*[right=$\mathtt{wf_d{\to}}$]
-    {\Om\vdash A\quad \Om\vdash B}
-    {\Om\vdash A\to B}
-\qquad
-\inferrule*[right=$\mathtt{wf_d\forall}$]
-    {\Om, a\vdash A}
-    {\Om\vdash \forall a. A}
-\end{gather*}
+%\centering \framebox{$\Psi \vdash A$}
+%\begin{gather*}
+%\inferrule*[right=$\mathtt{wf_d unit}$]
+%    {~}{\Psi\vdash 1}
+%\qquad
+%\inferrule*[right=$\mathtt{wf_d var}$]
+%    {a\in\Psi}{\Psi\vdash a}
+%\qquad
+%\inferrule*[right=$\mathtt{wf_d{\to}}$]
+%    {\Psi\vdash A\quad \Psi\vdash B}
+%    {\Psi\vdash A\to B}
+%\qquad
+%\inferrule*[right=$\mathtt{wf_d\forall}$]
+%    {\Psi, a\vdash A}
+%    {\Psi\vdash \forall a. A}
+%\end{gather*}
 
-\centering \framebox{$\Om \vdash A \le B$}
+\centering \framebox{$\Psi \vdash A \le B$}
 \begin{gather*}
 \inferrule*[right=$\mathtt{{\le}Var}$]
-    {a\in\Om}{\Om\vdash a\le a}
+    {a\in\Psi}{\Psi\vdash a\le a}
 \qquad
 \inferrule*[right=$\mathtt{{\le}Unit}$]
-    {~}{\Om \vdash 1 \le 1}
+    {~}{\Psi \vdash 1 \le 1}
 \qquad
 \inferrule*[right=$\mathtt{{\le}{\to}}$]
-    {\Om \vdash B_1 \le A_1 \quad \Om \vdash A_2 \le B_2}
-    {\Om\vdash A_1\to A_2 \le B_1\to B_2}
+    {\Psi \vdash B_1 \le A_1 \quad \Psi \vdash A_2 \le B_2}
+    {\Psi\vdash A_1\to A_2 \le B_1\to B_2}
 \\
 \inferrule*[right=$\mathtt{{\le}\forall L}$]
-    {\Om\vdash \tau \quad \Om\vdash [\tau/a] A \le B}
-    {\Om\vdash \forall a. A \le B}
+    {\Psi\vdash \tau \quad \Psi\vdash [\tau/a] A \le B}
+    {\Psi\vdash \forall a. A \le B}
 \qquad
 \inferrule*[right=$\mathtt{{\le}\forall R}$]
-    {\Om, a\vdash A\le B}
-    {\Om\vdash A \le \forall a. B}
+    {\Psi, a\vdash A\le B}
+    {\Psi\vdash A \le \forall a. B}
 \end{gather*}
-\caption{Well-formedness of Declarative Types and Declarative Subtyping}\label{fig:decl:wf_sub}
+\caption{%Well-formedness of Declarative Types and 
+Declarative Subtyping}\label{fig:decl:sub}
 \end{figure}
 
-In Figure~\ref{fig:decl:wf_sub},
-we give the well-formedness and subtyping relation for the declarative system.
+\begin{figure}[t]
+\centering \framebox{$\Psi \vdash A \Lto B$} $e$ checks against input type $A$.\\
+\centering \framebox{$\Psi \vdash A \To B$} $e$ synthesizes output type $A$.\\
+\centering \framebox{$\Psi \vdash \appInf{A}{e}{C}$} Applying a function of type $A$ to $e$ synthesizes type $C$.
+\begin{gather*}
+\inferrule*[right=$\mathtt{DeclVar}$]
+    {(x:A)\in\Psi}{\Psi\vdash x\To A}
+\qquad
+\inferrule*[right=$\mathtt{DeclSub}$]
+    {\Psi\vdash e\To A \quad \Psi\vdash A\le B}{\Psi \vdash e\Lto B}
+\qquad
+\inferrule*[right=$\mathtt{DeclAnno}$]
+    {\Psi \vdash A \quad \Psi\vdash e\Lto A}
+    {\Psi\vdash (e:A)\To A}
+\\
+\inferrule*[right=$\mathtt{Decl1I}$]
+    {~}{\Psi\vdash () \Lto 1}
+\qquad
+\inferrule*[right=$\mathtt{Decl1I\To}$]
+    {~}{\Psi\vdash () \To 1}
+\qquad
+\inferrule*[right=$\mathtt{Decl\forall I}$]
+    {\Psi,a \vdash e \Lto A}
+    {\Psi\vdash e\Lto \all A}
+\\
+\inferrule*[right=$\mathtt{Decl\forall App}$]
+    {\Psi \vdash \tau \quad \Psi\vdash \appInf{[\tau/a]A}{e}{C} }
+    {\Psi\vdash \appInf{\all A}{e}{C}}
+\qquad
+\inferrule*[right=$\mathtt{Decl\to I}$]
+    {\Psi,x:A \vdash e\Lto B}
+    {\Psi\vdash \lam e \Lto A \to B}
+\\
+\inferrule*[right=$\mathtt{Decl\to I\To}$]
+    {\Psi\vdash \sigma\to\tau \quad \Psi,x:\sigma \vdash e\Lto \tau}
+    {\Psi\vdash \lam e \To \sigma\to\tau}
+\qquad
+\inferrule*[right=$\mathtt{Decl\to E}$]
+    {\Psi\vdash e_1\To A \quad \Psi\vdash \appInf{A}{e_2}{C}}
+    {\Psi\vdash e_1~e_2 \To C}
+\\
+\inferrule*[right=$\mathtt{Decl\to App}$]
+    {\Psi\vdash e \Lto A}
+    {\Psi\vdash \appInf{A \to C}{e}{C}}
+\end{gather*}
+\caption{Declarative Typing}\label{fig:decl:typing}
+\end{figure}
+
+In Figures~\ref{fig:decl:sub} and \ref{fig:decl:typing},
+we give the subtyping and typing relations for the declarative system.
 

### 2019-01-25 sections/Metatheory.tex
Symbol changes; worklist extension relation

@@
 \begin{definition}[Worklist Measures]
-\[|\Gm|_e = \sum_{e\Lto A\in\Gm}|e|_e + \sum_{e\To_a \jg\in\Gm}|e|_e + \sum_{\appInfAlg{A}{e}\in\Gm}|e|_e\]
-\[|\Gm|_\Leftrightarrow = \#_{e\Lto A\in\Gm}
-    + 2\nil\#_{e\To_a \jg\in\Gm} + 3\nil\#_{\appInfAlg{A}{e}\in\Gm}\]
-\[|\Gm|_\forall = \sum_{e\Lto A\in\Gm}|A|_\forall + \sum_{\appInfAlg{A}{e}\in\Gm}|A|_\forall +
-    \sum_{A\le B\in\Gm} |A|_\forall + |B|_\forall\]
-\[|\Gm|_\to = \sum_{e\Lto A\in\Gm}|A|_\to + \sum_{\appInfAlg{A}{e}\in\Gm}|A|_\to +
-    \sum_{A\le B\in\Gm} |A|_\to + |B|_\to\]
-\[|\Gm|_{\al} = \#_{\al\in\Gm}\]
+\begin{gather*}
+\begin{aligned}
+|\Gm|_e &= \textstyle\sum_{e\Lto A\in\Gm}|e|_e + \sum_{e\To_a \jg\in\Gm}|e|_e +
+    \sum_{\appInfAlg{A}{e}\in\Gm}|e|_e\\
+|\Gm|_\Leftrightarrow &= \#_{e\Lto A\in\Gm} +
+    2\nil\#_{e\To_a \jg\in\Gm} + 3\nil\#_{\appInfAlg{A}{e}\in\Gm}\\
+|\Gm|_\forall &= \textstyle\sum_{e\Lto A\in\Gm}|A|_\forall + \sum_{\appInfAlg{A}{e}\in\Gm}|A|_\forall +
+    \sum_{A\le B\in\Gm} |A|_\forall + |B|_\forall\\
+|\Gm|_\to &= \textstyle\sum_{e\Lto A\in\Gm}|A|_\to + \sum_{\appInfAlg{A}{e}\in\Gm}|A|_\to +
+    \sum_{A\le B\in\Gm} |A|_\to + |B|_\to\\
+|\Gm|_{\al} &= \#_{\al\in\Gm}\\
+\end{aligned}
+\end{gather*}
 \end{definition}
 
+\begin{figure}
+\framebox{$\Gm\jExt\Gm'$}
+\begin{gather*}
+\inferrule*[right=$\mathtt{\jExt\nil}$]
+    {~}{\Gm\jExt\Gm}
+\qquad
+\inferrule*[right=$\mathtt{\jExt solve}$]
+    {|A|_\forall = 0 \quad \Gm_L,[A/\al]\Gm_R \jExt \Gm'}{\Gm_L,\al,\Gm_R \jExt \Gm'}
+\\
+\inferrule*[right=$\mathtt{\jExt \al}$]
+    {\Gm_L,\al,\Gm_R \jExt \Gm'}{\Gm_L,\Gm_R \jExt \Gm'}
+\qquad
+\inferrule*[right=$\mathtt{\jExt inst}$]
+    {\Gm_L,\al[1],\al[2],[\al[1]\to\al[2]/\al]\Gm_R \jExt \Gm'}{\Gm_L,\al,\Gm_R \jExt \Gm'}
+\end{gather*}
+\caption{Worklist Extension}\label{fig:worklist_ext}
+\end{figure}
+
+Shown in Figure~\ref{fig:worklist_ext},
+worklist extension relates minor changes of worklist during type-checking.
+When the algorithm processes the worklist, the first judgment is processed,
+and the rest of the worklist roughly remains its shape.
+Worklist extension formally defines the minor changes of the shape by three cases:
+existential variable solving ($\mathtt{\jExt solve}$), existential variable addition ($\mathtt{\jExt \al}$)
+and existential variable instantiation ($\mathtt{\jExt inst}$). \jimmy{Is the 3rd case redundant?}
+
 \subsection{Abella}

### 2019-01-25 sections/algorithm.tex
Decl. subtyping figure; Vdash symbol; transfer rules
Resolve overlappings by adding restriction to both decl and alg (proven)
Symbol changes; worklist extension relation
Algorithmic rule description

@@
 \text{Mono-types}\qquad&\tau &::=&\quad 1 \mid a \mid \tau_1\to \tau_2\\
-%\text{Context}\qquad&\Psi &::=&\quad \cdot \mid \Psi, a \mid \Psi, x:A' \color{red} \\[2mm]
-% \text{Declarative Judgements}\qquad&\jo &::=&\quad \cdot \mid \P\vdash A' \le B' \jc \jo\\
+%\text{Context}\qquad&\Psi &::=&\quad \nil \mid \Psi, a \mid \Psi, x:A' \color{red} \\[2mm]
+% \text{Declarative Judgements}\qquad&\jo &::=&\quad \nil \mid \P\vdash A' \le B' \jc \jo\\
 % &&&\quad \mid \P\vdash e\Leftarrow A' \jc \jo \mid \P\vdash e\To A' \jc \jo \mid \P\vdash A' \bullet e\TTo C' \jc \jo\\
@@
 \\[2mm]
-\text{Algorithmic worklist}\qquad&\Gm &::=&\quad \cdot \mid \Gm, a \mid \Gm, \al \mid \Gm, x: A \mid \Gm \vdash \jg\\
-\text{Declarative worklist}\qquad&\Om &::=&\quad \cdot \mid \Gm, a \mid \Gm, x: A \mid \Gm \vdash \jg
+\text{Algorithmic worklist}\qquad&\Gm &::=&\quad \nil \mid \Gm, a \mid \Gm, \al \mid \Gm, x: A \mid \Gm \Vdash \jg\\
+\text{Declarative worklist}\qquad&\Om &::=&\quad \nil \mid \Gm, a \mid \Gm, x: A \mid \Gm \Vdash \jg
 \end{aligned}
@@
 \begin{gather*}
-\inferrule*[right=$\mathtt{wf\cdot}$]
-{~}{\text{wf }\cdot}
+\inferrule*[right=$\mathtt{wf\nil}$]
+{~}{\text{wf }\nil}
 \qquad
@@
 {\text{wf }\Gm \\ \Gm\vdash\jg}
-{\text{wf }\Gm \vdash\jg}
+{\text{wf }\Gm \Vdash\jg}
 \end{gather*}
@@
 % TODO rules too long
-\begin{figure}[t]
+\begin{figure}[htp]
 \framebox{$\Gm\rto \Gm'$} $\Gm$ reduces to $\Gm'$.
@@
 \\[2mm]
-\Gm \vdash 1\le 1 &\algrule \Gm\\
-\Gm \vdash a\le a &\algrule \Gm\\
-\Gm \vdash \al\le \al &\algrule \Gm\\
-\Gm \vdash A_1\to A_2 \le B_1\to B_2 &\algrule \Gm \vdash A_2 \le B_2 \vdash B_1\le A_1\\
-\Gm \vdash \forall a. A\le B &\algrule \Gm,\al \vdash [\al/a]A\le B \quad\text{\jimmy{No more exvar leaks}}\\
-\Gm \vdash A\le \forall b. B &\algrule \Gm,b \vdash A\le B \quad\text{\jimmy{No more typevar leaks}}
+\Gm \Vdash 1\le 1 &\algrule \Gm\\
+\Gm \Vdash a\le a &\algrule \Gm\\
+\Gm \Vdash \al\le \al &\algrule \Gm\\
+\Gm \Vdash A_1\to A_2 \le B_1\to B_2 &\algrule \Gm \Vdash A_2 \le B_2 \Vdash B_1\le A_1\\
+\Gm \Vdash \forall a. A\le B &\algrule \Gm,\al \Vdash [\al/a]A\le B \quad\text{\jimmy{No more exvar leaks}}\\
+\Gm \Vdash A\le \forall b. B &\algrule \Gm,b \Vdash A\le B \quad\text{\jimmy{No more typevar leaks}}
 \\[2mm]
 %\\
-%\text{Let } \color{red}\Gm[\al \toto B // G_M] &:= \Gm_L, G_M, [B/\al]\Gm_R, \text{ when } \Gm[\al] = \Gm_L,\al,\Gm_R \ (G_M\text{ defaults to }\cdot)\\
-\Gm[\al] \vdash \al \le A\to B &\algrule [\al[1]\to\al[2]/\al] (\Gm[\al[1], \al[2]] \vdash \al[1]\to \al[2] \le A \to B)\\
+%\text{Let } \color{red}\Gm[\al \toto B // G_M] &:= \Gm_L, G_M, [B/\al]\Gm_R, \text{ when } \Gm[\al] = \Gm_L,\al,\Gm_R \ (G_M\text{ defaults to }\nil)\\
+\Gm[\al] \Vdash \al \le A\to B &\algrule [\al[1]\to\al[2]/\al] (\Gm[\al[1], \al[2]] \Vdash \al[1]\to \al[2] \le A \to B)\\
  &\qquad \text{when }\al\notin FV(A)\cup FV(B)\\
-\Gm[\al] \vdash A\to B \le \al &\algrule [\al[1]\to \al[2]/\al] (\Gm[\al[1], \al[2]] \vdash A \to B \le \al[1]\to \al[2])\\
+\Gm[\al] \Vdash A\to B \le \al &\algrule [\al[1]\to \al[2]/\al] (\Gm[\al[1], \al[2]] \Vdash A \to B \le \al[1]\to \al[2])\\
  &\qquad \text{when }\al\notin FV(A)\cup FV(B)
  \\[2mm]
-\Gm[\al][\bt] \vdash \al \le \bt &\algrule [\al/\bt](\Gm[\al][])\\
-\Gm[\al][\bt] \vdash \bt \le \al &\algrule [\al/\bt](\Gm[\al][])\\
-\Gm[a][\bt] \vdash a \le \bt &\algrule [a/\bt](\Gm[a][])\\
-\Gm[a][\bt] \vdash \bt \le a &\algrule [a/\bt](\Gm[a][])\\
-\Gm[\bt] \vdash 1 \le \bt &\algrule [1/\bt](\Gm[])\\
-\Gm[\bt] \vdash \bt \le 1 &\algrule [1/\bt](\Gm[])
+\Gm[\al][\bt] \Vdash \al \le \bt &\algrule [\al/\bt](\Gm[\al][])\\
+\Gm[\al][\bt] \Vdash \bt \le \al &\algrule [\al/\bt](\Gm[\al][])\\
+\Gm[a][\bt] \Vdash a \le \bt &\algrule [a/\bt](\Gm[a][])\\
+\Gm[a][\bt] \Vdash \bt \le a &\algrule [a/\bt](\Gm[a][])\\
+\Gm[\bt] \Vdash 1 \le \bt &\algrule [1/\bt](\Gm[])\\
+\Gm[\bt] \Vdash \bt \le 1 &\algrule [1/\bt](\Gm[])
 \\[2mm]
-\Gm \vdash e \Lto B &\algrule \Gm \vdash e\To_a a\le B\\
-\Gm \vdash () \Lto 1 &\algrule \Gm\\
-\Gm \vdash e\Lto \forall a. A &\algrule \Gm,a \vdash e\Lto A\\
-\Gm \vdash \lambda x. e \Lto A\to B &\algrule \Gm, x:A  \vdash e \Lto B\\
-\Gm[\al] \vdash \lambda x. e \Lto \al &\algrule [\al[1]\to \al[2] / \al](\Gm[\al[1],\al[2]], x:\al[1] \vdash e \Lto \al[2]) \quad\text{\jimmy{Additional}}
+\Gm \Vdash e \Lto B &\algrule \Gm \Vdash e\To_a a\le B \quad \text{when } e \neq \lam e', B \neq \all B'\\
+% \Gm \Vdash () \Lto 1 &\algrule \Gm\\
+\Gm \Vdash e\Lto \forall a. A &\algrule \Gm,a \Vdash e\Lto A\\
+\Gm \Vdash \lambda x. e \Lto A\to B &\algrule \Gm, x:A  \Vdash e \Lto B\\
+\Gm[\al] \Vdash \lambda x. e \Lto \al &\algrule [\al[1]\to \al[2] / \al](\Gm[\al[1],\al[2]], x:\al[1] \Vdash e \Lto \al[2]) \quad\text{\jimmy{Additional}}
 \\[2mm]
-\Gm \vdash x\To_a \jg &\algrule \Gm \vdash [A/a] \jg \qquad (x:A)\in \Gm\\
-\Gm \vdash (e:A)\To_a \jg &\algrule \Gm \vdash [A/a]\jg \vdash e \Lto A\\
-\Gm \vdash ()\To_a \jg &\algrule \Gm \vdash [1/a]\jg \qquad\jimmy{\downarrow\text{Precise control on exvars}}\\
-\Gm \vdash \lambda x. e \To_a \jg &\algrule
-	\Gm,\al,\bt \vdash [\al\to\bt/a]\jg, x:\al \vdash e\Lto \bt \quad\jimmy{\text{differs}}\\
-\Gm \vdash e_1\ e_2 \To_a \jg &\algrule \Gm \vdash e_1\To_{a_1} (\appInfAlg{a_1}{e_2})
+\Gm \Vdash x\To_a \jg &\algrule \Gm \Vdash [A/a] \jg \quad \text{when } (x:A)\in \Gm\\
+\Gm \Vdash (e:A)\To_a \jg &\algrule \Gm \Vdash [A/a]\jg \Vdash e \Lto A\\
+\Gm \Vdash ()\To_a \jg &\algrule \Gm \Vdash [1/a]\jg \qquad\jimmy{\downarrow\text{Precise control on exvars}}\\
+\Gm \Vdash \lambda x. e \To_a \jg &\algrule
+	\Gm,\al,\bt \Vdash [\al\to\bt/a]\jg, x:\al \Vdash e\Lto \bt \quad\jimmy{\text{differs}}\\
+\Gm \Vdash e_1\ e_2 \To_a \jg &\algrule \Gm \Vdash e_1\To_{a_1} (\appInfAlg{a_1}{e_2})
 \\[2mm]
-\Gm \vdash \appInfAlg{\forall a. A}{e} &\algrule \Gm,\al \vdash \appInfAlg{[\al/a]A}{e}\\
-\Gm \vdash \appInfAlg{A\to C}{e} &\algrule \Gm \vdash [C/a]\jg \vdash e \Lto A\\
-\Gm[\al] \vdash \appInfAlg{\al}{e} &\algrule
-	[\al[1]\to\al[2]/\al](\Gm[\al[1], \al[2]] \vdash [\al[2]/a]\jg \vdash e\Lto \al[1])\\
+\Gm \Vdash \appInfAlg{\forall a. A}{e} &\algrule \Gm,\al \Vdash \appInfAlg{[\al/a]A}{e}\\
+\Gm \Vdash \appInfAlg{A\to C}{e} &\algrule \Gm \Vdash [C/a]\jg \Vdash e \Lto A\\
+\Gm[\al] \Vdash \appInfAlg{\al}{e} &\algrule
+	[\al[1]\to\al[2]/\al](\Gm[\al[1], \al[2]] \Vdash [\al[2]/a]\jg \Vdash e\Lto \al[1])\\
  &\color{magenta} \makebox[0pt]{\qquad or} \phantom{{}\rrule{}{}}
- 	[\al[1]\to\al[2]/\al](\Gm[\al[1], \al[2]] \vdash \appInfAlg{\al[1]\to\al[2]}{e})
+ 	[\al[1]\to\al[2]/\al](\Gm[\al[1], \al[2]] \Vdash \appInfAlg{\al[1]\to\al[2]}{e})
 \end{aligned}
@@
 therefore we preserve well-formedness during reduction.
+
+Secondly, Rules 10 and 11 are algorithmic extensions to Rule $\mathtt{{\le}{\to}}$.
+Both of them partially instantiate $\al$ to function types.
+The domain $\al[1]$ and range $\al[2]$ of the new function type are not determined:
+they are fresh existential variables with the same scope as $\al$.
+We remove $\al$ from the worklist, and insert $\al[1], \al[2]$ back to the same position.
+To propagate the instantiation to the rest of the worklist and maintain the wel-formedness,
+every $\al$ is replaced by $\al[1] \to \al[2]$.
+The \emph{occurs-check} condition is to prevent a diverging infinite loop.
+Without such check, $\al \le 1 \to \al$, as an example, would diverge.
+
+Thirdly, Rules 12-17 are base cases where an existential variable is solved.
+Each of the rules removes and existential variable, substitutes the variable with its solution and continues.
+
+Fourthly, Rules 18-21 deal with checking judgments.
+Rule 18 is similar to $\mathtt{DeclSub}$, and rewrited in the continuation-passing-style.
+Rules 19 and 20 adapts the worklist style from their corresponding declarative rules.
+Rule 21 is a special case of $\mathtt{Decl\to I}$.
+Instantiation technique similar to Rules 10 and 11 applys.
+The declarative checking rule $\mathtt{Decl1I}$ is not translated to the algorithm,
+since combining Rule 18 and Rule 24 would give the same result.
+
+Fifthly, Rules 22-26 deal with inference judgments.
+Inference judgment chains are written in continuation-passing-style.
+When an inference judgment succeeds with type $A$,
+the algorithm continues to work on the inner-chain $\jg$ by assigning $A$ to its placeholder varible $a$.
+Rule 23 infers an annotated expression by changing into checking mode,
+therefore another judgment chain is created.
+Rule 25 infers the type of a lambda expression by introducing $\al, \bt$
+as the input and output types of the function, respectively.
+After checking the body $e$ under the assumption $x:\al$,
+the return type might reflect more information than simply $\al \to \bt$.
+The variable scopings are well-maintained during the process:
+the assumption of argument type ($x:\al$) is recycled after checking against the function body;
+the declaration of the function type ($\al,\bt$) is only visible to
+the checking and the remaining chain $[\al\to\bt/a]\jg$.
+
+Finally, Rules 27-29 deal with application inference judgments.
+Rules 27 and 28 behaves similarly to declarative rules $\mathtt{Decl\forall App}$ and $\mathtt{Decl\to App}$.
+Rule 29 instantiate $\al$ to the function type $\al[1] \to \al[2]$, just like Rules 10, 11 and 21.

### 2019-01-28 sections/introduction.tex
Intro WIP

@@
 Modern functional programming languages, such as Haskell or OCaml,
-use sophisticated forms of type inference.
+use sophisticated forms of type inference. The type systems of these languages are
+descendents of Hindley-Milner~\cite{}, which was revolutionary at the
+time in allowing type-inference to proceed without any type
+annotation. The traditional Hindley-Milner type system supports
+top-level \emph{implicit (parametric) polymorphism}. With implicit
+polymorphism, type arguments of polymorphic functions are
+automatically instantiated. Thus the Hindley-Milner type system 
+strikes a great balance between expressive power and usability.
 
-However, while being an
-important topic in the Programming Languages research, there is little
-work on the mechanization of the metatheory of type inference on
-theorem provers. In particular we are unaware of any complete
-formalization of the type inference algorithms that are the backbone of modern 
-functional languages.
+As functional languages evolved the need for more expressive
+power has motivated language designers to look beyond Hindley-Milner.
+In particular one popular direction has been to allow 
+\emph{higher-ranked polymorphism} where polymorphic types can
+occur anywhere in a type signature.  
+An important challenge is that full type inference for higher-ranked
+polymorphism is know to be undecidable~\cite{}. Therefore some type
+annotations are necessary to guide type inference. In response to this 
+challenge several decidable type systems requiring some annotations 
+have been proposed~\cite{}. Two closely related type systems that 
+support \emph{predicative} higher-ranked type inference were proposed 
+by Peyton Jones et al.~\cite{} and Dunfield and
+Krishnaswami (henceforth denoted as DK). 
+These type systems are
+popular among language designer and their ideas have been adopted by
+several modern functional languages, including Haskell and PureScript.
+In those type systems
+type annotations are required for polymorphic arguments of functions.
+but other type annotations can be omitted. A canonical example is:
+\bruno{the following text needs to be changed, since it is basically
+  as in ITP.}
+\begin{verbatim}
+hpoly = (\f :: forall a. a -> a) -> (f 1, f 'c')
+\end{verbatim}
+The function \verb|hpoly| cannot be
+type-checked in Hindley-Milner.  The type of \verb|hpoly| is 
+\verb|(forall a. a -> a) -> (Int, Char)|. The single universal quantifier
+does not appear at the top-level. Instead it is used to quantify a
+type variable \verb|a| used in the first argument of the
+function.
+Notably \verb|hpoly| requires a type annotation for the first
+argument (\verb|forall a. a -> a|). 
+Despite these additional annotations,
+the type-inference algorithm employed by GHC Haskell~\cite{jones2007practical} preserves 
+many of the desirable properties of Hindley-Milner.
+
+Despite the importance of type inference in practice and also in
+academic research, there is little work in mechanically formalizing
+such advanced forms of type inference in theorem provers. This is at
+odds with a current trend in programming language research. In
+particular both the POPLMark challenge~\cite{} and CompCert ~\cite{}
+have significantly promoted the use of theorem provers to model
+variaous aspects of programming languages. Indeed today there are
+routinely papers in top conferences that use theorem provers to
+mechanically formalize various aspects of programming languages,
+including: \emph{dynamic and static semantics} and their
+correctness properties~\cite{}, \emph{compiler correctness}~\cite{},
+\emph{correctness of optimizations}~\cite{} or even certain forms of
+\emph{abstract interpretation}. Mechanical formalizations have
+numerous advantages~\cite{read to make sure I do not miss any!}.
 

### 2019-01-28 sections/Metatheory.tex
Simplifies worklist extension (same with proof)

@@
     {|A|_\forall = 0 \quad \Gm_L,[A/\al]\Gm_R \jExt \Gm'}{\Gm_L,\al,\Gm_R \jExt \Gm'}
-\\
+\qquad
 \inferrule*[right=$\mathtt{\jExt \al}$]
     {\Gm_L,\al,\Gm_R \jExt \Gm'}{\Gm_L,\Gm_R \jExt \Gm'}
-\qquad
-\inferrule*[right=$\mathtt{\jExt inst}$]
-    {\Gm_L,\al[1],\al[2],[\al[1]\to\al[2]/\al]\Gm_R \jExt \Gm'}{\Gm_L,\al,\Gm_R \jExt \Gm'}
 \end{gather*}
@@
 and the rest of the worklist roughly remains its shape.
-Worklist extension formally defines the minor changes of the shape by three cases:
-existential variable solving ($\mathtt{\jExt solve}$), existential variable addition ($\mathtt{\jExt \al}$)
-and existential variable instantiation ($\mathtt{\jExt inst}$). \jimmy{Is the 3rd case redundant?}
+Worklist extension formally defines the minor changes of the shape by two cases:
+existential variable solving ($\mathtt{\jExt solve}$) and
+existential variable addition ($\mathtt{\jExt \al}$).
+Existential variable instantiation,
+i.e. $\Gm_L,\al,\Gm_R \jExt \Gm_L, \al[1], \al[2], [\al[1] \to \al[2] / \al]\Gm_R$,
+can be derived with the combination of the 3 rules.
 

### 2019-01-30 sections/Overview.tex
Resolve overlappings by adding restriction to both decl and alg (proven)
Declarative overlapping rules

@@
 \inferrule*[right=$\mathtt{DeclSub}$]
-    {\Psi\vdash e\To A \quad \Psi\vdash A\le B}{\Psi \vdash e\Lto B}
-\qquad
+    {e \neq \lam e' \quad B \neq \all B' \quad \Psi\vdash e\To A \quad \Psi\vdash A\le B}
+    {\Psi \vdash e\Lto B}
+\\
 \inferrule*[right=$\mathtt{DeclAnno}$]
@@
     {\Psi\vdash (e:A)\To A}
-\\
+\qquad
 \inferrule*[right=$\mathtt{Decl1I}$]
@@
     {~}{\Psi\vdash () \To 1}
-\qquad
+\\
 \inferrule*[right=$\mathtt{Decl\forall I}$]
@@
     {\Psi\vdash e\Lto \all A}
-\\
+\qquad
 \inferrule*[right=$\mathtt{Decl\forall App}$]
@@
     {\Psi\vdash \appInf{\all A}{e}{C}}
-\qquad
+\\
 \inferrule*[right=$\mathtt{Decl\to I}$]
@@
     {\Psi\vdash \lam e \Lto A \to B}
-\\
+\qquad
 \inferrule*[right=$\mathtt{Decl\to I\To}$]
@@
     {\Psi\vdash \lam e \To \sigma\to\tau}
-\qquad
+\\
 \inferrule*[right=$\mathtt{Decl\to E}$]
@@
     {\Psi\vdash e_1~e_2 \To C}
-\\
+\qquad
 \inferrule*[right=$\mathtt{Decl\to App}$]
@@
 we give the subtyping and typing relations for the declarative system.
+
+\paragraph{Overlapping rules}
+Some of the declarative rules overlaps with each other.
+Declarative subtyping rules $\mathtt{{\le}\forall L}$ and $\mathtt{{\le}\forall R}$
+both match the conclusion $\Psi\vdash \forall a. A \le \forall a. B$.
+In such case, choose $\mathtt{{\le}\forall R}$ first is always better,
+since we introduce the type variable $a$ to the context earlier,
+which gives more flexibility on the choice of $\tau$.
+
+Declarative typing rule $\mathtt{DeclSub}$ overlaps with
+both $\mathtt{Decl\forall I}$ and $\mathtt{Decl\to I}$.
+However, we could argue that more specific rules are always the best choices,
+i.e. $\mathtt{Decl\forall I}$ and $\mathtt{Decl\to I}$ should have
+higher priority to $\mathtt{DeclSub}$.
+For example, $\Psi\vdash \lam x \Lto \all a\to a$ succeeds if derived from
+Rule $\mathtt{Decl\forall I}$, but fails when applied to $\mathtt{DeclSub}$:
+$$
+\inferrule*[right={$\mathtt{Decl\forall I}$}]
+	{\inferrule*[Right={$\mathtt{Decl\to I}$}]
+		{\Psi,a,x:a \vdash x \Lto a}
+		{\Psi,a \vdash \lam x \Lto a \to a}
+	}
+	{\Psi\vdash \lam x \Lto \forall a \to a}
+$$
+$$
+\inferrule*[right={$\mathtt{DeclSub}$}]
+	{
+		\inferrule*[right=$\mathtt{Decl\to I\To}$]
+			{\Psi \vdash \blue\sigma\to \blue\tau \quad \Psi,x:\blue\sigma\vdash e \Lto \blue\tau}
+			{\Psi \vdash \lam x \To \blue\sigma\to \blue\tau}\quad
+		\inferrule*[Right=$\mathtt{{\le}\forall R}$]
+			{\inferrule*[Right=$\mathtt{{\le}{\to}}$]
+				{
+					\inferrule*[Right=$?$]
+						{\text{Impossible! }\blue\sigma \neq a}
+						{\Psi,a \vdash a \le \blue\sigma}
+					\quad \Psi,a \vdash \blue\tau \le a
+				}
+				{\Psi,a \vdash \blue\sigma\to \blue\tau \le a \to a}
+			}
+			{\Psi\vdash \blue\sigma\to \blue\tau\le \all a \to a}
+	}
+{\Psi\vdash \lam x \Lto \forall a \to a}
+$$
+
+Rule $\mathtt{Decl\to I}$ is also better at handling higher-order types.
+When the lambda-expression to be inferred has a polymorphic input type,
+such as $\all a \to a$,
+$\mathtt{DeclSub}$ may not derive some judgments.
+For example, $\Psi,id:\all a\to a \vdash \lam[f] f~id~(f~()) \Lto (\all a\to a) \to 1$
+requires the argument of the lambda-expression to be a polymorphic type,
+otherwise it could not be applied to both $id$ and $()$.
+If Rule $\mathtt{DeclSub}$ was chosen for derivation,
+the type of the argument is restricted by Rule $\mathtt{Decl\to I\To}$,
+which is not a polymorphic type.
+By contrast,
+Rule $\mathtt{Decl\to I}$ keeps the polymorphic argument type $\all a\to a$,
+and will successfully derive the judgment.
 

### 2019-01-30 sections/algorithm.tex
Simplifies worklist extension (same with proof)
Exvar and Hole notation

@@
 
-Figure~\ref{fig:alg:syntax} shows the
-syntax and the well-formedness judgement.
+Figure~\ref{fig:alg:syntax} shows the syntax and the well-formedness judgement.
 
@@
 
-\paragraph{Existential Variables} ...
+\paragraph{Existential Variables} In order to find the unknown type $\tau$ in the declarative system,
+\emph{existential variables} $\al, \bt$ are introduced to the algorithmic system.
+They behave similarly to unification variables,
+but their scoping is restricted by the position located in the worklist,
+rather than globally defined
+Any existential variable $\al$ should only be solved to
+a well-formed type under the worklist before $\al$ is defined.
+Thus the mono-type $\tau$ represented by the corresponding existential variable
+is always well-formed under the corresponding declarative context.
+In other words, the position of $\al$'s reflects the well-formedness restriction of $\tau$'s.
 
@@
 
-\paragraph{Hole Notation}...
+\paragraph{Hole Notation}
+We use the syntax $\Gm[\Gm_M]$ to denote the worklist $\Gm_L,\Gm_M,\Gm_R$,
+where $\Gm$ is the worklist $\Gm_L,\bullet,\Gm_R$ with a hole ($\bullet$).
+Hole notations with the same name implicitly shares the same structure $\Gm_L$ and $\Gm_R$.
+A multi-hole notation splits the worklist into more parts.
+For example, $\Gm[\al][\bt]$ means $\Gm_1,\al,\Gm_2,\bt,\Gm_3$.
 

### 2019-01-31 sections/Metatheory.tex
transfer

@@
 \subsection{Transfer to the Declarative System}
-\jimmy{The transfer rule outputs declarative judgment chains, which are later processed by another declarative relation, handling all the guessing.}
-
-The transfer rules are shown in Figure~\ref{fig:trans}.
-The rules recursively replace the first existential variable to a well-scoped mono-type,
-so the algorithmic worklist will finally become a declarative one.
-In order to maintain well-formedness,
-the substitution should affect all the judgments and term variable bindings in the worklist.
+%\jimmy{The transfer rule outputs declarative judgment chains, which are later processed by another declarative relation, handling all the guessing.}
 
 \begin{figure}[t]
-\framebox{$\Gm \sto \Om$} $\Gm$ transfers to $\Om$. \jimmy{Always substitute from the very first existential variable. Types in contexts ($\Gm,x:A,\ldots$) should be taken care.}
+\framebox{$\Gm \sto \Om$} $\Gm$ transfers to $\Om$.
 \begin{gather*}
@@
 
+The transfer rules are shown in Figure~\ref{fig:trans}.
+Rule $\mathtt{{\sto}\al}$ recursively replace the first existential variable to a well-scoped mono-type
+($\Om$, as a declarative worklist, contains no existential variables),
+so the algorithmic worklist will finally become a declarative one.
+In order to maintain well-formedness,
+the substitution should affect all the judgments and term variable bindings in the scope of $\al$.
+
+The following lemmas~\ref{lem:insert} and \ref{lem:extract}
+generalize Rule $\mathtt{{\sto}\al}$ from substituting the first existential variable
+to substituting any existential variable.
+
+\begin{lemma}[Insert]\label{lem:insert}
+If $\Gm_L\vdash \tau$ and $\Gm_L, [\tau/\al]\Gm_R \sto \Om$
+, then $\Gm_L, \al, \Gm_R \sto \Om$.
+\end{lemma}
+\begin{lemma}[Extract]\label{lem:extract}
+If $\Gm_L, \al, \Gm_R \sto \Om$
+, then $\exists \tau$ s.t. $\Gm_L\vdash\tau$ and $\Gm_L, [\tau/\al]\Gm_R \sto \Om$.
+\end{lemma}
+
 \begin{figure}[t]
@@
 \begin{aligned}
-\Om,a &\rto \Om \qquad \Om,\al \rto \Om \qquad \Om,x:A \rto \Om\\
+\Om,a &\rto \Om \qquad \Om \qquad \Om,x:A \rto \Om\\
 \Om\Vdash A\le B &\rto \Om \text{ when } \Om\vdash A\le B\\
@@
 \end{gather*}
-\caption{Algorithmic Worklist Typing Rules}
-\label{fig:alg:typing}
+\caption{Declarative Worklist Typing}
+\label{fig:decl:worklist}
 \end{figure}
 
-\begin{lemma}[Insert]
-If $\Gm_L\vdash \tau$ and $\Gm_L, [\tau/\al]\Gm_R \sto \Om$
-, then $\Gm_L, \al, \Gm_R \sto \Om$.
-\end{lemma}
-\begin{lemma}[Extract]
-If $\Gm_L, \al, \Gm_R \sto \Om$
-, then $\exists \tau$ s.t. $\Gm\vdash\tau$ and $\Gm_L, [\tau/\al]\Gm_R \sto \Om$.
-\end{lemma}
+Figure~\ref{fig:decl:worklist} relates the worklist-style declarative system
+to the original declarative system.
+A successful reduction means a valid variable declaration,
+a valid subtyping or checking judgment, or a successful guessing for inferences.
+Additionally, judgment chains are unfolded to match the original declarative rules.
+Although $\Om$ is defined differently from $\Psi$,
+our declarative worklist $\Om$ contains a ordered collection of variable definitions,
+which naturally fits declarative subtyping and typing.
 

### 2019-01-31 sections/Metatheory.tex
soundness and completeness

@@
 the substitution should affect all the judgments and term variable bindings in the scope of $\al$.
+
+The transfer $\Gm\sto\Om$ is not a function.
+On the one hand, there are infinite possibilities to instantiate an existential variable,
+thus a $\Gm$ could be transferred to different $\Om$'s.
+On the other hand, any valid mono-type in $\Om$ can be represented by an existential variable in $\Gm$,
+thus different $\Gm$'s could be transferred to a single $\Om$.
 
@@
 
+Our algorithm is sound with respect to the declarative system.
+For any worklist $\Gm$ that reduces successfully,
+we can find a valid transfer $\Gm\sto\Om$ that solves all the existential variables,
+such that $\Om$ also reduces successfully.
 \begin{theorem}[Soundness]
@@
 
+The proof proceeds by induction on derivation of $\Gm\redto\nil$.
+Proper applications of lemmas \ref{lem:insert} and \ref{lem:extract} finish off the theorem.
+
 \subsection{Completeness}
+
+Completeness of the algorithm means that any declarative system has an algorithmic counterpart.
 
@@
 
+We prove completeness by induction on the derivation of $\Om\redto\nil$.
+Since the declarative worklist is reduced judgment-by-judgment,
+the induction always analyse the first judgment until it is satisfied.
+As the algorithmic system introduces existential variables,
+a declarative rule may correspond to multiple algorithmic rules,
+and thus we analyse each of the possible cases.
+
+For subtyping, algorithmic Rules 10 and 11 worth special treatment.
+When the induction reaches the $\mathtt{{\le}{\to}}$ case,
+the first judgment is of shape $A_1 \to A_2 \le B_1 \to B_2$.
+One of the corresponding algorithmic judgment is $\al \le A \to B$.
+However, the case analysis does not imply that $\al$ is fresh in $A$ and $B$,
+therefore Rule~10 cannot be applied and the proof get stuck.
+The following lemma helps us out in those cases:
+the success in declarative subtyping indicates the freshness of $\al$ in $A$ and $B$
+in its corresponding algorithmic judgment.
+In other words, the declarative system do not accept infinite types,
+thus, as an example, we cannot find a mono-type $\tau$ such that $\tau\le 1\to \tau$,
+which could be transferred by $\al\le 1\to\al$.
+
 \begin{lemma}[Prune Transfer for Instantiation]
-If $(\Gm \Vdash \al \le A \to B) \sto (\Om \Vdash C \le A_1 \to B_1)$ and $\Om \vdash C \le A_1 \to B_1$,
-then $\al\notin FV(A) \cup FV(B)$.
+If $(\Gm \Vdash \al \le A \to B) \sto (\Om \Vdash C \le A_1 \to B_1)$ and
+$\Om \vdash C \le A_1 \to B_1$, then $\al\notin FV(A) \cup FV(B)$.
 \end{lemma}
+
+A symmetric lemma holds for $A\to B \le \al$.
 
@@
 \begin{gather*}
-\inferrule*[right=$\mathtt{\jExt\nil}$]
+\inferrule*[right=$\mathtt{\jExt id}$]
     {~}{\Gm\jExt\Gm}

### 2019-01-31 sections/Metatheory.tex
Some def and lemmas for decidability

@@
 \begin{theorem}[Decidability]
-For all well-formed $\Gm$, it is decidable whether $\Gm\redto\nil$ or not.
+Given \emph{wf }$\Gm$, it is decidable whether $\Gm\redto\nil$ or not.
 \end{theorem}
@@
 |\Gm|_\Leftrightarrow &= \#_{e\Lto A\in\Gm} +
-    2\nil\#_{e\To_a \jg\in\Gm} + 3\nil\#_{\appInfAlg{A}{e}\in\Gm}\\
+    2\cdot\#_{e\To_a \jg\in\Gm} + 3\cdot\#_{\appInfAlg{A}{e}\in\Gm}\\
 |\Gm|_\forall &= \textstyle\sum_{e\Lto A\in\Gm}|A|_\forall + \sum_{\appInfAlg{A}{e}\in\Gm}|A|_\forall +
@@
 |\Gm|_{\al} &= \#_{\al\in\Gm}\\
+\end{aligned}
+\end{gather*}
+\end{definition}
+
+\begin{definition}[$\Gm_\le, \Gm_i$]
+\begin{gather*}
+\begin{aligned}
+\Gm_\le &:= \nil \mid \Gm_\le, a \mid \Gm_\le, \al \mid \Gm_\le, A\le B\\
+\Gm_i &:= \nil \mid \Gm_i, \al\le A \mid \Gm_i, A\le\al \quad
+    \text{where } \al\notin FV(A) \cup FV(\Gm_i)
 \end{aligned}
@@
 
+\newcommand{\equivGm}[1]{|\Gm|_{#1} = |\Gm'|_{#1}}
+\begin{lemma}[Measure Invariants of Worklist Extension]
+If $\Gm\sto\Gm'$ then $\equivGm{e}$ and $\equivGm\Leftrightarrow$ and $\equivGm\forall$.
+\end{lemma}
+
+\begin{theorem}[Instantiation Decidability]
+For any well-formed algorithmic worklist $(\Gm, \Gm_i)$,
+\begin{enumerate}[1)]
+    \item When $|\Gm_i|_\forall = 0$, $(\Gm, \Gm_i)$ reduces to $\Gm'$
+        with $|\Gm_i|$ existential varaibles solved, where $\Gm\sto\Gm'$.
+    \item When $|\Gm_i|_\forall > 0$, $(\Gm, \Gm_i)$ reduces to $\Gm', \Gm_\le$,
+        where $|\Gm_\le|_\forall = |\Gm_i|_\forall - 1$ and $\Gm\sto\Gm'$.
+\end{enumerate}
+\end{theorem}
+
 \subsection{Abella}

### 2019-01-31 sections/Metatheory.tex
decidability

@@
 
-lexicographic group of induction measurements
+We have proven this theorem by a lexicographic group of induction measurements\\
 $\langle |\Gm|_e, |\Gm|_\Leftrightarrow, |\Gm|_\forall, |\Gm|_{\al}, |\Gm|_\to + |\Gm| \rangle$
+on the worklist $\Gm$.
+$|\cdot|_e$ and $|\cdot|_\Leftrightarrow$ measure the total ``size'' of teams
+and the total ``difficulty'' of judgments, respectively;
+$|\cdot|_\forall$ and $|\cdot|_\to$ count the total number of
+universal quantifiers and function types, respectively;
+$|\cdot|_{\al}$ counts the number of existential variables in the worklist.
 
@@
     \sum_{\appInfAlg{A}{e}\in\Gm}|e|_e\\
+\text{where } |x|_e &= |()|_e = 1 \qquad |\lam e|_e = |e|_e + 1\\
+|e_1~e_2|_e &= |e_1|_e + |e_2|_e + 1 \qquad |e:A|_e = |e| + 1\\[2mm]
+%
 |\Gm|_\Leftrightarrow &= \#_{e\Lto A\in\Gm} +
-    2\cdot\#_{e\To_a \jg\in\Gm} + 3\cdot\#_{\appInfAlg{A}{e}\in\Gm}\\
+    2\cdot\#_{e\To_a \jg\in\Gm} + 3\cdot\#_{\appInfAlg{A}{e}\in\Gm}\\[2mm]
 |\Gm|_\forall &= \textstyle\sum_{e\Lto A\in\Gm}|A|_\forall + \sum_{\appInfAlg{A}{e}\in\Gm}|A|_\forall +
     \sum_{A\le B\in\Gm} |A|_\forall + |B|_\forall\\
+\text{where } |1|_\forall &= |a|_\forall = |\al|_\forall = 1\\
+|A\to B|_\forall &= |A|_\forall + |B|_\forall \qquad |\all A|_\forall = |A|_\forall + 1\\[2mm]
+%
 |\Gm|_\to &= \textstyle\sum_{e\Lto A\in\Gm}|A|_\to + \sum_{\appInfAlg{A}{e}\in\Gm}|A|_\to +
     \sum_{A\le B\in\Gm} |A|_\to + |B|_\to\\
+\text{where } |1|_\to &= |a|_\to = |\al|_\to = 1\\
+|A\to B|_\to &= |A|_\to + |B|_\to + 1 \qquad |\all A|_\to = |A|_\to\\[2mm]
+%
 |\Gm|_{\al} &= \#_{\al\in\Gm}\\
@@
 \end{definition}
+
+It is not difficult to see that all but two algorithmic reduction rules
+decreases the group of measurements.
+(The result of Rule 29 could be directly reduced by Rule 28, which decreases the measurements.)
+The two exceptions are Rules 10 and 11; they increases the number of existential variables
+without decreasing the number of universal quantifiers.
+However, at the same time, they both break the subtyping problem into two sub-problems
+(Rule 7 applied right after Rule 10 or 11), in smaller sizes.
+Therefore we argue that such split makes the worklist easier, by analyzing a subset of worklist,
+which we call \emph{instantiation judgments} $\Gm_i$, with the help of $\Gm_\le$ for the proof.
 
@@
 \begin{figure}
-\framebox{$\Gm\jExt\Gm'$}
+\framebox{$\Gm\jExt\Gm'$} $\Gm$ extends to $\Gm'$.
 \begin{gather*}
@@
 
+\paragraph{Intuition of instantiation decidability}
+These instantiation judgments are the ones consumed and produced by Rules 10 and 11.
+When there are universal quantifiers in the initial instantiation judgments $\Gm_i$,
+the worklist $\Gm,\Gm_i$ will be multi-reduced to match Rule 8 or 9,
+which decreases the number of universal quantifiers by 1, and returns a $\Gm',\Gm_\le$.
+Otherwise, no more variables will be introduced,
+and each instantiation judgment solves one existential variable,
+thus finally there will be $|\Gm_i|$ existential variables solved,
+with the worklist reduced to $\Gm'$.
+$\Gm'$ basically refers to the original part of worklist $\Gm$.
+During the reduction, minor changes to worklist might occur, such as
+existential variable solving, insertion and instantiation,
+but the main structure of it remains unchanged.
+
 Shown in Figure~\ref{fig:worklist_ext},
-worklist extension relates minor changes of worklist during type-checking.
+worklist extension formally describe the minor changes to worklist during algorithmic reduction.
 When the algorithm processes the worklist, the first judgment is processed,
 and the rest of the worklist roughly remains its shape.
-Worklist extension formally defines the minor changes of the shape by two cases:
-existential variable solving ($\mathtt{\jExt solve}$) and
-existential variable addition ($\mathtt{\jExt \al}$).
-Existential variable instantiation,
+Worklist extension is formally defined by two cases:
+existential variable solving ($\mathtt{\jExt solve}$) and insertion ($\mathtt{\jExt \al}$).
+Existential variable instantiation, as another common form of worklist manipulation,
 i.e. $\Gm_L,\al,\Gm_R \jExt \Gm_L, \al[1], \al[2], [\al[1] \to \al[2] / \al]\Gm_R$,
 can be derived with the combination of the 3 rules.
+Within expectation, some worklist measurements are not changed modulo worklist extension.
 
@@
 
-\begin{theorem}[Instantiation Decidability]
+With all the definitions, we are able to prove the following instantiation decidability lemma.
+
+\begin{lemma}[Instantiation Decidability]\label{lemma:inst:decidable}
 For any well-formed algorithmic worklist $(\Gm, \Gm_i)$,
 \begin{enumerate}[1)]
-    \item When $|\Gm_i|_\forall = 0$, $(\Gm, \Gm_i)$ reduces to $\Gm'$
-        with $|\Gm_i|$ existential varaibles solved, where $\Gm\sto\Gm'$.
-    \item When $|\Gm_i|_\forall > 0$, $(\Gm, \Gm_i)$ reduces to $\Gm', \Gm_\le$,
+    \item When $|\Gm_i|_\forall = 0$, $(\Gm, \Gm_i) \redto \Gm'$,
+        where $|\Gm'|_{\al} = |\Gm|_{\al} - |\Gm_i|$ and $\Gm\sto\Gm'$.
+    \item When $|\Gm_i|_\forall > 0$, $(\Gm, \Gm_i) \redto (\Gm', \Gm_\le)$,
         where $|\Gm_\le|_\forall = |\Gm_i|_\forall - 1$ and $\Gm\sto\Gm'$.
 \end{enumerate}
-\end{theorem}
+\end{lemma}
+
+In short, after reducing any instantiation judgment prefix $\Gm_i$,
+the algorithm either decreases the number of $\forall$'s
+or solves one existential variable per instantiation judgment.
+The proof of this lemma is by induction on the measure $2*|\Gm_i|_\to + |\Gm_i|$
+of the instantiation judgment list $\Gm_i$.
+
+In summary, the decidability theorem is shown through a lexicographic group of induction measurements
+$\langle |\Gm|_e, |\Gm|_\Leftrightarrow, |\Gm|_\forall, |\Gm|_{\al}, |\Gm|_\to + |\Gm| \rangle$.
+The critical case is that, whenever we encounter an instantiation judgment at the top of the worklist,
+we make use of Lemma~\ref{lemma:inst:decidable}, which reduces $|\Gm|_{\al}$ or $|\Gm|_\forall$
+by consuming that instantiation judgment. Other cases are relatively straightforward.
 

### 2019-02-05 sections/algorithm.tex
transfer
soundness and completeness
Some def and lemmas for decidability
decidability
alg. sample derivation
Section 3 WIP

@@
-\section{Algorithm}
+\section{Algorithmic System}
+
+This section introduces a novel algorithmic system that implements the
+DK's declarative specification. The algorithmic system comes with
+several novelties.\bruno{fill in/summarize novelties}
 
@@
 \text{Algorithmic worklist}\qquad&\Gm &::=&\quad \nil \mid \Gm, a \mid \Gm, \al \mid \Gm, x: A \mid \Gm \Vdash \jg\\
-\text{Declarative worklist}\qquad&\Om &::=&\quad \nil \mid \Gm, a \mid \Gm, x: A \mid \Gm \Vdash \jg
+\text{Declarative worklist}\qquad&\Om &::=&\quad \nil \mid \Om, a \mid \Om, x: A \mid \Om \Vdash \jg
 \end{aligned}
@@
 \paragraph{Existential Variables} In order to find the unknown type $\tau$ in the declarative system,
-\emph{existential variables} $\al, \bt$ are introduced to the algorithmic system.
+\emph{existential variables} $\al, \bt$ are introduced in the algorithmic system.
 They behave similarly to unification variables,
-but their scoping is restricted by the position located in the worklist,
-rather than globally defined
+but their scoping is restricted by the position located in the
+worklist\bruno{be careful to ensure that the reader knows about
+  worklist at this point},
+rather than globally defined.
 Any existential variable $\al$ should only be solved to
@@
 
-\paragraph{Judgments and judgment chains} Judgment is a core component of algorithmic type-checking.
-There are four kinds of judgments in our system: subtyping, checking, inference and application inference.
-Subtyping and checking are relatively simple, since the result could be either success or fail, and no values will be returned.
-For inference and application inference, they would return some type that leads to further computation, so we use continuation-passing-style encodings for them.
-For example, a judgment chain $e \To_a (a \le B)$ contains two judgments: first we want to infer the type of the expression $e$, and then check if that type is a subtype of $B$.
-In the rest of the paper, we usually refer to ``judgment chains'' by ``judgments''.
+\paragraph{Monotypes, types and terms} The algorithm system also has a
+notion of monotypes, which is identical to the declarative system.
+\bruno{Since monotypes are identical to declarative system, maybe they
+do not need to be presented again. Perhaps drop their definition from 
+the syntax?} Algorithmic types include all the declarative types plus
+existential variables $\al$. \bruno{Aren't terms identical to the
+  Declarative system too? If they are briefly state that terms are
+  identical and do not show them here. Otherwise say what is different.}
+
+%% A \le B \mid e\Lto A \mid e\To_{a} \jg \mid \appInfAlg{A}{e}
+
+\paragraph{Judgments and judgment chains} 
+A \emph{judgment chain} $\jg$ is a core component of algorithmic
+type-checking. For conciseness, in the rest of the paper, we usually
+refer to ``judgment chains'' simply by ``judgments''.  There are four kinds
+of judgments in our system: subtyping ($A \le B$), checking ($e\Lto
+A$), inference ($e\To_{a} \j$) and
+application inference ($\appInfAlg{A}{e}$).  Subtyping and checking are relatively simple,
+since the result could be either success or fail, and no other output
+is returned.  However inference and application inference return some
+type that leads to further computation. We use a
+continuation-passing-style encoding for them.  For example, a judgment
+chain $e \To_a (a \le B)$ contains two judgments: first we want to
+infer the type of the expression $e$, and then check if that type is a
+subtype of $B$. The \emph{unknown} type of $e$ is represented by a
+type variable $a$, which is used in the second judgement to denote the 
+type of $e$.
 
@@
 \\[2mm]
-\Gm \Vdash e \Lto B &\algrule \Gm \Vdash e\To_a a\le B \quad \text{when } e \neq \lam e', B \neq \all B'\\
+\Gm \Vdash e \Lto B &\algrule \Gm \Vdash e\To_a a\le B \quad
+    \text{when } e \neq \lam e' \text{ and } B \neq \all B'\\
 % \Gm \Vdash () \Lto 1 &\algrule \Gm\\
@@
     \Gm,\al,\bt \Vdash [\al\to\bt/a]\jg, x:\al \Vdash e\Lto \bt \quad\jimmy{\text{differs}}\\
-\Gm \Vdash e_1\ e_2 \To_a \jg &\algrule \Gm \Vdash e_1\To_{a_1} (\appInfAlg{a_1}{e_2})
+\Gm \Vdash e_1\ e_2 \To_a \jg &\algrule \Gm \Vdash e_1\To_b (\appInfAlg{b}{e_2})
 \\[2mm]
@@
 \Gm[\al] \Vdash \appInfAlg{\al}{e} &\algrule
-	[\al[1]\to\al[2]/\al](\Gm[\al[1], \al[2]] \Vdash [\al[2]/a]\jg \Vdash e\Lto \al[1])\\
- &\color{magenta} \makebox[0pt]{\qquad or} \phantom{{}\rrule{}{}}
     [\al[1]\to\al[2]/\al](\Gm[\al[1], \al[2]] \Vdash \appInfAlg{\al[1]\to\al[2]}{e})
+%	[\al[1]\to\al[2]/\al](\Gm[\al[1], \al[2]] \Vdash [\al[2]/a]\jg \Vdash e\Lto \al[1])\\
+% &\color{magenta} \makebox[0pt]{\qquad or} \phantom{{}\rrule{}{}}
 \end{aligned}
@@
 
-The algorithmic typing reduction rules, defined in Figure~\ref{fig:alg}, has the form $\Gm \rto \Gm'$,
-where the worklists $\Gm, \Gm'$ contains variable declarations and judgments.
-In every step it pops the first judgment from the worklist for processing,
+The algorithmic typing reduction rules, defined in Figure~\ref{fig:alg}, have the form $\Gm \rto \Gm'$.
+The worklists $\Gm, \Gm'$ contain both variable declarations and judgments.
+In every step reduction pops the first judgment from the worklist for processing,
 possibly pushes new judgments to the worklist, and repeats this process until the list is empty.
-A worklist that could be finally reduced to empty is considered to be successfully type-checked.
-When the first element is a variable declaration, we simply continue the process (Rules 1-3).
+A worklist that finally reduces to empty is considered to be successfully type-checked.
+When the first element is a variable declaration, we simply continue
+the process (Rules 1-3).
+Note that those rules about to garbage collecting variables that are
+out-of-scope for the next judgement being processed.
 The other rules always deal with the first judgment in the worklist. Logically we can discern 6 group of rules.
 
-Firstly, we have six rules (Rules 4-9) that are similar to those of declarative subtyping.
-For instance, Rule 7 consumes one subtyping judgment and pushes two back to the worklist.
-Rule 8 differs from declarative Rule ${\le}{\forall}L$ by introducing an existential variable $\al$ instead of guessing the mono-type $\tau$ for instantiation.
-Each existential variable is later solved to a mono-type $\tau$ with the same context,
-so the final solution of $\al$, $\tau$, should be well-formed under context $\Gm$.
+Firstly, we have six rules (Rules 4-9) that are similar to those of
+declarative subtyping.  For instance, Rule 7 consumes one subtyping
+judgment and pushes two back to the worklist.  Rule 8 differs from
+declarative Rule ${\le}{\forall}L$ by introducing an existential
+variable $\al$ instead of guessing the monotype $\tau$ for
+instantiation.  Each existential variable is later solved to a
+monotype $\tau$ with the same context, so the final solution of
+$\al$, $\tau$, should be well-formed under context $\Gm$.
 
 \paragraph{Variable scopings within worklist}
-Rules 8 and 9 involve variable declaration, which demonstrates how our worklist treats variable scopings.
-Rule 8 introduces an existential variable $\al$ that is only visible to the judgment $[\al/a]A \le B$.
-Then reduction continues until all the subtyping judgments after $\al$ are satisfied.
-Finally we can safely remove $\al$ since no occurrence of $\al$ could leak to its left in the worklist.
-On the other hand, removing $\al$ according to the algorithm carbage-collects the variable at the perfect time: it leaves the environment immediately after unreferenced.
+Rules 8 and 9 involve variable declaration, which demonstrates how our
+worklist treats variable scopings. Rule 8 introduces an existential
+variable $\al$ that is only visible to the judgment $[\al/a]A \le B$.
+Then reduction continues until all the subtyping judgments after $\al$
+are satisfied.  Finally we can safely remove $\al$ since no occurrence
+of $\al$ could leak to its left in the worklist.  On the other hand,
+removing $\al$ according to the algorithm garbage-collects the
+variable at the perfect time: it leaves the environment immediately
+after being unreferenced.\bruno{Perhaps this is better illustrated
+  with a simple example?}
 
-The design of our judgment chain is closely related to the property of the judgments.
-Subtyping and checking does not return anything, so variables cannot leak anyway, as applied to Rules 8 and 9.
+The design of our judgment chain is closely related to the property of
+the judgments\bruno{What property? Not very clear}.
+Subtyping and checking do not return anything, so variables cannot
+leak anyway, as applied to Rules 8 and 9.
+\bruno{Is the discussion that follows in the right place? We just to
+  rule 26. Perhaps we can wait until we talk about inference to
+  discuss those issues?}
 However, inference and application inference may return a type that contains new variables.
@@
 Rule 29 instantiate $\al$ to the function type $\al[1] \to \al[2]$, just like Rules 10, 11 and 21.
+
+The syntax $\Gm\redto \Gm'$ denotes multiple reduction.
+If $\Gm\redto\nil$, we say that $\Gm$ reduces successfully.
+
+\paragraph{Example}
+
+\begin{figure}
+\begin{gather*}
+\begin{aligned}
+           & \nil \Vdash (\lam x)~() \Lto 1\\
+\rrule{18} & \nil \Vdash (\lam x)~() \To_a a\le 1\\
+\rrule{26} & \nil \Vdash (\lam x) \To_b (\appInfAlg{b}{()}[a][a\le 1])\\
+\rrule{25} & \nil,\al,\bt \Vdash \appInfAlg{\al\to\bt}{()}[a][a\le 1], x:\al \Vdash x\Lto \bt\\
+\rrule{18} & \nil,\al,\bt \Vdash \appInfAlg{\al\to\bt}{()}[a][a\le 1], x:\al \Vdash x\To_b b\le \bt\\
+\rrule{22} & \nil,\al,\bt \Vdash \appInfAlg{\al\to\bt}{()}[a][a\le 1], x:\al \Vdash \al\le \bt\\
+\rrule{12} & \nil,\al \Vdash \appInfAlg{\al\to\al}{()}[a][a\le 1], x:\al\\
+\rrule{3}  & \nil,\al \Vdash \appInfAlg{\al\to\al}{()}[a][a\le 1]\\
+\rrule{28} & \nil,\al \Vdash \al\le 1 \Vdash () \Lto \al\\
+\rrule{18} & \nil,\al \Vdash \al\le 1 \Vdash () \To_a a\le \al\\
+\rrule{24} & \nil,\al \Vdash \al\le 1 \Vdash 1\le \al\\
+\rrule{16} & \nil \Vdash 1\le 1\\
+\rrule{4}  & \nil
+\end{aligned}
+\end{gather*}
+\caption{A Sample Derivation for Algorithmic Typing}
+\label{fig:alg:sample}
+\end{figure}
+
+Figure~\ref{fig:alg:sample} shows a sample derivation of the algorithm.

### 2019-02-07 sections/algorithm.tex
Section 3

@@
 DK's declarative specification. The algorithmic system comes with
-several novelties.\bruno{fill in/summarize novelties}
+several novelties.\bruno{fill in/summarize novelties. I think
+  novelties include: precise scoping control; CPS-style; dealing with
+  inference judgements.}
 
@@
 leak anyway, as applied to Rules 8 and 9.
-\bruno{Is the discussion that follows in the right place? We just to
+\bruno{Is the discussion that follows in the right place? We just jump
+  to
   rule 26. Perhaps we can wait until we talk about inference to
@@
 
+\paragraph{Existential decomposition}
 Secondly, Rules 10 and 11 are algorithmic extensions to Rule $\mathtt{{\le}{\to}}$.
@@
 We remove $\al$ from the worklist, and insert $\al[1], \al[2]$ back to the same position.
-To propagate the instantiation to the rest of the worklist and maintain the wel-formedness,
+To propagate the instantiation to the rest of the worklist and maintain the well-formedness,
 every $\al$ is replaced by $\al[1] \to \al[2]$.
@@
 
-Thirdly, Rules 12-17 are base cases where an existential variable is solved.
-Each of the rules removes and existential variable, substitutes the variable with its solution and continues.
+\paragraph{Solving existentials.} Thirdly, Rules 12-17 are base cases where an existential variable is solved.
+Each of the rules removes and existential variable, substitutes the
+variable with its solution and continues. Importantly the rules
+respect the scoping of existential variables. For example, rule 12 
+states that an existential variable $\al$ can be solved with another
+existential variable $\bt$ only if $\bt$ occurs after $\al$.
 
+\paragraph{Checking judgements}
 Fourthly, Rules 18-21 deal with checking judgments.
-Rule 18 is similar to $\mathtt{DeclSub}$, and rewrited in the continuation-passing-style.
+Rule 18 is similar to $\mathtt{DeclSub}$, and rewrited in the
+continuation-passing-style.
+\bruno{The side condition needs to be explained, and some mention of
+  the overlapping is probably needed here.}
 Rules 19 and 20 adapts the worklist style from their corresponding declarative rules.
-Rule 21 is a special case of $\mathtt{Decl\to I}$.
-Instantiation technique similar to Rules 10 and 11 applys.
+Rule 21 is a special case of $\mathtt{Decl\to I}$.\bruno{Rule 21
+  deserves some more explanation. What is the motivation for this
+  rule? You may want to give some concrete example to motivate it.}
+Instantiation technique similar to Rules 10 and 11 applies.
 The declarative checking rule $\mathtt{Decl1I}$ is not translated to the algorithm,
@@
 
+\paragraph{Inference judgments}
 Fifthly, Rules 22-26 deal with inference judgments.
@@
 the checking and the remaining chain $[\al\to\bt/a]\jg$.
+\bruno{You have some note ``differs'' on the rules. Should that
+  deserve an explanation here?}
 
+\paragraph{Application inferences judgements}
 Finally, Rules 27-29 deal with application inference judgments.
@@
 
+\subsection{Multiple reduction}
 The syntax $\Gm\redto \Gm'$ denotes multiple reduction.
@@
 
-Figure~\ref{fig:alg:sample} shows a sample derivation of the algorithm.
+Figure~\ref{fig:alg:sample} shows a sample derivation of the
+algorithm.
+\bruno{Derivation should be explained briefly at least.}

### 2019-02-07 sections/Metatheory.tex
Upto Section 4.4

@@
 \section{Metatheory}
+
+This section presents the metatheory of the algorithmic system
+presented in Section~\ref{}. We show that three main results hold: 
+\emph{soundness}, \emph{completeness} and \emph{decidability}.
+All the results have been mechanically formalized and proved in the 
+Abella theorem prover.
 
@@
 \end{gather*}
-\caption{Transfer Rules}
+\caption{Transfer rules}
 \label{fig:trans}
@@
 
+To aid formalizing the correspondence between the declarative and
+algorithmic systems, we use a relation that transfers algorithmic
+worklist judgements into corresponding declarative worklist judgements. In essence this
+judgment instantiates every existential variable with a well-formed monotype.
+
 The transfer rules are shown in Figure~\ref{fig:trans}.
-Rule $\mathtt{{\sto}\al}$ recursively replace the first existential variable to a well-scoped mono-type
-($\Om$, as a declarative worklist, contains no existential variables),
-so the algorithmic worklist will finally become a declarative one.
+Rule $\mathtt{{\sto}\al}$ recursively replaces the first existential variable to a well-scoped monotype.
+Note that $\Om$, as a declarative worklist, contains no existential
+variables. Therefore by recursively applying this rule, the
+algorithmic worklist will finally become a declarative one.
 In order to maintain well-formedness,
@@
 
-The transfer $\Gm\sto\Om$ is not a function.
+The transfer $\Gm\sto\Om$ is not a function.\bruno{Do you mean here
+  ``is not deterministic''? Would that be more appropriate?}
 On the one hand, there are infinite possibilities to instantiate an existential variable,
-thus a $\Gm$ could be transferred to different $\Om$'s.
-On the other hand, any valid mono-type in $\Om$ can be represented by an existential variable in $\Gm$,
+thus a judment $\Gm$ could be transferred to different $\Om$'s.
+On the other hand, any valid monotype in $\Om$ can be represented by an existential variable in $\Gm$,
 thus different $\Gm$'s could be transferred to a single $\Om$.
 
-The following lemmas~\ref{lem:insert} and \ref{lem:extract}
+Lemmas~\ref{lem:insert} and \ref{lem:extract}
 generalize Rule $\mathtt{{\sto}\al}$ from substituting the first existential variable
@@
 \end{gather*}
-\caption{Declarative Worklist Typing}
+\caption{Declarative Worklist Typing.\bruno{I think your names for the
+  two relations/definitions aren't great. I think what you're defining
+here should be called transfer. The previous relation should probably
+be called something different. Maybe something like Declarative
+worklist instantiation/realization for the previous relation?}}
 \label{fig:decl:worklist}
@@
 which naturally fits declarative subtyping and typing.
+\bruno{I think that to make things precise what you want to show is
+  that you can always define a erasure function that drops the
+  judgements from the worklist, and retains only the variable
+  declarations (thus becomes a declarative environment). Isn't this
+  what you do in your formalization?}
 
@@
 The proof proceeds by induction on derivation of $\Gm\redto\nil$.
-Proper applications of lemmas \ref{lem:insert} and \ref{lem:extract} finish off the theorem.
+Proper applications of lemmas \ref{lem:insert} and \ref{lem:extract}
+finish off the theorem.
+\bruno{Give a bit more informal explanation here. Something like: If
+  $\Gm$ is reducible then we can show that there is always an $\Om$
+  where all the corresponding declarative judgements hold.}
+\bruno{I wonder if 1 or 2 examples (concrete instantiations of the
+  theorem) would be helpful to complement the explanation.}
 
@@
 Since the declarative worklist is reduced judgment-by-judgment,
-the induction always analyse the first judgment until it is satisfied.
+the induction always analyses the first judgment until it is satisfied.
 As the algorithmic system introduces existential variables,
@@
 
-For subtyping, algorithmic Rules 10 and 11 worth special treatment.
+For subtyping, algorithmic Rules 10 and 11 require special treatment.
 When the induction reaches the $\mathtt{{\le}{\to}}$ case,
@@
 However, the case analysis does not imply that $\al$ is fresh in $A$ and $B$,
-therefore Rule~10 cannot be applied and the proof get stuck.
+therefore Rule~10 cannot be applied and the proof gets stuck.
 The following lemma helps us out in those cases:
@@
 in its corresponding algorithmic judgment.
-In other words, the declarative system do not accept infinite types,
-thus, as an example, we cannot find a mono-type $\tau$ such that $\tau\le 1\to \tau$,
+In other words, the declarative system does not accept infinite types,
+thus, as an example, we cannot find a monotype $\tau$ such that $\tau\le 1\to \tau$,
 which could be transferred by $\al\le 1\to\al$.

### 2019-02-07 main.tex
Intro WIP
Abstract small revision

@@
 Modern functional programming languages, such as Haskell or OCaml,
-use sophisticated forms of type inference. Yet, while it is an
+use sophisticated forms of type inference. While an
 important topic in the Programming Languages research, there is little
@@
 not clear how to replace these by other valid proofs/lemmas. 
-\bruno{others} Despite the use of a different algorithm we prove the
+\bruno{others?} Despite the use of a different algorithm we prove the
 same results as DK, although with significantly different proofs and
-proof techniques.  
+proof techniques. Since such type inference algorithms are quite
+subtle and have a complex metatheory, we believe that mechanical
+formalizations are an important advance in type-inference research.
 \end{abstract}

### 2019-02-10 sections/Metatheory.tex
a few edits in section 4

@@
 This section presents the metatheory of the algorithmic system
-presented in Section~\ref{}. We show that three main results hold: 
+presented in the previous section. We show that three main results hold: 
 \emph{soundness}, \emph{completeness} and \emph{decidability}.
-All the results have been mechanically formalized and proved in the 
-Abella theorem prover.
+These three results have been mechanically formalized and proved in the 
+Abella theorem prover~\cite{}.
 
@@
 The transfer rules are shown in Figure~\ref{fig:trans}.
-Rule $\mathtt{{\sto}\al}$ recursively replaces the first existential variable to a well-scoped monotype.
+Rule $\mathtt{{\sto}\al}$ replaces the first existential variable with a well-scoped monotype and
+repeats the process in the resulting worklist.
 Note that $\Om$, as a declarative worklist, contains no existential

### 2019-02-10 sections/Metatheory.tex
measurements -> measures

@@
 
-We have proven this theorem by a lexicographic group of induction measurements\\
+We have proven this theorem by a lexicographic group of induction measures\\
 $\langle |\Gm|_e, |\Gm|_\Leftrightarrow, |\Gm|_\forall, |\Gm|_{\al}, |\Gm|_\to + |\Gm| \rangle$
@@
 It is not difficult to see that all but two algorithmic reduction rules
-decreases the group of measurements.
-(The result of Rule 29 could be directly reduced by Rule 28, which decreases the measurements.)
+decreases the group of measures.
+(The result of Rule 29 could be directly reduced by Rule 28, which decreases the measures.)
 The two exceptions are Rules 10 and 11; they increases the number of existential variables
@@
 can be derived with the combination of the 3 rules.
-Within expectation, some worklist measurements are not changed modulo worklist extension.
+Within expectation, some worklist measures are not changed modulo worklist extension.
 
@@
 
-In summary, the decidability theorem is shown through a lexicographic group of induction measurements
+In summary, the decidability theorem is shown through a lexicographic group of induction measures
 $\langle |\Gm|_e, |\Gm|_\Leftrightarrow, |\Gm|_\forall, |\Gm|_{\al}, |\Gm|_\to + |\Gm| \rangle$.

### 2019-02-11 sections/introduction.tex
Intro WIP

@@
 In those type systems
-type annotations are required for polymorphic arguments of functions.
+type annotations are required for polymorphic arguments of functions,
 but other type annotations can be omitted. A canonical example is:
@@
 The function \verb|hpoly| cannot be
-type-checked in Hindley-Milner.  The type of \verb|hpoly| is 
-\verb|(forall a. a -> a) -> (Int, Char)|. The single universal quantifier
+type-checked in Hindley-Milner.  The type of \verb|hpoly| is the rank-2 type:
+\verb|(forall a. a -> a) -> (Int, Char)|. Notably (and unlike
+Hindley-Milner) the lambda argument \verb|f| requires a
+\emph{polymorphic} type annotation.
+This annotation is needed because the single universal quantifier
 does not appear at the top-level. Instead it is used to quantify a
@@
 function. 
-Notably \verb|hpoly| requires a type annotation for the first
-argument (\verb|forall a. a -> a|). 
-Despite these additional annotations,
-the type-inference algorithm employed by GHC Haskell~\cite{jones2007practical} preserves 
-many of the desirable properties of Hindley-Milner.
+Despite these additional annotations, Peyton Jones et al. and DK's
+type inference algorithms preserve many of the desirable properties 
+of Hindley-Milner. For example the applications of \verb|f| implicitly 
+instantiate the polymorphic type arguments of \verb|f|.
 
@@
 such advanced forms of type inference in theorem provers. This is at
-odds with a current trend in programming language research. In
-particular both the POPLMark challenge~\cite{} and CompCert ~\cite{}
-have significantly promoted the use of theorem provers to model
-variaous aspects of programming languages. Indeed today there are
-routinely papers in top conferences that use theorem provers to
-mechanically formalize various aspects of programming languages,
-including: \emph{dynamic and static semantics} and their
-correctness properties~\cite{}, \emph{compiler correctness}~\cite{},
-\emph{correctness of optimizations}~\cite{} or even certain forms of
-\emph{abstract interpretation}. Mechanical formalizations have
-numerous advantages~\cite{read to make sure I do not miss any!}.
+odds with the current trend of mechanical formalizations in
+programming language research. In particular both the POPLMark
+challenge~\cite{} and CompCert ~\cite{} have significantly promoted
+the use of theorem provers to model various aspects of programming
+languages. Indeed today there are routinely papers in top conferences
+that use theorem provers to mechanically formalize: \emph{dynamic and
+  static semantics} and their correctness properties~\cite{},
+\emph{compiler correctness}~\cite{}, \emph{correctness of
+  optimizations}~\cite{}, certain forms of \emph{abstract
+  interpretation} or proofs involving \emph{logical relations}. The
+main argument for mechanical formalizations is a simple one. Proofs
+for programming languages tend to be \emph{long}, \emph{tedious} and
+\emph{error-prone}. In such proofs it is very easy to make mistakes
+that invalidate the whole development. Furthermore, readers and
+reviewers often do not have time to look at the proofs carefuly to
+check their correctnessness. Therefore errors can go unnoticed for a
+long time.  Mechanical formalizations provide, in principle, a natural
+solution for such problems. Theorem provers can automaticaly check and
+validate the proofs, which removes the burden from checking from both
+the person doing the proofs as well as readers or reviewers.
+
+DK problem in lemma ??. Castagna problem in a recent type inference paper.
+
+with the exception of some work
+done on the formalization of 
+certain aspects of Hindley-Milner type inference, we do not know
 

### 2019-02-11 sections/algorithm.tex

@@
 
-This section introduces a novel algorithmic system that implements the
+This section introduces a novel algorithmic system that implements 
 DK's declarative specification. The algorithmic system comes with
@@
 
-\paragraph{Existential Variables} In order to find the unknown type $\tau$ in the declarative system,
-\emph{existential variables} $\al, \bt$ are introduced in the algorithmic system.
+\paragraph{Existential Variables} In order to find unknown types $\tau$ in the declarative system,
+the algorithmic system introduces \emph{existential variables} $\al, \bt$.
 They behave similarly to unification variables,
-but their scoping is restricted by the position located in the
+but their scoping is restricted by their position in the
 worklist\bruno{be careful to ensure that the reader knows about
   worklist at this point},
-rather than globally defined.
+rather than being global.
 Any existential variable $\al$ should only be solved to
-a well-formed type under the worklist before $\al$ is defined.
-Thus the mono-type $\tau$ represented by the corresponding existential variable
+a type that is well-formed with respect to the worklist to which $\al$ has been added.
+The point is that the mono-type $\tau$ represented by the corresponding existential variable
 is always well-formed under the corresponding declarative context.
@@
 \paragraph{Judgments and judgment chains} 
-A \emph{judgment chain} $\jg$ is a core component of algorithmic
+\emph{Judgment chain}s $\jg$ are the core components of our algorithmic
 type-checking. For conciseness, in the rest of the paper, we usually
@@
 of judgments in our system: subtyping ($A \le B$), checking ($e\Lto
-A$), inference ($e\To_{a} \j$) and
+A$), inference ($e\To_{a} \jg$) and
 application inference ($\appInfAlg{A}{e}$).  Subtyping and checking are relatively simple,
-since the result could be either success or fail, and no other output
-is returned.  However inference and application inference return some
-type that leads to further computation. We use a
-continuation-passing-style encoding for them.  For example, a judgment
+since their result is only success or failure. However both inference and
+application inference return a type that is used in further computation. We use
+continuation-passing-style encoding to accomplish this. For example, the judgment
 chain $e \To_a (a \le B)$ contains two judgments: first we want to
@@
 each variable is visible only after it is declared.
-For example, $[a,x:a,x \Lto a]$ is a valid worklist, but $[\textcolor{red}{x} \Lto a, x:\textcolor{red}{a},a]$ is not (red-colored symbols are out-of-scope).
+For example, $[a,x:a,x \Lto a]$ is a valid worklist, but $[\textcolor{red}{x} \Lto a, x:\textcolor{red}{a},a]$ is not (the red-colored symbols are out of scope).
 
@@
 The worklists $\Gm, \Gm'$ contain both variable declarations and judgments.
-In every step reduction pops the first judgment from the worklist for processing,
-possibly pushes new judgments to the worklist, and repeats this process until the list is empty.
-A worklist that finally reduces to empty is considered to be successfully type-checked.
-When the first element is a variable declaration, we simply continue
-the process (Rules 1-3).
-Note that those rules about to garbage collecting variables that are
-out-of-scope for the next judgement being processed.
-The other rules always deal with the first judgment in the worklist. Logically we can discern 6 group of rules.
+The reduction process treats the worklist as a stack.
+In every step it pops the first judgment from the worklist for processing,
+possibly pushes new judgments onto the worklist, and repeats this process until the list is empty.
+A worklist that fully reduces to empty corresponds to a successful type-checking.
 
-Firstly, we have six rules (Rules 4-9) that are similar to those of
-declarative subtyping.  For instance, Rule 7 consumes one subtyping
+Rules 1-3 pop a variable declaration; this garbage collects variables that are
+out of scope for the remaining judgements.
+
+All other rules concern a judgement at the front of the worklist. Logically we can discern 6 group of rules.
+
+Firstly, we have six subtyping rules (Rules 4-9) that have similar 
+declarative counterparts.  For instance, Rule 7 consumes a subtyping
 judgment and pushes two back to the worklist.  Rule 8 differs from
@@
 instantiation.  Each existential variable is later solved to a
-monotype $\tau$ with the same context, so the final solution of
-$\al$, $\tau$, should be well-formed under context $\Gm$.
+monotype $\tau$ with the same context, so the final solution $\tau$ of
+$\al$ should be well-formed under context $\Gm$.
 
 \paragraph{Variable scopings within worklist}
-Rules 8 and 9 involve variable declaration, which demonstrates how our
+Rules 8 and 9 involve variable declarations and demonstrate how our
 worklist treats variable scopings. Rule 8 introduces an existential
 variable $\al$ that is only visible to the judgment $[\al/a]A \le B$.
-Then reduction continues until all the subtyping judgments after $\al$
+Reduction continues until all the subtyping judgments after $\al$
 are satisfied.  Finally we can safely remove $\al$ since no occurrence
-of $\al$ could leak to its left in the worklist.  On the other hand,
-removing $\al$ according to the algorithm garbage-collects the
+of $\al$ could have leaked into the left part of the worklist.  Moreover,
+the algorithm garbage-collects the $\al$
 variable at the perfect time: it leaves the environment immediately
 after being unreferenced.\bruno{Perhaps this is better illustrated
-  with a simple example?}
+  with a simple example?}\tom{Yes, it would be good to show one or more
+example derivations to illustrate the key points.}
 
@@
 The \emph{occurs-check} condition is to prevent a diverging infinite loop.
-Without such check, $\al \le 1 \to \al$, as an example, would diverge.
+For example, without such check $\al \le 1 \to \al$ would diverge.
 
 \paragraph{Solving existentials.} Thirdly, Rules 12-17 are base cases where an existential variable is solved.
-Each of the rules removes and existential variable, substitutes the
-variable with its solution and continues. Importantly the rules
+Each of the rules removes an existential variable and substitutes the
+variable with its solution in the remaining worklist. Importantly the rules
 respect the scoping of existential variables. For example, rule 12 
@@
 Fourthly, Rules 18-21 deal with checking judgments.
-Rule 18 is similar to $\mathtt{DeclSub}$, and rewrited in the
+Rule 18 is similar to $\mathtt{DeclSub}$, but rewritten in the
 continuation-passing-style.
@@
   the overlapping is probably needed here.}
-Rules 19 and 20 adapts the worklist style from their corresponding declarative rules.
+Rules 19 and 20 adapt their declarative counterparts to the worklist style.
 Rule 21 is a special case of $\mathtt{Decl\to I}$.\bruno{Rule 21
@@
   rule? You may want to give some concrete example to motivate it.}
-Instantiation technique similar to Rules 10 and 11 applies.
+The same instantiation technique as in Rules 10 and 11 applies.
 The declarative checking rule $\mathtt{Decl1I}$ is not translated to the algorithm,
@@
 When an inference judgment succeeds with type $A$,
-the algorithm continues to work on the inner-chain $\jg$ by assigning $A$ to its placeholder varible $a$.
+the algorithm continues to work on the inner-chain $\jg$ by assigning $A$ to its placeholder variable $a$.
 Rule 23 infers an annotated expression by changing into checking mode,
@@
 the return type might reflect more information than simply $\al \to \bt$.
-The variable scopings are well-maintained during the process:
+The variable scopings are maintained during the process:
 the assumption of argument type ($x:\al$) is recycled after checking against the function body;
@@
 \subsection{Multiple reduction}
-The syntax $\Gm\redto \Gm'$ denotes multiple reduction.
+The syntax $\Gm\redto \Gm'$ denotes multiple reduction steps.
 If $\Gm\redto\nil$, we say that $\Gm$ reduces successfully.

### 2019-02-11 sections/Overview.tex
DK's declarative rules
soundness and completeness
bad lemmas in DK

@@
 \subsection{DK's Declarative System}
-Subtyping and Typing. \jimmy{Can you prepare the figures. Subtyping is already in ITP.}
+% Subtyping and Typing. \jimmy{Can you prepare the figures. Subtyping is already in ITP.}
 
@@
 \text{Monotypes}\qquad&\tau,\sigma &::=&\quad 1 \mid a \mid \tau\to \sigma\\
-\text{Contexts}\qquad&\Psi &::=&\quad \cdot \mid \Psi, a\\
-%\text{Judgments}\qquad&\exps &::=&\quad \cdot \mid A \le B : \exps
+\text{Contexts}\qquad&\Psi &::=&\quad \nil \mid \Psi, a\\
+%\text{Judgments}\qquad&\exps &::=&\quad \nil \mid A \le B : \exps
 \end{array}
@@
 \inferrule*[right=$\mathtt{DeclSub}$]
-    {e \neq \lam e' \quad B \neq \all B' \quad \Psi\vdash e\To A \quad \Psi\vdash A\le B}
+%e \neq \lam e' \quad B \neq \all B' \quad 
+    {\Psi\vdash e\To A \quad \Psi\vdash A\le B}
     {\Psi \vdash e\Lto B}
@@
 If Rule $\mathtt{DeclSub}$ was chosen for derivation,
-the type of the argument is restricted by Rule $\mathtt{Decl\to I\To}$,
+the type of its argument is restricted by Rule $\mathtt{Decl\to I\To}$,
 which is not a polymorphic type.
@@
 
-\paragraph{Metatheory} Talk about the lemmas that are false, show counterexample.
+\paragraph{Metatheory} % Talk about the lemmas that are false, show counterexample.
+We have tried to follow DK's proof for formalization,
+but failed due to incorrect proof techniques or logic errors.
+The proof of Lemma 19 (Extension Equality Preservation) applies a wrong
+induction hypothesis for the ${\longrightarrow}\mathtt{Uvar}$ case.
+Fortunately the lemma could be proven without the well-formedness conditions
+$\Gm\vdash A$ and $\Gm\vdash B$.
+Similar problem happens for Lemma 14 (Subsumption), where many applications of
+induction hypotheses are not correct.
+For manual proofs, induction hypotheses are not automatically generated by programs,
+thus are easily misused.
+
+However, Lemma 29 (Parallel Admissibility) is false.
+The first conclusion of the lemma says that if $\Gm_L\longrightarrow \Delta_R$ and
+$\Gm_L,\Gm_R\longrightarrow \Delta_L,\Delta_R$ then
+$\Gm_L,\al,\Gm_R\longrightarrow \Delta_L,\al,\Delta_R$.
+If we pick $\Gm_L = \nil, \Gm_R = \bt, \Delta_L = \bt, \Delta_R = \nil$, the conditions
+$\nil\longrightarrow \bt$ and $\bt\longrightarrow \bt$ hold, but the conclusion
+$\al,\bt \longrightarrow \bt,\al$ does not hold.
 

### 2019-02-11 sections/introduction.tex
Intro WIP

@@
 polymorphism, type arguments of polymorphic functions are
-automatically instantiated. Thus the Hindley-Milner type system 
+automatically instantiated. Thus implicit polymorphism and the absense of
+type annotations mean that the Hindley-Milner type system 
 strikes a great balance between expressive power and usability.
@@
 These type systems are
-popular among language designer and their ideas have been adopted by
+popular among language designers and their ideas have been adopted by
 several modern functional languages, including Haskell and PureScript.
@@
 type annotations are required for polymorphic arguments of functions,
-but other type annotations can be omitted. A canonical example is:
-\bruno{the following text needs to be changed, since it is basically
-  as in ITP.}
+but other type annotations can be omitted. A canonical example (here written in Haskell) is:
 \begin{verbatim}
@@
 
-This paper presents the first full mechanical formalization of the metatheory for
-higher-ranked polymorphic type inference. The system that we formalize
-is the bidirectional type system by Dunfield and
-Krishnaswami (DK). The DK type system has two variants (a declarative
-and an algorithmic one) that have been been \emph{manually} proved to be
-\emph{sound}, \emph{complete} and \emph{decidable}.
+This paper presents the first full mechanical formalization of the
+metatheory for higher-ranked polymorphic type inference.
+The system
+that we formalize is the bidirectional type system by Dunfield and
+Krishnaswami (DK)~\cite{}. We chose DK's type system because it is
+quite elegant, well-documented and it comes with detailed manually
+written proofs. Furthermore the system is adopted in practice by a few
+real implementations of functional languages, including PureScript and
+\bruno{others}.  The DK type system has two variants: a declarative
+and an algorithmic one. The two variants have been been
+\emph{manually} proved to be \emph{sound}, \emph{complete} and
+\emph{decidable}.
+We present a mechanical formalization in the Abella theorem prover for
+DK's declarative type system using a different algorithm. While our
+initial goal was to formalize both DK's declarative and algorithmic
+versions, we faced technical challenges with the later, prompting us to find
+an alternative.
 
-We present 
-a mechanical formalization in the Abella theorem prover 
-for DK's declarative type system using a different algorithm. We use a
-new algorithm for a few different reasons. Firstly, our new algorithm
-employs \emph{worklist judgements}, which allow precise capture of the
-scope of vairables and simplify the formalization of scoping in a
-theorem prover. Secondly, while DK's original formalization comes with very
-well-written manual proofs, there are several details missing,
-which complicate the task of writing a formla proof. 
-Perhaps more importantly, some proofs/lemmas are wrong and it is
-not clear how to replace these by other valid proofs/lemmas. 
-\bruno{others}
+One key challenge in type inference is variable binding,
+because the algorithms typically do not rely simply on local
+environments, but instead propagate information across judgements.
+Yet, there is little work on how to deal with these complex forms of
+binding in theorem provers. To keep track of variable scoping, DK's
+algorithmic version employs input and output contexts to track
+information that is discovered throught type inference. However
+modelling output contexts in a theorem prover is non-trivial. Our
+work takes a different approach by refining the idea of
+\emph{worklists judgments}~\cite{}, proposed recently to mechanically
+formalize an algorithm for \emph{polymorphic subtyping}.
+However, a key challenge not addressed previously with
+worklist judgements was how to deal with inference judgements.
+In our work we show how worklist judgements can be extended to
+deal with the inference judgements that are necessary for type-inference.
+Furthermore we also present new techniques that enable precise tracking
+of the scope of variables.
+
+%% Polymorphic subtyping is one component of higher-ranked type inference, and
+%% it is also a key part of DK's system.
+
+Another challenge that we faced were missing details and sometimes
+incorrect proofs/lemmas in DK's formalization. While DK's original
+formalization comes with very well-written manual proofs, there are
+several details missing, which complicate the task of writing a
+mechanically verified proof. Perhaps more importantly, some
+proofs/lemmas are wrong and it is not clear how to replace these by
+other valid proofs/lemmas.
 
@@
 same results by DK, although with significantly different proofs and
-proof techniques.
-
-The calculus and its metatheory
-have been fully formalized in the Abella theorem prover\cite{AbellaDesc}. % Add at least one ref to make compile work normally
+proof techniques. The calculus and its metatheory
+have been fully formalized in the Abella theorem prover~\cite{AbellaDesc}. % Add at least one ref to make compile work normally
 
@@
 \item {\bf A full mechanical formalization of type inference with
-    higher-ranked types:}
+  higher-ranked types:} Our work presents the first full mechanical formalization
+  for type inference of higher ranked types. The formalization is done in the
+  Abella theorem prover\cite{AbellaDesc} and it is available
+  online\footnote{{\{bf Note to reviewers:} Due to the anonymous submission process
+  the url is sent as part of the supplementary materials.}
 
-\item {\bf A new algorithm for DK:}
+\item {\bf A new algorithm for DK:} Our work proposes a novel algorithm that implements
+  DK's declaratiuve bidirectional type systems. We prove the soundness ...
 
@@
 \item {\bf Abella formalization:} The calculus and its metatheory
-have been fully formalized in the Abella theorem prover\cite{AbellaDesc}.
+have been fully formalized in the .
 

### 2019-02-11 sections/Overview.tex

@@
 %-------------------------------------------------------------------------------
-\subsection{Small-Step Unification} Mention some algorithms for unification for 
-dependent types that use a small-step approach. Credit them later for some ideas that 
-we also employ.
+\subsection{Small-Step Unification}
+% Mention some algorithms for unification for 
+% dependent types that use a small-step approach. Credit them later for some ideas that 
+% we also employ.
+There are literals that make use of small-step unification for dependently typed
+inference and reconstruction algorithms~\cite{}.
+These approaches collect a list of constraints and process one at a time.
+Similar to DK's algorithm, unification variables are used to represent the types to guess.
+In order to solve unification variables to correctly scoped types
+(typically constrained by the context when the variable is introduced),
+the context information should be kept with the variables.
+Judgments that represent unifications of terms might also require such a context
+to keep the constraint collection well-formed.
 

### 2019-02-11 sections/Metatheory.tex
0211 TODOs

@@
 
-The transfer $\Gm\sto\Om$ is not a function.\bruno{Do you mean here
-  ``is not deterministic''? Would that be more appropriate?}
+The transfer $\Gm\sto\Om$ is not deterministic.
+% \bruno{Do you mean here
+%   ``is not deterministic''? Would that be more appropriate?}
 On the one hand, there are infinite possibilities to instantiate an existential variable,
@@
   what you do in your formalization?}
-
+\jimmy{TODO: a erasure function would make things more clear...}
 \subsection{Soundness}

### 2019-02-11 sections/Overview.tex
0211 TODOs

@@
 %-------------------------------------------------------------------------------
+\subsection{DK's Algorithm}
+\jimmy{TODO: some of the text above about DK's algorithm may be moved here}
+
+%-------------------------------------------------------------------------------
 \subsection{Small-Step Unification}
@@
 
+\jimmy{TODO explain the duplicated context with some examples (their judgment form)}
 Some issues to point out: Duplicated contexts (rather than shared contexts), which make 
@@
 subtyping judgment. This approach overcomes the problem's in DK's formulation
-by using a \emph{worklist}-based judgement of the form $\Gamma \vdash \Omega$
+by using a \emph{worklist}-based judgement of the form $$\Gamma \vdash \Omega$$
+\jimmy{emphasize the form of judgment}
 where $\Omega$ is a worklist, i.e. a sequence, of subtyping problems of the

### 2019-02-11 sections/conclusion.tex
0211 TODOs

@@
-\section{Conclusion and Future Work}
+\section{Conclusion}
 
-In this paper we have shown how to mechanise an algorithmic subtyping relation
-for higher-order polymorphism, together with its proofs of soundness,
-completeness and decidability, in the Abella proof assistant. 
-In ongoing work we are extending our mechanisation with a bidirectional type
-inference algorithm. The main difficulty there is communicating the
-instantiations of existential variables from the subtyping algorithm to the
-type inference. To make this possible we are exploring a continuation passing
-style formulation, which generalises the worklist approach.
-Another possible extension is to have the algorithm return an explicit witness
-for the subtyping as part of type-directed elaboration into System F.
+This is the conclusion.
+
+\jimmy{Discuss the efficient algorithm and its equivalence}

### 2019-02-12 sections/introduction.tex
0211 TODOs
Full Intro

@@
 \section{Introduction}
+\bruno{Jimmy can you please add the references? Let me know if you have questions.}
 
@@
 power has motivated language designers to look beyond Hindley-Milner.
-In particular one popular direction has been to allow 
+In particular one popular direction is to allow 
 \emph{higher-ranked polymorphism} where polymorphic types can
@@
 challenge several decidable type systems requiring some annotations 
-have been proposed~\cite{}. Two closely related type systems that 
+have been proposed~\cite{}\bruno{works on impredicative type inference + DK and Peyton Jones}.
+Two closely related type systems that 
 support \emph{predicative} higher-ranked type inference were proposed 
 by Peyton Jones et al.~\cite{} and Dunfield and
-Krishnaswami (henceforth denoted as DK). 
+Krishnaswami~\cite{} (henceforth denoted as DK). 
 These type systems are
@@
 academic research, there is little work in mechanically formalizing
-such advanced forms of type inference in theorem provers. This is at
+such advanced forms of type inference in theorem provers.
+The remarkable exception is some work done on the formalization of 
+certain parts of Hindley-Milner type inference~\cite{}\bruno{all previous work on mechanizing HM}. However
+there is still no formalization of the higher-ranked type systems
+that are employeed by modern languages like Haskell.
+This is at
 odds with the current trend of mechanical formalizations in
@@
 the use of theorem provers to model various aspects of programming
-languages. Indeed today there are routinely papers in top conferences
+languages. Today there are routinely papers in top conferences
 that use theorem provers to mechanically formalize: \emph{dynamic and
@@
   optimizations}~\cite{}, certain forms of \emph{abstract
-  interpretation} or proofs involving \emph{logical relations}. The
+  interpretation}\bruno{van Horn and colleagues} or proofs involving \emph{logical relations}\bruno{See Andreas Abel new paper}. The
 main argument for mechanical formalizations is a simple one. Proofs
@@
 \emph{error-prone}. In such proofs it is very easy to make mistakes
-that invalidate the whole development. Furthermore, readers and
+that may invalidate the whole development. Furthermore, readers and
 reviewers often do not have time to look at the proofs carefuly to
@@
 the person doing the proofs as well as readers or reviewers.
-
-DK problem in lemma ??. Castagna problem in a recent type inference paper.
-
-with the exception of some work
-done on the formalization of 
-certain aspects of Hindley-Milner type inference, we do not know
 
@@
 
-One key challenge in type inference is variable binding,
-because the algorithms typically do not rely simply on local
-environments, but instead propagate information across judgements.
-Yet, there is little work on how to deal with these complex forms of
-binding in theorem provers. To keep track of variable scoping, DK's
-algorithmic version employs input and output contexts to track
-information that is discovered throught type inference. However
-modelling output contexts in a theorem prover is non-trivial. Our
-work takes a different approach by refining the idea of
-\emph{worklists judgments}~\cite{}, proposed recently to mechanically
-formalize an algorithm for \emph{polymorphic subtyping}.
-However, a key challenge not addressed previously with
-worklist judgements was how to deal with inference judgements.
-In our work we show how worklist judgements can be extended to
-deal with the inference judgements that are necessary for type-inference.
-Furthermore we also present new techniques that enable precise tracking
-of the scope of variables.
+One key challenge in type inference is variable binding, because the
+algorithms typically do not rely simply on local environments, but
+instead propagate information across judgements.  Yet, there is little
+work on how to deal with these complex forms of binding in theorem
+provers. To keep track of variable scoping, DK's algorithmic version
+employs input and output contexts to track information that is
+discovered throught type inference. However modelling output contexts
+in a theorem prover is non-trivial. Our work takes a different
+approach by refining and extending the idea of \emph{worklists
+  judgments}~\cite{}, proposed recently to mechanically formalize an
+algorithm for \emph{polymorphic subtyping}.  However, a key challenge
+not addressed previously with worklist judgements was how to deal with
+\emph{inference judgements}.  In our work we show how worklist
+judgements can be extended to deal with the inference judgements that
+are necessary for type-inference.  The idea is to use a continuation
+passing style to enable the transfer of inferred information across
+judgements. A further refinement to the idea of worklist judgements is
+the \emph{unification between ordered contexts~\cite{}\bruno{DK and the work by McBride (see ITP citations)} and
+  worklists}. This enables precise scope tracking of free variables in
+judgements. Furthermore it avoids the duplication of context
+information across judgments in worklists that occurs in other
+techniques~\cite{}\bruno{cite the two works we cited in ITP (and maybe others)}.
 
-%% Polymorphic subtyping is one component of higher-ranked type inference, and
-%% it is also a key part of DK's system.
-
-Another challenge that we faced were missing details and sometimes
-incorrect proofs/lemmas in DK's formalization. While DK's original
-formalization comes with very well-written manual proofs, there are
-several details missing, which complicate the task of writing a
-mechanically verified proof. Perhaps more importantly, some
-proofs/lemmas are wrong and it is not clear how to replace these by
-other valid proofs/lemmas.
+Finally, another challenge that we faced were missing details and
+sometimes incorrect proofs/lemmas in DK's formalization. While DK's
+original formalization comes with very well written manual proofs,
+there are still several details missing. These complicate the task of
+writing a mechanically verified proof. Moreover some proofs and/or
+lemmas are wrong and sometimes it is not clear how to replace these by
+other valid lemmas or proofs. Although there are problems in DK's manual formalization
+we believe that their results are true and these problems do not
+invalidate their work. In fact we have nothing but praise to their detailed
+and clearly written metatheory and proofs, which provided invaluable
+help to our own work.
+With some additional thought we expect that the issues
+are fixable. We believe that for most non-trivial manual
+proofs similar problems exist, so this should not be understood as a sign of sloppiness
+on their part. Instead it should be an indicator that reinforces the arguments
+for mechanical formalizations: manual formalizations are error-prone due to the multiple
+tedious details involved in them.
+There are several other examples of manual formalizations that were found to have
+similar problems. For example, Klein et al.~\cite{}\bruno{Run your research} mechanized formalizations
+in Redex for nine ICFP 2009 papers and all were found to have mistakes.
 
@@
   for type inference of higher ranked types. The formalization is done in the
-  Abella theorem prover\cite{AbellaDesc} and it is available
-  online\footnote{{\{bf Note to reviewers:} Due to the anonymous submission process
+  Abella theorem prover~\cite{AbellaDesc} and it is available
+  online.\footnote{{\bf Note to reviewers:} Due to the anonymous submission process
   the url is sent as part of the supplementary materials.}
 
-\item {\bf A new algorithm for DK:} Our work proposes a novel algorithm that implements
-  DK's declaratiuve bidirectional type systems. We prove the soundness ...
+\item {\bf A new algorithm for DK's type system:} Our work proposes a novel algorithm that implements
+  DK's declarative bidirectional type system. We prove
+  \emph{soundness}, \emph{completeness} and
+  \emph{decidability}.
 
-\item {\bf Worlists with inference judgments:}
+\item {\bf Worklists with inference judgments:} One technical contribution is the
+  support for inference judgments using worklists. The idea is to
+  use a continuation passing style to enable the transfer of inferred information across
+  judgements.
 
-\item {\bf Abella formalization:} The calculus and its metatheory
-have been fully formalized in the .
+\item {\bf Unified judgment worklists and contexts:} Another technical contribution is the unification
+  between ordered contexts and judgement worklists. This enables precise scope tracking
+  of variables in judgements, and avoids the duplication of context information across
+  judgments in worklists.
 

### 2019-02-12 sections/algorithm.tex
Quick responses to comments; Judgment erasure (metatheory)

@@
 variable at the perfect time: it leaves the environment immediately
-after being unreferenced.\bruno{Perhaps this is better illustrated
-  with a simple example?}\tom{Yes, it would be good to show one or more
+after being unreferenced.
+\jimmy{An example on subtyping: $\nil\vdash (1\to 1)\to 1 \le (\forall a. a\to a) \to 1$
+TODO derivation and explanation}
+\bruno{Perhaps this is better illustrated with a simple example?}
+\tom{Yes, it would be good to show one or more
 example derivations to illustrate the key points.}
@@
 The design of our judgment chain is closely related to the property of
-the judgments\bruno{What property? Not very clear}.
+the judgments
+\jimmy{requires further clarification}
+\bruno{What property? Not very clear}.
 Subtyping and checking do not return anything, so variables cannot
@@
 However, inference and application inference may return a type that contains new variables.
-Take Rule 26 as an example, if it simply returns $\al \to \bt$, passes that to the next judgment and continues the type-checking process,
+Take Rule 26 as an example, if it simply returns $\al \to \bt$,
+passes that to the next judgment and continues the type-checking process,
 variables $\al$ and $\bt$ in the next judgment are out of scope

### 2019-02-12 sections/introduction.tex
Quick responses to comments; Judgment erasure (metatheory)
intro most citations

@@
 use sophisticated forms of type inference. The type systems of these languages are
-descendents of Hindley-Milner~\cite{}, which was revolutionary at the
+descendents of Hindley-Milner~\cite{hindley1969principal,milner1978theory,damas1982principal}
+, which was revolutionary at the
 time in allowing type-inference to proceed without any type
 annotation. The traditional Hindley-Milner type system supports
-top-level \emph{implicit (parametric) polymorphism}. With implicit
+top-level \emph{implicit (parametric) polymorphism}~\cite{reynolds1983types}. With implicit
 polymorphism, type arguments of polymorphic functions are
@@
 An important challenge is that full type inference for higher-ranked
-polymorphism is know to be undecidable~\cite{}. Therefore some type
+polymorphism is know to be undecidable~\cite{wells1999typability}. Therefore some type
 annotations are necessary to guide type inference. In response to this 
 challenge several decidable type systems requiring some annotations 
-have been proposed~\cite{}\bruno{works on impredicative type inference + DK and Peyton Jones}.
+have been proposed~\cite{,dunfield2013complete,jones2007practical}
+\bruno{works on impredicative type inference + DK and Peyton Jones}.
 Two closely related type systems that 
 support \emph{predicative} higher-ranked type inference were proposed 
-by Peyton Jones et al.~\cite{} and Dunfield and
-Krishnaswami~\cite{} (henceforth denoted as DK). 
+by Peyton Jones et al.~\cite{jones2007practical} and Dunfield and
+Krishnaswami~\cite{dunfield2013complete} (henceforth denoted as DK). 
 These type systems are
@@
 The remarkable exception is some work done on the formalization of 
-certain parts of Hindley-Milner type inference~\cite{}\bruno{all previous work on mechanizing HM}. However
+certain parts of Hindley-Milner type inference~\cite{naraschewski1999type,
+dubois2000proving,dubois1999certification,urban2008nominal,
+garrigue2015certified}\bruno{all previous work on mechanizing HM}. However
 there is still no formalization of the higher-ranked type systems
@@
 programming language research. In particular both the POPLMark
-challenge~\cite{} and CompCert ~\cite{} have significantly promoted
+challenge~\cite{aydemir2005mechanized} and
+CompCert ~\cite{leroy2012compcert} have significantly promoted
 the use of theorem provers to model various aspects of programming
@@
 reviewers often do not have time to look at the proofs carefuly to
-check their correctnessness. Therefore errors can go unnoticed for a
+check their correctness. Therefore errors can go unnoticed for a
 long time.  Mechanical formalizations provide, in principle, a natural
@@
 that we formalize is the bidirectional type system by Dunfield and
-Krishnaswami (DK)~\cite{}. We chose DK's type system because it is
+Krishnaswami (DK)~\cite{dunfield2013complete}.
+We chose DK's type system because it is
 quite elegant, well-documented and it comes with detailed manually
@@
 approach by refining and extending the idea of \emph{worklists
-  judgments}~\cite{}, proposed recently to mechanically formalize an
+  judgments}~\cite{itp2018}, proposed recently to mechanically formalize an
 algorithm for \emph{polymorphic subtyping}.  However, a key challenge
@@
 judgements. A further refinement to the idea of worklist judgements is
-the \emph{unification between ordered contexts~\cite{}\bruno{DK and the work by McBride (see ITP citations)} and
-  worklists}. This enables precise scope tracking of free variables in
+the \emph{unification between ordered contexts~\cite{gundry2010type,dunfield2013complete}
+\bruno{DK and the work by McBride (see ITP citations)} and
+  worklists}.
+This enables precise scope tracking of free variables in
 judgements. Furthermore it avoids the duplication of context
 information across judgments in worklists that occurs in other
-techniques~\cite{}\bruno{cite the two works we cited in ITP (and maybe others)}.
+techniques~\cite{Reed2009,Abel2011higher}
+\bruno{cite the two works we cited in ITP (and maybe others)}.
 
@@
 There are several other examples of manual formalizations that were found to have
-similar problems. For example, Klein et al.~\cite{}\bruno{Run your research} mechanized formalizations
+similar problems. For example, Klein et al.~\cite{KleinRunYourResearch}
+\bruno{Run your research} mechanized formalizations
 in Redex for nine ICFP 2009 papers and all were found to have mistakes.
@@
 proof techniques. The calculus and its metatheory
-have been fully formalized in the Abella theorem prover~\cite{AbellaDesc}. % Add at least one ref to make compile work normally
+have been fully formalized in the Abella theorem prover~\cite{AbellaDesc}.
 
@@
 
+\jimmy{Notes @20190211 4 points of novalty:\\
+1) Dealing with inference judgements and CPS-style chains\\
+2) The form of the judgment itself with a single shared context\\
+3) The way we deal with scope (which may follow from 2)\\
+4) Immediate substitution (judgment list)
+}
+
 \end{itemize}

### 2019-02-12 sections/Overview.tex
small step unification

@@
 There are literals that make use of small-step unification for dependently typed
-inference and reconstruction algorithms~\cite{}.
+inference and reconstruction algorithms~\cite{Reed2009,Abel2011higher}.
 These approaches collect a list of constraints and process one at a time.
@@
 
-\jimmy{TODO explain the duplicated context with some examples (their judgment form)}
-Some issues to point out: Duplicated contexts (rather than shared contexts), which make 
-the formalization harder since requires ``synchronizing'' things ...
+As an example, Abel et al. defines their constraint $K$ by
+\begin{gather*}
+\begin{aligned}
+K &::= {\top} \mid {\bot} &&\text{Trivial constraint and inconsistency.}\\
+    & \ \mid \  \Psi\vdash M = N : C &&\text{Unify term $M$ with $N$.}\\
+    & \ \mid \  \Psi\mid R:A \vdash E = E' &&\text{Unify evaluation context $E$ with $E'$.}\\
+    & \ \mid \  \Psi\vdash u\leftarrow M: C &&\text{Solution for $u$ (metavariable) found.}
+\end{aligned}
+\end{gather*}
+where the unification constraints and metavariable solutions are bound by a proper context.
+A unification problem in their system is described by $\Delta \Vdash \overline{K}$,
+where $\Delta$ is a collection of metavariables with their defined scopes.
+This approach clearly states the scroping of each unification problem,
+therefore rules out invalid instantiations to metavariables.
+
+However, such ``duplicated contexts'' are not ideal for our formalization.
+Following DK's algorithm, an existential variable $\al$ defined in the context
+could be decomposed into a function type $\al[1] \to \al[2]$,
+so the declaration of $\al$ should be replaced by two declarations $\al[1], \al[2]$.
+Such operation causes all the context that contains $\al$ to change,
+which brings difficulty on the synchronization of multiple contexts,
+as variable declarations are not centralized.
+
+We find a way of encoding multiple contexts by ``compressing'' them into a single worklist.
+As our type inference rules applied to a judgment,
+we typically keep the context or add variable declarations to the old context
+before analysing its sub-judgment(s).
+This enables us to write the judgment worklist in such a form:
+$$\Gm, \{\text{variable declarations}\}, \{\text{judgment}\},
+\{\text{variable declarations}\}, \{\text{judgment}\}, \ldots .$$
+Typically, we simplify the right-most judgment and push back smaller tasks,
+where the old ``context'' automatically gets inherited.
+When some rules solve or partially solve an existential variable,
+we could easily propagate the solution to all the judgments,
+and safely modify its declaration, such as remove from the worklist.
+New judgment(s) may also introduce new \emph{local} variables by declaring
+that variable right before the new judgment(s).
+After the new judgments are satisfied, these local variables are properly recycled.
+
+% \jimmy{TODO explain the duplicated context with some examples (their judgment form)}
+% Some issues to point out: Duplicated contexts (rather than shared contexts), which make 
+% the formalization harder since requires ``synchronizing'' things ...
 

### 2019-02-12 sections/introduction.tex
small edits in the introduction

@@
 use sophisticated forms of type inference. The type systems of these languages are
-descendents of Hindley-Milner~\cite{hindley1969principal,milner1978theory,damas1982principal}
+descendants of Hindley-Milner~\cite{hindley1969principal,milner1978theory,damas1982principal}
 , which was revolutionary at the
@@
 
-Despite the importance of type inference in practice and also in
+Although type inference is important in practice and also in
 academic research, there is little work in mechanically formalizing
@@
 there is still no formalization of the higher-ranked type systems
-that are employeed by modern languages like Haskell.
+that are employed by modern languages like Haskell.
 This is at
@@
 the use of theorem provers to model various aspects of programming
-languages. Today there are routinely papers in top conferences
-that use theorem provers to mechanically formalize: \emph{dynamic and
+languages. Today papers in top conferences routinely
+use theorem provers to mechanically formalize: \emph{dynamic and
   static semantics} and their correctness properties~\cite{},
@@
 that may invalidate the whole development. Furthermore, readers and
-reviewers often do not have time to look at the proofs carefuly to
+reviewers often do not have time to look at the proofs carefully to
 check their correctness. Therefore errors can go unnoticed for a
 long time.  Mechanical formalizations provide, in principle, a natural
-solution for such problems. Theorem provers can automaticaly check and
+solution for these problems. Theorem provers can automatically check and
 validate the proofs, which removes the burden from checking from both
@@
 initial goal was to formalize both DK's declarative and algorithmic
-versions, we faced technical challenges with the later, prompting us to find
-an alternative.
+versions, we faced technical challenges with the latter, prompting us to find
+an alternative formulation formulation.
 
@@
 
-Finally, another challenge that we faced were missing details and
-sometimes incorrect proofs/lemmas in DK's formalization. While DK's
+Finally, another challenge that we faced were missing details as well as
+a few incorrect proofs and lemmas in DK's formalization. While DK's
 original formalization comes with very well written manual proofs,
@@
 writing a mechanically verified proof. Moreover some proofs and/or
-lemmas are wrong and sometimes it is not clear how to replace these by
-other valid lemmas or proofs. Although there are problems in DK's manual formalization
-we believe that their results are true and these problems do not
-invalidate their work. In fact we have nothing but praise to their detailed
+lemmas are wrong and it is not in all cases clear to us how to fix them. Yet,
+despite these problems in DK's manual formalization,
+we believe that these problems do not
+invalidate their work and that their results are still true. In fact we have nothing but praise for their detailed
 and clearly written metatheory and proofs, which provided invaluable
@@
 Despite the use of a different algorithm we prove the
-same results by DK, although with significantly different proofs and
+same results as DK, although with significantly different proofs and
 proof techniques. The calculus and its metatheory
@@
   online.\footnote{{\bf Note to reviewers:} Due to the anonymous submission process
-  the url is sent as part of the supplementary materials.}
+  it is sent as part of the supplementary materials.}
 

### 2019-02-12 sections/introduction.tex
and/or -> and
footnote

@@
 there are still several details missing. These complicate the task of
-writing a mechanically verified proof. Moreover some proofs and/or
-lemmas are wrong and it is not in all cases clear to us how to fix them. Yet,
-despite these problems in DK's manual formalization,
+writing a mechanically verified proof. Moreover some proofs and
+lemmas are wrong and it is not in all cases clear to us how to fix them.\footnote{
+Perhaps, with some additional thought, a work-around can be found, but, as our algorithm
+differs significantly from DK's, it would not further our mechanisation goal and so we did not pursue this.}
+Despite the problems in DK's manual formalization,
 we believe that these problems do not
@@
 help to our own work.
-With some additional thought we expect that the issues
-are fixable. We believe that for most non-trivial manual
+We believe that for most non-trivial manual
 proofs similar problems exist, so this should not be understood as a sign of sloppiness

### 2019-02-12 sections/Discussion.tex
typos and discussion

\section{Discussion}

\subsection{Overlapping Rules}

\subsection{Implementation}
Anything to say about the implementation? Do we have one?

\subsection{Making DK's scoping more precise}

\subsection{Scoped type variables}

\subsection{Elaboration}

OTHERS?

### 2019-02-12 sections/Overview.tex
move algorithm paragraphs from 2.1 to 2.2
DK algorithm description

@@
 
-\paragraph{Metatheory} % Talk about the lemmas that are false, show counterexample.
-We have tried to follow DK's proof for formalization,
-but failed due to incorrect proof techniques or logic errors.
+%-------------------------------------------------------------------------------
+\subsection{DK's Algorithm}
+
+The DK algorithm revolves around their notion of \emph{algorithmic context}.
+\[
+\begin{array}{l@{\qquad}lcl}
+\text{Algorithmic Contexts}\qquad&\Gamma,\Delta,\Theta &::=&\quad \nil \mid
+\Gamma, a \mid \Gamma, \hat{\alpha} \mid \Gamma, \hat{\alpha} = \tau \mid
+\Gamma, \blacktriangleright_{\hat{\alpha}}
+\end{array}
+\]
+In addition to the regular (universally quantified) type variables $a$, the
+algorithmic context also contains \emph{existential} type variables
+$\hat{\alpha}$. These are placeholders for monotypes $\tau$ that are still to
+be determined by the inference algorithm. When the existential variable is
+``solved'', its entry in the context is replaced by the assignment
+$\hat{\alpha} = \tau$. With the notation $[\Gamma]$ an algorithmic context
+$\Gamma$ denotes the substitution that replaces all its solved existential type
+variables with their solution.
+
+All algorithmic judgments thread an algorithmic context. They have the form
+$\Gamma \vdash \ldots \dashv \Delta$, where $\Gamma$ is the input context and
+$\Delta$ is the output context:
+$\Gamma \vdash A \le B \dashv \Delta$  for subtyping, 
+$\Gamma \vdash e \Leftarrow A \dashv \Delta$  for type checking, and so on. 
+The output context is a functional update of the input context that records newly
+introduced existentials and solutions.
+
+Solutions are incrementally propagated by applying the algorithmic output
+context of a previous task as substitution to the next task. This can be seen
+in the subsumption rule:
+\[
+\inferrule*
+  {\Gamma \vdash e \Rightarrow A \dashv \Theta \\ 
+   \Theta \vdash [\Theta]A \le [\Theta]B \dashv \Delta
+  }
+  { \Gamma \vdash e \Leftarrow B \dashv \Delta}
+\]
+The inference task yields an output context $\Theta$ which is applied as a substitution
+to the types $A$ and $B$ before performing the subtyping check to propagate any solutions
+of existential variables that appear in $A$ and $B$.
+
+The sequential order of entries in the algoritmic context, in combination with
+the threading of contexts,  does not perfectly capture the scoping of all
+existential variables. For this reason the DK algorithm uses scope markers
+$\blacktriangleright_{\hat{\alpha}}$ in a few places. An example is given in the following
+rule:
+\[
+\inferrule*
+  {\Gamma,\blacktriangleright_{\hat{\alpha}}, \hat{\alpha} \vdash [\hat{\alpha}/a]A \le B \dashv \Delta,\blacktriangleright_{\hat{\alpha}},\Theta}
+  {\Gamma \vdash \forall a. A \le B \dashv \Delta}
+\]
+To indicate that the scope of $\hat{\alpha}$ is local to the subtyping check
+$[\hat{\alpha}/a]A \le B$, the marker is pushed onto its input stack and popped
+from the output stack together with the subsequent part $\Theta$, which may
+refer to $\hat{\alpha}$. (Remember that later entries may refer to earlier
+ones, but not vice versa.) This way $\hat{\alpha}$ does not escape its scope.
+
+At first sight, the DK algorithm would seem a good basis for mechanisation. After
+all it comes with a careful description and extensive manual proofs.
+Unfortunately, we ran into several obstacles that have prompted us to formulate
+a different, more mechanisation-friendly algorithm.
+
+%- - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - 
+\paragraph{Broken Metatheory} % Talk about the lemmas that are false, show counterexample.
+While going through the manual proofs of DK's algorithm, we found several problems.
 The proof of Lemma 19 (Extension Equality Preservation) applies a wrong
@@
 mechanised setting.
-
-%-------------------------------------------------------------------------------
-\subsection{DK's Algorithm}
-\jimmy{TODO: some of the text above about DK's algorithm may be moved here}
 

### 2019-02-12 sections/introduction.tex

@@
 use sophisticated forms of type inference. The type systems of these languages are
-descendants of Hindley-Milner~\cite{hindley1969principal,milner1978theory,damas1982principal}
-, which was revolutionary at the
+descendants of
+Hindley-Milner~\cite{hindley1969principal,milner1978theory,damas1982principal}, 
+which was revolutionary at the
 time in allowing type-inference to proceed without any type
@@
 versions, we faced technical challenges with the latter, prompting us to find
-an alternative formulation formulation.
+an alternative formulation.
 

### 2019-02-12 main.tex
Intro WIP
Lemma 29 counterexample move to appendix

@@
 which complicate the task of writing a formal proof. 
-Perhaps more importantly, some proofs/lemmas are wrong and it is
-not clear how to replace these by other valid proofs/lemmas. 
+Moreover, some proofs/lemmas are wrong and it is
+not clear how to replace these by other valid ones. 
 \bruno{others?} Despite the use of a different algorithm we prove the
@@
 
-Text of appendix \ldots
+\jimmy{Lemma 29 is false. TODO structure and more text}
+The first conclusion of the lemma says that if $\Gm_L\longrightarrow \Delta_R$ and
+$\Gm_L,\Gm_R\longrightarrow \Delta_L,\Delta_R$ then
+$\Gm_L,\al,\Gm_R\longrightarrow \Delta_L,\al,\Delta_R$.
+If we pick $\Gm_L = \nil, \Gm_R = \bt, \Delta_L = \bt, \Delta_R = \nil$, the conditions
+$\nil\longrightarrow \bt$ and $\bt\longrightarrow \bt$ hold, but the conclusion
+$\al,\bt \longrightarrow \bt,\al$ does not hold.
 

### 2019-02-12 sections/Overview.tex
quick notes
Lemma 29 counterexample move to appendix

@@
 \paragraph{Overlapping rules}
+\jimmy{Move to the Alg. Section}
 Some of the declarative rules overlaps with each other.
@@
 
-However, Lemma 29 (Parallel Admissibility) is false.
-The first conclusion of the lemma says that if $\Gm_L\longrightarrow \Delta_R$ and
-$\Gm_L,\Gm_R\longrightarrow \Delta_L,\Delta_R$ then
-$\Gm_L,\al,\Gm_R\longrightarrow \Delta_L,\al,\Delta_R$.
-If we pick $\Gm_L = \nil, \Gm_R = \bt, \Delta_L = \bt, \Delta_R = \nil$, the conditions
-$\nil\longrightarrow \bt$ and $\bt\longrightarrow \bt$ hold, but the conclusion
-$\al,\bt \longrightarrow \bt,\al$ does not hold.
+Moreover, Lemma 29 (Parallel Admissibility) is false.
+One may find a counterexample with detailed explaination in the appendix.
+Lemma 29 is an important lemma for the proof of soundness and completeness,
+which are two main results of their metatheory.
+Instantiation soundness (i.e. a part of subtyping soundness) and typing completeness
+both directly require Lemma 29.
+Lemma 54 (Typing Extension) requires Lemma 29 and is applied 13 times
+in the proof of typing soundness and completeness.
 
@@
 
+\jimmy{Compare with DK's algo: subsumption rule might cause leak}
+
 \paragraph{Inference Judgements}

### 2019-02-13 sections/Overview.tex
Declarative subtyping description

@@
 The syntax of this declarative system is shown in Figure~\ref{fig:decl:syntax}.
+A declarative type could be a unit type 1, a type variable $a$,
+a universal quantification $\forall a. A$ or a function type $A \to B$.
+Nested universal quantifiers are allowed for types,
+but monotypes do not have any universal quantifier.
+Context $\Psi$ is just a collection of type variable declarations.
 
@@
 In Figures~\ref{fig:decl:sub} and \ref{fig:decl:typing},
-we give the subtyping and typing relations for the declarative system.
+we give the subtyping and typing relations for the declarative system,
+which is exactly the same as DK's.
+The subtyping relations originate from Odersky and La\"ufer~\cite{odersky1996putting},
+and is adopted by DK as a part of their typing system.
 
-\paragraph{Overlapping rules}
+\paragraph{Declarative Subtyping}
+For implicit polymorphic type systems,
+subtyping relation compares the degree of polymorphism of types.
+In short, if a polymorphic type $A$ always succeed to instantiate to any instantiation of $B$,
+then A is ``at least as polymorphic as'' $B$,
+or we just say that $A$ is ``more polymorphic than'' $B$, written as $A \le B$.
+
+Subtyping rules $\mathtt{{\le}Var}$, $\mathtt{{\le}Unit}$ and $\mathtt{{\le}{\to}}$
+handle simple cases where no universal quantifiers get involved.
+The subtyping rule for function types $\mathtt{{\le}{\to}}$ is standard,
+being covariant on the return type and contravariant on the argument type.
+Rule $\mathtt{{\le}\forall L}$ states that if $A$ is a subtype of $B$
+under the context ($\Psi, a$), where $a$ is fresh in $A$, then $A\le\all B$.
+Intuitively, if $A$ is more general than $\all B$ (where the universal quantifier
+already indicates that $\all B$ is a general type),
+then $A$ must instantiate to $[\tau/a]B$ for every $\tau$.
+
+The most interesting rule is $\mathtt{{\le}\forall R}$.
+If any instantiation of $\all A$, $[\tau/a]A$, is a subtype of $B$,
+then $\all A \le B$.
+The monotype $\tau$ we used to instantiation is \emph{guessed} for the declarative rule,
+but the algorithmic system should be able to compute the guess properly.
+The fact that $\tau$ is a monotype rules out the possibility of
+polymorphic (or impredicative) instantiation.
+Peyton Jones et al.~\cite{} also has such restriction of
+predicative instantiation in their type system.
+Both systems are adopt by several practical programming languages.
+
+\paragraph{Declarative Typing}
+\jimmy{TODO decl. typing description}
+
+\paragraph{Overlapping Rules}
 \jimmy{Move to the Alg. Section}
@@
 Declarative subtyping rules $\mathtt{{\le}\forall L}$ and $\mathtt{{\le}\forall R}$
-both match the conclusion $\Psi\vdash \forall a. A \le \forall a. B$.
+both match the conclusion $\Psi\vdash \all A \le \all B$.
 In such case, choose $\mathtt{{\le}\forall R}$ first is always better,

### 2019-02-13 sections/Overview.tex
polish language in 2.1 + comment

--- 2019-02-12 sections/Overview.tex
@@
 The syntax of this declarative system is shown in Figure~\ref{fig:decl:syntax}.
+A declarative type $A$ is either the unit type $1$, a type variable $a$,
+a universal quantification $\forall a. A$ or a function type $A \to B$.
+Nested universal quantifiers are allowed in types,
+but monotypes $\tau$ do not contain any universal quantifier.
+Contexts $\Psi$ are just sequences of type variable declarations.
 
@@
 
-In Figures~\ref{fig:decl:sub} and \ref{fig:decl:typing},
-we give the subtyping and typing relations for the declarative system.
+Figures~\ref{fig:decl:sub} and \ref{fig:decl:typing},
+give the subtyping and typing relations for the declarative DK system.
+The DK subtyping relation was adopted from \citet{odersky1996putting}.
 
-\paragraph{Overlapping rules}
+\paragraph{Declarative Subtyping}
+Figure~\ref{fig:decl:sub} shows DK's declarative subtyping judgment $\Psi \vdash A \le B$,
+which was adopted from \citet{odersky1996putting}. It compares the
+degree of polymorphism between between $A$ and $B$ in DK's implicit polymorphic type system. 
+Essentially, if $A$ can always be instantiated to match any instantiation of $B$,
+then A is ``at least as polymorphic as'' $B$. We also 
+say that $A$ is ``more polymorphic than'' $B$ and write $A \le B$.
+
+Subtyping rules $\mathtt{{\le}Var}$, $\mathtt{{\le}Unit}$ and $\mathtt{{\le}{\to}}$
+handle simple cases that do not involve universal quantifiers.
+The subtyping rule for function types $\mathtt{{\le}{\to}}$ is standard,
+being covariant on the return type and contravariant on the argument type.
+Rule $\mathtt{{\le}\forall R}$ states that if $A$ is a subtype of $B$
+in the context $\Psi, a$, where $a$ is fresh in $A$, then $A\le\all B$.
+Intuitively, if $A$ is more general than $\all B$ (where the universal quantifier
+already indicates that $\all B$ is a general type),
+then $A$ must instantiate to $[\tau/a]B$ for every $\tau$.
+
+The most interesting rule is $\mathtt{{\le}\forall L}$.
+If some instantiation of $\all A$, $[\tau/a]A$, is a subtype of $B$,
+then $\all A \le B$.
+The monotype $\tau$ we used to instantiate $a$ is \emph{guessed} in this
+declarative rule, but the algorithmic system does not guess and defers the
+instantiation until it can determine the monotype deterministically.
+The fact that $\tau$ is a monotype rules out the possibility of
+polymorphic (or impredicative) instantiation.\tom{say why this restriction is imposed}
+Peyton Jones et al.~\cite{} also impose the restriction of
+predicative instantiation in their type system.
+Both systems are adopt by several practical programming languages.
+
+\paragraph{Declarative Typing}
+\jimmy{TODO decl. typing description}
+
+\paragraph{Overlapping Rules}
 \jimmy{Move to the Alg. Section}

### 2019-02-13 sections/Overview.tex
revise 2.2 + comment

@@
 \paragraph{Broken Metatheory} % Talk about the lemmas that are false, show counterexample.
-While going through the manual proofs of DK's algorithm, we found several problems.
-The proof of Lemma 19 (Extension Equality Preservation) applies a wrong
-induction hypothesis for the ${\longrightarrow}\mathtt{Uvar}$ case.
-Fortunately the lemma could be proven without the well-formedness conditions
-$\Gm\vdash A$ and $\Gm\vdash B$.
-Similar problem happens for Lemma 14 (Subsumption), where many applications of
-induction hypotheses are not correct.
-For manual proofs, induction hypotheses are not automatically generated by programs,
-thus are easily misused.
+While going through the manual proofs of DK's algorithm, we found several
+problems.  Indeed, two proofs of lemmas---Lemma 19 (Extension Equality
+Preservation) and Lemma 14 (Subsumption)--- wrongly apply induction hypotheses
+in several cases. Fortunately, we have found simple work-arounds that fix these
+proofs without affecting the appeals to these lemmas.
 
-Moreover, Lemma 29 (Parallel Admissibility) is false.
-One may find a counterexample with detailed explaination in the appendix.
-Lemma 29 is an important lemma for the proof of soundness and completeness,
-which are two main results of their metatheory.
-Instantiation soundness (i.e. a part of subtyping soundness) and typing completeness
-both directly require Lemma 29.
-Lemma 54 (Typing Extension) requires Lemma 29 and is applied 13 times
-in the proof of typing soundness and completeness.
+% The proof of Lemma 19 (Extension Equality Preservation) applies a wrong
+% induction hypothesis for the ${\longrightarrow}\mathtt{Uvar}$ case.
+% Fortunately the lemma could be proven without the well-formedness conditions
+% $\Gm\vdash A$ and $\Gm\vdash B$.
+% Similar problem happens for Lemma 14 (Subsumption), where many applications of
+% induction hypotheses are not correct.
+% For manual proofs, induction hypotheses are not automatically generated by programs,
+% thus are easily misused.
+
+More seriously, we have also found a lemma that simply does not hold, 
+Lemma 29 (Parallel Admissibility). See Appendix~\ref{} for a detailed explanation
+and counterexample. This lemma is a cornerstone of the two metatheoretical results 
+of the algorithm, soundness and completeness with respect to the declarative system.
+In particular, both instantiation soundness (i.e. a part of subtyping
+soundness) and typing completeness both directly require the broken lemma.
+Moreover, Lemma 54 (Typing Extension) also requires the broken lemma and is
+itself used 13 times in the proof of typing soundness and completeness.
+Unfortunately, we have not yet found a way to fix this problem.
+\tom{You have communicated this to the authors and not received a fix, right?
+     We could mention that too then.}
 
@@
 traditional constraint-based type-inference algorithms like Algorithm W which
-features an implicit global scope for all type variables. Indeed, DK uses an
-explicit and ordered context that tracks the relative scope of universal and
+features an implicit global scope for all type variables. Indeed, DK uses its
+explicit and ordered context $\Gamma$ that tracks the relative scope of universal and
 existential variables and it is careful to only instantiate existential
@@
 Moreover, DK's algorithm carefully propagates instantiations by recording them
-into the context and threading this context through all judgments.  means of
-two This means that every judgments takes two contexts---an input and an output
-context---rather than the conventional single context. The output context
-records any new variable instantations; to propagate these instantiations to
-the remaining judgements, their predecessor's output context---which is their
-input context---is applied to them as a substitution.
-
-While it works, DK's approach is still fairly involved and thus hard to reason
-about in a mechanised setting. Indeed, the instantiations have to be recorded
-in the context and are applied incrementally to each remaining judgment in
-turn, rather than immediately to all remaining judgments at once. Also the
-stack discipline of the ordered contexts does not mesh well with the use of
-output contexts; explicit marker entries are needed in two places to
-demarcate the end of an existential variable's scope. These are compelling 
-reasons to look for a simpler algorithm that is easier to reason about in a 
-mechanised setting.
+into the context and threading this context through all judgments. 
+% two This means that every judgments takes two contexts---an input and an output
+% context---rather than the conventional single context. The output context
+% records any new variable instantations; to propagate these instantiations to
+% the remaining judgements, their predecessor's output context---which is their
+% input context---is applied to them as a substitution.
+While this works well on paper, this approach is still fairly involved and thus
+hard to reason about in a mechanised setting. Indeed, the instantiations have
+to be recorded in the context and are applied incrementally to each remaining
+judgment in turn, rather than immediately to all remaining judgments at once.
+Also, as we have mentioned above, the stack discipline of the ordered contexts
+does not mesh well with the use of output contexts; explicit marker entries are
+needed in two places to demarcate the end of an existential variable's scope.
+These are compelling reasons to look for a simpler algorithm that is easier to
+reason about in a mechanised setting.
 

### 2019-02-14 sections/Overview.tex
revised 2.3

@@
 %-------------------------------------------------------------------------------
-\subsection{Small-Step Unification}
-% Mention some algorithms for unification for 
-% dependent types that use a small-step approach. Credit them later for some ideas that 
-% we also employ.
-There are literals that make use of small-step unification for dependently typed
-inference and reconstruction algorithms~\cite{Reed2009,Abel2011higher}.
-These approaches collect a list of constraints and process one at a time.
-Similar to DK's algorithm, unification variables are used to represent the types to guess.
-In order to solve unification variables to correctly scoped types
-(typically constrained by the context when the variable is introduced),
-the context information should be kept with the variables.
-Judgments that represent unifications of terms might also require such a context
-to keep the constraint collection well-formed.
+\subsection{Judgment Lists}
 
-As an example, Abel et al. defines their constraint $K$ by
-\begin{gather*}
-\begin{aligned}
-K &::= {\top} \mid {\bot} &&\text{Trivial constraint and inconsistency.}\\
-    & \ \mid \  \Psi\vdash M = N : C &&\text{Unify term $M$ with $N$.}\\
-    & \ \mid \  \Psi\mid R:A \vdash E = E' &&\text{Unify evaluation context $E$ with $E'$.}\\
-    & \ \mid \  \Psi\vdash u\leftarrow M: C &&\text{Solution for $u$ (metavariable) found.}
-\end{aligned}
-\end{gather*}
-where the unification constraints and metavariable solutions are bound by a proper context.
-A unification problem in their system is described by $\Delta \Vdash \overline{K}$,
-where $\Delta$ is a collection of metavariables with their defined scopes.
-This approach clearly states the scroping of each unification problem,
-therefore rules out invalid instantiations to metavariables.
+To avoid the problem of incrementally applying a substitution to remaining
+tasks, we can find inspiration in the formulation of constraint solving
+algorithms. For instance, the well-known unification
+algorithm by \citet{unification} decomposes the problem of unifying two terms $s \stackrel{.}{=} t$ into a set
+of related unification problems between pairs of terms $s_i \stackrel{.}{=} t_i)$. These smaller
+problems are not tackled independently, but kept together in a set $S$. 
+The algorithm itself is typically formulated as a small-step-style state
+transition system $S \rightarrowtail S'$ that rewrites the set of unification
+problems until it is in solved form or until a contradiction has been found.
+For instance, the variable elimination rule is written as:
+\[
+   x \stackrel{.}{=} t, S  ~~\rightarrowtail~~  x \stackrel{.}{=} t, [t/x]S   \qquad\qquad{if}~x \not\in t~\text{and}~{x \in S}
+\]
+Because the whole set is explicitly available, the variable $x$ can be
+simultaneously substituted everywhere.
 
-However, such ``duplicated contexts'' are not ideal for our formalization.
-Following DK's algorithm, an existential variable $\al$ defined in the context
-could be decomposed into a function type $\al[1] \to \al[2]$,
-so the declaration of $\al$ should be replaced by two declarations $\al[1], \al[2]$.
-Such operation causes all the context that contains $\al$ to change,
-which brings difficulty on the synchronization of multiple contexts,
-as variable declarations are not centralized.
+In the above standard unification problem, all variables are implicitly in the same
+global scope. In contrast, recent unification algorithms for dependently-typed languages
+are more explicit about scopes. For instance, \citet{Reed2009} represents a unification
+problem as $\Delta \vdash P$ where $P$ is a set of equations to be solved and $\Delta$ is
+a (modal) context. \citet{Abel2011higher} even use multiple contexts within a unification problem.
+Such a problem is denoted $\Delta \Vdash \mathcal{K}$ where the meta-context
+$\Delta$ contains all the typings of meta-variables in the constraint set
+$\mathcal{K}$. The latter consists of constraints like $\Psi \vdash M = N : C$
+that are equipped with their individual context $\Psi$. While acurrately tracking
+the scoping of regular and meta-variables, this approach is not ideal because it
+repeatedly copies contexts when decomposing a unification problem, and later has to 
+substitute solutions into these copies.
 
-We find a way of encoding multiple contexts by ``compressing'' them into a single worklist.
-As our type inference rules applied to a judgment,
-we typically keep the context or add variable declarations to the old context
-before analysing its sub-judgment(s).
-This enables us to write the judgment worklist in such a form:
-$$\Gm, \{\text{variable declarations}\}, \{\text{judgment}\},
-\{\text{variable declarations}\}, \{\text{judgment}\}, \ldots .$$
-Typically, we simplify the right-most judgment and push back smaller tasks,
-where the old ``context'' automatically gets inherited.
-When some rules solve or partially solve an existential variable,
-we could easily propagate the solution to all the judgments,
-and safely modify its declaration, such as remove from the worklist.
-New judgment(s) may also introduce new \emph{local} variables by declaring
-that variable right before the new judgment(s).
-After the new judgments are satisfied, these local variables are properly recycled.
-
-% \jimmy{TODO explain the duplicated context with some examples (their judgment form)}
-% Some issues to point out: Duplicated contexts (rather than shared contexts), which make 
-% the formalization harder since requires ``synchronizing'' things ...
-
+% %-------------------------------------------------------------------------------
+% \subsection{Small-Step Unification}
+% % Mention some algorithms for unification for 
+% % dependent types that use a small-step approach. Credit them later for some ideas that 
+% % we also employ.
+% There are literals that make use of small-step unification for dependently typed
+% inference and reconstruction algorithms~\cite{Reed2009,Abel2011higher}.
+% These approaches collect a list of constraints and process one at a time.
+% Similar to DK's algorithm, unification variables are used to represent the types to guess.
+% In order to solve unification variables to correctly scoped types
+% (typically constrained by the context when the variable is introduced),
+% the context information should be kept with the variables.
+% Judgments that represent unifications of terms might also require such a context
+% to keep the constraint collection well-formed.
+% 
+% As an example, Abel et al. defines their constraint $K$ by
+% \begin{gather*}
+% \begin{aligned}
+% K &::= {\top} \mid {\bot} &&\text{Trivial constraint and inconsistency.}\\
+%     & \ \mid \  \Psi\vdash M = N : C &&\text{Unify term $M$ with $N$.}\\
+%     & \ \mid \  \Psi\mid R:A \vdash E = E' &&\text{Unify evaluation context $E$ with $E'$.}\\
+%     & \ \mid \  \Psi\vdash u\leftarrow M: C &&\text{Solution for $u$ (metavariable) found.}
+% \end{aligned}
+% \end{gather*}
+% where the unification constraints and metavariable solutions are bound by a proper context.
+% A unification problem in their system is described by $\Delta \Vdash \overline{K}$,
+% where $\Delta$ is a collection of metavariables with their defined scopes.
+% This approach clearly states the scroping of each unification problem,
+% therefore rules out invalid instantiations to metavariables.
+% 
+% However, such ``duplicated contexts'' are not ideal for our formalization.
+% Following DK's algorithm, an existential variable $\al$ defined in the context
+% could be decomposed into a function type $\al[1] \to \al[2]$,
+% so the declaration of $\al$ should be replaced by two declarations $\al[1], \al[2]$.
+% Such operation causes all the context that contains $\al$ to change,
+% which brings difficulty on the synchronization of multiple contexts,
+% as variable declarations are not centralized.
+% 
+% We find a way of encoding multiple contexts by ``compressing'' them into a single worklist.
+% As our type inference rules applied to a judgment,
+% we typically keep the context or add variable declarations to the old context
+% before analysing its sub-judgment(s).
+% This enables us to write the judgment worklist in such a form:
+% $$\Gm, \{\text{variable declarations}\}, \{\text{judgment}\},
+% \{\text{variable declarations}\}, \{\text{judgment}\}, \ldots .$$
+% Typically, we simplify the right-most judgment and push back smaller tasks,
+% where the old ``context'' automatically gets inherited.
+% When some rules solve or partially solve an existential variable,
+% we could easily propagate the solution to all the judgments,
+% and safely modify its declaration, such as remove from the worklist.
+% New judgment(s) may also introduce new \emph{local} variables by declaring
+% that variable right before the new judgment(s).
+% After the new judgments are satisfied, these local variables are properly recycled.
+% 
+% % \jimmy{TODO explain the duplicated context with some examples (their judgment form)}
+% % Some issues to point out: Duplicated contexts (rather than shared contexts), which make 
+% % the formalization harder since requires ``synchronizing'' things ...
+% 
 %-------------------------------------------------------------------------------
-\subsection{Worklist Algorithm for Subtyping}
+\subsection{Single-Context Worklist Algorithm for Subtyping}
 

### 2019-02-18 sections/Overview.tex
Definition of declarative terms

--- 2019-02-13 sections/Overview.tex
@@
 \begin{array}{l@{\qquad}lcl}
-\text{Type variables}\qquad&a, b\\[3mm]
+\text{Type variables}\qquad&a, b
+\\[3mm]
 \text{Types}\qquad&A, B, C &::=&\quad 1 \mid a \mid \forall a. A \mid A\to B\\
-\text{Monotypes}\qquad&\tau,\sigma &::=&\quad 1 \mid a \mid \tau\to \sigma\\
-\text{Contexts}\qquad&\Psi &::=&\quad \nil \mid \Psi, a\\
-%\text{Judgments}\qquad&\exps &::=&\quad \nil \mid A \le B : \exps
+\text{Monotypes}\qquad&\tau,\sigma &::=&\quad 1 \mid a \mid \tau\to \sigma
+\\[3mm]
+\text{Expressions}\qquad&e &::=&\quad x \mid () \mid \lam e \mid e_1~e_2 \mid (e:A)
+\\[3mm]
+\text{Contexts}\qquad&\Psi &::=&\quad \nil \mid \Psi, a \mid \Psi, x:A
 \end{array}
@@
 The syntax of this declarative system is shown in Figure~\ref{fig:decl:syntax}.
-A declarative type $A$ is either the unit type $1$, a type variable $a$,
+A declarative type could be a unit type 1, a type variable $a$,
 a universal quantification $\forall a. A$ or a function type $A \to B$.
-Nested universal quantifiers are allowed in types,
-but monotypes $\tau$ do not contain any universal quantifier.
-Contexts $\Psi$ are just sequences of type variable declarations.
+Nested universal quantifiers are allowed for types,
+but monotypes do not have any universal quantifier.
+Terms include a unit term $()$, variables $x$, lambda-functions $\lam e$,
+applications $e_1~e_2$ and annotations $(e:A)$.
+Context $\Psi$ is a collection of type variable declarations and
+term variables with its type declared $x:A$.
 
@@
 
-Figures~\ref{fig:decl:sub} and \ref{fig:decl:typing},
-give the subtyping and typing relations for the declarative DK system.
-The DK subtyping relation was adopted from \citet{odersky1996putting}.
+In Figures~\ref{fig:decl:sub} and \ref{fig:decl:typing},
+we give the subtyping and typing relations for the declarative system,
+which is exactly the same as DK's.
+The subtyping relations originate from Odersky and La\"ufer~\cite{odersky1996putting},
+and is adopted by DK as a part of their typing system.
 
 \paragraph{Declarative Subtyping}
-Figure~\ref{fig:decl:sub} shows DK's declarative subtyping judgment $\Psi \vdash A \le B$,
-which was adopted from \citet{odersky1996putting}. It compares the
-degree of polymorphism between between $A$ and $B$ in DK's implicit polymorphic type system. 
-Essentially, if $A$ can always be instantiated to match any instantiation of $B$,
-then A is ``at least as polymorphic as'' $B$. We also 
-say that $A$ is ``more polymorphic than'' $B$ and write $A \le B$.
+For implicit polymorphic type systems,
+subtyping relation compares the degree of polymorphism of types.
+In short, if a polymorphic type $A$ always succeed to instantiate to any instantiation of $B$,
+then A is ``at least as polymorphic as'' $B$,
+or we just say that $A$ is ``more polymorphic than'' $B$, written as $A \le B$.
 
 Subtyping rules $\mathtt{{\le}Var}$, $\mathtt{{\le}Unit}$ and $\mathtt{{\le}{\to}}$
-handle simple cases that do not involve universal quantifiers.
+handle simple cases where no universal quantifiers get involved.
 The subtyping rule for function types $\mathtt{{\le}{\to}}$ is standard,
 being covariant on the return type and contravariant on the argument type.
-Rule $\mathtt{{\le}\forall R}$ states that if $A$ is a subtype of $B$
-in the context $\Psi, a$, where $a$ is fresh in $A$, then $A\le\all B$.
+Rule $\mathtt{{\le}\forall L}$ states that if $A$ is a subtype of $B$
+under the context ($\Psi, a$), where $a$ is fresh in $A$, then $A\le\all B$.
 Intuitively, if $A$ is more general than $\all B$ (where the universal quantifier
@@
 
-The most interesting rule is $\mathtt{{\le}\forall L}$.
-If some instantiation of $\all A$, $[\tau/a]A$, is a subtype of $B$,
+The most interesting rule is $\mathtt{{\le}\forall R}$.
+If any instantiation of $\all A$, $[\tau/a]A$, is a subtype of $B$,
 then $\all A \le B$.
-The monotype $\tau$ we used to instantiate $a$ is \emph{guessed} in this
-declarative rule, but the algorithmic system does not guess and defers the
-instantiation until it can determine the monotype deterministically.
+The monotype $\tau$ we used to instantiation is \emph{guessed} for the declarative rule,
+but the algorithmic system should be able to compute the guess properly.
 The fact that $\tau$ is a monotype rules out the possibility of
-polymorphic (or impredicative) instantiation.\tom{say why this restriction is imposed}
-Peyton Jones et al.~\cite{} also impose the restriction of
+polymorphic (or impredicative) instantiation.
+Peyton Jones et al.~\cite{} also has such restriction of
 predicative instantiation in their type system.

### 2019-02-18 sections/Overview.tex

--- 2019-02-14 sections/Overview.tex
@@
 a universal quantification $\forall a. A$ or a function type $A \to B$.
-Nested universal quantifiers are allowed in types,
-but monotypes $\tau$ do not contain any universal quantifier.
-Contexts $\Psi$ are just sequences of type variable declarations.
+
+Nested universal quantifiers are allowed for types,
+but monotypes $\tau$ do not have any universal quantifier.
+Terms include a unit term $()$, variables $x$, lambda-functions $\lam e$,
+applications $e_1~e_2$ and annotations $(e:A)$.
+Contexts $\Psi$ are sequences of type variable declarations and
+term variables with its type declared $x:A$.
 

### 2019-02-19 sections/algorithm.tex
some non-overlapping text

@@
 \bruno{Derivation should be explained briefly at least.}
+
+\subsection{Non-overlapping and Deterministic Reduction}
+An interesting and important aspect of our algorithmic rules is that
+they are directly implementable.
+Although written in the form of reduction rules, they do not overlap with each other.
+\jimmy{TODO: subtyping: $\forall$L with $\forall$R}
+Rule 18 is the most general case for checking,
+where its corresponding rule in DK's algorithm overlaps with all other checking rules.
+We prevent the non-determinism by restricting the condition of Rule 18,
+namely $e\neq\lam e'$ and $B\neq\all B'$.
+In other words, Rules 19, 20 and 21 have higher priority to Rule 18 when implemented.
+Thus our algorithm runs in polymorphic time, which is suitable for practical scenarios.
+
+\jimmy{Describe how to achieve the non-overlapping: a proof of declarative system,
+and then link to the metatheory for more details.}

### 2019-02-20 sections/Metatheory.tex
Quick responses to comments; Judgment erasure (metatheory)
Minor fixes for metatheory and algorithm

@@
 \begin{figure}[t]
+\framebox{$\|\Om\|$} Judgment erasure for $\Om$, returns a declarative context $\Psi$.
+\begin{gather*}
+\begin{aligned}
+\|\nil\| &= \nil\\
+\|\Om,a\| &= \|\Om\|, a\\
+\|\Om,x:A\| &= \|\Om\|, x:A\\
+\|\Om\Vdash\jg\| &= \|\Om\|
+\end{aligned}
+\end{gather*}
+
 \framebox{$\Om \rto \Om'$} $\Om$ reduces to $\Om'$.
@@
 \begin{aligned}
-\Om,a &\rto \Om \qquad \Om \qquad \Om,x:A \rto \Om\\
-\Om\Vdash A\le B &\rto \Om \text{ when } \Om\vdash A\le B\\
-\Om\Vdash e\Lto A &\rto \Om \text{ when } \Om\vdash e\Lto A\\
-\Om\Vdash e\To_a \jg &\rto \Om\Vdash[A/a]\jg \text{ when } \Om\vdash e\To A\\
-\Om\Vdash \appInfAlg{A}{e} &\rto \Om\Vdash[C/a]\jg \text{ when } \Om\vdash \appInf{A}{e}{C}\\
+\Om,a &\rto \Om \qquad \Om,x:A \rto \Om\\
+\Om\Vdash A\le B &\rto \Om \text{ when } \|\Om\| \vdash A\le B\\
+\Om\Vdash e\Lto A &\rto \Om \text{ when } \|\Om\| \vdash e\Lto A\\
+\Om\Vdash e\To_a \jg &\rto \Om\Vdash[A/a]\jg \text{ when } \|\Om\| \vdash e\To A\\
+\Om\Vdash \appInfAlg{A}{e} &\rto \Om\Vdash[C/a]\jg \text{ when } \|\Om\| \vdash \appInf{A}{e}{C}\\
 \end{aligned}
@@
 \jimmy{TODO: a erasure function would make things more clear...}
+
+\subsection{Non-overlapping Declarative System}
+\jimmy{Talk about the modified subsumption rule, reasoning about the unit rule,
+and the subtyping $\forall$ overlapping}
+
+%-------------------------------------------------------------------------------
 \subsection{Soundness}
@@
 
+\begin{corollary}[Soundness, single judgment form]
+
+\end{corollary}
+
 \subsection{Completeness}
@@
 If \emph{wf }$\Gm$ and $\exists \Om$ s.t. $\Gm\sto\Om$ and $\Om\redto\nil$, then $\Gm \redto \nil$.
+\label{thm:completeness}
 \end{theorem}
@@
 and thus we analyse each of the possible cases.
+\jimmy{Cite DK's proof as the sketch.}
 
@@
 \end{lemma}
+A symmetric lemma holds for $A\to B \le \al$.
 
-A symmetric lemma holds for $A\to B \le \al$.
+All other cases are handled easily... \jimmy{more details if necessary}
+
+The following corollary is derived immediately from Theorem~\ref{thm:completeness}.
+\begin{corollary}[Completeness, single judgment form]
+\end{corollary}
 
@@
 %
-|\Gm|_\Leftrightarrow &= \#_{e\Lto A\in\Gm} +
-    2\cdot\#_{e\To_a \jg\in\Gm} + 3\cdot\#_{\appInfAlg{A}{e}\in\Gm}\\[2mm]
+|\Gm|_\Leftrightarrow &= 2\cdot\#_{e\Lto A\in\Gm} +
+    \#_{e\To_a \jg\in\Gm} + 3\cdot\#_{\appInfAlg{A}{e}\in\Gm}\\[2mm]
 |\Gm|_\forall &= \textstyle\sum_{e\Lto A\in\Gm}|A|_\forall + \sum_{\appInfAlg{A}{e}\in\Gm}|A|_\forall +
@@
 
+Combining all the three main results, we know that the declarative system is decidable,
+with the help of our algorithm.
+\begin{corollary}[Decidability of Declarative Typing]
+\end{corollary}
 \subsection{Abella}

### 2019-02-20 sections/algorithm.tex
Minor fixes for metatheory and algorithm

@@
 \Gm \Vdash A_1\to A_2 \le B_1\to B_2 &\algrule \Gm \Vdash A_2 \le B_2 \Vdash B_1\le A_1\\
-\Gm \Vdash \forall a. A\le B &\algrule \Gm,\al \Vdash [\al/a]A\le B \quad\text{\jimmy{No more exvar leaks}}\\
+\Gm \Vdash \forall a. A\le B &\algrule \Gm,\al \Vdash [\al/a]A\le B \quad\text{when } B \neq \all B' \quad\text{\jimmy{No more exvar leaks}}\\
 \Gm \Vdash A\le \forall b. B &\algrule \Gm,b \Vdash A\le B \quad\text{\jimmy{No more typevar leaks}}
@@
 Although written in the form of reduction rules, they do not overlap with each other.
-\jimmy{TODO: subtyping: $\forall$L with $\forall$R}
+
+For subtyping, Rules 8 and 9 corresponds to declarative rules
+$\mathtt{{\le}\forall L}$ and $\mathtt{{\le}\forall R}$,
+where both of them matches the goal $\all A\le \all[b]B$.
+We eliminate the overlapping by restricting Rule 8 ($B\neq\all B'$)
+and thus always applying Rule 9 first for algorithmic reduction.
+
 Rule 18 is the most general case for checking,
 where its corresponding rule in DK's algorithm overlaps with all other checking rules.
-We prevent the non-determinism by restricting the condition of Rule 18,
+We prevent the non-determinism by adding conditions to Rule 18,
 namely $e\neq\lam e'$ and $B\neq\all B'$.

### 2019-02-20 sections/Overview.tex
declarative typing description draft

@@
 a universal quantification $\forall a. A$ or a function type $A \to B$.
-
 Nested universal quantifiers are allowed for types,
@@
 \begin{figure}[t]
-\centering \framebox{$\Psi \vdash A \Lto B$} $e$ checks against input type $A$.\\
-\centering \framebox{$\Psi \vdash A \To B$} $e$ synthesizes output type $A$.\\
-\centering \framebox{$\Psi \vdash \appInf{A}{e}{C}$} Applying a function of type $A$ to $e$ synthesizes type $C$.
+\begin{tabular}{rl}
+    \framebox{$\Psi \vdash A \Lto B$} & $e$ checks against input type $A$.\\[0.5mm]
+    \framebox{$\Psi \vdash A \To B$} & $e$ synthesizes output type $A$.\\[0.5mm]
+    \framebox{$\Psi \vdash \appInf{A}{e}{C}$} & Applying a function of type $A$ to $e$ synthesizes type $C$.
+\end{tabular}
 \begin{gather*}
@@
 \paragraph{Declarative Typing}
-\jimmy{TODO decl. typing description}
+The bidirectional type system has three judgments shown in Figure~\ref{fig:decl:typing}.
+The checking judgment $\Psi\vdash e\Lto A$ checks $e$ against the type $A$ under the context $\Psi$.
+The synthesis judgment $\Psi\vdash e\To A$ means that we can infer the type of $e$ under the context $\Psi$.
+The application judgment $\Psi\vdash \appInf{A}{e}{C}$ means that if a function of type $A$
+(which could be a polymorphic one) is applied to the argument $e$, then the whole application
+synthesizes type $C$ as its return type.
+
+Some of the rules handles simple cases with common approaches.
+Rule $\mathtt{DeclVar}$ looks term variables up in the context.
+Rules $\mathtt{Decl1I}$ and $\mathtt{Decl1I{\To}}$ checks and synthesizes the unit type, respectively.
+Rule $\mathtt{DeclAnno}$ synthesizes the type of the annotated expression $(e:A)$
+to its annotated type $A$, and requires $e$ to be checked agains $A$.
+Checking an expression $e$ against a polymorphic type $\all A$ under the context $\Psi$ succeeds
+if $e$ checks against A under an extended context $(\Psi, a)$, where $a$ is picked as a free variable.
+
+The subsumption rule $\mathtt{DeclSub}$ depends on the subtyping relation,
+and changes mode from checking to synthesis: if $e$ synthesizes $A$ and $A\le B$
+($A$ is more polymorphic than $B$), then $e$ can be checked against $B$.
+If a checking problem does not match any other rules,
+$\mathtt{DeclSub}$ can be applied and try to synthesize a type
+before checking whether the synthesized one is more general.
+
+Lambda-functions is a critical component of bidirectional type system, and they require much attention.
+Checking $\lam e$ against function type $A\to B$ is simple:
+we check the body $e$ against $B$ under the context extended with $x:A$.
+However, synthesizing a lambda-function is not easy in general,
+and this \emph{predicative} declarative system is able to
+synthesize to a suitable monotype $\sigma\to\tau$.
+In other words, if no type information is given for a lambda function,
+we will synthesize it to the best mono-typed function according to the context,
+which also requires some guessing similar to the subtyping rule $\mathtt{\forall L}$.
+
+Application $(e_1~e_2)$ is handled by Rule $\mathtt{Decl{\to}E}$,
+which firstly synthesize type $A$ for the function $e_1$.
+If $A$ is a function type $B\to C$, Rule $\mathtt{Decl{\to}App}$ is applied,
+by checking the argument $e_2$ against $B$ and returns type $C$.
+The only other possibile kind of types that
+the function $e_1$ could be synthesized is polymorphic types $\all A$.
+In that case, we instantiate $A$ to $[\tau/a]A$ with a monotype $\tau$ (which is also guessed)
+according to Rule $\mathtt{Decl{\to}I{\To}}$.
+If $[\tau/a]A$ is a function type, Rule $\mathtt{Decl{\to}App}$ proceeds;
+if $[\tau/a]A$ is still polymorphic, Rule $\mathtt{Decl{\to}I{\To}}$ is recursively applied.
+
+\jimmy{Possibly some discussions about the typing rules}
 
 \paragraph{Overlapping Rules}
-\jimmy{Move to the Alg. Section}
+\jimmy{Mention the resolution of overlapping briefly, and then point to algorithm and metatheory}
 Some of the declarative rules overlaps with each other.
@@
 	}
-	{\Psi\vdash \lam x \Lto \forall a \to a}
+	{\Psi\vdash \lam x \Lto \all a \to a}
 $$
@@
 	}
-{\Psi\vdash \lam x \Lto \forall a \to a}
+{\Psi\vdash \lam x \Lto \all a \to a}
 $$

### 2019-02-20 sections/introduction.tex
Abstract and Intro

@@
 \section{Introduction}
-\bruno{Jimmy can you please add the references? Let me know if you have questions.}
 
@@
 have been proposed~\cite{,dunfield2013complete,jones2007practical}
-\bruno{works on impredicative type inference + DK and Peyton Jones}.
+\bruno{works on impredicative type inference are still missing}.
 Two closely related type systems that 
@@
 popular among language designers and their ideas have been adopted by
-several modern functional languages, including Haskell and PureScript.
+several modern functional languages, including Haskell, PureScript~\cite{} and
+Unison~\bruno{Add references. Here is the one for unison\url{http://unisonweb.org}} among others.
 In those type systems
@@
 dubois2000proving,dubois1999certification,urban2008nominal,
-garrigue2015certified}\bruno{all previous work on mechanizing HM}. However
+garrigue2015certified}. However
 there is still no formalization of the higher-ranked type systems
@@
 the use of theorem provers to model various aspects of programming
-languages. Today papers in top conferences routinely
+languages. Today papers in various programming language venues routinely
 use theorem provers to mechanically formalize: \emph{dynamic and
-  static semantics} and their correctness properties~\cite{},
-\emph{compiler correctness}~\cite{}, \emph{correctness of
-  optimizations}~\cite{}, certain forms of \emph{abstract
-  interpretation}\bruno{van Horn and colleagues} or proofs involving \emph{logical relations}\bruno{See Andreas Abel new paper}. The
+  static semantics} and their correctness properties~\cite{aydemir2008engineering},
+\emph{compiler correctness}~\cite{leroy2012compcert}, \emph{correctness of
+  optimizations}~\cite{Bertot04}, \emph{program analisys}~\cite{Chang2006}
+or proofs involving \emph{logical relations}~\cite{abel2018}. The
 main argument for mechanical formalizations is a simple one. Proofs
@@
 real implementations of functional languages, including PureScript and
-\bruno{others}.  The DK type system has two variants: a declarative
+Unison. The DK type system has two variants: a declarative
 and an algorithmic one. The two variants have been been
@@
 
-One key challenge in type inference is variable binding, because the
-algorithms typically do not rely simply on local environments, but
-instead propagate information across judgements.  Yet, there is little
-work on how to deal with these complex forms of binding in theorem
-provers. To keep track of variable scoping, DK's algorithmic version
-employs input and output contexts to track information that is
-discovered throught type inference. However modelling output contexts
-in a theorem prover is non-trivial. Our work takes a different
-approach by refining and extending the idea of \emph{worklists
-  judgments}~\cite{itp2018}, proposed recently to mechanically formalize an
-algorithm for \emph{polymorphic subtyping}.  However, a key challenge
-not addressed previously with worklist judgements was how to deal with
-\emph{inference judgements}.  In our work we show how worklist
-judgements can be extended to deal with the inference judgements that
-are necessary for type-inference.  The idea is to use a continuation
-passing style to enable the transfer of inferred information across
-judgements. A further refinement to the idea of worklist judgements is
-the \emph{unification between ordered contexts~\cite{gundry2010type,dunfield2013complete}
-\bruno{DK and the work by McBride (see ITP citations)} and
-  worklists}.
-This enables precise scope tracking of free variables in
-judgements. Furthermore it avoids the duplication of context
-information across judgments in worklists that occurs in other
-techniques~\cite{Reed2009,Abel2011higher}
-\bruno{cite the two works we cited in ITP (and maybe others)}.
-
-Finally, another challenge that we faced were missing details as well as
+The first challenge that we faced were missing details as well as
 a few incorrect proofs and lemmas in DK's formalization. While DK's
@@
 writing a mechanically verified proof. Moreover some proofs and
-lemmas are wrong and it is not in all cases clear to us how to fix them.\footnote{
-Perhaps, with some additional thought, a work-around can be found, but, as our algorithm
-differs significantly from DK's, it would not further our mechanisation goal and so we did not pursue this.}
+lemmas are wrong and, in some cases, it is not clear to us how to fix them.
+%%\footnote{
+%%Perhaps, with some additional thought, a work-around can be found, but, as our algorithm
+%%differs significantly from DK's, it would not further our mechanisation goal and so we did not pursue this.}
 Despite the problems in DK's manual formalization,
@@
 help to our own work.
-We believe that for most non-trivial manual
+We expect that for most non-trivial manual
 proofs similar problems exist, so this should not be understood as a sign of sloppiness
@@
 similar problems. For example, Klein et al.~\cite{KleinRunYourResearch}
-\bruno{Run your research} mechanized formalizations
+mechanized formalizations
 in Redex for nine ICFP 2009 papers and all were found to have mistakes.
 
+Another challenge was variable binding. Type inference algorithms
+typically do not rely simply on local environments but instead
+propagate information across judgements. While local environments are
+well-studied in mechanical formalizations, there is little work on how
+to deal with the complex forms of binding employed by type inference algorithms
+in theorem provers. To
+keep track of variable scoping, DK's algorithmic version employs input
+and output contexts to track information that is discovered throught
+type inference. However modelling output contexts in a theorem prover
+is non-trivial.
+
+Due to those two challenges, our work takes a different approach by refining and
+extending the idea of \emph{worklists judgments}~\cite{itp2018},
+proposed recently to mechanically formalize an algorithm for
+\emph{polymorphic subtyping}. A key innovation in our work is how
+to adapt the idea or worklist judgments with
+\emph{inference judgements}, which are not needed for polymorphic subtyping.
+In our work we show how worklist
+judgements can be extended to deal with the inference judgements that
+are necessary for type-inference.  The idea is to use a continuation
+passing style to enable the transfer of inferred information across
+judgements. A further refinement to the idea of worklist judgements is
+the \emph{unification between ordered
+  contexts~\cite{gundry2010type,dunfield2013complete} and worklists}.  This
+enables precise scope tracking of free variables in
+judgements. Furthermore it avoids the duplication of context
+information across judgments in worklists that occurs in other
+techniques~\cite{Reed2009,Abel2011higher}.
 Despite the use of a different algorithm we prove the
@@
 
+\begin{comment}
 \jimmy{Notes @20190211 4 points of novalty:\\
@@
 }
+\end{comment}
 

### 2019-02-20 sections/Overview.tex
language in 2.1

--- 2019-02-18 sections/Overview.tex
@@
 \paragraph{Declarative Typing}
-\jimmy{TODO decl. typing description}
+The bidirectional type system, shown in Figure~\ref{fig:decl:typing}, has three judgments.
+The checking judgment $\Psi\vdash e\Lto A$ checks expression $e$ against the type $A$ in the context $\Psi$.
+The synthesis judgment $\Psi\vdash e\To A$ synthesizes the type $A$ of expression $e$ in the context $\Psi$.
+The application judgment $\Psi\vdash \appInf{A}{e}{C}$ synthesizes the type $C$ of the application of a function of type $A$
+(which could be polymorphic) to the argument $e$.
+
+Many rules are standard.
+Rule $\mathtt{DeclVar}$ looks up term variables in the context.
+Rules $\mathtt{Decl1I}$ and $\mathtt{Decl1I{\To}}$ respectively check and synthesize the unit type.
+Rule $\mathtt{DeclAnno}$ synthesizes the annotated type $A$ of the annotated expression $(e:A)$
+and checks that $e$ has type $A$.
+Checking an expression $e$ against a polymorphic type $\all A$ in the context $\Psi$ succeeds
+if $e$ checks against $A$ in the extended context $(\Psi, a)$.
+
+The subsumption rule $\mathtt{DeclSub}$ depends on the subtyping relation,
+and changes mode from checking to synthesis: if $e$ synthesizes type $A$ and $A\le B$
+($A$ is more polymorphic than $B$), then $e$ checks against $B$.
+If a checking problem does not match any other rules,
+this rule can be applied to synthesize a type instead and then
+check whether the synthesized type entails the checked type.
+
+The Lambda abstraction is the most complicated construct of the bidirectional
+type system. 
+Checking $\lam e$ against function type $A\to B$ is easy:
+we check the body $e$ against $B$ in the context extended with $x:A$.
+However, synthesizing a lambda-function is a lot less easy, and 
+this type system only synthesizes monotypes $\sigma\to\tau$.
+% and this \emph{predicative} declarative system is able to
+% synthesize to a suitable monotype $\sigma\to\tau$.
+% In other words, if no type information is given for a lambda function,
+% we will synthesize it to the best mono-typed function according to the context,
+% which also requires some guessing similar to the subtyping rule $\mathtt{\forall L}$.
+
+Application $(e_1~e_2)$ is handled by Rule $\mathtt{Decl{\to}E}$,
+which first synthesizes the type $A$ of the function $e_1$.
+If $A$ is a function type $B\to C$, Rule $\mathtt{Decl{\to}App}$ is applied;
+it checks the argument $e_2$ against $B$ and returns type $C$.
+The synthesized type of function $e_1$ can also be polymorphic, of the form $\all A$.
+In that case, we instantiate $A$ to $[\tau/a]A$ with a monotype $\tau$ % (which is also guessed)
+using according to Rule $\mathtt{Decl{\to}I{\To}}$.
+If $[\tau/a]A$ is a function type, Rule $\mathtt{Decl{\to}App}$ proceeds;
+if $[\tau/a]A$ has another universal quantifier, Rule $\mathtt{Decl{\to}I{\To}}$ is recursively applied.
+
+\jimmy{Possibly some discussions about the typing rules}
 
 \paragraph{Overlapping Rules}
-\jimmy{Move to the Alg. Section}
+\jimmy{Mention the resolution of overlapping briefly, and then point to algorithm and metatheory}
 Some of the declarative rules overlaps with each other.

### 2019-02-20 main.tex
Abstract and Intro
3 keywords

@@
 not clear how to replace these by other valid ones. 
-\bruno{others?} Despite the use of a different algorithm we prove the
+Despite the use of a different algorithm we prove the
 same results as DK, although with significantly different proofs and
@@
 %% comma separated list
-\keywords{keyword1, keyword2, keyword3}  %% \keywords are mandatory in final camera-ready submission
+\keywords{type inference, higher-rank polymorphism, mechanisation}  %% \keywords are mandatory in final camera-ready submission
 

### 2019-02-20 main.tex
Sections 2.1 and 2.2

@@
 proof techniques. Since such type inference algorithms are quite
-subtle and have a complex metatheory, we believe that mechanical
+subtle and have a complex metatheory, mechanical
 formalizations are an important advance in type-inference research.
@@
 %% comma separated list
-\keywords{type inference, higher-rank polymorphism, mechanisation}  %% \keywords are mandatory in final camera-ready submission
+\keywords{keyword1, keyword2, keyword3}  %% \keywords are mandatory in final camera-ready submission
 

### 2019-02-20 sections/Overview.tex
Sections 2.1 and 2.2

--- 2019-02-20 sections/Overview.tex
@@
 
-\bruno{I think some text/ideas-for-text can be borrowed from ITP.}
+This section starts by introducing DK's declarative type system. Then
+it discusses several techniques that have been used in algorithmic
+formulations, which have influenced our own algorithmic design.
+Finally we introduce the novelties of our new algorithm.
+In particular the support for inference judgements in
+judgement worklists, and a new form of judgement worklist
+that the unifies \emph{ordered contexts} and the worklists themselves.
 
@@
 
-The syntax of this declarative system is shown in Figure~\ref{fig:decl:syntax}.
+\paragraph{Syntax.}
+The syntax of DK's declarative system~\cite{dunfield2013complete} is shown in Figure~\ref{fig:decl:syntax}.
 A declarative type $A$ is either the unit type $1$, a type variable $a$,
@@
 \end{figure}
+\bruno{Some of the rule names are messed up.}
 
-Figures~\ref{fig:decl:sub} and \ref{fig:decl:typing},
-give the subtyping and typing relations for the declarative DK system.
-The DK subtyping relation was adopted from \citet{odersky1996putting}.
+%The DK subtyping relation was adopted from \citet{odersky1996putting}.
 
@@
 The fact that $\tau$ is a monotype rules out the possibility of
-polymorphic (or impredicative) instantiation.\tom{say why this restriction is imposed}
-Peyton Jones et al.~\cite{} also impose the restriction of
+polymorphic (or impredicative) instantiation.
+However this restriction ensures that the subtyping relation remains
+decidable. Allowing an arbitrary type (rather than a monotype) in rule $\mathtt{{\le}\forall L}$
+is know to give rise to an undecidable subtyping relation~\cite{tiuryn1996subtyping}.
+Peyton Jones et al.~\cite{jones2007practical} also impose the restriction of
 predicative instantiation in their type system.
-Both systems are adopt by several practical programming languages.
+Both systems are adopted by several practical programming languages.
 
@@
 
-Some of the rules handles simple cases with common approaches.
+Some of the rules handle simple cases with common approaches.
 Rule $\mathtt{DeclVar}$ looks term variables up in the context.
@@
 Rule $\mathtt{DeclAnno}$ synthesizes the type of the annotated expression $(e:A)$
-to its annotated type $A$, and requires $e$ to be checked agains $A$.
+to its annotated type $A$, and requires $e$ to be checked against $A$.
 Checking an expression $e$ against a polymorphic type $\all A$ under the context $\Psi$ succeeds
@@
 The subsumption rule $\mathtt{DeclSub}$ depends on the subtyping relation,
-and changes mode from checking to synthesis: if $e$ synthesizes $A$ and $A\le B$
+and changes the mode from checking to synthesis: if $e$ synthesizes $A$ and $A\le B$
 ($A$ is more polymorphic than $B$), then $e$ can be checked against $B$.
@@
 
-Lambda-functions is a critical component of bidirectional type system, and they require much attention.
+Lambdas are critical components of the bidirectional type system, and they require much attention.
 Checking $\lam e$ against function type $A\to B$ is simple:
 we check the body $e$ against $B$ under the context extended with $x:A$.
-However, synthesizing a lambda-function is not easy in general,
-and this \emph{predicative} declarative system is able to
-synthesize to a suitable monotype $\sigma\to\tau$.
+However, synthesizing a the type of an unannotated lambda is not easy. 
+Nevertheless the \emph{predicative} declarative system is able to
+synthesize to a suitable monotype $\sigma\to\tau$, if one exists.
 In other words, if no type information is given for a lambda function,
@@
 
-Application $(e_1~e_2)$ is handled by Rule $\mathtt{Decl{\to}E}$,
-which firstly synthesize type $A$ for the function $e_1$.
+Applications $e_1~e_2$ are handled by Rule $\mathtt{Decl{\to}E}$,
+which firstly synthesizes type $A$ for the function $e_1$.
 If $A$ is a function type $B\to C$, Rule $\mathtt{Decl{\to}App}$ is applied,
 by checking the argument $e_2$ against $B$ and returns type $C$.
-The only other possibile kind of types that
-the function $e_1$ could be synthesized is polymorphic types $\all A$.
+The only other possibile kind of type that
+the function $e_1$ could be synthesized to is a polymorphic type $\all A$.
 In that case, we instantiate $A$ to $[\tau/a]A$ with a monotype $\tau$ (which is also guessed)
@@
 both match the conclusion $\Psi\vdash \all A \le \all B$.
-In such case, choose $\mathtt{{\le}\forall R}$ first is always better,
+In such case, choosing $\mathtt{{\le}\forall R}$ first is always better,
 since we introduce the type variable $a$ to the context earlier,
@@
 
-The DK algorithm revolves around their notion of \emph{algorithmic context}.
+DK's algorithm version revolves around their notion of \emph{algorithmic context}.
 \[
@@
 
+\paragraph{Markers for scoping.}
 The sequential order of entries in the algoritmic context, in combination with
@@
 In particular, both instantiation soundness (i.e. a part of subtyping
-soundness) and typing completeness both directly require the broken lemma.
+soundness) and typing completeness directly require the broken lemma.
 Moreover, Lemma 54 (Typing Extension) also requires the broken lemma and is
@@
 Unfortunately, we have not yet found a way to fix this problem.
+\footnote{We found the issue with Lemma 29 in 2016 in an earlier attempt to mechanically
+  formalize DK's algorithm. The authors acknowledged the problem after we contacted them through email.
+  Although they briefly mentioned that it should be possible to use a weaker lemma instead they did
+  not go into details.
+}
+\bruno{We should acknowledge Ningning for finding this problem, but I think
+  this may be against the anonymous submission? Perhaps we do not mention her name
+in the submission, and if accepted we mention her?}
 \tom{You have communicated this to the authors and not received a fix, right?
@@
 
-Besides the problematic lemmas in the DK metatheory, there are other reasons to
+Besides the problematic lemmas in DK's metatheory, there are other reasons to
 look for an alternative algorithmic formulation of the type system that is more

### 2019-02-20 sections/introduction.tex

@@
 polymorphism, type arguments of polymorphic functions are
-automatically instantiated. Thus implicit polymorphism and the absense of
+automatically instantiated. Thus implicit polymorphism and the absence of
 type annotations mean that the Hindley-Milner type system 
@@
 
-Although type inference is important in practice and also in
-academic research, there is little work in mechanically formalizing
+Although type inference is important in practice and receives a lot of
+attention in
+academic research, there is little work on mechanically formalizing
 such advanced forms of type inference in theorem provers.
@@
 \emph{polymorphic subtyping}. A key innovation in our work is how
-to adapt the idea or worklist judgments with
+to adapt the idea of worklist judgments to
 \emph{inference judgements}, which are not needed for polymorphic subtyping.
@@
 
-\item {\bf Unified judgment worklists and contexts:} Another technical contribution is the unification
+\item {\bf Unification of judgment worklists and contexts:} Another technical contribution is the unification
   between ordered contexts and judgement worklists. This enables precise scope tracking

### 2019-02-20 sections/Overview.tex
merge

--- 2019-02-20 sections/Overview.tex
@@
 this type system only synthesizes monotypes $\sigma\to\tau$.
-% and this \emph{predicative} declarative system is able to
-% synthesize to a suitable monotype $\sigma\to\tau$.
-% In other words, if no type information is given for a lambda function,
-% we will synthesize it to the best mono-typed function according to the context,
-% which also requires some guessing similar to the subtyping rule $\mathtt{\forall L}$.
 
-Application $(e_1~e_2)$ is handled by Rule $\mathtt{Decl{\to}E}$,
+Application $e_1~e_2$ is handled by Rule $\mathtt{Decl{\to}E}$,
 which first synthesizes the type $A$ of the function $e_1$.

### 2019-02-20 sections/Overview.tex
some small fixed in text after merge
Sections 2.3 and 2.4

@@
 if $e$ checks against $A$ in the extended context $(\Psi, a)$.
-
 The subsumption rule $\mathtt{DeclSub}$ depends on the subtyping relation,
@@
 
-The Lambda abstraction is the most complicated construct of the bidirectional
-type system. 
+Lambda abstractions are the hardest construct of the bidirectional
+type system to deal with. 
 Checking $\lam e$ against function type $A\to B$ is easy:
@@
 algorithm by \citet{unification} decomposes the problem of unifying two terms $s \stackrel{.}{=} t$ into a set
-of related unification problems between pairs of terms $s_i \stackrel{.}{=} t_i)$. These smaller
+of related unification problems between pairs of terms $s_i \stackrel{.}{=} t_i$. These smaller
 problems are not tackled independently, but kept together in a set $S$. 
@@
 
-In recent work~\cite{itp2018}, we have shown how to mechanise a variant of DK's
+In recent work, \citep{itp2018} have shown how to mechanise a variant of DK's
 subtyping algorithm and shown it correct with respect to DK's declarative
@@
 by using a \emph{worklist}-based judgement of the form $$\Gamma \vdash \Omega$$
-\jimmy{emphasize the form of judgment}
 where $\Omega$ is a worklist, i.e. a sequence, of subtyping problems of the

### 2019-02-20 Proto.hs
Prototype implementation WIP

@@ -1,0 +1,118 @@
+import Prelude hiding (drop)
+
+-- Algorithmic Types
+data Typ =
+    TInt
+  | TVar String
+  | EVar String
+  | Forall String Typ
+  | FType Typ Typ
+  deriving (Show, Eq)
+
+-- Terms
+data Exp =
+    Var String
+  | Lit Int
+  | Lam String Exp
+  | App Exp Exp
+  | Ann Exp Typ
+  deriving (Show, Eq)
+
+-- Algorithmic Chain
+data Chain =
+    Sub Typ Typ
+  | Chk Exp Typ
+  | Inf Exp (Typ -> Chain)
+  | AInf Typ Exp (Typ -> Chain)
+
+-- Worklist
+data Worklist =
+    Empty
+  | ConsTVar Worklist String
+  | ConsEVar Worklist String
+  | ConsBind Worklist String Typ
+  | ConsJug Worklist Chain
+
+esubst :: String -> Typ -> Exp -> Exp
+esubst e s (Lam x b) = Lam x (esubst e s b)
+esubst e s (App e1 e2) = App (esubst e s e1) (esubst e s e2)
+esubst e s (Ann e1 t) = Ann (esubst e s e1) (tsubst e s t)
+esubst e s t = t
+
+tsubst :: String -> Typ -> Typ -> Typ
+tsubst s t TInt = TInt
+tsubst s t (TVar a) = TVar a
+tsubst s t (EVar x)
+  | s == x     = t
+  | otherwise  = EVar x
+tsubst s t (Forall a b)
+  | s == a     = Forall a b
+  | otherwise  = Forall a (tsubst s t b)
+tsubst s t (FType t1 t2) =
+  FType (tsubst s t t1) (tsubst s t t2)
+
+csubst :: String -> Typ -> Chain -> Chain
+csubst s t (Sub a b) = Sub (tsubst s t a) (tsubst s t b)
+csubst s t (Chk a b) = Chk (esubst s t a) (tsubst s t b)
+csubst s t (Inf e f) = Inf (esubst s t e) (\t1 -> csubst s t (f t1))
+csubst s t (AInf t1 e f) = AInf (tsubst s t t1) (esubst s t e) (\t1 -> csubst s t (f t1))
+
+wsubst :: String -> Typ -> Worklist -> Worklist
+wsubst = error "TODO"
+
+replace :: String -> String -> String -> Worklist -> Worklist
[…]

### 2019-02-21 sections/Discussion.tex
Minor fixes for metatheory and algorithm
DK's scoping control issue

@@
 
-\subsection{Overlapping Rules}
+% \subsection{Overlapping Rules} Already discussed in previous sections
 
@@
 
-\subsection{Making DK's scoping more precise}
+\subsection{Making DK's scoping more precise}\label{sec:discussion:scoping}
+
+DK's algorithm deals with garbage variables in a different approach,
+as discussed in Section~\ref{ssec:DK_Algorithm},
+through type variables or existential variable markers.
+However, there is still a chance that
+unused existential variables leak to output context and accumulate infinitely.
+For example, the derivation of the judgment $(\lam x)~() \Lto 1$ is as follows
+$$
+\inferrule*
+{
+    \inferrule*
+    {
+        \inferrule*
+        {
+            \inferrule*
+            {\ldots x \To \al \ldots \\ \ldots \al \le \bt \ldots}
+            {\Gm, \al, \bt, x:\al \vdash x \Lto \bt \dashv \Gm_1, x:\al}
+        }
+        {\Gm \vdash \lam x \To \al \to \bt \dashv \Gm_1}
+        \\
+        \inferrule*
+        {\ldots () \Lto \al \ldots}
+        {\Gm_1 \vdash \appInf{\al \to \al}{()}{\al} \dashv \Gm_2}
+    }
+    {\Gm \vdash (\lam x)~() \To \al \dashv \Gm_2}
+    \\
+    \inferrule*
+    {~}
+    {\Gm_2 \vdash 1 \le 1 \dashv \Gm_2}
+}
+{\Gm \vdash (\lam x)~() \Lto 1 \dashv \Gm, \al = 1, \bt = \al}
+$$
+where $\Gm_1 := (\Gm, \al, \bt = \al$) solves $\bt$,
+and $\Gm_2 := (\Gm, \al = 1, \bt = \al$) solves both $\al$ and $\bt$.
+
+If the reader is not familiar with DK's algorithm,
+he/she might be confused about the inconsistency across judgment.
+As an example, $(\lam x)~()$ synthesizes $\al$,
+but the second premise of the subsumption rule uses $1$ for the result.
+This is because a context application $[\Gm,\al=1,\bt=\al]\al = 1$ happens between the premises.
+
+The existential variables $\al$ and $\bt$ are clearly not used after the subsumption rule,
+but according to the algorithm, they are kept in the context
+until some parent judgment recycles a block of variables,
+or to the very end of a type inference task.
+In that sense, DK's algorithm does not control the scoping of variables precisely enough.
+However, it is a minor issue that does not affect the soundness and completeness properties.
 

### 2019-02-21 sections/Overview.tex
Prototype implementation WIP
DK's scoping control issue

@@
 it discusses several techniques that have been used in algorithmic
-formulations, which have influenced our own algorithmic design.
+formulations, and which have influenced our own algorithmic design.
 Finally we introduce the novelties of our new algorithm.
@@
 \qquad
-\inferrule*[right=$\mathtt{Decl1I\To}$]
+\inferrule*[right=$\mathtt{Decl1I{\To}}$]
     {~}{\Psi\vdash () \To 1}
@@
 \\
-\inferrule*[right=$\mathtt{Decl\to I}$]
+\inferrule*[right=$\mathtt{Decl{\to}I}$]
     {\Psi,x:A \vdash e\Lto B}
@@
 \qquad
-\inferrule*[right=$\mathtt{Decl\to I\To}$]
+\inferrule*[right=$\mathtt{Decl{\to}I{\To}}$]
     {\Psi\vdash \sigma\to\tau \quad \Psi,x:\sigma \vdash e\Lto \tau}
@@
 \\
-\inferrule*[right=$\mathtt{Decl\to E}$]
+\inferrule*[right=$\mathtt{Decl{\to} E}$]
     {\Psi\vdash e_1\To A \quad \Psi\vdash \appInf{A}{e_2}{C}}
@@
 \qquad
-\inferrule*[right=$\mathtt{Decl\to App}$]
+\inferrule*[right=$\mathtt{Decl{\to}App}$]
     {\Psi\vdash e \Lto A}
@@
 %-------------------------------------------------------------------------------
-\subsection{DK's Algorithm}
+\subsection{DK's Algorithm}\label{ssec:DK_Algorithm}
 
@@
 from the output stack together with the subsequent part $\Theta$, which may
-refer to $\hat{\alpha}$. (Remember that later entries may refer to earlier
+refer to $\hat{\alpha}$. \jimmy{which may contain definition/solution of $\al$ and references to $\al$}
+(Remember that later entries may refer to earlier
 ones, but not vice versa.) This way $\hat{\alpha}$ does not escape its scope.
+
+\jimmy{Do we talk about scoping control without markers as well here?\\
+e.g. $\forall I$ ($e \Lto \all A$) or $\le\forall R$ ($A \le \all B$)\\
+Text about garbage refers to this point!}
+
+Due to the complicated scoping control method,
+it is hard to argue the preciseness of scoping for each variable when designing algorithms.
+We found a scoping issue related to the subsumption rule,
+which might cause existential variable leakage across judgments.
+A detailed discussion on this issue with a possible fix is in Section~\ref{sec:discussion:scoping}.
 
@@
 
-\jimmy{Compare with DK's algo: subsumption rule might cause leak}
-
 \paragraph{Inference Judgements}

### 2019-02-22 sections/Discussion.tex
S2.4 scoping; hat alpha format

@@
 If the reader is not familiar with DK's algorithm,
-he/she might be confused about the inconsistency across judgment.
+he/she might be confused about the inconsistent types across judgment.
 As an example, $(\lam x)~()$ synthesizes $\al$,
@@
 In that sense, DK's algorithm does not control the scoping of variables precisely enough.
-However, it is a minor issue that does not affect the soundness and completeness properties.
+%However, it is a minor issue that does not affect the soundness and completeness properties.
 

### 2019-02-22 sections/Overview.tex
S2.4 scoping; hat alpha format

@@
 
-\jimmy{Possibly some discussions about the typing rules}
-
 \paragraph{Overlapping Rules}
@@
 \text{Algorithmic Contexts}\qquad&\Gamma,\Delta,\Theta &::=&\quad \nil \mid
-\Gamma, a \mid \Gamma, \hat{\alpha} \mid \Gamma, \hat{\alpha} = \tau \mid
-\Gamma, \blacktriangleright_{\hat{\alpha}}
+\Gamma, a \mid \Gamma, \al \mid \Gamma, \al = \tau \mid
+\Gamma, \blacktriangleright_{\al}
 \end{array}
@@
 algorithmic context also contains \emph{existential} type variables
-$\hat{\alpha}$. These are placeholders for monotypes $\tau$ that are still to
+$\al$. These are placeholders for monotypes $\tau$ that are still to
 be determined by the inference algorithm. When the existential variable is
 ``solved'', its entry in the context is replaced by the assignment
-$\hat{\alpha} = \tau$. With the notation $[\Gamma]$ an algorithmic context
+$\al = \tau$. With the notation $[\Gamma]$ an algorithmic context
 $\Gamma$ denotes the substitution that replaces all its solved existential type
@@
 existential variables. For this reason the DK algorithm uses scope markers
-$\blacktriangleright_{\hat{\alpha}}$ in a few places. An example is given in the following
+$\blacktriangleright_{\al}$ in a few places. An example is given in the following
 rule:
@@
 \inferrule*
-  {\Gamma,\blacktriangleright_{\hat{\alpha}}, \hat{\alpha} \vdash [\hat{\alpha}/a]A \le B \dashv \Delta,\blacktriangleright_{\hat{\alpha}},\Theta}
+  {\Gamma,\blacktriangleright_{\al}, \al \vdash [\al/a]A \le B \dashv \Delta,\blacktriangleright_{\al},\Theta}
   {\Gamma \vdash \forall a. A \le B \dashv \Delta}
 \]
-To indicate that the scope of $\hat{\alpha}$ is local to the subtyping check
-$[\hat{\alpha}/a]A \le B$, the marker is pushed onto its input stack and popped
+To indicate that the scope of $\al$ is local to the subtyping check
+$[\al/a]A \le B$, the marker is pushed onto its input stack and popped
 from the output stack together with the subsequent part $\Theta$, which may
-refer to $\hat{\alpha}$. \jimmy{which may contain definition/solution of $\al$ and references to $\al$}
+refer to $\al$. \jimmy{which may contain definition/solution of $\al$ and references to $\al$}
 (Remember that later entries may refer to earlier
-ones, but not vice versa.) This way $\hat{\alpha}$ does not escape its scope.
+ones, but not vice versa.) This way $\al$ does not escape its scope.
 
-\jimmy{Do we talk about scoping control without markers as well here?\\
-e.g. $\forall I$ ($e \Lto \all A$) or $\le\forall R$ ($A \le \all B$)\\
-Text about garbage refers to this point!}
+A type variable may have a similar functionality to the scoping markers.
+An example rule that checks an expression against a polymorphic type is as follows:
+$$
+\inferrule*
+{\Gm, a \vdash e \Lto A \dashv \Delta, a, \Theta}
+{\Gm \vdash e \Lto \all A \dashv \Delta}
+$$
 
-Due to the complicated scoping control method,
+Due to the complication of its scoping control method,
 it is hard to argue the preciseness of scoping for each variable when designing algorithms.
-We found a scoping issue related to the subsumption rule,
+Actually we found a scoping issue related to the subsumption rule,
 which might cause existential variable leakage across judgments.

### 2019-02-22 sections/Related.tex
part of RW

\section{Related Work}

\paragraph{Polymorphic Type Inference Algorithms}
Higher-order polymorphism is a convenient and practical feature of programming language.
Since full type-inference for System F is undecidable~\cite{wells1999typability},
various decidable partial type-inference algorithms were developed.
The declarative system of this paper, proposed by DK, is a \emph{predicative} one,
where $\forall$'s only instantiate to mono-types.
The mono-type restriction on instantiation is considered reasonable and practical for most programs,
except for those that require sophisticated form of higher-order polymorphism.
In those cases, the bidirectional system accepts guidance through type annotations,
which also improves readability of the program and is typically not considered as a burden.

Impredicative System F allows instantiation with polymorphic types,
but unfortunately its subtyping system is already undecidable~\cite{tiuryn1996subtyping}.
Works on partial impredicative type-inference algorithms\cite{le2003ml,leijen2008hmf,vytiniotis2008fph}
navigate a variety of design tradeoffs for a decidable algorithm.
As a result, such algorithms tend to be more complicated, and thus less adopted in practise.

\jimmy{Guarded Impredicative}

\jimmy{Dunfield's new work}

### 2019-02-23 sections/Overview.tex
Section 2.4 and 2.5

@@
 the scoping of regular and meta-variables, this approach is not ideal because it
-repeatedly copies contexts when decomposing a unification problem, and later has to 
+repeatedly copies contexts when decomposing a unification problem, and
+later it has to 
 substitute solutions into these copies.
@@
 
-In recent work, \citep{itp2018} have shown how to mechanise a variant of DK's
+In recent work, \citet{itp2018} have shown how to mechanise a variant of DK's
 subtyping algorithm and shown it correct with respect to DK's declarative
@@
 addresses the open problems. 
+\bruno{I think this section needs more detail here. The textual
+  description alone does not convey much about the key ideas of the
+  approach. I suggest showing the forms of the judgements and maybe
+  illustrating with some examples why the blend between ordered
+  contexts and worklists leads to precise scope traking. This can be
+  done just by illustrating a concrete instance of an algorithmic
+  context (no derivation is actually needed). Similar things should be
+done for inference judgements. Actually chains needs some careful
+explanation, I think since they are the hardest to get aspect of the formalization.}
 

### 2019-02-23 sections/Overview.tex
Example for S2.5

@@
 
+Take the following worklist as an example
+$$\al \Vdash \blue{(\lam x)~() \To_a a \le \al} , x:\al, \bt \Vdash \blue{\al \le \bt} \Vdash \blue{\bt \le 1}$$
+A worklist is structurally a stack with its top element to the right.
+There are three (blue-colored) judgment chains in the worklist,
+where the first and second judgment chains are two subtyping judgments,
+and the third judgment chain, $(\lam x)~() \To_a a \le \al$,
+is a sequence of an inference judgment followed by a subtyping judgment.
+
+The algorithm first analyse the two subtyping judgments and
+will find the best solutions $\al = \bt = 1$
+(please refer to Figure~\ref{fig:alg} for detailed derivations).
+Then we substitute every instance of $\al$ and $\bt$ to $1$,
+so the variable declarations can be safely removed from the worklist.
+Now we reduce the worklist to the following state
+$$\nil \Vdash \blue{(\lam x)~() \To_a a \le 1}, x:1$$
+which has a term variable declaration as the top element.
+Because no judgments before the declaration refer to $x$,
+as a basic property of an ordered context
+(Figure~\ref{fig:alg:syntax} shows its well-formedness criteria),
+we remove all declarations before the first judgment chain to collect unused variables.
+As a result all garbage variables are collected
+once the judgment chains refering to them are processed.
+
+After that, we proceed the last judgment chain from the left,
+namely an inference judgment $(\lam x)~()\To~?$, with the unit type $1$ as its result.
+Finally the last judgment becomes $1 \le 1$, and is satisfied as a base case.
+
 % Our algorithm borrows some ideas from previous work, while adding new ones. 

### 2019-02-23 sections/Overview.tex
S2.5 one more example

@@
 
+\jimmy{Some high level description for the structure}
+A worklist is structurally a stack with its top element to the right.
+$$\begin{aligned}
+\text{Algorithmic judgement chain}\qquad&\jg &::=&\quad A \le B \mid e\Lto A \mid e\To_{a} \jg \mid \appInfAlg{A}{e}\\
+\text{Algorithmic worklist}\qquad&\Gm &::=&\quad \nil \mid \Gm, a \mid \Gm, \al \mid \Gm, x: A \mid \Gm \Vdash \jg\\
+\end{aligned}$$
+
 \paragraph{Scope Tracking}
@@
 garbage builds up.
+
+For example, the top judgment of the following worklist
+checks the identity function against $\all a \to a$:
+$$\Gm \Vdash \lam x \Lto \all a \to a$$
+Then, two auxiliary variables $a$ and $x$ are introduced to help the type checking:
+$$\Gm, a, x:a \Vdash x \Lto a$$
+which will be satisfied after several steps, and the worklist becomes
+$$\Gm, a, x:a$$
+Since the variable declarations $a, x:a$ are only used for a judgment already processed,
+they can be safely removed, leaving the remaining worklist $\Gm$ to be further reduced.
+
+% $$\Gm \Vdash \all a \to a \le \all a \to a$$
+% $$\Gm, a, \al \Vdash \al \to \al \le a \to a$$
 
@@
 $$\al \Vdash \blue{(\lam x)~() \To_a a \le \al} , x:\al, \bt \Vdash \blue{\al \le \bt} \Vdash \blue{\bt \le 1}$$
-A worklist is structurally a stack with its top element to the right.
 There are three (blue-colored) judgment chains in the worklist,

### 2019-02-23 sections/Overview.tex
S2.5 examples with more explaination

@@
 
-\jimmy{Some high level description for the structure}
-A worklist is structurally a stack with its top element to the right.
-$$\begin{aligned}
-\text{Algorithmic judgement chain}\qquad&\jg &::=&\quad A \le B \mid e\Lto A \mid e\To_{a} \jg \mid \appInfAlg{A}{e}\\
-\text{Algorithmic worklist}\qquad&\Gm &::=&\quad \nil \mid \Gm, a \mid \Gm, \al \mid \Gm, x: A \mid \Gm \Vdash \jg\\
-\end{aligned}$$
-
 \paragraph{Scope Tracking}
@@
 
+$$\begin{aligned}
+\text{Algorithmic judgement chain}\qquad&\jg &::=&\quad A \le B \mid e\Lto A \mid e\To_{a} \jg \mid \appInfAlg{A}{e}\\
+\text{Algorithmic worklist}\qquad&\Gm &::=&\quad \nil \mid \Gm, a \mid \Gm, \al \mid \Gm, x: A \mid \Gm \Vdash \jg\\
+\end{aligned}$$
+
 For example, the top judgment of the following worklist
@@
 $$\Gm \Vdash \lam x \Lto \all a \to a$$
-Then, two auxiliary variables $a$ and $x$ are introduced to help the type checking:
+To proceed, two auxiliary variables $a$ and $x$ are introduced to help the type checking:
 $$\Gm, a, x:a \Vdash x \Lto a$$
@@
 they can be safely removed, leaving the remaining worklist $\Gm$ to be further reduced.
+
+Our worklist can be seen as an all-in-one stack,
+containing variable declarations and subtyping/typing judgments.
+The stack is a enriched form of ordered context,
+and it has a similar variable scoping scheme.
+The well-formedness condition, shown in Figure~\ref{fig:alg:syntax},
+states that every variable must be declared prior to it usage.
+Therefore, if any variable declaration is the top element,
+then no usage should appear in the rest of the worklist, so it can be removed.
 
@@
 As a result all garbage variables are collected
-once the judgment chains refering to them are processed.
+once the judgment chains refering to them are processed,
+just as what we have mentioned above about the scope tracking.
 

### 2019-02-23 sections/Overview.tex
Section 2.5

--- 2019-02-23 sections/Overview.tex
@@
 addresses the open problems.
-\bruno{I think this section needs more detail here. The textual
-  description alone does not convey much about the key ideas of the
-  approach. I suggest showing the forms of the judgements and maybe
-  illustrating with some examples why the blend between ordered
-  contexts and worklists leads to precise scope traking. This can be
-  done just by illustrating a concrete instance of an algorithmic
-  context (no derivation is actually needed). Similar things should be
-done for inference judgements. Actually chains needs some careful
-explanation, I think since they are the hardest to get aspect of the formalization.}
 
-\jimmy{Some high level description for the structure}
-A worklist is structurally a stack with its top element to the right.
+\paragraph{Scope Tracking}
+We avoid scoping garbage by blending the ordered context and the
+worklist into a single syntactic sort $\Gamma$, our algorithmic worklist. This
+algorithmic worklist interleaves (type and term) variables with \emph{work}
+like checking or inferring types of expressions. The interleaving keeps track
+of the variable scopes in the usual, natural way: each variable is in scope of
+anything in front of it in the worklist. If there is nothing in front, the
+variable is no longer needed and can be popped from the worklist. This way, no
+garbage (i.e. variables out-of-scope) builds up.
+
 $$\begin{aligned}
@@
 
-\paragraph{Scope Tracking}
-Firstly, we avoid scoping garbage by blending the ordered context and the
-worklist into a single syntactic sort $\Gamma$, our algorithmic worklist. This
-algorithmic worklist interleaves (type and term) variables with \emph{work}
-like checking or inferring types of expressions. The interleaving keeps track
-of the variable scopes in the usual, natural way: each variables is in scope of
-anything in front of it in the worklist. If there is nothing in front, the
-variable is no longer needed and can be popped from the worklist. This way, no
-garbage builds up.
-
-For example, the top judgment of the following worklist
+For example, suppose that the top judgment of the following worklist
 checks the identity function against $\all a \to a$:
 $$\Gm \Vdash \lam x \Lto \all a \to a$$
-Then, two auxiliary variables $a$ and $x$ are introduced to help the type checking:
+To proceed, two auxiliary variables $a$ and $x$ are introduced to help the type checking:
 $$\Gm, a, x:a \Vdash x \Lto a$$
@@
 
+Our worklist can be seen as an all-in-one stack,
+containing variable declarations and subtyping/typing judgments.
+The stack is an enriched form of ordered context,
+and it has a similar variable scoping scheme.
+%The well-formedness condition, shown in Figure~\ref{fig:alg:syntax},
+%states that every variable must be declared prior to it usage.
+%Therefore, if any variable declaration is the top element,
+%then no usage should appear in the rest of the worklist, so it can be removed.
+
 % $$\Gm \Vdash \all a \to a \le \all a \to a$$
@@
 \paragraph{Inference Judgements}
-To express the DK inference judgements, we use a novel form of work entries in
+To express the DK's inference judgements, we use a novel form of work entries in
 the worklist: our algorithmic judgement chains $\omega$. In its simplest form,
 such a judgement chain is just a check, like a subtyping check $A \leq B$ or a
-term typecheck $e \Leftarrow A$.  The non-trivial forms of chains capture an
+term typecheck $e \Leftarrow A$. 
+However, the non-trivial forms of chains capture an
 inference task together with the the work that depends on its outcome. For
 instance, a type inference task takes the form $e \Rightarrow_a \omega$. It
-denotes that we need to infer the type, say $A$, of expression $e$ and use it
-in the chain $\omega$ by subsituting it for the placeholder type variable $a$.
+denotes that we need to infer the type, say $A$, for expression $e$ and use it
+in the chain $\omega$ by substituting it for the placeholder type variable $a$.
 
@@
 
-The algorithm first analyse the two subtyping judgments and
+The algorithm first analyses the two subtyping judgments and
 will find the best solutions $\al = \bt = 1$
@@
 which has a term variable declaration as the top element.
-Because no judgments before the declaration refer to $x$,
-as a basic property of an ordered context
-(Figure~\ref{fig:alg:syntax} shows its well-formedness criteria),
-we remove all declarations before the first judgment chain to collect unused variables.
-As a result all garbage variables are collected
-once the judgment chains refering to them are processed.
-
-After that, we proceed the last judgment chain from the left,
+%Because no judgments before the declaration refer to $x$,
+%as a basic property of an ordered context
+%(Figure~\ref{fig:alg:syntax} shows its well-formedness criteria),
+%we remove all declarations before the first judgment chain to collect unused variables.
+%As a result all garbage variables are collected
+%once the judgment chains refering to them are processed,
+%just as what we have mentioned above about the scope tracking.
+After removing the garbage term variable declaration from the worklist, we process the last judgment chain from the left,
 namely an inference judgment $(\lam x)~()\To~?$, with the unit type $1$ as its result.

### 2019-02-24 sections/Related.tex
RW

@@
 
+\jimmy{Dunfield's new work}
+DK's algorithm is shown to be sound, complete and decidable with 70 pages of manual proofs.
+Though carefully written, some of the proofs are incorrect,
+thus we find it difficult to formalize in a proof assistant.
+Their follow-up work~\cite{} enriches the bidirectional higher-rank system with
+existentials and indexed types.
+With a more complex declarative system, they developed a proof of over 150 pages.
+It is even more difficult to argue its correctness for every single detail within such a big development.
+Unfortunately, we find its Lemma 26 (Parallel Admissibility) has the same issue with
+a lemma of the same name in their previous work, which has a false conclusion.
+
 Impredicative System F allows instantiation with polymorphic types,
@@
 \jimmy{Guarded Impredicative}
+A recent work proposed Guarded Impredicative Polymorphism~\cite{},
+as an improvement on GHC's type inference algorithm with impredicative instantiation.
+They make use of local informations in $n$-ary applications to
+infer polymorphic instantiations with a relatively simple specification and unification algorithm.
+Although not all impredicative instantiation can be handled perfectly,
+their algorithm is already practically useful.
+Unfortunately, no mechanical formalization is provided for this work.
 
-\jimmy{Dunfield's new work}
+\paragraph{Formalizations of Type-Inference Algorithms}
+Since the publication of the {\sc POPLMark} challenge~\cite{aydemir2005mechanized},
+many theorem provers and packages provide new methods for dealing
+with variable binding~\cite{aydemir2008engineering,urban2008nominalTech,chlipala2008parametric},
+and more and more type systems are formalized with these tools.
+However, mechanising algorithmic aspects, like unification and
+constraint solving, have received very little attention and is still challenging.
+Moreover, while most tools supports local (input) contexts in a neat way, which have a simple
+binding and scoping structure, many practical type-inference algorithms require
+more complex binding structures with output contexts or various form of constraint solving procedure.
+
+Algorithm $\mathcal{W}$,
+as one of the classic type inference algorithms for polymorphic type systems,
+has been manually proven to be sound and complete
+with respect to the Hindley-Milner type system~\cite{}.
+After around 15 years, the algorithm is foramlly verified by
+Naraschewski and Nipkow~\cite{naraschewski1999type} in Isabelle/HOL~\cite{nipkow2002isabelle}.
+The treatment of new variables is tricky at that time, while the overall structure follow the
+structure of Damas's manual proof closely.
+%Another complication is the encoding of substitutions,
+%which they chose to formalize as a function instead of an association list.
+Later on, other researchers~\cite{dubois2000proving,dubois1999certification}
+formalized algorithm $\mathcal{W}$ in Coq~\cite{Coq}.
+Nominal techniques~\cite{urban2008nominalTech} in Isabelle/HOL have been
+developed to help programming language formalizations, and are used for a similar
+verification~\cite{urban2008nominal}. Moreover, Garrigue~\cite{garrigue2015certified}
+mechanized a type inference algorithm,
+with the help of locally nameless cofinite quantification~\cite{},
+for Core ML extended with structural polymorphism and recursion.

### 2019-02-24 sections/Related.tex
Related Work

--- 2019-02-22 sections/Related.tex
@@
 
-\paragraph{Polymorphic Type Inference Algorithms}
-Higher-order polymorphism is a convenient and practical feature of programming language.
-Since full type-inference for System F is undecidable~\cite{wells1999typability},
-various decidable partial type-inference algorithms were developed.
-The declarative system of this paper, proposed by DK, is a \emph{predicative} one,
-where $\forall$'s only instantiate to mono-types.
-The mono-type restriction on instantiation is considered reasonable and practical for most programs,
-except for those that require sophisticated form of higher-order polymorphism.
-In those cases, the bidirectional system accepts guidance through type annotations,
-which also improves readability of the program and is typically not considered as a burden.
+\bruno{Jimmy, please fill in missing references.}
 
+\paragraph{Predicative Higher-Ranked Polymorphism Type Inference Algorithms}
+Higher-ranked polymorphism is a convenient and practical feature of
+programming languages.  Since full type-inference for System F is
+undecidable~\cite{wells1999typability}, various decidable partial
+type-inference algorithms were developed~\cite{}.  The declarative system of
+this paper, proposed by \citet{dunfield2013complete}, is \emph{predicative}: 
+$\forall$'s only instantiate to monotypes.  The monotype restriction
+on instantiation is considered reasonable and practical for most
+programs, except for those that require sophisticated forms of
+higher-order polymorphism.  In those cases, the bidirectional system
+accepts guidance through type annotations, which allow polymorphic types.
+Such annotations also improve
+readability of the program, and are not much of a burden in practice.
+
+\jimmy{Dunfield's new work}
+DK's algorithm is shown to be sound, complete and decidable in 70 pages of manual proofs.
+Though carefully written, some of the proofs are incorrect (see discussion in Section~\ref{}),
+which creates difficulties when formalizing them in a proof assistant.
+Their follow-up work~\cite{} enriches the bidirectional higher-rank system with
+existentials and indexed types.
+With a more complex declarative system, they developed a proof of over 150 pages.
+It is even more difficult to argue its correctness for every single detail within such a big development.
+Unfortunately, we find that their Lemma 26 (Parallel Admissibility) appears to have the same issue 
+as lemma 29 discussed in Section~\ref{}: the conclusion is false.
+\bruno{I think you will need to justify this better and in the appendix you must show
+the lemma and counter-example.}
+\bruno{Missing Simon Peyton Jones work: which is the other predicative work.
+  Maybe the OutsideIn work can also be briefly mentioned (you can ask Tom for details
+  of that work).
+ }
+
+\paragraph{Impredicative Higher-Ranked Polymorphism Type Inference Algorithms}
 Impredicative System F allows instantiation with polymorphic types,
@@
 As a result, such algorithms tend to be more complicated, and thus less adopted in practise.
+\jimmy{Guarded Impredicative}
+A recent work proposed Guarded Impredicative Polymorphism~\cite{},
+as an improvement on GHC's type inference algorithm with impredicative instantiation.
+They make use of local informations in $n$-ary applications to
+infer polymorphic instantiations with a relatively simple specification and unification algorithm.
+Although not all impredicative instantiations can be handled well,
+their algorithm is already quite useful in practice.
 
-\jimmy{Guarded Impredicative}
+\paragraph{Mechanical Formalization of Polymorphic Subtyping}
+In all previous work on type inference for higher-ranked polymorphism
+(predicative and impredicative) discussed above, all proofs and
+metatheory are manual. The only partial effort on mechanizing aspects of type inference
+for higher-ranked types is
+the Abella formalization of \emph{polymorphic subtyping} by
+\citet{itp2018}. ... \bruno{a comparison with our own ITP paper here!}
 
-\jimmy{Dunfield's new work}
+\paragraph{Mechanical Formalizations of Other Type-Inference Algorithms}
+Since the publication of the {\sc POPLMark} challenge~\cite{aydemir2005mechanized},
+many theorem provers and packages provide new methods for dealing
+with variable binding~\cite{aydemir2008engineering,urban2008nominalTech,chlipala2008parametric}.
+More and more type systems are formalized with these tools.
+However, mechanising certain algorithmic aspects, like unification and
+constraint solving, has received very little attention and are still challenging.
+Moreover, while most tools supports local (input) contexts in a neat way, which have a simple
+binding and scoping structure, many practical type-inference algorithms require
+more complex binding structures with output contexts or various forms of constraint solving procedures.
+
+Algorithm $\mathcal{W}$,
+as one of the classic type inference algorithms for polymorphic type systems,
+has been manually proven to be sound and complete
+with respect to the Hindley-Milner type system~\cite{}.
+After around 15 years, the algorithm was formally verified by
+\citet{naraschewski1999type} in Isabelle/HOL~\cite{nipkow2002isabelle}.
+The treatment of new variables was tricky at that time, while the overall structure follow the
+structure of Damas's manual proof closely.
+%Another complication is the encoding of substitutions,
+%which they chose to formalize as a function instead of an association list.
+Later on, other researchers~\cite{dubois2000proving,dubois1999certification}
+formalized algorithm $\mathcal{W}$ in Coq~\cite{Coq}.
+Nominal techniques~\cite{urban2008nominalTech} in Isabelle/HOL have been
+developed to help programming language formalizations, and are used for a similar
+verification~\cite{urban2008nominal}. Moreover, Garrigue~\cite{garrigue2015certified}
+mechanized a type inference algorithm,
+with the help of locally nameless cofinite quantification~\cite{},
+for Core ML extended with structural polymorphism and recursion.
+
+\paragraph{Ordered Contexts in Type Inference}
+
+\paragraph{Lists of Judgements in Unification}
+
+\bruno{Missing: work on type inference and unification of dependent types.
+  Also missing: discussion about ordered contexts and the work by Gundry (in addition to DK).
+  Regarding these two, since we have an extended discussion in Section 2 you can have
+  a summarized discussion here, pointoing to Section 2 for further deatils.}
+
+\bruno{Final comment: You should essentially have all the discussion that we have in ITP +
+  some more. So, double-check carefully that all the works we cite in our related work in ITP
+are cited here, and that we include all discussions!}

### 2019-02-25 sections/Related.tex
Summary sentence
RW ITP paper

@@
 \section{Related Work}
+
+Throughout the paper we have already discussed much of the closest related work.
+In this section we summarize the key differences and novelties, and we discuss some other
+related work.
 
@@
 DK's algorithm is shown to be sound, complete and decidable in 70 pages of manual proofs.
-Though carefully written, some of the proofs are incorrect (see discussion in Section~\ref{}),
+Though carefully written, some of the proofs are incorrect (see discussion in Appendix),
 which creates difficulties when formalizing them in a proof assistant.
@@
 Unfortunately, we find that their Lemma 26 (Parallel Admissibility) appears to have the same issue 
-as lemma 29 discussed in Section~\ref{}: the conclusion is false.
+as lemma 29 discussed in Appendix: the conclusion is false.
 \bruno{I think you will need to justify this better and in the appendix you must show
@@
 for higher-ranked types is
-the Abella formalization of \emph{polymorphic subtyping} by
-\citet{itp2018}. ... \bruno{a comparison with our own ITP paper here!}
+the Abella formalization of \emph{polymorphic subtyping} by \citet{itp2018}.
+The judgment form of worklist $\Gm \vdash \Omega$ used in the formalization simplifies
+the propagation of existential variable instantiations.
+However, the approach has two main drawbacks;
+it does not collect unused variable declarations effectively,
+and the simple form of judgment cannot handle inference modes, which outputs types.
+The new worklist introduced in this paper inherits the simplicity of propagating instantiations,
+but overcome both of the issues by mixing judgments with declarations
+and using the continuation-passing-style judgment chains.
+\bruno{a comparison with our own ITP paper here!}
 

### 2019-02-25 sections/Related.tex
RW lists of judgments

@@
 \paragraph{Ordered Contexts in Type Inference}
+Gundry et al.~\cite{gundry2010type}
 
 \paragraph{Lists of Judgements in Unification}
+Some work~\cite{Reed2009,Abel2011higher} adopts a similar idea to this paper
+in dependently typed inference and reconstruction.
+Their algorithms collect unification constraints and simplify them progressively in a single-step style.
+In comparison, our algorithm mixes variable declarations with judgments,
+resulting in a simpler judgment form,
+while processing them in a similar way to those algorithms.
+A detailed discussion is in Section~\ref{sec:overview:list}.
 

### 2019-02-25 sections/Related.tex
RW Ordered contexts

@@
 \paragraph{Ordered Contexts in Type Inference}
-Gundry et al.~\cite{gundry2010type}
+Gundry et al.~\cite{gundry2010type} revisit algorithm $\mathcal{W}$ and
+propose a new unification algorithm with the help of ordered contexts.
+Similar to DK's algorithm, information of meta variables flow from input contexts to output contexts.
+Not surprisingly, its information increase relation acts as a similar role to DK's context extension.
+Our algorithm, in contrast,
+eliminates the use of output contexts and their variable solution records ($\al = \tau$),
+simplifying the information propagation process through immediate substitution
+by collecting all the judgments in a single worklist.
 

### 2019-02-26 sections/Related.tex
Related work

@@
 the propagation of existential variable instantiations.
-However, the approach has two main drawbacks;
-it does not collect unused variable declarations effectively,
-and the simple form of judgment cannot handle inference modes, which outputs types.
+However, the approach has two main drawbacks:
+it does not collect unused variable declarations effectively;
+and the simple form of judgment cannot handle inference modes, which output types.
 The new worklist introduced in this paper inherits the simplicity of propagating instantiations,
-but overcome both of the issues by mixing judgments with declarations
-and using the continuation-passing-style judgment chains.
-\bruno{a comparison with our own ITP paper here!}
+but overcomes both of the issues by mixing judgments with declarations
+and using the continuation-passing-style judgment chains. Furthermore,
+we formalize the complete bidirectional type system by
+\citet{dunfield2013complete}, whereas Zhao et al. only formalize
+the subtyping relation.
 
@@
 while processing them in a similar way to those algorithms.
+\bruno{Still missing: 1) they duplicate context information; 2) their
+  focus is on unification, so they do not have to deal with inference
+  judgements as we do.}
 A detailed discussion is in Section~\ref{sec:overview:list}.
-
-\bruno{Missing: work on type inference and unification of dependent types.
-  Also missing: discussion about ordered contexts and the work by Gundry (in addition to DK).
-  Regarding these two, since we have an extended discussion in Section 2 you can have
-  a summarized discussion here, pointoing to Section 2 for further deatils.}
 

### 2019-02-26 sections/Metatheory.tex
S4 Metatheory: some corollaries, synced with Abella script

@@
 \begin{corollary}[Soundness, single judgment form]
-
+Given \emph{wf }$\Psi$ and $\Gm = \Psi$:
+\begin{enumerate}
+    \item If $\Gm \Vdash A \le B \redto \nil$ then $\Psi \vdash A \le B$;
+    \item If $\Gm \Vdash e \Lto A \redto \nil$ then $\Psi \vdash e \Lto A$;
+    \item If $\Gm \Vdash e \To_a \jg \redto \nil$ for any $\jg$, then there exists $A$
+        s.t. $\Psi \vdash e \To A$;
+    \item If $\Gm \Vdash \appInfAlg{A}{e} \redto \nil$ for any $\jg$, then there exists $C$
+        s.t. $\Psi \vdash \appInf{A}{e}{C}$.
+\end{enumerate}
 \end{corollary}
@@
 \begin{theorem}[Completeness]
-If \emph{wf }$\Gm$ and $\exists \Om$ s.t. $\Gm\sto\Om$ and $\Om\redto\nil$, then $\Gm \redto \nil$.
+If \emph{wf }$\Gm$ and $\Gm\sto\Om$ and $\Om\redto\nil$, then $\Gm \redto \nil$.
 \label{thm:completeness}
@@
 \begin{corollary}[Completeness, single judgment form]
+Given \emph{wf }$\Psi$ and $\Gm = \Psi$:
+\begin{enumerate}
+    \item If $\Psi \vdash A \le B$ then $\Gm \Vdash A \le B \redto \nil$;
+    \item If $\Psi \vdash e \Lto A$ then $\Gm \Vdash e \Lto A \redto \nil$;
+    \item If $\Psi \vdash e \To A$ then $\Gm \Vdash e \To_a 1 \le 1 \redto \nil$;
+    \item If $\Psi \vdash \appInf{A}{e}{C}$ then $\Gm \Vdash \appInfAlg{A}{e}[a][1 \le 1] \redto \nil$.
+\end{enumerate}
 \end{corollary}
@@
 on the worklist $\Gm$.
-$|\cdot|_e$ and $|\cdot|_\Leftrightarrow$ measure the total ``size'' of teams
+$|\cdot|_e$ and $|\cdot|_\Leftrightarrow$ measure the total ``size'' of terms
 and the total ``difficulty'' of judgments, respectively;
@@
 \begin{corollary}[Decidability of Declarative Typing]
+Given \emph{wf }$\Om$, it is decidable whether $\Om\redto\nil$ or not.
 \end{corollary}

### 2019-02-26 sections/Metatheory.tex
S4 except for 4.6 Abella

@@
 
+The original DK's declarative system, shown in Figures \ref{fig:decl:sub} and \ref{fig:decl:typing},
+has a few overlapping rules.
+We propose a slightly modified declarative system, by changing a few rules in DK's declarative system:
+\begin{enumerate}
+\item Add a constraint to the subtyping rule $\mathtt{\forall L}$
+$$
+\inferrule*[right=$\mathtt{\forall L'}$]
+{B \neq \all[b]B' \\ \Psi \vdash \tau \\ \Psi \vdash [\tau/a]A \le B}
+{\Psi \vdash \all A \le B}
+$$
+\item Remove the redundant rule $\mathtt{Decl1I}$,
+which could be easily derived by a combination of
+$\mathtt{DeclSub}$, $\mathtt{Decl1I{\To}}$ and $\mathtt{{\le}Unit}$:
+$$
+\inferrule*[right=$\mathtt{DeclSub}$]
+{
+    \inferrule*[right=$\mathtt{Decl1I{\To}}$]
+    {~}
+    {\Psi \vdash () \To 1}
+    \\
+    \inferrule*[right=$\mathtt{{\le}Unit}$]
+    {~}
+    {\Psi \vdash 1 \le 1}
+}
+{\Psi \vdash () \Lto 1}
+$$
+\item Add constraints to the subsumption rule $\mathtt{DeclSub}$
+$$
+\inferrule*[right=$\mathtt{DeclSub'}$]
+{e \neq \lam e' \\ A \neq \all A' \\ \Psi \vdash e \To A \\ \Psi \vdash A \le B}
+{\Psi \vdash e \Lto B}
+$$
+\end{enumerate}
+
+Now the new declarative system has no overlapping rules
+and becomes more similar to the algorithmic system,
+as the algorithmic rules contain similar constraints.
+In addition, we proved that the modified declarative system is equivalent to the original one.
+
+Modification (2) is relatively easy to justify, with the derivation given above:
+the rule is redundant and can be replaced by a combination of three other rules.
+
+Modificaitons (1) and (3) require inversion lemmas regarding rules that overlaps with them.
+Rule $\mathtt{\forall L}$ initially overlaps with $\mathtt{\forall R}$ when facing judgment
+$\Psi \vdash \all A \le \all[b]B$.
+The following inversion lemma for Rule $\mathtt{\forall R}$ settles the overlapping:
+\begin{lemma}[Invertibility of $\mathtt{\forall R}$]
+If $\Psi \vdash A \le \all B$ then $\Psi, a \vdash A \le B$.
+\end{lemma}
+The lemma implies that giving a higher priority to $\mathtt{\forall R}$ does not affect the derivation.
+Therefore the restriction $B \neq \all[b]B'$ is reasonable.
+
+Rule $\mathtt{DeclSub}$ overlaps with both $\mathtt{Decl\forall I}$ and $\mathtt{Decl{\to}I}$.
+Two similar inversion lemmas should be proven:
+\begin{lemma}[Invertibility of $\mathtt{Decl\forall I}$]
+If $\Psi \vdash e \Lto \all A$ then $\Psi, a \vdash e \Lto A$.
+\end{lemma}
+\begin{lemma}[Invertibility of $\mathtt{Decl{\to}I}$]
+If $\Psi \vdash \lam e \Lto A \to B$ then $\Psi, x:A \vdash e \Lto B$.
+\end{lemma}
+These lemmas basically state that applying the specific rules is always no worse than
+applying the more general Rule $\mathtt{DeclSub}$.
+Proving both of the lemmas require an important property of the declarative system,
+the subsumption lemma.
+We first define the context subtyping relation in Figure~\ref{fig:context_subtyping},
+and then show the subsumption lemma as follows.
+
+\begin{figure}[t]
+\framebox{$\Psi' \le \Psi$}
+$$\begin{aligned}
+\inferrule*[right=$\mathtt{CtxSubEmpty}$]
+{~}{\nil \le \nil}\qquad
+\inferrule*[right=$\mathtt{CtxSubTyVar}$]
+{\Psi' \le \Psi}{\Psi', a \le \Psi, a}\qquad
+\inferrule*[right=$\mathtt{CtxSubTmVar}$]
+{\Psi' \le \Psi \\ \Psi \vdash A' \le A}{\Psi', x:A' \le \Psi, x:A}\qquad
+\end{aligned}$$
+\caption{Context Subtyping}
+\label{fig:context_subtyping}
+\end{figure}
+
+\begin{lemma}[Subsumption]
+Given $\Psi' \le \Psi$:
+\begin{enumerate}
+    \item If $\Psi \vdash e \Lto A$ and $\Psi \vdash A \le A'$ then $\Psi' \vdash e \Lto A$;
+    \item If $\Psi \vdash e \To A$ then there exists
+        $A'$ s.t. $\Psi \vdash A' \le A$ and $\Psi' \vdash e \To A$;
+    \item If $\Psi \vdash \appInf{C}{e}{A}$ and $\Psi \vdash C' \le C$,
+        then there exists $A'$ s.t. $\Psi \vdash A' \le A$ and $\Psi' \vdash \appInf{C'}{e}{A'}$.
+\end{enumerate}
+\end{lemma}
+
+We tried to follow DK's proof on the subsumption lemma,
+but the manual proof has several problems and is not able to be formalized.
+Fortunately we fixed the issues and get the lemma proven.
+Details about this issue can be found in Appendix~\ref{}.
+
+\paragraph{Proof Sketch for Soundness and Completeness Theorems}
+We now have three systems defined: the DK's declarative system,
+the modified non-overlapping declarative system, and the algorithmic system.
+First, we have got the statement that the non-overlapping declarative system,
+as a restricted version, is equivalent to the DK's declarative system.
+Then, the soundness of the algorithm can be directly shown against the original declarative system.
+Finally, the completeness of the algorithm is shown against the non-overlapping declarative system,
+which is equivalent the DK's declarative system.
+As a result, we show that our algorithm is sound and complete with respect to DK's declarative system.
+
 %-------------------------------------------------------------------------------
@@
 The proof proceeds by induction on derivation of $\Gm\redto\nil$.
+Interesting cases are the those involving existential variable instantiations,
+including Rules 10, 11, 21 and 29.
 Proper applications of lemmas \ref{lem:insert} and \ref{lem:extract}
-finish off the theorem.
+help analyse the full instantiation of those existential variables.
+For example, when $\al$ is solved to be $\al[1] \to \al[2]$ in the algorithm,
+applying Extract gives two instantiations $\al[1] = \sigma$ and $\al[2] = \tau$.
+Inserting back $\al = \sigma \to \tau$ for the induction case
+matches the shape of induction hypothesis and thus finishes the corresponding case.
+
 \bruno{Give a bit more informal explanation here. Something like: If
@@
 We prove completeness by induction on the derivation of $\Om\redto\nil$.
-Since the declarative worklist is reduced judgment-by-judgment,
+Here we refer to the non-overlapping declarative system without chaning any syntax.
+Since the declarative worklist is reduced judgment-by-judgment
+(shown in Figure~\ref{fig:decl:worklist}),
 the induction always analyses the first judgment until it is satisfied.
+In Abella, we use a different relation to get the correct induction scheme.
+
 As the algorithmic system introduces existential variables,
@@
 
-All other cases are handled easily... \jimmy{more details if necessary}
+All other cases are relatively easy to prove.
+The Insert and Extract lemmas are applied when the algorithm uses existential variables,
+but transferred to a monotype for the declarative system,
+such as Rules 6, 10, 11, 12-17, 21 and 29.
 

### 2019-02-26 sections/Related.tex
Introduction

@@
 \jimmy{Guarded Impredicative}
-A recent work proposed Guarded Impredicative Polymorphism~\cite{},
+A recent work proposed Guarded Impredicative Polymorphism~\cite{Serrano2018},
 as an improvement on GHC's type inference algorithm with impredicative instantiation.

### 2019-02-26 sections/introduction.tex
Introduction

@@
 challenge several decidable type systems requiring some annotations 
-have been proposed~\cite{,dunfield2013complete,jones2007practical}
-\bruno{works on impredicative type inference are still missing}.
+have been proposed~\cite{,dunfield2013complete,jones2007practical,Serrano2018,le2003ml,leijen2008hmf,vytiniotis2008fph}.
 Two closely related type systems that 
@@
 popular among language designers and their ideas have been adopted by
-several modern functional languages, including Haskell, PureScript~\cite{} and
-Unison~\bruno{Add references. Here is the one for unison\url{http://unisonweb.org}} among others.
+several modern functional languages, including Haskell, PureScript~\cite{PureScript} and
+Unison~\cite{Unison} among others.
 In those type systems
@@
 such advanced forms of type inference in theorem provers.
-The remarkable exception is some work done on the formalization of 
+The remarkable exception is work done on the formalization of 
 certain parts of Hindley-Milner type inference~\cite{naraschewski1999type,
@@
 proposed recently to mechanically formalize an algorithm for
-\emph{polymorphic subtyping}. A key innovation in our work is how
+\emph{polymorphic subtyping}~\cite{odersky1996putting}. A key innovation in our work is how
 to adapt the idea of worklist judgments to
-\emph{inference judgements}, which are not needed for polymorphic subtyping.
-In our work we show how worklist
-judgements can be extended to deal with the inference judgements that
-are necessary for type-inference.  The idea is to use a continuation
-passing style to enable the transfer of inferred information across
+\emph{inference judgements}, which are not needed for polymorphic
+subtyping, but are necessary for type-inference.  The idea is to use a \emph{continuation
+passing style} to enable the transfer of inferred information across
 judgements. A further refinement to the idea of worklist judgements is
@@
 
-\item {\bf Unification of judgment worklists and contexts:} Another technical contribution is the unification
-  between ordered contexts and judgement worklists. This enables precise scope tracking
+\item {\bf Unification of worklists and contexts:} Another technical contribution is the unification
+  between ordered contexts and worklists. This enables precise scope tracking
   of variables in judgements, and avoids the duplication of context information across

### 2019-02-26 sections/algorithm.tex
revision of section 3, wip

--- 2019-02-12 sections/algorithm.tex
@@
 
-Figure~\ref{fig:alg:syntax} shows the syntax and the well-formedness judgement.
+Figure~\ref{fig:alg:syntax} shows the syntax---with accompanying well-formedness judgement---used by the algorithm.
 
@@
 \begin{aligned}
-\text{Type variables}\qquad&a, b\\
+% \text{Type variables}\qquad&a, b\\
 \text{Existential variables}\qquad&\al, \bt
@@
 %\text{Types}\qquad&A', B', C' &::=&\quad 1 \mid a \mid \forall x. A' \mid A'\to B'\\
-\text{Mono-types}\qquad&\tau &::=&\quad 1 \mid a \mid \tau_1\to \tau_2\\
+% \text{Mono-types}\qquad&\tau &::=&\quad 1 \mid a \mid \tau_1\to \tau_2\\
 %\text{Context}\qquad&\Psi &::=&\quad \nil \mid \Psi, a \mid \Psi, x:A' \color{red} \\[2mm]
@@
 % \noalign{\jimmy{\text{What do the markers do: $\G \vdash,a \vdash,b\vdash \j_1;\j_2;\j_3$ means $\G,a,b\vdash \j_1 \land \G,a\vdash \j_1,\j_2 \land \G\vdash \j_3$}}} \\[2mm]
-\text{Algorithmic types}\qquad&A, B, C &::=&\quad 1 \mid a \mid \al \mid \forall a. A \mid A\to B
+\text{Algorithmic types}\qquad&A, B, C &::=& \quad \ldots \mid \al % \quad 1 \mid a \mid \al \mid \forall a. A \mid A\to B
 \\[2mm]
-\text{Terms}\qquad&e&::=&\quad x \mid () \mid \lam{e} \mid e_1~e_2 \mid (e:A)
-\\[2mm]
+% \text{Terms}\qquad&e&::=&\quad x \mid () \mid \lam{e} \mid e_1~e_2 \mid (e:A)
+% \\[2mm]
 \text{Algorithmic judgement chain}\qquad&\jg &::=&\quad A \le B \mid e\Lto A \mid e\To_{a} \jg \mid \appInfAlg{A}{e}
@@
 \end{gather*}
-\caption{Syntax and Well-Formedness for the Algorithmic System}\label{fig:alg:syntax}
+\caption{Extended Syntax and Well-Formedness for the Algorithmic System}\label{fig:alg:syntax}
 \end{figure}
 
-\paragraph{Existential Variables} In order to find unknown types $\tau$ in the declarative system,
-the algorithmic system introduces \emph{existential variables} $\al, \bt$.
-They behave similarly to unification variables,
-but their scoping is restricted by their position in the
+\paragraph{Existential Variables} 
+The algorithmic system inherits the syntax of terms and types from 
+the declartive system. It only introduces one additional feature.
+In order to find unknown types $\tau$ in the declarative system, the
+algorithmic system extends the declarative types $A$ with \emph{existential variables} $\al, \bt$.
+They behave like unification variables,
+but their scope is restricted by their position in the
 worklist\bruno{be careful to ensure that the reader knows about
@@
 
-\paragraph{Monotypes, types and terms} The algorithm system also has a
-notion of monotypes, which is identical to the declarative system.
-\bruno{Since monotypes are identical to declarative system, maybe they
-do not need to be presented again. Perhaps drop their definition from 
-the syntax?} Algorithmic types include all the declarative types plus
-existential variables $\al$. \bruno{Aren't terms identical to the
-  Declarative system too? If they are briefly state that terms are
-  identical and do not show them here. Otherwise say what is different.}
-
 %% A \le B \mid e\Lto A \mid e\To_{a} \jg \mid \appInfAlg{A}{e}
 
-\paragraph{Judgments and judgment chains} 
-\emph{Judgment chain}s $\jg$ are the core components of our algorithmic
-type-checking. For conciseness, in the rest of the paper, we usually
-refer to ``judgment chains'' simply by ``judgments''.  There are four kinds
-of judgments in our system: subtyping ($A \le B$), checking ($e\Lto
+\paragraph{Judgment Chains} 
+Judgment chains, or judgments for short, $\jg$ are the core components of our algorithmic
+type-checking. There are four kinds
+of these judgments in our system: subtyping ($A \le B$), checking ($e\Lto
 A$), inference ($e\To_{a} \jg$) and
@@
 since their result is only success or failure. However both inference and
-application inference return a type that is used in further computation. We use
+application inference return a type that is used in further computation. We use a
 continuation-passing-style encoding to accomplish this. For example, the judgment
@@
 subtype of $B$. The \emph{unknown} type of $e$ is represented by a
-type variable $a$, which is used in the second judgement to denote the 
+type variable $a$, which is used as a placeholder in the second judgement to denote the 
 type of $e$.
@@
 \paragraph{Worklist Judgments} Our algorithm has a non-standard form.
-We combine the traditional contexts and judgment(s) into a single worklist.
-The worklist collects variable bindings and judgments in \emph{order}, which represents their scopings:
-each variable is visible only after it is declared.
-For example, $[a,x:a,x \Lto a]$ is a valid worklist, but $[\textcolor{red}{x} \Lto a, x:\textcolor{red}{a},a]$ is not (the red-colored symbols are out of scope).
+We combine traditional contexts and judgment(s) into a single object, the \emph{worklist} $\Gm$.
+The worklist is an \emph{ordered} collection of both variable bindings and judgments. The order captures the scope:
+only the objects that come after a variable's binding in the worklist can refer to it.
+For example, $[a,x:a,x \Lto a]$ is a valid worklist, but $[\textcolor{red}{x} \Lto a, x:\textcolor{red}{a},a]$ is not (the red symbols refer to out-of-scope variables).
 
@@
 where $\Gm$ is the worklist $\Gm_L,\bullet,\Gm_R$ with a hole ($\bullet$).
-Hole notations with the same name implicitly shares the same structure $\Gm_L$ and $\Gm_R$.
+Hole notations with the same name implicitly share the same structure $\Gm_L$ and $\Gm_R$.
 A multi-hole notation splits the worklist into more parts.
@@
 
-The algorithmic typing reduction rules, defined in Figure~\ref{fig:alg}, have the form $\Gm \rto \Gm'$.
-The worklists $\Gm, \Gm'$ contain both variable declarations and judgments.
-The reduction process treats the worklist as a stack.
-In every step it pops the first judgment from the worklist for processing,
-possibly pushes new judgments onto the worklist, and repeats this process until the list is empty.
-A worklist that fully reduces to empty corresponds to a successful type-checking.
+The algorithmic typing reduction rules, defined in Figure~\ref{fig:alg}, have
+the form $\Gm \rto \Gm'$.
+% The worklists $\Gm, \Gm'$ contain both variable declarations and judgments.
+The reduction process treats the worklist as a stack.  In every step it pops
+the first judgment from the worklist for processing and possibly pushes new
+judgments onto the worklist.  The syntax $\Gm\redto \Gm'$ denotes multiple
+reduction steps. If $\Gm\redto\nil$, we say that $\Gm$ reduces successfully;
+this corresponds to a successful type checking.
 
+Let us explain the rulres in more detail:
 Rules 1-3 pop a variable declaration; this garbage collects variables that are
 out of scope for the remaining judgements.
+All other rules concern a judgement at the front of the worklist. Logically we
+can discern 6 group of rules.
 
-All other rules concern a judgement at the front of the worklist. Logically we can discern 6 group of rules.
-
-Firstly, we have six subtyping rules (Rules 4-9) that have similar 
+\begin{description}
+\item[1. Declarative Rules]
+We have six subtyping rules (Rules 4-9) that are similar to their
 declarative counterparts. For instance, Rule 7 consumes a subtyping
@@
 
-\paragraph{Variable scopings within worklist}
+\paragraph{Worklist Variable Scoping}
 Rules 8 and 9 involve variable declarations and demonstrate how our
@@
 
-\paragraph{Existential decomposition}
-Secondly, Rules 10 and 11 are algorithmic extensions to Rule $\mathtt{{\le}{\to}}$.
-Both of them partially instantiate $\al$ to function types.
+\item[2. Existential decomposition]
+Rules 10 and 11 are algorithmic versions of Rule $\mathtt{{\le}{\to}}$; they
+both partially instantiate $\al$ to function types.
 The domain $\al[1]$ and range $\al[2]$ of the new function type are not determined:
 they are fresh existential variables with the same scope as $\al$.
-We remove $\al$ from the worklist, and insert $\al[1], \al[2]$ back to the same position.
-To propagate the instantiation to the rest of the worklist and maintain the well-formedness,
-every $\al$ is replaced by $\al[1] \to \al[2]$.
-The \emph{occurs-check} condition is to prevent a diverging infinite loop.
-For example, without such check $\al \le 1 \to \al$ would diverge.
+We remove replace $\al$ in the worklist with  $\al[1], \al[2]$.
+To propagate the instantiation to the rest of the worklist and maintain well-formedness,
+every reference to $\al$ is replaced by $\al[1] \to \al[2]$.
+The \emph{occurs-check} condition prevents divergence as usual.
+For example, without it $\al \le 1 \to \al$ would diverge.
 
-\paragraph{Solving existentials.} Thirdly, Rules 12-17 are base cases where an existential variable is solved.
-Each of the rules removes an existential variable and substitutes the
+\item[3. Solving existentials] Rules 12-17 are base cases where an existential variable is solved.
+They all remove an existential variable and substitute the
 variable with its solution in the remaining worklist. Importantly the rules
-respect the scoping of existential variables. For example, rule 12 
+respect the scope of existential variables. For example, rule 12 
 states that an existential variable $\al$ can be solved with another
@@
 
-\paragraph{Checking judgements}
-Fourthly, Rules 18-21 deal with checking judgments.
+\item[4. Checking judgements]
+Rules 18-21 deal with checking judgments.
 Rule 18 is similar to $\mathtt{DeclSub}$, but rewritten in the
@@
 The same instantiation technique as in Rules 10 and 11 applies.
-The declarative checking rule $\mathtt{Decl1I}$ is not translated to the algorithm,
-since combining Rule 18 and Rule 24 would give the same result.
+The declarative checking rule $\mathtt{Decl1I}$ does not have a direct counterpart in the algorithm, 
+because Rules 18 and 24 can be combined to give the same result.
 
-\paragraph{Inference judgments}
-Fifthly, Rules 22-26 deal with inference judgments.
+\item[5. Inference judgments]
+Rules 22-26 deal with inference judgments.
 Inference judgment chains are written in continuation-passing-style.
@@
 the return type might reflect more information than simply $\al \to \bt$.
-The variable scopings are maintained during the process:
+The variable scopes are maintained during the process:
 the assumption of argument type ($x:\al$) is recycled after checking against the function body;
@@
 
-\paragraph{Application inferences judgements}
+\item[6. Application inferences judgements]
 Finally, Rules 27-29 deal with application inference judgments.
 Rules 27 and 28 behaves similarly to declarative rules $\mathtt{Decl\forall App}$ and $\mathtt{Decl\to App}$.
-Rule 29 instantiate $\al$ to the function type $\al[1] \to \al[2]$, just like Rules 10, 11 and 21.
+Rule 29 instantiates $\al$ to the function type $\al[1] \to \al[2]$, just like Rules 10, 11 and 21.
 
-\subsection{Multiple reduction}
-The syntax $\Gm\redto \Gm'$ denotes multiple reduction steps.
-If $\Gm\redto\nil$, we say that $\Gm$ reduces successfully.
+\end{description}
 
@@
 \bruno{Derivation should be explained briefly at least.}
+
+\paragraph{Non-overlapping and Deterministic Reduction}
+An important feature of our algorithmic rules is that they are directly
+implementable. Indeed, although written in the form of reduction rules, they do
+not overlap and are thus deterministic.  This has the added benefit that our
+algorithm runs in polynomial time, which makes it suitable for practical use.
+
+Consider in particular Rules 8 and 9, which correspond to the declarative rules
+$\mathtt{{\le}\forall L}$ and $\mathtt{{\le}\forall R}$. While those
+declarative rules both match the goal $\all A\le \all[b]B$,
+we have eliminated this overlap in the algorithm by restricting Rule 8
+($B\neq\all B'$) and thus always applying Rule 9 to $\all A\le \all[b]B$.
+
+Similarly, the declarative rule $\mathtt{DeclSub}$ overlaps highly with the
+other checking rules. Its algorithmic counterpart is Rule 18. Yet, we have
+avoided the overlap with other algorithmic checking rules by adding
+side-conditions to Rule 18, namely $e\neq\lam e'$ and $B\neq\all B'$.
+
+\jimmy{Describe how to achieve the non-overlapping: a proof of declarative system,
+and then link to the metatheory for more details.}

### 2019-02-26 sections/Related.tex
remove Outside todo from related work
RW regarding comments

--- 2019-02-26 sections/Related.tex
@@
 undecidable~\cite{wells1999typability}, various decidable partial
-type-inference algorithms were developed~\cite{}.  The declarative system of
-this paper, proposed by \citet{dunfield2013complete}, is \emph{predicative}: 
+type-inference algorithms were developed.
+% No need to cite here: the following text is describing the point ~\cite{}.
+The declarative system of this paper,
+proposed by \citet{dunfield2013complete}, is \emph{predicative}: 
 $\forall$'s only instantiate to monotypes.  The monotype restriction
@@
 
-\jimmy{Dunfield's new work}
 DK's algorithm is shown to be sound, complete and decidable in 70 pages of manual proofs.
@@
 which creates difficulties when formalizing them in a proof assistant.
-Their follow-up work~\cite{} enriches the bidirectional higher-rank system with
+Their follow-up work~\cite{DunfieldIndexed} enriches the bidirectional higher-rank system with
 existentials and indexed types.
@@
 the lemma and counter-example.}
-\bruno{Missing Simon Peyton Jones work: which is the other predicative work.
-  Maybe the OutsideIn work can also be briefly mentioned (you can ask Tom for details
-  of that work).
- }
+
+\bruno{Missing Simon Peyton Jones work: which is the other predicative work.}
+\citet{jones2007practical} Developed another higher-rank predicative bidirectional type system.
+Their subtyping relation is enriched with deep skolemisation,
+which is more general than ours and allows more valid relations.
+In comparison to DK's system, they do not use the application inference judgment,
+resulting in complicated implicit instantiation taken care by the unification process for the algorithm.
+Although a hand-written proof is given, showing that the algorithm is sound and
+complete with respect to their declarative specification, no machanized proof is provided.
 
@@
 As a result, such algorithms tend to be more complicated, and thus less adopted in practise.
-\jimmy{Guarded Impredicative}
-A recent work proposed Guarded Impredicative Polymorphism~\cite{},
+A recent work proposed Guarded Impredicative Polymorphism~\cite{GuardedImpredicative},
 as an improvement on GHC's type inference algorithm with impredicative instantiation.
@@
 mechanized a type inference algorithm,
-with the help of locally nameless cofinite quantification~\cite{},
+with the help of locally nameless cofinite quantification~\cite{LocallyNameless},
 for Core ML extended with structural polymorphism and recursion.
@@
 in dependently typed inference and reconstruction.
-Their algorithms collect unification constraints and simplify them progressively in a single-step style.
+Either algorithm simplifies a collection of unification constraints progressively in a single-step style.
 In comparison, our algorithm mixes variable declarations with judgments,
@@
 while processing them in a similar way to those algorithms.
+Duplicated context in their unification judgments complicates the unification process,
+instead we make use of the nature of ordered context to control the scoping of unification variables.
+While their algorithms focus only on unification,
+our algorithm also deals with other type of judgments like type inference.
+A detailed discussion is in Section~\ref{sec:overview:list}.
+
 \bruno{Still missing: 1) they duplicate context information; 2) their
@@
   judgements as we do.}
-A detailed discussion is in Section~\ref{sec:overview:list}.
 

### 2019-02-26 sections/Related.tex
Merged RW minor conflicts

@@
 As a result, such algorithms tend to be more complicated, and thus less adopted in practise.
-A recent work proposed Guarded Impredicative Polymorphism~\cite{GuardedImpredicative},
+A recent work proposed Guarded Impredicative Polymorphism~\cite{Serrano2018},
 as an improvement on GHC's type inference algorithm with impredicative instantiation.

### 2019-02-26 sections/Metatheory.tex
4.1 revision, wip

@@
 These three results have been mechanically formalized and proved in the 
-Abella theorem prover~\cite{}.
+Abella theorem prover~\cite{AbellaDesc}.
 
-\subsection{Transfer to the Declarative System}
+%-------------------------------------------------------------------------------
+\subsection{Declarative Worklist}
 %\jimmy{The transfer rule outputs declarative judgment chains, which are later processed by another declarative relation, handling all the guessing.}
@@
 \begin{figure}[t]
-\framebox{$\Gm \sto \Om$} $\Gm$ transfers to $\Om$.
+\begin{gather*}
+\begin{aligned}
+\text{Declarative worklist}\qquad&\Om &::=&\quad \nil \mid \Om, a \mid \Om, x: A \mid \Om \Vdash \jg
+\end{aligned}
+\end{gather*}
+\hfill \framebox{$\Gm \sto \Om$} \hfill $\Gm$ instantiates to $\Om$.
 \begin{gather*}
@@
 \end{gather*}
-\caption{Transfer rules}
+\caption{Declarative Worklists and Instantiation}
 \label{fig:trans}
@@
 
-To aid formalizing the correspondence between the declarative and
-algorithmic systems, we use a relation that transfers algorithmic
-worklist judgements into corresponding declarative worklist judgements. In essence this
-judgment instantiates every existential variable with a well-formed monotype.
+To aid formalizing the correspondence between the declarative and algorithmic
+systems, we introduce the notion of declarative worklist $\Om$, define in
+Figure~\ref{fig:trans}. A declarative worklist $\Om$ has the same structure as an
+algorithmic worklist $\Gm$, but does not contain any extistential variables $\al$.
 
-The transfer rules are shown in Figure~\ref{fig:trans}.
+The relation $\Gm \sto \Om$ expresses that the algorithmic worklist $\Gm$ can 
+be instiantiated to the declarative worklist $\Om$, namely by appropriately instantiating
+all exitential variables $\al$ in $\Gm$ with well-scoped monotypes $\tau$.
+The rules of this instantiation relation are shown in Figure~\ref{fig:trans} too.
 Rule $\mathtt{{\sto}\al}$ replaces the first existential variable with a well-scoped monotype and
-repeats the process in the resulting worklist.
-Note that $\Om$, as a declarative worklist, contains no existential
-variables. Therefore by recursively applying this rule, the
-algorithmic worklist will finally become a declarative one.
-In order to maintain well-formedness,
-the substitution should affect all the judgments and term variable bindings in the scope of $\al$.
+repeats the process on the resulting worklist until no existential variable remains
+and thus the algorithmic worklist has become a declarative one.
+In order to maintain well-scopedness
+the substitution is applied to all the judgments and term variable bindings in the scope of $\al$.
 
-The transfer $\Gm\sto\Om$ is not deterministic.
+Observe that the instantiation $\Gm\sto\Om$ is not deterministic.
 % \bruno{Do you mean here
 %   ``is not deterministic''? Would that be more appropriate?}
-On the one hand, there are infinite possibilities to instantiate an existential variable,
-thus a judment $\Gm$ could be transferred to different $\Om$'s.
-On the other hand, any valid monotype in $\Om$ can be represented by an existential variable in $\Gm$,
-thus different $\Gm$'s could be transferred to a single $\Om$.
+From left to right, there are infinitely many possibilities to instantiate an existential variable and
+thus infinitely many declarative worklists that one can get from an algorithmic one.
+In the other direction, any valid monotype in $\Om$ can be abstracted to an
+existential variable in $\Gm$. Thus different $\Gm$'s can be instantiated to the same
+$\Om$.
 
@@
 \begin{figure}[t]
-\framebox{$\|\Om\|$} Judgment erasure for $\Om$, returns a declarative context $\Psi$.
+\hfill \framebox{$\|\Om\|$} \hfill Judgment erasure.
 \begin{gather*}
@@
 
-\framebox{$\Om \rto \Om'$} $\Om$ reduces to $\Om'$.
+\hfill \framebox{$\Om \rto \Om'$} \hfill $\Om$ reduces to $\Om'$.
 \begin{gather*}
@@
 
-Figure~\ref{fig:decl:worklist} relates the worklist-style declarative system
-to the original declarative system.
-A successful reduction means a valid variable declaration,
-a valid subtyping or checking judgment, or a successful guessing for inferences.
-Additionally, judgment chains are unfolded to match the original declarative rules.
-Although $\Om$ is defined differently from $\Psi$,
-our declarative worklist $\Om$ contains a ordered collection of variable definitions,
-which naturally fits declarative subtyping and typing.
-\bruno{I think that to make things precise what you want to show is
-  that you can always define a erasure function that drops the
-  judgements from the worklist, and retains only the variable
-  declarations (thus becomes a declarative environment). Isn't this
-  what you do in your formalization?}
-\jimmy{TODO: a erasure function would make things more clear...}
+Figure~\ref{fig:decl:worklist} defines a declarative version $\Om \sto \Om'$ of
+the algorithmic reduction rules, which relates the declarative worklists to the
+declarative type system, by checking every judgement entry in the worklist with
+the help of the corresponding conventional declarative judgement. Recall that
+the worklist consists of an interleaving of the typing context with judgments.
+By means of the auxiliary function $\|\Om\|$ we can extracts the typing context
+out of it, for use in the conventional declarative judgments. What is does to
+obtain the typing context is simply erase the judgment entries from the
+worklist.
 

### 2019-02-26 sections/algorithm.tex
wibble
4.1 revision, wip
S3 last paragraph, pointing to S4.2

@@
 \begin{figure}[htp]
-\framebox{$\Gm\rto \Gm'$} $\Gm$ reduces to $\Gm'$.
+\hfill \framebox{$\Gm\rto \Gm'$} \hfill $\Gm$ reduces to $\Gm'$.
 %\jimmy{Whenever there is a change on judgment count, marker should be properly added/removed. Decidability: all valid judgment chain should be reduced to nil.}
@@
 
-\jimmy{Describe how to achieve the non-overlapping: a proof of declarative system,
-and then link to the metatheory for more details.}
+Such restrictions are not arbitrarily imposed:
+we foramlly proved that the restricted algorithm is still complete.
+In Section~\ref{sec:metatheory:non-overlapping} we discuss the relavant metatheory,
+with the help of a restricted version of declarative system.

### 2019-02-26 sections/Metatheory.tex
S3 last paragraph, pointing to S4.2
further revision of 4.1

@@
 
-\hfill \framebox{$\Om \rto \Om'$} \hfill $\Om$ reduces to $\Om'$.
+\hfill \framebox{$\Om \rto \Om'$} \hfill Declarative transfer.
 \begin{gather*}
 \begin{aligned}
-\Om,a &\rto \Om \qquad \Om,x:A \rto \Om\\
-\Om\Vdash A\le B &\rto \Om \text{ when } \|\Om\| \vdash A\le B\\
-\Om\Vdash e\Lto A &\rto \Om \text{ when } \|\Om\| \vdash e\Lto A\\
-\Om\Vdash e\To_a \jg &\rto \Om\Vdash[A/a]\jg \text{ when } \|\Om\| \vdash e\To A\\
-\Om\Vdash \appInfAlg{A}{e} &\rto \Om\Vdash[C/a]\jg \text{ when } \|\Om\| \vdash \appInf{A}{e}{C}\\
+\Om,a &\rto \Om \\  \Om,x:A & \rto \Om\\
+\Om\Vdash A\le B &\rto \Om &\text{ when } \|\Om\| \vdash A\le B\\
+\Om\Vdash e\Lto A &\rto \Om & \text{ when } \|\Om\| \vdash e\Lto A\\
+\Om\Vdash e\To_a \jg &\rto \Om\Vdash[A/a]\jg & \text{ when } \|\Om\| \vdash e\To A\\
+\Om\Vdash \appInfAlg{A}{e} &\rto \Om\Vdash[C/a]\jg & \text{ when } \|\Om\| \vdash \appInf{A}{e}{C}\\
 \end{aligned}
 \end{gather*}
-\caption{Declarative Worklist Typing.\bruno{I think your names for the
-  two relations/definitions aren't great. I think what you're defining
-here should be called transfer. The previous relation should probably
-be called something different. Maybe something like Declarative
-worklist instantiation/realization for the previous relation?}}
+\caption{Declarative Transfer.}
 \label{fig:decl:worklist}
@@
 Figure~\ref{fig:decl:worklist} defines a declarative version $\Om \sto \Om'$ of
-the algorithmic reduction rules, which relates the declarative worklists to the
+the algorithmic reduction rules, which transfers judgements in the declarative worklists to the
 declarative type system, by checking every judgement entry in the worklist with

### 2019-02-27 sections/Metatheory.tex
Section 4

@@
 %-------------------------------------------------------------------------------
-\subsection{Declarative Worklist}
+\subsection{Declarative Worklist and Transfer}
 %\jimmy{The transfer rule outputs declarative judgment chains, which are later processed by another declarative relation, handling all the guessing.}
@@
 To aid formalizing the correspondence between the declarative and algorithmic
-systems, we introduce the notion of declarative worklist $\Om$, define in
+systems, we introduce the notion of a declarative worklist $\Om$, defined in
 Figure~\ref{fig:trans}. A declarative worklist $\Om$ has the same structure as an
@@
 
+\paragraph{Worklist instantiation.}
 The relation $\Gm \sto \Om$ expresses that the algorithmic worklist $\Gm$ can 
-be instiantiated to the declarative worklist $\Om$, namely by appropriately instantiating
-all exitential variables $\al$ in $\Gm$ with well-scoped monotypes $\tau$.
+be instiantiated to the declarative worklist $\Om$, by appropriately instantiating
+all existential variables $\al$ in $\Gm$ with well-scoped monotypes $\tau$.
 The rules of this instantiation relation are shown in Figure~\ref{fig:trans} too.
@@
 
-Figure~\ref{fig:decl:worklist} defines a declarative version $\Om \sto \Om'$ of
-the algorithmic reduction rules, which transfers judgements in the declarative worklists to the
-declarative type system, by checking every judgement entry in the worklist with
-the help of the corresponding conventional declarative judgement. Recall that
-the worklist consists of an interleaving of the typing context with judgments.
-By means of the auxiliary function $\|\Om\|$ we can extracts the typing context
-out of it, for use in the conventional declarative judgments. What is does to
-obtain the typing context is simply erase the judgment entries from the
-worklist.
+\paragraph{Declarative transfer.}
+Figure~\ref{fig:decl:worklist} defines a relation $\Om \rto \Om'$,
+which transfers all judgements in the declarative worklists to the
+declarative type system. This relation checks that every judgement entry in the worklist
+holds using a corresponding conventional declarative judgement. 
+The typing contexts of declarative judgements are recovered using an
+auxiliary erasure function $\|\Om\|$. The erasure function
+simply drops all judgment entries from the worklist, keeping only variable
+and type variable declarations.
 
-\subsection{Non-overlapping Declarative System}
-\jimmy{Talk about the modified subsumption rule, reasoning about the unit rule,
-and the subtyping $\forall$ overlapping}
+\subsection{Non-Overlapping Declarative System}
+\label{sec:metatheory:non-overlapping}
 
-The original DK's declarative system, shown in Figures \ref{fig:decl:sub} and \ref{fig:decl:typing},
-has a few overlapping rules.
-We propose a slightly modified declarative system, by changing a few rules in DK's declarative system:
+The original DK's declarative system, shown in Figures
+\ref{fig:decl:sub} and \ref{fig:decl:typing}, has a few overlapping
+rules.
+In contrast our algorithm removes all overlapping and makes it
+clear when to apply each rule.
+It is possible to create a slightly modified declarative system,
+by changing a few rules in DK's declarative system, that is completely
+non-overlapping as well. The changes are as follows:
 \begin{enumerate}
@@
 as the algorithmic rules contain similar constraints.
-In addition, we proved that the modified declarative system is equivalent to the original one.
-
+In addition, we proved soundness and completeness lemmas that show that
+the modified declarative system is equivalent to the original one. Thus the
+``restrictions'' do not change the expressive power of the system.
 Modification (2) is relatively easy to justify, with the derivation given above:
 the rule is redundant and can be replaced by a combination of three other rules.
-
-Modificaitons (1) and (3) require inversion lemmas regarding rules that overlaps with them.
+Modifications (1) and (3) require inversion lemmas regarding rules that overlaps with them.
 Rule $\mathtt{\forall L}$ initially overlaps with $\mathtt{\forall R}$ when facing judgment
@@
 \end{lemma}
-These lemmas basically state that applying the specific rules is always no worse than
+These lemmas basically state that applying the specific rules is never worse than
 applying the more general Rule $\mathtt{DeclSub}$.
@@
 We tried to follow DK's proof on the subsumption lemma,
-but the manual proof has several problems and is not able to be formalized.
-Fortunately we fixed the issues and get the lemma proven.
-Details about this issue can be found in Appendix~\ref{}.
+but the manual proof had several problems and we were not able to formalize it directly.
+Fortunately we found a different way todo the proof and get the lemma proven.
+Details about this issue can be found in Appendix~\ref{}.\bruno{reference}
 
-\paragraph{Proof Sketch for Soundness and Completeness Theorems}
+\paragraph{Three-Way Soundness and Completeness Theorems}
 We now have three systems defined: the DK's declarative system,
@@
 For any worklist $\Gm$ that reduces successfully,
-we can find a valid transfer $\Gm\sto\Om$ that solves all the existential variables,
-such that $\Om$ also reduces successfully.
+we can find a valid instantiation $\Gm\sto\Om$ that solves all the existential variables,
+such that $\Om$ also transfers all judgements successfully.
 \begin{theorem}[Soundness]
@@
 matches the shape of induction hypothesis and thus finishes the corresponding case.
-
-\bruno{Give a bit more informal explanation here. Something like: If
-  $\Gm$ is reducible then we can show that there is always an $\Om$
-  where all the corresponding declarative judgements hold.}
-\bruno{I wonder if 1 or 2 examples (concrete instantiations of the
-  theorem) would be helpful to complement the explanation.}
+Some immediate corollaries, which show the soundness for specific judgement forms are:
 
@@
 \end{corollary}
+\bruno{I do not think the corollary is precisely stated. I think you want todo
+  instantiation on $\Gm$ and then use erasure on the judgements, right? As it stands
+  it is wrong because there is a ``type error''. Worklists and declarative typing contexts are of different shapes
+so the equality does not make sense.}
 
@@
 We prove completeness by induction on the derivation of $\Om\redto\nil$.
-Here we refer to the non-overlapping declarative system without chaning any syntax.
+Here we refer to the non-overlapping declarative system.
 Since the declarative worklist is reduced judgment-by-judgment
@@
 
-For subtyping, algorithmic Rules 10 and 11 require special treatment.
+Most cases are relatively easy to prove.
+The Insert and Extract lemmas are applied when the algorithm uses existential variables,
+but transferred to a monotype for the declarative system,
+such as Rules 6, 10, 11, 12-17, 21 and 29.
+Algorithmic Rules 10 and 11 require special treatment.
 When the induction reaches the $\mathtt{{\le}{\to}}$ case,
 the first judgment is of shape $A_1 \to A_2 \le B_1 \to B_2$.
-One of the corresponding algorithmic judgment is $\al \le A \to B$.
+One of the corresponding algorithmic judgments is $\al \le A \to B$.
 However, the case analysis does not imply that $\al$ is fresh in $A$ and $B$,
@@
 in its corresponding algorithmic judgment.
-In other words, the declarative system does not accept infinite types,
-thus, as an example, we cannot find a monotype $\tau$ such that $\tau\le 1\to \tau$,
-which could be transferred by $\al\le 1\to\al$.
+In other words, the declarative system does not accept infinite types.
+A symmetric lemma holds for $A\to B \le \al$.
+%thus, as an example, we cannot find a monotype $\tau$ such that $\tau\le 1\to \tau$,
+%which could be transferred by $\al\le 1\to\al$.
 
@@
 \end{lemma}
-A symmetric lemma holds for $A\to B \le \al$.
-
-All other cases are relatively easy to prove.
-The Insert and Extract lemmas are applied when the algorithm uses existential variables,
-but transferred to a monotype for the declarative system,
-such as Rules 6, 10, 11, 12-17, 21 and 29.
 
@@
 \end{corollary}
+\bruno{Similar problem to the soundness corollaries here.}
 
 \subsection{Decidability}
+
+\bruno{Tom, this is the section that I feel less helpful. Maybe
+you can look at this one more closely?}
+
+The decidability theorem is:
+
 \begin{theorem}[Decidability]
@@
 $|\cdot|_e$ and $|\cdot|_\Leftrightarrow$ measure the total ``size'' of terms
-and the total ``difficulty'' of judgments, respectively;
-$|\cdot|_\forall$ and $|\cdot|_\to$ count the total number of
-universal quantifiers and function types, respectively;
+and the total ``difficulty'' of judgments, respectively.
+$|\cdot|_\forall$ and $|\cdot|_\to$ counts the total number of
+universal quantifiers and function types, respectively. Finally,
 $|\cdot|_{\al}$ counts the number of existential variables in the worklist.
@@
 \end{definition}
+\bruno{The definitions look a bit intimidating. I do not have immediate suggestions,
+  but perhaps Tom has suggestions on how to improve the presentation here.}
 
 It is not difficult to see that all but two algorithmic reduction rules
-decreases the group of measures.
+decrease the group of measures.
 (The result of Rule 29 could be directly reduced by Rule 28, which decreases the measures.)
-The two exceptions are Rules 10 and 11; they increases the number of existential variables
+The two exceptions are Rules 10 and 11. Both rules increase the number of existential variables
 without decreasing the number of universal quantifiers.
@@
 (Rule 7 applied right after Rule 10 or 11), in smaller sizes.
-Therefore we argue that such split makes the worklist easier, by analyzing a subset of worklist,
-which we call \emph{instantiation judgments} $\Gm_i$, with the help of $\Gm_\le$ for the proof.
+Therefore we argue that such split makes the worklist easier, by analyzing a subset of the worklist,
+which we call \emph{instantiation judgments} $\Gm_i$ (defined with the help of $\Gm_\le$ for the proof).
 
@@
 $\Gm'$ basically refers to the original part of worklist $\Gm$.
-During the reduction, minor changes to worklist might occur, such as
-existential variable solving, insertion and instantiation,
-but the main structure of it remains unchanged.
+During the reduction, minor changes to the worklist may occur, such as
+existential variable solving or insertion and instantiation.
+Nevertheless, the main structure of remains unchanged.
 
 Shown in Figure~\ref{fig:worklist_ext},
-worklist extension formally describe the minor changes to worklist during algorithmic reduction.
+worklist extension formally describes the minor changes to worklist during algorithmic reduction.
 When the algorithm processes the worklist, the first judgment is processed,
-and the rest of the worklist roughly remains its shape.
+and the rest of the worklist roughly preserves its shape.
 Worklist extension is formally defined by two cases:

### 2019-02-27 sections/algorithm.tex
S3 algorithm: exaplaination for the sample derivation

@@
 \\[2mm]
-\text{Algorithmic worklist}\qquad&\Gm &::=&\quad \nil \mid \Gm, a \mid \Gm, \al \mid \Gm, x: A \mid \Gm \Vdash \jg\\
-\text{Declarative worklist}\qquad&\Om &::=&\quad \nil \mid \Om, a \mid \Om, x: A \mid \Om \Vdash \jg
+\text{Algorithmic worklist}\qquad&\Gm &::=&\quad \nil \mid \Gm, a \mid \Gm, \al \mid \Gm, x: A \mid \Gm \Vdash \jg%\\
+% \text{Declarative worklist}\qquad&\Om &::=&\quad \nil \mid \Om, a \mid \Om, x: A \mid \Om \Vdash \jg
 \end{aligned}
@@
 
-Figure~\ref{fig:alg:sample} shows a sample derivation of the
-algorithm.
+Figure~\ref{fig:alg:sample} shows a sample derivation of the algorithm.
+The judgment given is to check the application $(\lam x)~()$ against the unit type.
+According to the algorithm, we apply Rule 18 (subsumption), changing to inference mode.
+Type inference of the application breaks into two steps by Rule 26:
+firstly we infer the type of the function,
+and then the application inference judgment help determine the return type.
+The type of the identity function, $\lam x$, is inferred to be $\al \to \al$,
+which can be interpreted as $\tau \to \tau$ for any valid monotype $\tau$,
+after the following 5 steps:
+a checking on the body of the lambda function (Rule 25),
+calling the inference mode for the checking (Rule 18)
+a term variable inference (Rule 22),
+a subtyping on existential variables (Rule 12) and
+a garbage collection of the term variable $x$ (Rule 3).
+
+After that, Rule 28 changes the application inference judgment to
+a checking on the argument against the input type $\al$ and returns the output type $\al$.
+Checking $()$ against the existential variable $\al$ solves $\al$ to the unit type $1$
+through Rules 18, 24 and 16.
+Immediately after $\al$ is solved, the algorithm substitute every occurrence of $\al$ to $1$.
+Therefore the worklist remains a simple subtyping judgment $1 \le 1$, which finishes off by Rule 4.
+Finally we get an empty worklist, indicating the success of the whole derivation.
+
+To summarize, our algorithm accepts the type checking $(\lam x)~() \Lto 1$.
+
 \bruno{Derivation should be explained briefly at least.}

### 2019-02-27 sections/Related.tex
Related Work

@@
 DK's algorithm is shown to be sound, complete and decidable in 70 pages of manual proofs.
-Though carefully written, some of the proofs are incorrect (see discussion in Appendix),
+Though carefully written, some of the proofs are incorrect (see discussion in Appendix~\ref{}),
 which creates difficulties when formalizing them in a proof assistant.
-Their follow-up work~\cite{DunfieldIndexed} enriches the bidirectional higher-rank system with
+In their follow-up work \citet{DunfieldIndexed} enrich the bidirectional higher-rank system with
 existentials and indexed types.
-With a more complex declarative system, they developed a proof of over 150 pages.
+With a more complex declarative system, they developed a proofs of over 150 pages.
 It is even more difficult to argue its correctness for every single detail within such a big development.
 Unfortunately, we find that their Lemma 26 (Parallel Admissibility) appears to have the same issue 
-as lemma 29 discussed in Appendix: the conclusion is false.
+as lemma 29 in \citet{dunfield2013complete}: the conclusion is false. We also discuss
+the issue in more detail in Appendix~\ref{}. 
 \bruno{I think you will need to justify this better and in the appendix you must show
@@
 
-\bruno{Missing Simon Peyton Jones work: which is the other predicative work.}
-\citet{jones2007practical} Developed another higher-rank predicative bidirectional type system.
-Their subtyping relation is enriched with deep skolemisation,
+\citet{jones2007practical} developed another higher-rank predicative bidirectional type system.
+Their subtyping relation is enriched with \emph{deep skolemisation},
 which is more general than ours and allows more valid relations.
 In comparison to DK's system, they do not use the application inference judgment,
-resulting in complicated implicit instantiation taken care by the unification process for the algorithm.
-Although a hand-written proof is given, showing that the algorithm is sound and
-complete with respect to their declarative specification, no machanized proof is provided.
+resulting in a complicated mechanism for implicit instantiation taken care by the unification process for the algorithm.
+A manual proof is given, showing that the algorithm is sound and
+complete with respect to their declarative specification.
 
@@
 but unfortunately its subtyping system is already undecidable~\cite{tiuryn1996subtyping}.
-Works on partial impredicative type-inference algorithms\cite{le2003ml,leijen2008hmf,vytiniotis2008fph}
+Works on partial impredicative type-inference algorithms~\cite{le2003ml,leijen2008hmf,vytiniotis2008fph}
 navigate a variety of design tradeoffs for a decidable algorithm.
-As a result, such algorithms tend to be more complicated, and thus less adopted in practise.
-A recent work proposed Guarded Impredicative Polymorphism~\cite{Serrano2018},
+As a result, such algorithms tend to be more complicated, and thus less adopted in practice.
+Recent work proposed \emph{Guarded Impredicative Polymorphism}~\cite{Serrano2018},
 as an improvement on GHC's type inference algorithm with impredicative instantiation.
-They make use of local informations in $n$-ary applications to
+They make use of local information in $n$-ary applications to
 infer polymorphic instantiations with a relatively simple specification and unification algorithm.
@@
 In all previous work on type inference for higher-ranked polymorphism
-(predicative and impredicative) discussed above, all proofs and
+(predicative and impredicative) discussed above, proofs and
 metatheory are manual. The only partial effort on mechanizing aspects of type inference
@@
 However, mechanising certain algorithmic aspects, like unification and
-constraint solving, has received very little attention and are still challenging.
-Moreover, while most tools supports local (input) contexts in a neat way, which have a simple
-binding and scoping structure, many practical type-inference algorithms require
+constraint solving, has received very little attention and is still challenging.
+Moreover, while most tools supports local (input) contexts in a neat way,
+many practical type-inference algorithms require
 more complex binding structures with output contexts or various forms of constraint solving procedures.
@@
 has been manually proven to be sound and complete
-with respect to the Hindley-Milner type system~\cite{}.
+with respect to the Hindley-Milner type system~\cite{hindley1969principal,milner1978theory,damas1982principal}.
 After around 15 years, the algorithm was formally verified by
@@
 Similar to DK's algorithm, information of meta variables flow from input contexts to output contexts.
-Not surprisingly, its information increase relation acts as a similar role to DK's context extension.
+Not surprisingly, its information increase relation has a similar role to DK's context extension.
 Our algorithm, in contrast,
@@
 Some work~\cite{Reed2009,Abel2011higher} adopts a similar idea to this paper
-in dependently typed inference and reconstruction.
-Either algorithm simplifies a collection of unification constraints progressively in a single-step style.
+in work on unification for dependently typed languages. Similarly to our work
+the algorithms need to be very careful about scoping, since the order of variable
+declarations is fundamental in a dependently typed setting. 
+Their algorithms simplify a collection of unification constraints progressively in a single-step style.
 In comparison, our algorithm mixes variable declarations with judgments,
@@
 while processing them in a similar way to those algorithms.
-Duplicated context in their unification judgments complicates the unification process,
-instead we make use of the nature of ordered context to control the scoping of unification variables.
+One important difference is that contexts are
+duplicated in their unification judgments, which complicates the unification process,
+since the information of each local context needs to be synchronized.
+Instead we make use of the natural of ordered context to control the scoping of unification variables.
 While their algorithms focus only on unification,
@@
 A detailed discussion is in Section~\ref{sec:overview:list}.
-
-\bruno{Still missing: 1) they duplicate context information; 2) their
-  focus is on unification, so they do not have to deal with inference
-  judgements as we do.}
-
-\bruno{Final comment: You should essentially have all the discussion that we have in ITP +
-  some more. So, double-check carefully that all the works we cite in our related work in ITP
-are cited here, and that we include all discussions!}

### 2019-02-27 main.tex
create the appendix tex file

@@
 \usepackage{graphicx}
+
+%\usepackage{tikz}
 
@@
 \appendix
-\section{Appendix}
-
-\jimmy{Lemma 29 is false. TODO structure and more text}
-The first conclusion of the lemma says that if $\Gm_L\longrightarrow \Delta_R$ and
-$\Gm_L,\Gm_R\longrightarrow \Delta_L,\Delta_R$ then
-$\Gm_L,\al,\Gm_R\longrightarrow \Delta_L,\al,\Delta_R$.
-If we pick $\Gm_L = \nil, \Gm_R = \bt, \Delta_L = \bt, \Delta_R = \nil$, the conditions
-$\nil\longrightarrow \bt$ and $\bt\longrightarrow \bt$ hold, but the conclusion
-$\al,\bt \longrightarrow \bt,\al$ does not hold.
+\input{sections/Appendix.tex}
 

### 2019-02-27 sections/Appendix.tex
create the appendix tex file

\section{Appendix}

\jimmy{Lemma 29 is false. TODO structure and more text}
The first conclusion of the lemma says that if $\Gm_L\longrightarrow \Delta_R$ and
$\Gm_L,\Gm_R\longrightarrow \Delta_L,\Delta_R$ then
$\Gm_L,\al,\Gm_R\longrightarrow \Delta_L,\al,\Delta_R$.
If we pick $\Gm_L = \nil, \Gm_R = \bt, \Delta_L = \bt, \Delta_R = \nil$, the conditions
$\nil\longrightarrow \bt$ and $\bt\longrightarrow \bt$ hold, but the conclusion
$\al,\bt \longrightarrow \bt,\al$ does not hold.

### 2019-02-27 sections/Metatheory.tex
S4.3-4.4 corollaries

@@
 \begin{corollary}[Soundness, single judgment form]
-Given \emph{wf }$\Psi$ and $\Gm = \Psi$:
+Given \emph{wf }$\Gm$:
 \begin{enumerate}
-    \item If $\Gm \Vdash A \le B \redto \nil$ then $\Psi \vdash A \le B$;
-    \item If $\Gm \Vdash e \Lto A \redto \nil$ then $\Psi \vdash e \Lto A$;
-    \item If $\Gm \Vdash e \To_a \jg \redto \nil$ for any $\jg$, then there exists $A$
-        s.t. $\Psi \vdash e \To A$;
-    \item If $\Gm \Vdash \appInfAlg{A}{e} \redto \nil$ for any $\jg$, then there exists $C$
-        s.t. $\Psi \vdash \appInf{A}{e}{C}$.
+    \item If $\Gm \Vdash A \le B \redto \nil$\\
+        then there exist $A', B', \Om$ s.t.
+        $\Gm \Vdash A \le B \sto \Om \Vdash A' \le B'$ and $\|\Om\| \vdash A' \le B'$;
+    \item If $\Gm \Vdash e \Lto A \redto \nil$\\
+        then there exist $e', A', \Om$ s.t.
+        $\Gm \Vdash e \Lto A \sto \Om \Vdash e \Lto A'$ and $\|\Om\| \vdash e \Lto A'$;
+    \item If $\Gm \Vdash e \To_a \jg \redto \nil$ for any $\jg$\\
+        then there exists $\Om, \jg', A$ s.t.
+        $\Gm \sto \Om$ and $\|\Om\| \vdash e \To A$;
+    \item If $\Gm \Vdash \appInfAlg{A}{e} \redto \nil$ for any $\jg$\\
+        then there exists $\Om, \jg', A', C$ s.t.
+        $\Gm \Vdash \appInfAlg{A}{e} \sto \Om \Vdash \appInfAlg{A'}{e}[a][\jg']$
+            and $\|\Om\| \vdash \appInf{A'}{e}{C}$.
 \end{enumerate}
@@
 If $(\Gm \Vdash \al \le A \to B) \sto (\Om \Vdash C \le A_1 \to B_1)$ and
-$\Om \vdash C \le A_1 \to B_1$, then $\al\notin FV(A) \cup FV(B)$.
+$\|\Om\| \vdash C \le A_1 \to B_1$, then $\al\notin FV(A) \cup FV(B)$.
 \end{lemma}
@@
 \begin{corollary}[Completeness, single judgment form]
-Given \emph{wf }$\Psi$ and $\Gm = \Psi$:
+Given \emph{wf }$\Gm$ containing no judgment chains:
 \begin{enumerate}
-    \item If $\Psi \vdash A \le B$ then $\Gm \Vdash A \le B \redto \nil$;
-    \item If $\Psi \vdash e \Lto A$ then $\Gm \Vdash e \Lto A \redto \nil$;
-    \item If $\Psi \vdash e \To A$ then $\Gm \Vdash e \To_a 1 \le 1 \redto \nil$;
-    \item If $\Psi \vdash \appInf{A}{e}{C}$ then $\Gm \Vdash \appInfAlg{A}{e}[a][1 \le 1] \redto \nil$.
+    \item If $\Gm \Vdash A \le B \sto \Om \Vdash A' \le B'$ and $\|\Om\| \vdash A' \le B'$
+        then $\Gm \Vdash A \le B \redto \nil$;
+    \item If $\Gm \Vdash e \Lto A \sto \Om \Vdash e \Lto A'$ and $\|\Om\| \vdash e \Lto A'$
+        then $\Gm \Vdash e \Lto A \redto \nil$;
+    \item If $\Gm \sto \Om$ and exists $A$ s.t. $\|\Om\| \vdash e \To A$
+        then $\Gm \Vdash e \To_a 1 \le 1 \redto \nil$;
+    \item If $\Gm \Vdash \appInfAlg{A}{e}[a][1 \to 1] \sto \Om \Vdash \appInfAlg{A'}{e}[a][1 \to 1]$
+        and exists $C$ s.t. $\|\Om\| \vdash \appInf{A'}{e}{C}$
+        \\
+        then $\Gm \Vdash \appInfAlg{A}{e}[a][1 \le 1] \redto \nil$.
 \end{enumerate}

### 2019-02-27 sections/Metatheory.tex
revision of 4.2

--- 2019-02-27 sections/Metatheory.tex
@@
 
+%-------------------------------------------------------------------------------
 \subsection{Non-Overlapping Declarative System}
@@
 
-The original DK's declarative system, shown in Figures
-\ref{fig:decl:sub} and \ref{fig:decl:typing}, has a few overlapping
-rules.
-In contrast our algorithm removes all overlapping and makes it
-clear when to apply each rule.
-It is possible to create a slightly modified declarative system,
-by changing a few rules in DK's declarative system, that is completely
-non-overlapping as well. The changes are as follows:
+DK's declarative system, shown in Figures \ref{fig:decl:sub} and
+\ref{fig:decl:typing}, has a few overlapping rules. In contrast, our algorithm
+has removed all overlap; at most one rule applies in any given situation.
+This discrepancy makes it more difficult to relate the two systems.
+
+To simplify matters, we introduce an intermediate system that is still declarative
+in nature, but has no overlap. This intermediate system differs only in a few 
+rules from DK's declarative system; these are the changes:
 \begin{enumerate}
-\item Add a constraint to the subtyping rule $\mathtt{\forall L}$
+\item Restrict the shape of $B$ in the rule $\mathtt{\forall L}$ subtyping rule:
 $$
@@
 $$
-\item Remove the redundant rule $\mathtt{Decl1I}$,
-which could be easily derived by a combination of
+\item Drop the redundant rule $\mathtt{Decl1I}$,
+which can be easily derived by a combination of
 $\mathtt{DeclSub}$, $\mathtt{Decl1I{\To}}$ and $\mathtt{{\le}Unit}$:
@@
 $$
-\item Add constraints to the subsumption rule $\mathtt{DeclSub}$
+\item Restrict the shapes of $e$ and $A$ in the subsumption rule $\mathtt{DeclSub}$:
 $$
@@
 \end{enumerate}
+The resulting declarative system has no overlapping rules
+and more closely resembles the aglorithmic system, which contains
+the same shape constraints.
 
-Now the new declarative system has no overlapping rules
-and becomes more similar to the algorithmic system,
-as the algorithmic rules contain similar constraints.
-In addition, we proved soundness and completeness lemmas that show that
-the modified declarative system is equivalent to the original one. Thus the
-``restrictions'' do not change the expressive power of the system.
+We have proven soundness and completeness of the non-overlapping declarative
+system with respect to the overlapping one to establish their equivalence.
+Thus the restrictions do not change the expressive power of the system.
 Modification (2) is relatively easy to justify, with the derivation given above:
 the rule is redundant and can be replaced by a combination of three other rules.
-Modifications (1) and (3) require inversion lemmas regarding rules that overlaps with them.
-Rule $\mathtt{\forall L}$ initially overlaps with $\mathtt{\forall R}$ when facing judgment
+Modifications (1) and (3) require inversion lemmas for the rules that overlap.
+Firstly, Rule $\mathtt{\forall L}$ overlaps with Rule $\mathtt{\forall R}$ for the judgment
 $\Psi \vdash \all A \le \all[b]B$.
-The following inversion lemma for Rule $\mathtt{\forall R}$ settles the overlapping:
+The following inversion lemma for Rule $\mathtt{\forall R}$ resolves the overlap:
 \begin{lemma}[Invertibility of $\mathtt{\forall R}$]
@@
 \end{lemma}
-The lemma implies that giving a higher priority to $\mathtt{\forall R}$ does not affect the derivation.
-Therefore the restriction $B \neq \all[b]B'$ is reasonable.
+The lemma implies that preferring Rule $\mathtt{\forall R}$ does not affect the derivability of the judgment.
+Therefore the restriction $B \neq \all[b]B'$ in $\mathtt{\forall L'}$ is valid.
 
-Rule $\mathtt{DeclSub}$ overlaps with both $\mathtt{Decl\forall I}$ and $\mathtt{Decl{\to}I}$.
-Two similar inversion lemmas should be proven:
+Secondly, Rule $\mathtt{DeclSub}$ overlaps with both $\mathtt{Decl\forall I}$ and $\mathtt{Decl{\to}I}$.
+We have proven two inversion lemmas for these overlaps:
 \begin{lemma}[Invertibility of $\mathtt{Decl\forall I}$]
@@
 \end{lemma}
-These lemmas basically state that applying the specific rules is never worse than
-applying the more general Rule $\mathtt{DeclSub}$.
-Proving both of the lemmas require an important property of the declarative system,
-the subsumption lemma.
-We first define the context subtyping relation in Figure~\ref{fig:context_subtyping},
-and then show the subsumption lemma as follows.
+These lemmas express that applying the more specific rules, rather than the more general rule $\mathtt{DeclSub}$, does not reduce the expressive power.
+
+The proofs of the above two lemmas rely on an important property of the
+declarative system, the \emph{subsumption lemma}. To be able to formulate this
+lemma, Figure~\ref{fig:context_subtyping} introduces the \emph{context
+subtyping relation} $\Psi \le \Psi'$. Context $\Psi$ subsumes context $\Psi'$ if they
+bind the same variables in the same order, but the types $A$ of the term variables $x$
+in the former are subtypes of types $A'$ assigned to those term variables in the latter.
+Now we can state the subsumption lemma:
 
@@
 \end{lemma}
+This lemma expresses that any derivation in a context $\Psi$ has a corresponding derivation in any context
+$\Psi'$ that it subsumes.
 
-We tried to follow DK's proof on the subsumption lemma,
-but the manual proof had several problems and we were not able to formalize it directly.
-Fortunately we found a different way todo the proof and get the lemma proven.
-Details about this issue can be found in Appendix~\ref{}.\bruno{reference}
+We have tried to follow DK's manual proof of this lemma,
+but we discovered several problems in their reasoning that we have been able to address.
+Fortunately we have found a different way to prove the lemma.
+The details of this issue can be found in Appendix~\ref{}.\bruno{reference}
 
 \paragraph{Three-Way Soundness and Completeness Theorems}
-We now have three systems defined: the DK's declarative system,
-the modified non-overlapping declarative system, and the algorithmic system.
-First, we have got the statement that the non-overlapping declarative system,
-as a restricted version, is equivalent to the DK's declarative system.
-Then, the soundness of the algorithm can be directly shown against the original declarative system.
-Finally, the completeness of the algorithm is shown against the non-overlapping declarative system,
-which is equivalent the DK's declarative system.
-As a result, we show that our algorithm is sound and complete with respect to DK's declarative system.
+We now have three systems that can be related: DK's overlapping declarative system,
+our non-overlapping declarative system, and our algorithmic system.
+We have already establishes the first relation, that the two declarative
+systems are equivalent.
+
+In what follows, we will establish the soundness of our algorithm directly
+against the original overlapping declarative system. However, we have found
+that showing completeness of the algorithm is easier against the
+non-overlapping declarative system. Of course, as a corollary, it follows that
+our algorithm is also complete with respect to DK's declarative system.
 

### 2019-02-27 sections/Overview.tex
RW lists of judgments
Section 2

@@
 \end{figure}
-\bruno{Some of the rule names are messed up.}
 
@@
 is know to give rise to an undecidable subtyping relation~\cite{tiuryn1996subtyping}.
-Peyton Jones et al.~\cite{jones2007practical} also impose the restriction of
+\citet{jones2007practical} also impose the restriction of
 predicative instantiation in their type system.
@@
 check whether the synthesized type entails the checked type.
-
 Lambda abstractions are the hardest construct of the bidirectional
@@
 \paragraph{Overlapping Rules}
-\jimmy{Mention the resolution of overlapping briefly, and then point to algorithm and metatheory}
 Some of the declarative rules overlaps with each other.
@@
 which gives more flexibility on the choice of $\tau$.
-
 Declarative typing rule $\mathtt{DeclSub}$ overlaps with
@@
 i.e. $\mathtt{Decl\forall I}$ and $\mathtt{Decl\to I}$ should have
-higher priority to $\mathtt{DeclSub}$.
+higher priority to $\mathtt{DeclSub}$. We will come back to this
+topic in Section~\ref{sec:metatheory}.
+\begin{comment}
 For example, $\Psi\vdash \lam x \Lto \all a\to a$ succeeds if derived from
@@
 and will successfully derive the judgment.
+\end{comment}
 
@@
 from the output stack together with the subsequent part $\Theta$, which may
-refer to $\al$. \jimmy{which may contain definition/solution of $\al$ and references to $\al$}
+refer to $\al$. 
 (Remember that later entries may refer to earlier
 ones, but not vice versa.) This way $\al$ does not escape its scope.
-
+\begin{comment}
 A type variable may have a similar functionality to the scoping markers.
@@
 $$
+\end{comment}
 
@@
 Actually we found a scoping issue related to the subsumption rule,
-which might cause existential variable leakage across judgments.
+which might cause existential variables to leak across judgments.
 A detailed discussion on this issue with a possible fix is in Section~\ref{sec:discussion:scoping}.
@@
 }
-\bruno{We should acknowledge Ningning for finding this problem, but I think
-  this may be against the anonymous submission? Perhaps we do not mention her name
-in the submission, and if accepted we mention her?}
-\tom{You have communicated this to the authors and not received a fix, right?
-     We could mention that too then.}
 
@@
 variables across judgments. 
-
 DK is already quite disciplined on these accounts, in particular compared to
-traditional constraint-based type-inference algorithms like Algorithm W which
+traditional constraint-based type-inference algorithms like Algorithm $\mathcal{W}$~\cite{milner1978theory} which
 features an implicit global scope for all type variables. Indeed, DK uses its
@@
 %-------------------------------------------------------------------------------
-\subsection{Judgment Lists}
+\subsection{Judgment Lists}\label{sec:overview:list}
 
@@
 
-In the above standard unification problem, all variables are implicitly in the same
-global scope. In contrast, recent unification algorithms for dependently-typed languages
+In the above unification problem, all variables are implicitly in the same
+global scope. Recent unification algorithms for dependently-typed languages
 are more explicit about scopes. For instance, \citet{Reed2009} represents a unification
@@
 by using a \emph{worklist}-based judgement of the form $$\Gamma \vdash \Omega$$
-where $\Omega$ is a worklist, i.e. a sequence, of subtyping problems of the
+where $\Omega$ is a worklist (or sequence) of subtyping problems of the
 form $A \leq B$.  The judgement is defined by case analysis on the first
@@
 Using both a single common ordered context $\Gamma$ and a worklist $\Omega$ greatly
-simplifies propagating the instantiation of type variables that happen in one
+simplifies propagating the instantiation of type variables in one
 subtyping problem to the remaining ones.
@@
 inference task together with the the work that depends on its outcome. For
-instance, a type inference task takes the form $e \Rightarrow_a \omega$. It
+instance, a type inference task takes the form $e \Rightarrow_a \omega$.
+This form models a \emph{continuation passing style}:
+it
 denotes that we need to infer the type, say $A$, for expression $e$ and use it
@@
 There are three (blue-colored) judgment chains in the worklist,
-where the first and second judgment chains are two subtyping judgments,
+where the first and second judgment chains (from the right) are two subtyping judgments,
 and the third judgment chain, $(\lam x)~() \To_a a \le \al$,

### 2019-02-27 sections/algorithm.tex

@@
 \begin{description}
-\item[1. Declarative Rules]
+\item[1. Declarative Subtyping]
 We have six subtyping rules (Rules 4-9) that are similar to their
@@
 after being unreferenced.
-\jimmy{An example on subtyping: $\nil\vdash (1\to 1)\to 1 \le (\forall a. a\to a) \to 1$
+
+\jimmy{New example on Feb 27}
+\paragraph{Example} An example derivation on the subtyping judgment
+$(1\to 1)\to 1 \le (\all 1\to 1) \to 1$ is shown as follows:
+\begin{gather*}
+\begin{aligned}
+  & \nil\vdash (1\to 1)\to 1 \le (\all 1 \to 1) \to 1\\
+\rrule{7}  & \nil \Vdash 1 \le 1 \Vdash \all 1 \to 1 \le 1 \to 1\\
+\rrule{8}  & \nil \Vdash 1 \le 1 ,\al \Vdash 1 \to 1 \le 1 \to 1\\
+\rrule{7}  & \nil \Vdash 1 \le 1 ,\al \Vdash 1 \to 1 \Vdash 1 \le 1\\
+\rrule{4} & \nil \Vdash 1 \le 1 ,\al \Vdash 1 \to 1\\
+\rrule{4} & \nil \Vdash 1 \le 1 ,\al\\
+\rrule{2} & \nil \Vdash 1 \le 1\\
+\rrule{4} & \nil
+\end{aligned}
+\end{gather*}
+Firstly, the subtyping of two function types is split into two judgments by Rule 7:
+a convariant subtyping on the return type and a contravariant subtyping on the argument type.
+Then we apply Rule 8 to reduce the $\forall$ quantifier on the left side.
+The type $\forall a. 1 \to 1$ is actually not a general type, which
+has an almost equivalent behavior to the type $1 \to 1$.
+But according to the rule, we still introduce an existential variable $\al$ to the context.
+In the following 3 steps we satisfy the judgment $1 \to 1 \le 1 \to 1$ by Rules 7, 4 and 4.
+
+Now the existential variable $\al$, being introduced before but remains unsolved,
+is on the top of the worklist. By Rule 2 it is effectively recycled.
+The process is carefully designed within the algorithmic rules:
+when $\al$ is introduced,
+we foresee the recycling of $\al$ after all the judgments requiring $\al$ processed.
+Finally Rule 4 reduces one of the base cases and finishes the subtyping derivation.
+
+\jimmy{An example on subtyping: $\nil\vdash (1\to 1)\to 1 \le (\forall a. 1 \to 1) \to 1$
 TODO derivation and explanation}
@@
 
-The design of our judgment chain is closely related to the property of
+The design of our judgment chain is closely related to the shape of
 the judgments

### 2019-02-27 sections/Metatheory.tex
Section 2
changes to 4.3, 4.4

@@
 
-Our algorithm is sound with respect to the declarative system.
+Our algorithm is sound with respect to DK's declarative system.
 For any worklist $\Gm$ that reduces successfully,
-we can find a valid instantiation $\Gm\sto\Om$ that solves all the existential variables,
-such that $\Om$ also transfers all judgements successfully.
+there is a valid instantiation $\Om$ that transfers all judgements successfully to the declarative system.
 \begin{theorem}[Soundness]
@@
 including Rules 10, 11, 21 and 29.
-Proper applications of lemmas \ref{lem:insert} and \ref{lem:extract}
+Applications of Lemmas \ref{lem:insert} and \ref{lem:extract}
 help analyse the full instantiation of those existential variables.
-For example, when $\al$ is solved to be $\al[1] \to \al[2]$ in the algorithm,
-applying Extract gives two instantiations $\al[1] = \sigma$ and $\al[2] = \tau$.
-Inserting back $\al = \sigma \to \tau$ for the induction case
-matches the shape of induction hypothesis and thus finishes the corresponding case.
-Some immediate corollaries, which show the soundness for specific judgement forms are:
+For example, when $\al$ is solved to $\al[1] \to \al[2]$ in the algorithm,
+applying the Extract lemma gives two instantiations $\al[1] = \sigma$ and $\al[2] = \tau$.
+It follows that $\al = \sigma \to \tau$, which enables the induction hypothesis
+and finishes the corresponding case. Some immediate corollaries which
+show the soundness for specific judgement forms are:
 
@@
 
+%-------------------------------------------------------------------------------
 \subsection{Completeness}
 
-Completeness of the algorithm means that any declarative system has an algorithmic counterpart.
+Completeness of our algorithm means that any derivation in the
+declarative system has an algorithmic counterpart.
 
@@
 
-We prove completeness by induction on the derivation of $\Om\redto\nil$.
-Here we refer to the non-overlapping declarative system.
+We prove completeness by induction on the derivation of $\Om\redto\nil$
+and use the non-overlapping declarative system.
 Since the declarative worklist is reduced judgment-by-judgment

### 2019-02-27 sections/Metatheory.tex
revision of 4.5

@@
-\section{Metatheory}
+\section{Metatheory}\label{sec:metatheory}
 
@@
 
-\subsection{Decidability}
+%-------------------------------------------------------------------------------
+\subsection{Termination and Decidability}
 
-\bruno{Tom, this is the section that I feel less helpful. Maybe
-you can look at this one more closely?}
-
-The decidability theorem is:
-
+Finally, we show that our algorithm terminates. In other words:
 \begin{theorem}[Decidability]
@@
 \end{theorem}
-
-We have proven this theorem by a lexicographic group of induction measures\\
+Our termination proof is based on a lexicographic group of induction measures\\
 $\langle |\Gm|_e, |\Gm|_\Leftrightarrow, |\Gm|_\forall, |\Gm|_{\al}, |\Gm|_\to + |\Gm| \rangle$
-on the worklist $\Gm$.
-$|\cdot|_e$ and $|\cdot|_\Leftrightarrow$ measure the total ``size'' of terms
-and the total ``difficulty'' of judgments, respectively.
-$|\cdot|_\forall$ and $|\cdot|_\to$ counts the total number of
+on the worklist $\Gm$. The definitions of these measures can be found in Figure~\ref{fig:measures}. The first two, 
+$|\cdot|_e$ and $|\cdot|_\Leftrightarrow$, measure the total size of terms
+and the total difficulty of judgments, respectively. In the latter, check judgements
+count for 2, inference judgments for 1 and function inference judgments for 3.
+The next two measures, $|\cdot|_\forall$ and $|\cdot|_\to$, count the total number of
 universal quantifiers and function types, respectively. Finally,
@@
 
-\begin{definition}[Worklist Measures]
+\begin{figure}[t]
+\hfill \framebox{$|\Gm|_e$} \hfill  \llap{Size measure.}
 \begin{gather*}
@@
 \text{where } |x|_e &= |()|_e = 1 \qquad |\lam e|_e = |e|_e + 1\\
-|e_1~e_2|_e &= |e_1|_e + |e_2|_e + 1 \qquad |e:A|_e = |e| + 1\\[2mm]
-%
+|e_1~e_2|_e &= |e_1|_e + |e_2|_e + 1 \qquad |e:A|_e = |e| + 1
+\end{aligned}
+\end{gather*}
+\hfill \framebox{$|\Gm|_\Leftrightarrow$} \hfill  \llap{Judgment measure.}
+\begin{gather*}
+\begin{aligned}
 |\Gm|_\Leftrightarrow &= 2\cdot\#_{e\Lto A\in\Gm} +
-    \#_{e\To_a \jg\in\Gm} + 3\cdot\#_{\appInfAlg{A}{e}\in\Gm}\\[2mm]
+    \#_{e\To_a \jg\in\Gm} + 3\cdot\#_{\appInfAlg{A}{e}\in\Gm}
+\end{aligned}
+\end{gather*}
+\hfill \framebox{$|\Gm|_\forall$} \hfill  \llap{Polymorphism measure.}
+\begin{gather*}
+\begin{aligned}
 |\Gm|_\forall &= \textstyle\sum_{e\Lto A\in\Gm}|A|_\forall + \sum_{\appInfAlg{A}{e}\in\Gm}|A|_\forall +
@@
 \text{where } |1|_\forall &= |a|_\forall = |\al|_\forall = 1\\
-|A\to B|_\forall &= |A|_\forall + |B|_\forall \qquad |\all A|_\forall = |A|_\forall + 1\\[2mm]
-%
+|A\to B|_\forall &= |A|_\forall + |B|_\forall \qquad |\all A|_\forall = |A|_\forall + 1
+\end{aligned}
+\end{gather*}
+\hfill \framebox{$|\Gm|_\to$} \hfill  \llap{Function measure.}
+\begin{gather*}
+\begin{aligned}
 |\Gm|_\to &= \textstyle\sum_{e\Lto A\in\Gm}|A|_\to + \sum_{\appInfAlg{A}{e}\in\Gm}|A|_\to +
@@
 \text{where } |1|_\to &= |a|_\to = |\al|_\to = 1\\
-|A\to B|_\to &= |A|_\to + |B|_\to + 1 \qquad |\all A|_\to = |A|_\to\\[2mm]
-%
+|A\to B|_\to &= |A|_\to + |B|_\to + 1 \qquad |\all A|_\to = |A|_\to
+\end{aligned}
+\end{gather*}
+\hfill \framebox{$|\Gm|_{\al}$} \hfill  \llap{Existential measure.}
+\begin{gather*}
+\begin{aligned}
 |\Gm|_{\al} &= \#_{\al\in\Gm}\\
@@
 \end{gather*}
-\end{definition}
-\bruno{The definitions look a bit intimidating. I do not have immediate suggestions,
-  but perhaps Tom has suggestions on how to improve the presentation here.}
+\caption{Worklist measures}
+\label{fig:measures}
+\end{figure}
 
 It is not difficult to see that all but two algorithmic reduction rules
-decrease the group of measures.
-(The result of Rule 29 could be directly reduced by Rule 28, which decreases the measures.)
-The two exceptions are Rules 10 and 11. Both rules increase the number of existential variables
-without decreasing the number of universal quantifiers.
-However, at the same time, they both break the subtyping problem into two sub-problems
-(Rule 7 applied right after Rule 10 or 11), in smaller sizes.
-Therefore we argue that such split makes the worklist easier, by analyzing a subset of the worklist,
-which we call \emph{instantiation judgments} $\Gm_i$ (defined with the help of $\Gm_\le$ for the proof).
+decrease the group of measures.  (The result of Rule 29 could be directly
+reduced by Rule 28, which decreases the measures.) The two exceptions are Rules
+10 and 11. Both rules increase the number of existential variables without
+decreasing the number of universal quantifiers. However, they are both
+immediately followed by Rule 7, which breaks the subtyping problem into two
+smaller problems of the form $\al \le A$ and $A \le \al$ which we call
+\emph{instantiation judgements}.
 
-\begin{definition}[$\Gm_\le, \Gm_i$]
+We now show that entirely reducing these smaller problems leaves the worklist
+in a state with an overall smaller measure. Our starting point is a 
+worklist of the form $\Gm,\Gm_i$ where $\Gm_i$ are instantiation judgments.
 \begin{gather*}
 \begin{aligned}
-\Gm_\le &:= \nil \mid \Gm_\le, a \mid \Gm_\le, \al \mid \Gm_\le, A\le B\\
 \Gm_i &:= \nil \mid \Gm_i, \al\le A \mid \Gm_i, A\le\al \quad
@@
 \end{gather*}
-\end{definition}
+Fully reducing these instantiation judgments at the top of the worklist
+has a twofold impact. Firstly, new entries may be pushed onto the worklist which
+are not instantiation judgments. This only happens when $\Gm_i$ contains a universal quantifier
+that is reduced by Rule 8 or 9. The new entries then are of the form $\Gm_\forall$:
+\begin{gather*}
+\begin{aligned}
+\Gm_\le &:= \nil \mid \Gm_\le, a \mid \Gm_\le, \al \mid \Gm_\le, A\le B
+\end{aligned}
+\end{gather*}
+Secondly, reducing the instantiation judgments may also affect the remainder of the worklist $\Gm$,
+by solving existing existentials and introducing new ones. This worklist update is captured in the 
+update judgement $\Gm \jExt \Gm'$ defined in Figure~\ref{fig:worklist_ext}.
+For instance, existential variable instantiation, 
+i.e. $\Gm_L,\al,\Gm_R \jExt \Gm_L, \al[1], \al[2], [\al[1] \to \al[2] / \al]\Gm_R$,
+can be derived as a combination of the three rules that define the update relation.
 
 \begin{figure}
-\framebox{$\Gm\jExt\Gm'$} $\Gm$ extends to $\Gm'$.
+\hfill \framebox{$\Gm\jExt\Gm'$} \hfill \llap{$\Gm$ updates to $\Gm'$}.
 \begin{gather*}
@@
 \end{gather*}
-\caption{Worklist Extension}\label{fig:worklist_ext}
+\caption{Worklist Update}\label{fig:worklist_ext}
 \end{figure}
 
-\paragraph{Intuition of instantiation decidability}
-These instantiation judgments are the ones consumed and produced by Rules 10 and 11.
-When there are universal quantifiers in the initial instantiation judgments $\Gm_i$,
-the worklist $\Gm,\Gm_i$ will be multi-reduced to match Rule 8 or 9,
-which decreases the number of universal quantifiers by 1, and returns a $\Gm',\Gm_\le$.
-Otherwise, no more variables will be introduced,
-and each instantiation judgment solves one existential variable,
-thus finally there will be $|\Gm_i|$ existential variables solved,
-with the worklist reduced to $\Gm'$.
-$\Gm'$ basically refers to the original part of worklist $\Gm$.
-During the reduction, minor changes to the worklist may occur, such as
-existential variable solving or insertion and instantiation.
-Nevertheless, the main structure of remains unchanged.
-
-Shown in Figure~\ref{fig:worklist_ext},
-worklist extension formally describes the minor changes to worklist during algorithmic reduction.
-When the algorithm processes the worklist, the first judgment is processed,
-and the rest of the worklist roughly preserves its shape.
-Worklist extension is formally defined by two cases:
-existential variable solving ($\mathtt{\jExt solve}$) and insertion ($\mathtt{\jExt \al}$).
-Existential variable instantiation, as another common form of worklist manipulation,
-i.e. $\Gm_L,\al,\Gm_R \jExt \Gm_L, \al[1], \al[2], [\al[1] \to \al[2] / \al]\Gm_R$,
-can be derived with the combination of the 3 rules.
-Within expectation, some worklist measures are not changed modulo worklist extension.
-
-\newcommand{\equivGm}[1]{|\Gm|_{#1} = |\Gm'|_{#1}}
+The good news is that worklist update does not affect the three main worklist measures:
+\newcommand{\equivGm}[1]{|\Gm|_{#1} = |\Gm'|_{#1}}\label{lemma:inst:invariant}
 \begin{lemma}[Measure Invariants of Worklist Extension]
@@
 
-With all the definitions, we are able to prove the following instantiation decidability lemma.
-
+Moreover, we can characterise the reduction of the instantiation judgments as follows.
 \begin{lemma}[Instantiation Decidability]\label{lemma:inst:decidable}
@@
 \end{lemma}
-
-In short, after reducing any instantiation judgment prefix $\Gm_i$,
+Hence, reducing the instantiation judgment prefix $\Gm_i$ either
+decreases the number of universal quantifiers or 
 the algorithm either decreases the number of $\forall$'s
 or solves one existential variable per instantiation judgment.
-The proof of this lemma is by induction on the measure $2*|\Gm_i|_\to + |\Gm_i|$
+The proof of this lemma proceeds by induction on the measure $2*|\Gm_i|_\to + |\Gm_i|$
 of the instantiation judgment list $\Gm_i$.
 
-In summary, the decidability theorem is shown through a lexicographic group of induction measures
-$\langle |\Gm|_e, |\Gm|_\Leftrightarrow, |\Gm|_\forall, |\Gm|_{\al}, |\Gm|_\to + |\Gm| \rangle$.
-The critical case is that, whenever we encounter an instantiation judgment at the top of the worklist,
-we make use of Lemma~\ref{lemma:inst:decidable}, which reduces $|\Gm|_{\al}$ or $|\Gm|_\forall$
-by consuming that instantiation judgment. Other cases are relatively straightforward.
+Let us go back to the whole algorithm and summarize our findings.
+The decidability theorem is shown through a lexicographic group of induction measures
+$\langle |\Gm|_e, |\Gm|_\Leftrightarrow, |\Gm|_\forall, |\Gm|_{\al}, |\Gm|_\to + |\Gm| \rangle$
+which is decreased by nearly all rules.
+In the exceptional case that the measure does not decrease immediately, 
+we encounter an instantiation judgment at the top of the worklist. We can then
+make use of Lemma~\ref{lemma:inst:decidable} to show that $|\Gm|_{\al}$ or $|\Gm|_\forall$ decreases
+when that instantiation judgment is consumed. Moreover, Lemma~\ref{lemma:inst:invariant} establishes
+that no higher-priority measure component increases. Hence, in the exceptional case we have an overall measure decrease too.
 
-Combining all the three main results, we know that the declarative system is decidable,
-with the help of our algorithm.
+Combining all three main results (soundness, completeness and termination), we conclude that the declarative system is decidable
+by means of our algorithm.
 \begin{corollary}[Decidability of Declarative Typing]

### 2019-02-27 sections/Metatheory.tex
section 4 clean-up

@@
 \end{corollary}
-\bruno{I do not think the corollary is precisely stated. I think you want todo
-  instantiation on $\Gm$ and then use erasure on the judgements, right? As it stands
-  it is wrong because there is a ``type error''. Worklists and declarative typing contexts are of different shapes
-so the equality does not make sense.}
 
@@
 and use the non-overlapping declarative system.
-Since the declarative worklist is reduced judgment-by-judgment
+Since the declarative worklist is reduced judgment by judgment
 (shown in Figure~\ref{fig:decl:worklist}),
 the induction always analyses the first judgment until it is satisfied.
-In Abella, we use a different relation to get the correct induction scheme.
 
@@
 \end{corollary}
-\bruno{Similar problem to the soundness corollaries here.}
 

### 2019-02-27 sections/Metatheory.tex
section 4 clean-up
S4: Minor typo fix

@@
 Finally, we show that our algorithm terminates. In other words:
-\begin{theorem}[Decidability]
-Given \emph{wf }$\Gm$, it is decidable whether $\Gm\redto\nil$ or not.
+\begin{theorem}[Termination]
+Given a well-formed input $\Gm$, the algorithm terminates.
 \end{theorem}
@@
 are not instantiation judgments. This only happens when $\Gm_i$ contains a universal quantifier
-that is reduced by Rule 8 or 9. The new entries then are of the form $\Gm_\forall$:
+that is reduced by Rule 8 or 9. The new entries then are of the form $\Gm_\le$:
 \begin{gather*}
@@
 \begin{lemma}[Measure Invariants of Worklist Extension]
-If $\Gm\sto\Gm'$ then $\equivGm{e}$ and $\equivGm\Leftrightarrow$ and $\equivGm\forall$.
+If $\Gm\jExt\Gm'$ then $\equivGm{e}$ and $\equivGm\Leftrightarrow$ and $\equivGm\forall$.
 \end{lemma}
@@
     \item When $|\Gm_i|_\forall = 0$, $(\Gm, \Gm_i) \redto \Gm'$,
-        where $|\Gm'|_{\al} = |\Gm|_{\al} - |\Gm_i|$ and $\Gm\sto\Gm'$.
+        where $|\Gm'|_{\al} = |\Gm|_{\al} - |\Gm_i|$ and $\Gm\jExt\Gm'$.
     \item When $|\Gm_i|_\forall > 0$, $(\Gm, \Gm_i) \redto (\Gm', \Gm_\le)$,
-        where $|\Gm_\le|_\forall = |\Gm_i|_\forall - 1$ and $\Gm\sto\Gm'$.
+        where $|\Gm_\le|_\forall = |\Gm_i|_\forall - 1$ and $\Gm\jExt\Gm'$.
 \end{enumerate}
@@
 Let us go back to the whole algorithm and summarize our findings.
-The decidability theorem is shown through a lexicographic group of induction measures
+The termination theorem is shown through a lexicographic group of induction measures
 $\langle |\Gm|_e, |\Gm|_\Leftrightarrow, |\Gm|_\forall, |\Gm|_{\al}, |\Gm|_\to + |\Gm| \rangle$

### 2019-02-27 sections/Metatheory.tex
fix broken sentence

@@
 decreases the number of universal quantifiers or 
-the algorithm either decreases the number of $\forall$'s
-or solves one existential variable per instantiation judgment.
+eliminates one existential variable per instantiation judgment.
 The proof of this lemma proceeds by induction on the measure $2*|\Gm_i|_\to + |\Gm_i|$

### 2019-02-28 sections/Appendix.tex
Appendix and all pointers

@@
 
-\jimmy{Lemma 29 is false. TODO structure and more text}
-The first conclusion of the lemma says that if $\Gm_L\longrightarrow \Delta_R$ and
-$\Gm_L,\Gm_R\longrightarrow \Delta_L,\Delta_R$ then
-$\Gm_L,\al,\Gm_R\longrightarrow \Delta_L,\al,\Delta_R$.
-If we pick $\Gm_L = \nil, \Gm_R = \bt, \Delta_L = \bt, \Delta_R = \nil$, the conditions
-$\nil\longrightarrow \bt$ and $\bt\longrightarrow \bt$ hold, but the conclusion
+\subsection{False Lemmas in DK's papers}
+\label{appendix:false_lemmas}
+
+We found false lemmas in the manual proofs of DK' two papers
+\cite{dunfield2013complete} and \cite{DunfieldIndexed}.
+\begin{itemize}
+    \item
+        In the first paper, Lemma 29 on page 9 of its appendix says:
+        \begin{lemma}[Parallel Admissibility of~\cite{dunfield2013complete}]~\\
+        If $\Gamma_L \longrightarrow \Delta_L$ and
+        $\Gamma_L, \Gamma_R \longrightarrow \Delta_L, \Delta_R$ then:
+        \begin{enumerate}
+            \item $\Gamma_L,\al,\Gamma_R \longrightarrow \Delta_L, \al, \Delta_R$
+            \item If $\Delta_L \vdash \tau'$ then
+                $\Gamma_L,\al,\Gamma_R \longrightarrow \Delta_L, \al = \tau', \Delta_R$.
+            \item If $\Gamma_L \vdash \tau$ and $\Delta_L \vdash \tau'$ and
+                $[\Delta_L]\tau = [\Delta_L]\tau'$, then
+                $\Gamma_L, \al = \tau, \Gamma_R \longrightarrow \Delta_L, \al = \tau', \Delta_R$.
+        \end{enumerate}
+        \end{lemma}
+
+        We give a counter-example to this lemma:\\
+        Pick $\Gm_L = \nil, \Gm_R = \bt, \Delta_L = \bt, \Delta_R = \nil$, then both conditions
+        $\nil\longrightarrow \bt$ and $\bt\longrightarrow \bt$ hold, but the first conclusion
         $\al,\bt \longrightarrow \bt,\al$ does not hold.
+    \item
+        In the second paper, as an extended work to the first paper, Lemma 26 on page 22 of its appendix says:
+        \begin{lemma}[Parallel Admissibility of~\cite{DunfieldIndexed}]~\\
+        If $\Gamma_L \longrightarrow \Delta_L$ and
+        $\Gamma_L, \Gamma_R \longrightarrow \Delta_L, \Delta_R$ then:
+        \begin{enumerate}
+            \item $\Gamma_L,\al:\kappa,\Gamma_R \longrightarrow \Delta_L, \al:\kappa, \Delta_R$
+            \item If $\Delta_L \vdash \tau' : \kappa$ then
+                $\Gamma_L,\al:\kappa,\Gamma_R \longrightarrow \Delta_L, \al:\kappa = \tau', \Delta_R$.
+            \item If $\Gamma_L \vdash \tau : \kappa$ and $\Delta_L \vdash \tau' types$ and
+                $[\Delta_L]\tau = [\Delta_L]\tau'$, then
+                $\Gamma_L, \al:\kappa = \tau, \Gamma_R \longrightarrow \Delta_L, \al:\kappa = \tau', \Delta_R$.
+        \end{enumerate}
+        \end{lemma}
+
+        A similar counter-example is given:\\
+        Pick $\Gm_L = \nil, \Gm_R = \bt:\star, \Delta_L = \bt:\star, \Delta_R = \nil$, then both conditions
+        $\nil\longrightarrow \bt:\star$ and $\bt:\star\longrightarrow \bt:\star$ hold, but the first conclusion
+        $\al:\kappa,\bt:\star \longrightarrow \bt:\star,\al:\kappa$ does not hold.
+\end{itemize}
+
+\subsection{A Fix to DK's Subsumption Lemma}
+\label{appendix:subsumption}
+
+Let's first recall the lemma:
+\begin{lemma}[Subsumption]
+Given $\Psi' \le \Psi$:
+\begin{enumerate}
+    \item If $\Psi \vdash e \Lto A$ and $\Psi \vdash A \le A'$ then $\Psi' \vdash e \Lto A$;
+    \item If $\Psi \vdash e \To B$ then there exists
+        $A'$ s.t. $\Psi \vdash B' \le B$ and $\Psi' \vdash e \To B$;
+    \item If $\Psi \vdash \appInf{C}{e}{A}$ and $\Psi \vdash C' \le C$,
+        then there exists $A'$ s.t. $\Psi \vdash A' \le A$ and $\Psi' \vdash \appInf{C'}{e}{A'}$.
+\end{enumerate}
+\end{lemma}
+
+\paragraph{Problem in DK's Manual Proof}
+We tried to follow DK's manual proof for this lemma.
+While being written in a clear format and providing enough details,
+the proof is not fully accepted by the proof assistant.
+Specifically, the first two applications of induction hypotheses on page 22 are not accepted.
+Either of them seems to use a slightly different ``induction hypothesis'' than the true one.
+
+\paragraph{A Fix to the Proof}
+We make use of our worklist measures for the induction.
+Recall that $|e|_e$ measures term size;
+and the judgment measure counts checking as 2, inference as 1 and application inference as 3;
+and $|A|_\forall$ counts the number of $\forall$'s in a type.
+
+The proof is by a nested mutual induction on the lexicographical order of the measures
+$$\langle |e|_e, |\cdot|_\Leftrightarrow, |A \to B|_\forall \rangle,$$
+where the second measure is 2 for checking (1), 1 for inference (2) and 3 for application inference (3);
+and the third measure does not apply to case (2) since no $A$ is used.
+
+All but two cases can be finished easily following the declarative typing derivation,
+and the proof shares a similar structure to DK's.
+One tricky case related to Rule $\mathtt{Decl\forall I}$ inducates that $A$ has the shape $\all A_0$,
+thus the subtyping relation derives from either $\mathtt{{\le}\forall L}$ or $\mathtt{{\le}\forall R}$.
+For each of the case, the third measure $|A \to B|_\forall$ decreases
+(the $\mathtt{{\le}\forall L}$ case requires a type substitution lemma obtaining
+$\Psi \vdash e \Lto [\tau/a]A_0$ from the typing derivation).
+
+Another tricky case is $\mathtt{Decl{\to}I}$.
+When the subtyping relation is derived from $\mathtt{{\le}{\to}}$,
+a simple application of induction hypothesis finishes the proof.
+When the subtyping relation is derived from $\mathtt{{\le}\forall R}$,
+$|B|_\forall$ decreases, and thus the induction hypothesis finishes this case.

### 2019-02-28 sections/Related.tex
Related Work
Appendix and all pointers

@@
 related work.
-
-\bruno{Jimmy, please fill in missing references.}
 
@@
 DK's algorithm is shown to be sound, complete and decidable in 70 pages of manual proofs.
-Though carefully written, some of the proofs are incorrect (see discussion in Appendix~\ref{}),
+Though carefully written, some of the proofs are incorrect
+(see discussion in Appendix~\ref{appendix:false_lemmas} and \ref{appendix:subsumption}),
 which creates difficulties when formalizing them in a proof assistant.
@@
 as lemma 29 in \citet{dunfield2013complete}: the conclusion is false. We also discuss
-the issue in more detail in Appendix~\ref{}. 
+the issue in more detail in Appendix~\ref{appendix:false_lemmas}. 
 \bruno{I think you will need to justify this better and in the appendix you must show

### 2019-02-28 sections/Discussion.tex
S3 inference explaination; S5.2 fix DK scoping

@@
 
+One may blame that the inference rule for lambda functions does not collect garbages correctly.
+Here we also show another rule for application inference with a similar problem:
+$$
+\inferrule*[right=$\mathtt{DK\_{\to}I{\To}}$]
+{\Gamma, \al, \bt, x:\al \vdash e \Lto \bt \dashv \Delta, x:\al, \Theta}
+{\Gamma \vdash \lam e \To \al \to \bt \dashv \Delta}
+\qquad
+\inferrule*[right=$\mathtt{DK\_\forall App}$]
+{\Gamma, \al \vdash [\al/a] \appInf{A}{e}{C} \dashv \Delta}
+{\Gamma \vdash \appInf{\all A}{e}{C} \dashv \Delta}
+$$
+But our algorithm collects the existential variables right after the next judgment chain:
+$$\Gm \Vdash \lambda x. e \To_a \jg \rrule{25}
+\Gm,\al,\bt \Vdash [\al\to\bt/a]\jg, x:\al \Vdash e\Lto \bt$$
+$$\Gm \Vdash \appInfAlg{\forall a. A}{e} \rrule{27} \Gm,\al \Vdash \appInfAlg{[\al/a]A}{e}$$
+But it seems impossible to propose a fix by only modifying this rule.
+Taking $\mathtt{DK\_{\to}I{\To}}$ as an example,
+the declaration or solution for $\al$ and $\bt$ may be referred by following judgments,
+therefore leaving $\al$ and $\bt$ in the output context is the only choice,
+when the next judgments are inaccessible to us at the moment.
+
+To fix the problem, we suggest a modification on the algorithmic subsumption rule,
+as garbage collection for a checking judgment is always safe:
+$$
+\inferrule*[right=$\mathtt{DK\_Sub}$]
+{\Gamma, \blacktriangleright_{\al} \vdash e \To A \dashv \Theta \\
+\Theta \vdash [\Theta]A \le [\Theta]B \dashv \Delta, \blacktriangleright_{\al}, \Delta'}
+{\Gamma \vdash e \Lto B \dashv \Delta}
+$$
+We pick this rule because this is the only one where a checking judgment calls an inference judgment.
+And that's the point where our algorithm recycles variables---right after a judgment chain is satisfied.
+After applying this patch, to our best knowledge,
+DK's algorithm behaves equivalently to our algorithm in terms of variable scoping.
+
 \subsection{Scoped type variables}

### 2019-02-28 sections/algorithm.tex
S3 inference explaination; S5.2 fix DK scoping

@@
 \Gm \Vdash A_1\to A_2 \le B_1\to B_2 &\algrule \Gm \Vdash A_2 \le B_2 \Vdash B_1\le A_1\\
-\Gm \Vdash \forall a. A\le B &\algrule \Gm,\al \Vdash [\al/a]A\le B \quad\text{when } B \neq \all B' \quad\text{\jimmy{No more exvar leaks}}\\
-\Gm \Vdash A\le \forall b. B &\algrule \Gm,b \Vdash A\le B \quad\text{\jimmy{No more typevar leaks}}
+\Gm \Vdash \forall a. A\le B &\algrule \Gm,\al \Vdash [\al/a]A\le B \quad\text{when } B \neq \all B'\\
+\Gm \Vdash A\le \forall b. B &\algrule \Gm,b \Vdash A\le B
 \\[2mm]
@@
 \Gm \Vdash \lambda x. e \Lto A\to B &\algrule \Gm, x:A  \Vdash e \Lto B\\
-\Gm[\al] \Vdash \lambda x. e \Lto \al &\algrule [\al[1]\to \al[2] / \al](\Gm[\al[1],\al[2]], x:\al[1] \Vdash e \Lto \al[2]) \quad\text{\jimmy{Additional}}
+\Gm[\al] \Vdash \lambda x. e \Lto \al &\algrule [\al[1]\to \al[2] / \al](\Gm[\al[1],\al[2]], x:\al[1] \Vdash e \Lto \al[2])
+% \quad\text{\jimmy{Additional}}
 \\[2mm]
@@
 \Gm \Vdash (e:A)\To_a \jg &\algrule \Gm \Vdash [A/a]\jg \Vdash e \Lto A\\
-\Gm \Vdash ()\To_a \jg &\algrule \Gm \Vdash [1/a]\jg \qquad\jimmy{\downarrow\text{Precise control on exvars}}\\
+\Gm \Vdash ()\To_a \jg &\algrule \Gm \Vdash [1/a]\jg\\
 \Gm \Vdash \lambda x. e \To_a \jg &\algrule
-    \Gm,\al,\bt \Vdash [\al\to\bt/a]\jg, x:\al \Vdash e\Lto \bt \quad\jimmy{\text{differs}}\\
+    \Gm,\al,\bt \Vdash [\al\to\bt/a]\jg, x:\al \Vdash e\Lto \bt\\
 \Gm \Vdash e_1\ e_2 \To_a \jg &\algrule \Gm \Vdash e_1\To_b (\appInfAlg{b}{e_2})
@@
 
-The design of our judgment chain is closely related to the shape of
-the judgments
-\jimmy{requires further clarification}
-\bruno{What property? Not very clear}.
-Subtyping and checking do not return anything, so variables cannot
-leak anyway, as applied to Rules 8 and 9.
-\bruno{Is the discussion that follows in the right place? We just jump
-  to
-  rule 26. Perhaps we can wait until we talk about inference to
-  discuss those issues?}
-However, inference and application inference may return a type that contains new variables.
-Take Rule 26 as an example, if it simply returns $\al \to \bt$,
-passes that to the next judgment and continues the type-checking process,
-variables $\al$ and $\bt$ in the next judgment are out of scope
-and will break the well-formedness of the worklist.
-By making use of the continuation-passing-style judgement chain,
-inner judgments always shares context with its parent judgment,
-therefore we preserve well-formedness during reduction.
-
 \item[2. Existential decomposition]
@@
 \item[5. Inference judgments]
-Rules 22-26 deal with inference judgments.
+Inference judgments behave differently compared with subtyping and checking judgments:
+they \emph{return} a type instead of only accepting or rejecting.
+For algorithmic system, where guessings are involved,
+it may happen that the output type of an inference judgment refers to new existential variables,
+such as Rule 25.
+In comparison to Rule 8 and 9, where new variables are only refered by the sub-derivation,
+Rule 25 introduces variables $\al, \bt$ that affect the remaining judgment chain.
+This rule is carefully designed so that the output variables are bounded by declarations before,
+thus the well-formedness of the worklist is preserved,
+and the garbage will be collected at the correct time.
+By making use of the continuation-passing-style judgement chain,
+inner judgments always shares context with its parent judgment,
+therefore we achieve all the desired property in such a neat way.
+
+\begin{comment}
+Old text:
+The design of our judgment chain is closely related to the shape of
+the judgments
+\jimmy{requires further clarification}
+\bruno{What property? Not very clear}.
+Subtyping and checking do not return anything, so variables cannot
+leak anyway, as applied to Rules 8 and 9.
+\bruno{Is the discussion that follows in the right place? We just jump
+  to
+  rule 26. Perhaps we can wait until we talk about inference to
+  discuss those issues?}
+However, inference and application inference may return a type that contains new variables.
+Take Rule 26 as an example, if it simply returns $\al \to \bt$,
+passes that to the next judgment and continues the type-checking process,
+variables $\al$ and $\bt$ in the next judgment are out of scope
+and will break the well-formedness of the worklist.
+\end{comment}
+
+Rules 22-26 deal with type inference judgments.
 Inference judgment chains are written in continuation-passing-style.
@@
 therefore another judgment chain is created.
+
 Rule 25 infers the type of a lambda expression by introducing $\al, \bt$
@@
 After checking the body $e$ under the assumption $x:\al$,
-the return type might reflect more information than simply $\al \to \bt$.
+the return type might reflect more information than simply $\al \to \bt$
+through propagation when existential varaibles are solved or partially solved.
 The variable scopes are maintained during the process:
@@
 the checking and the remaining chain $[\al\to\bt/a]\jg$.
+The recycling process of Rule 25 differs from DK's corresponding rule significantly,
+and we further discuss the differences in Section~\ref{sec:discussion:scoping}.
+
 \bruno{You have some note ``differs'' on the rules. Should that

### 2019-02-28 sections/algorithm.tex
Section 3

@@
 This section introduces a novel algorithmic system that implements 
-DK's declarative specification. The algorithmic system comes with
-several novelties.\bruno{fill in/summarize novelties. I think
-  novelties include: precise scoping control; CPS-style; dealing with
-  inference judgements.}
+DK's declarative specification. The new algorithm extends the idea
+of worklists proposed by \citet{itp2018} with two new features. Firstly,
+unlike \citet{itp2018}'s worklists, the scope of variables is precisely tracked
+and variables do not escape their scope. This is achieved by unifying algorithmic contexts and the worklists themselves.
+Secondly, our algorithm also
+accounts for the type system (and not just subtyping). To deal with
+inference judgments that arize in the type system we employ a \emph{continuation
+passing style} to enable the transfer of inferred information across
+judgements in a worklist.
 
@@
 
-Figure~\ref{fig:alg:syntax} shows the syntax---with accompanying well-formedness judgement---used by the algorithm.
+Figure~\ref{fig:alg:syntax} shows the syntax and well-formedness judgements used by the algorithm.
 
@@
 \text{Existential variables}\qquad&\al, \bt
-\\[2mm]
+\\
 %\text{Types}\qquad&A', B', C' &::=&\quad 1 \mid a \mid \forall x. A' \mid A'\to B'\\
@@
 % \noalign{\jimmy{\text{What do the markers do: $\G \vdash,a \vdash,b\vdash \j_1;\j_2;\j_3$ means $\G,a,b\vdash \j_1 \land \G,a\vdash \j_1,\j_2 \land \G\vdash \j_3$}}} \\[2mm]
-\text{Algorithmic types}\qquad&A, B, C &::=& \quad \ldots \mid \al % \quad 1 \mid a \mid \al \mid \forall a. A \mid A\to B
-\\[2mm]
+\text{Algorithmic types}\qquad&A, B, C &::=& \quad 1 \mid a \mid \forall a. A \mid A\to B \mid \al % \quad 1 \mid a \mid \al \mid \forall a. A \mid A\to B
+\\
 % \text{Terms}\qquad&e&::=&\quad x \mid () \mid \lam{e} \mid e_1~e_2 \mid (e:A)
@@
 \text{Algorithmic judgement chain}\qquad&\jg &::=&\quad A \le B \mid e\Lto A \mid e\To_{a} \jg \mid \appInfAlg{A}{e}
-\\[2mm]
+\\
 \text{Algorithmic worklist}\qquad&\Gm &::=&\quad \nil \mid \Gm, a \mid \Gm, \al \mid \Gm, x: A \mid \Gm \Vdash \jg%\\
@@
 but their scope is restricted by their position in the
-worklist\bruno{be careful to ensure that the reader knows about
-  worklist at this point},
-rather than being global.
+algorithmic worklist rather than being global.
 Any existential variable $\al$ should only be solved to
 a type that is well-formed with respect to the worklist to which $\al$ has been added.
-The point is that the mono-type $\tau$ represented by the corresponding existential variable
+The point is that the monotype $\tau$ represented by the corresponding existential variable
 is always well-formed under the corresponding declarative context.
@@
 type-checking. There are four kinds
-of these judgments in our system: subtyping ($A \le B$), checking ($e\Lto
+of judgments in our system: subtyping ($A \le B$), checking ($e\Lto
 A$), inference ($e\To_{a} \jg$) and
@@
 since their result is only success or failure. However both inference and
-application inference return a type that is used in further computation. We use a
+application inference return a type that is used in subsequent judgements. We use a
 continuation-passing-style encoding to accomplish this. For example, the judgment
@@
 \paragraph{Worklist Judgments} Our algorithm has a non-standard form.
-We combine traditional contexts and judgment(s) into a single object, the \emph{worklist} $\Gm$.
+We combine traditional contexts and judgment(s) into a single sort, the \emph{worklist} $\Gm$.
 The worklist is an \emph{ordered} collection of both variable bindings and judgments. The order captures the scope:
 only the objects that come after a variable's binding in the worklist can refer to it.
-For example, $[a,x:a,x \Lto a]$ is a valid worklist, but $[\textcolor{red}{x} \Lto a, x:\textcolor{red}{a},a]$ is not (the red symbols refer to out-of-scope variables).
+For example, $[a,x:a,x \Lto a]$ is a valid worklist, but $[{\bf x} \Lto a, x:{\bf a},a]$ is not
+(the symbols in bold refer to out-of-scope variables).
 
@@
 
-\begin{description}
-\item[1. Declarative Subtyping]
+\paragraph{{\bf 1. Declarative Subtyping}}
 We have six subtyping rules (Rules 4-9) that are similar to their
@@
 
-\item[2. Existential decomposition]
+\paragraph{\bf 2. Existential decomposition.}
 Rules 10 and 11 are algorithmic versions of Rule $\mathtt{{\le}{\to}}$; they
@@
 
-\item[3. Solving existentials] Rules 12-17 are base cases where an existential variable is solved.
+\paragraph{\bf 3. Solving existentials} Rules 12-17 are base cases where an existential variable is solved.
 They all remove an existential variable and substitute the
@@
 
-\item[4. Checking judgements]
+\paragraph{\bf 4. Checking judgements.}
 Rules 18-21 deal with checking judgments.
@@
 
-\item[5. Inference judgments]
+\paragraph{\bf 5. Inference judgments.}
 Inference judgments behave differently compared with subtyping and checking judgments:
@@
 
-\item[6. Application inferences judgements]
+\paragraph{\bf 6. Application inferences judgements}
 Finally, Rules 27-29 deal with application inference judgments.
@@
 Rule 29 instantiates $\al$ to the function type $\al[1] \to \al[2]$, just like Rules 10, 11 and 21.
-
-\end{description}
 

### 2019-02-28 sections/algorithm.tex
A few small improvements

@@
 DK's declarative specification. The new algorithm extends the idea
-of worklists proposed by \citet{itp2018} with two new features. Firstly,
+of worklists proposed by \citet{itp2018} in two ways. Firstly,
 unlike \citet{itp2018}'s worklists, the scope of variables is precisely tracked
@@
 The algorithmic system inherits the syntax of terms and types from 
-the declartive system. It only introduces one additional feature.
+the declarative system. It only introduces one additional feature.
 In order to find unknown types $\tau$ in the declarative system, the
@@
 a type that is well-formed with respect to the worklist to which $\al$ has been added.
-The point is that the monotype $\tau$ represented by the corresponding existential variable
+The point is that the monotype $\tau$, represented by the corresponding existential variable,
 is always well-formed under the corresponding declarative context.
@@
 \paragraph{Judgment Chains} 
-Judgment chains, or judgments for short, $\jg$ are the core components of our algorithmic
+Judgment chains $\jg$, or judgments for short, are the core components of our algorithmic
 type-checking. There are four kinds

### 2019-02-28 sections/algorithm.tex
S3 algorithm. 4 Checking judgments: discussion on Rule 21

@@
 continuation-passing-style.
+The side conditions $e \neq \lam e'$ and $B \neq \all B'$ is to
+prevent overlapping with Rules 19, 20 and 21,
+which are further discussed in the last part of this section.
 \bruno{The side condition needs to be explained, and some mention of
@@
 Rules 19 and 20 adapt their declarative counterparts to the worklist style.
-Rule 21 is a special case of $\mathtt{Decl\to I}$.\bruno{Rule 21
+Rule 21 is a special case of $\mathtt{Decl\to I}$,
+dealing with the case when the input type is an existential variable,
+representing a monotype \emph{function} as in the declarative system
+(it must be a function type, since the expression is a function).
+\bruno{Rule 21
   deserves some more explanation. What is the motivation for this
@@
 because Rules 18 and 24 can be combined to give the same result.
+
+\paragraph{Design Variant for Rule 21}
+A combination of Rules $18'$ (Rule 18 without the restriction on $e$) and 25 and further derivations
+may result in a similar effect to Rule 21.
+Rule 21 guesses a monotype for the lambda function $\lam e$, so as Rule 25.
+The only difference is that a subtyping relation is introduced by Rule $18'$,
+and is to be processed after the type inference:
+$$\Gm \Vdash \lam e \Lto \al \rrule{18'} \Gm \Vdash \lam e \To_a a \le \al
+\rrule{25} \Gm, \bt[1], \bt[2] \Vdash \bt[1] \to \bt[2] \le \al, x:\bt[1] \Vdash e \Lto \bt[2]$$
+Since two monotypes with a valid subtyping relation indicates equality,
+eventually $\bt[1] \to \bt[2]$ and $\al$ should be unified with each other.
+DK's algorithm takes a similar approach to such variant,
+and we believe that they are equivalent.
+
+Choosing Rule 21 simplifies the side condition on Rule 18,
+the metatheory and the design: whenever an input type could be a function type,
+the algorithmic rule should always take care of
+the possibility when the type being a single existential variable.
+This pattern applies to all the possible cases, including Rules 10, 11, 21 and 29.
 
@@
 the return type might reflect more information than simply $\al \to \bt$
-through propagation when existential varaibles are solved or partially solved.
+through propagation when existential variables are solved or partially solved.
 The variable scopes are maintained during the process:

### 2019-02-28 sections/Discussion.tex
S5.3 Elaboration

@@
 
-\subsection{Scoped type variables}
+% \subsection{Scoped type variables} Not discussed for this paper
 
+%-------------------------------------------------------------------------------
 \subsection{Elaboration}
+While our algorithmic reduction does not elaborate to System F,
+it is not difficult to extend the algorithm with a type-directed translation.
+Since the judgment form of our algorithmic worklist contains a collection of judgments,
+elaboration expressions are also generated as a list of equal length to
+the number of judgments (NOT judgment chains!) in the worklist.
+Within expectation, subtyping judgments translate to type coercions (denoted by $f$),
+all three other types of judgments translate to terms in System F (denoted by $t$).
 
-OTHERS?
+Let $\Phi$ be the elaboration list, containing translated type coercions and terms:
+$$\Phi ::= \nil \mid \Phi, f \mid \Phi, t$$
+
+And the form of our algorithmic judgment becomes:
+$$\Gamma \hookrightarrow \Phi$$
+
+We take Rule 18 as an example, rewritting small-step reduction into a relational style,
+$$
+\inferrule*[right=$\mathtt{Translation\_18}$]
+{\Gamma \Vdash e \To_a a \le B \hookrightarrow \Phi, f, t}
+{\Gamma \Vdash e \Lto B \hookrightarrow \Phi, f t}
+$$
+As is shown in the conclusion of the rule,
+a checking judgment on the top of the worklist corresponds to a top element for elaboration.
+The premise indicates that one judgment chain may relate to more than one elaboration elements,
+and that the outer judgment, being processed before inner ones,
+elaborates to the top element in the elaboration list.
+
+Existential variables need special treatment, since they may be solved at any step,
+or being recycled if not solved in its scope.
+If an existential variable is solved, we not only propagate the solution to the rest judgments,
+but also replace ocurrances in the elaboration list.
+If an existential variable is recycled, meaning that it is not constraint,
+we can pick any well-formed monotype as its solution.
+The unit type $1$, as the simplest type in the system, is a good choice.

### 2019-02-28 sections/algorithm.tex
further revision of S3

--- 2019-02-28 sections/algorithm.tex
@@
 accounts for the type system (and not just subtyping). To deal with
-inference judgments that arize in the type system we employ a \emph{continuation
+inference judgments that arise in the type system we employ a \emph{continuation
 passing style} to enable the transfer of inferred information across
@@
 only the objects that come after a variable's binding in the worklist can refer to it.
-For example, $[a,x:a,x \Lto a]$ is a valid worklist, but $[{\bf x} \Lto a, x:{\bf a},a]$ is not
-(the symbols in bold refer to out-of-scope variables).
+For example, $[a,x:a,x \Lto a]$ is a valid worklist, but $[\underline{x} \Lto a, x:\underline{a},a]$ is not
+(the underlined symbols refer to out-of-scope variables).
 
@@
 
-Let us explain the rulres in more detail:
+Let us explain the rules in more detail:
 Rules 1-3 pop a variable declaration; this garbage collects variables that are
@@
 declarative Rule ${\le}{\forall}L$ by introducing an existential
-variable $\al$ instead of guessing the monotype $\tau$ for
+variable $\al$ instead of guessing the monotype $\tau$
 instantiation. Each existential variable is later solved to a
 monotype $\tau$ with the same context, so the final solution $\tau$ of
-$\al$ should be well-formed under context $\Gm$.
+$\al$ should be well-formed under $\Gm$.
 
@@
 Rules 8 and 9 involve variable declarations and demonstrate how our
-worklist treats variable scopings. Rule 8 introduces an existential
+worklist treats variable scopes. Rule 8 introduces an existential
 variable $\al$ that is only visible to the judgment $[\al/a]A \le B$.
@@
 \jimmy{New example on Feb 27}
-\paragraph{Example} An example derivation on the subtyping judgment
-$(1\to 1)\to 1 \le (\all 1\to 1) \to 1$ is shown as follows:
+\paragraph{Example} The derivation of the subtyping judgment
+$(1\to 1)\to 1 \le (\all 1\to 1) \to 1$ proceeds as follows:
 \begin{gather*}
@@
 \rrule{8}  & \nil \Vdash 1 \le 1 ,\al \Vdash 1 \to 1 \le 1 \to 1\\
-\rrule{7}  & \nil \Vdash 1 \le 1 ,\al \Vdash 1 \to 1 \Vdash 1 \le 1\\
-\rrule{4} & \nil \Vdash 1 \le 1 ,\al \Vdash 1 \to 1\\
+\rrule{7}  & \nil \Vdash 1 \le 1 ,\al \Vdash 1 \le 1 \Vdash 1 \le 1\\
+\rrule{4} & \nil \Vdash 1 \le 1 ,\al \Vdash 1 \le 1\\
 \rrule{4} & \nil \Vdash 1 \le 1 ,\al\\
@@
 \end{gather*}
-Firstly, the subtyping of two function types is split into two judgments by Rule 7:
-a convariant subtyping on the return type and a contravariant subtyping on the argument type.
+First, the subtyping of two function types is split into two judgments by Rule 7:
+a covariant subtyping on the return type and a contravariant subtyping on the argument type.
 Then we apply Rule 8 to reduce the $\forall$ quantifier on the left side.
-The type $\forall a. 1 \to 1$ is actually not a general type, which
-has an almost equivalent behavior to the type $1 \to 1$.
-But according to the rule, we still introduce an existential variable $\al$ to the context.
+The rule introduces an existential variable $\al$ to the context (even though 
+the type $\forall a. 1 \to 1$ does not actually refer to the quantified type
+variable $a$).
 In the following 3 steps we satisfy the judgment $1 \to 1 \le 1 \to 1$ by Rules 7, 4 and 4.
 
-Now the existential variable $\al$, being introduced before but remains unsolved,
-is on the top of the worklist. By Rule 2 it is effectively recycled.
+Now the existential variable $\al$, introduced before but still unsolved,
+is at the top of the worklist and Rule 2 garbage collects it.
 The process is carefully designed within the algorithmic rules:
 when $\al$ is introduced,
-we foresee the recycling of $\al$ after all the judgments requiring $\al$ processed.
+we foresee the recycling of $\al$ after all the judgments requiring $\al$ have been processed.
 Finally Rule 4 reduces one of the base cases and finishes the subtyping derivation.
-
-\jimmy{An example on subtyping: $\nil\vdash (1\to 1)\to 1 \le (\forall a. 1 \to 1) \to 1$
-TODO derivation and explanation}
-\bruno{Perhaps this is better illustrated with a simple example?}
-\tom{Yes, it would be good to show one or more
-example derivations to illustrate the key points.}
 
@@
 variable with its solution in the remaining worklist. Importantly the rules
-respect the scope of existential variables. For example, rule 12 
+respect the scope of existential variables. For example, Rule 12 
 states that an existential variable $\al$ can be solved with another
@@
 continuation-passing-style.
-\bruno{The side condition needs to be explained, and some mention of
-  the overlapping is probably needed here.}
+The side conditions $e \neq \lam e'$ and $B \neq \all B'$ 
+prevent overlap with Rules 19, 20 and 21;
+this is further discussed at the end of this section.
+% \bruno{The side condition needs to be explained, and some mention of
+%   the overlapping is probably needed here.}
 Rules 19 and 20 adapt their declarative counterparts to the worklist style.
-Rule 21 is a special case of $\mathtt{Decl\to I}$.\bruno{Rule 21
-  deserves some more explanation. What is the motivation for this
-  rule? You may want to give some concrete example to motivate it.}
+Rule 21 is a special case of $\mathtt{Decl\to I}$,
+dealing with the case when the input type is an existential variable,
+representing a monotype \emph{function} as in the declarative system
+(it must be a function type, since the expression is a function).
+% \bruno{Rule 21
+%   deserves some more explanation. What is the motivation for this
+%   rule? You may want to give some concrete example to motivate it.}
 The same instantiation technique as in Rules 10 and 11 applies.
@@
 because Rules 18 and 24 can be combined to give the same result.
+
+\paragraph{Rule 21 Design Choice}
+The addition of Rule 21 is a design choice we have made to simplify the side
+condition of Rule 18 (which avoids overlap). It also streamlines the algorithm
+and the meatatheory as we now treat all cases
+where we can see that an existential variable should be instantiated to a
+function type  (i.e., Rules 10, 11, 21 and 29) uniformly.
+
+The alternative would have been to omit Rule 21 and drop the condition on $e$
+in Rule 18. The modified Rule 18 would then handle $\Gm \Vdash \lam e \Lto \al$
+and yield $\Gm \Vdash \lam e \To_a a \le \al$, which would be further processed
+by Rule 25 to $\Gm, \bt[1], \bt[2] \Vdash \bt[1] \to \bt[2] \le \al, x:\bt[1] \Vdash e \Lto \bt[2]$.
+As a subtyping constraint between two monotypes only succeeds when the types are equal, $\bt[1] \to \bt[2] \le \al$
+must end up equating $\bt[1] \to \bt[2]$ with $\al$ and thus have the same effect as Rule 21, but in a more roundabout
+fashion. DK's original algorithm also avoids this detour and contains a rule similar to our Rule 21.
+% A combination of Rules $18'$ (Rule 18 without the restriction on $e$) and 25 and further derivations
+% may result in a similar effect to Rule 21.
+% Rule 21 guesses a monotype for the lambda function $\lam e$, so as Rule 25.
+% The only difference is that a subtyping relation is introduced by Rule $18'$,
+% and is to be processed after the type inference:
+% $$\Gm \Vdash \lam e \Lto \al \rrule{18'} \Gm \Vdash \lam e \To_a a \le \al
+% \rrule{25} \Gm, \bt[1], \bt[2] \Vdash \bt[1] \to \bt[2] \le \al, x:\bt[1] \Vdash e \Lto \bt[2]$$
+% Since two monotypes with a valid subtyping relation indicates equality,
+% eventually $\bt[1] \to \bt[2]$ and $\al$ should be unified with each other.
+% DK's algorithm takes a similar approach to such variant,
+% and we believe that they are equivalent.
+% 
+% Choosing Rule 21 simplifies the side condition on Rule 18,
+% the metatheory and the design: whenever an input type could be a function type,
+% the algorithmic rule should always take care of
+% the possibility when the type being a single existential variable.
+% This pattern applies to all the possible cases, including Rules 10, 11, 21 and 29.
 
@@
 Rule 25 introduces variables $\al, \bt$ that affect the remaining judgment chain.
-This rule is carefully designed so that the output variables are bounded by declarations before,
+This rule is carefully designed so that the output variables are bound by earlier declarations,
 thus the well-formedness of the worklist is preserved,
@@
 
-\bruno{You have some note ``differs'' on the rules. Should that
-  deserve an explanation here?}
+% \bruno{You have some note ``differs'' on the rules. Should that
+%   deserve an explanation here?}
 
@@
 Figure~\ref{fig:alg:sample} shows a sample derivation of the algorithm.
-The judgment given is to check the application $(\lam x)~()$ against the unit type.
+It checks the application $(\lam x)~()$ against the unit type.
 According to the algorithm, we apply Rule 18 (subsumption), changing to inference mode.
 Type inference of the application breaks into two steps by Rule 26:
-firstly we infer the type of the function,
-and then the application inference judgment help determine the return type.
-The type of the identity function, $\lam x$, is inferred to be $\al \to \al$,
-which can be interpreted as $\tau \to \tau$ for any valid monotype $\tau$,
-after the following 5 steps:
-a checking on the body of the lambda function (Rule 25),
-calling the inference mode for the checking (Rule 18)
-a term variable inference (Rule 22),
-a subtyping on existential variables (Rule 12) and
-a garbage collection of the term variable $x$ (Rule 3).
+first we infer the type of the function,
+and then the application inference judgment helps determine the return type.
+In the following 5 steps The type of the identity function, $\lam x$, is inferred to be $\al \to \al$:
+checking the body of the lambda function (Rule 25),
+switching from check mode to inference mode (Rule 18),
+inferring the type of a term variable (Rule 22),
+solving a subtyping between existential variables (Rule 12) and
+garbage collecting the term variable $x$ (Rule 3).
 
 After that, Rule 28 changes the application inference judgment to
-a checking on the argument against the input type $\al$ and returns the output type $\al$.
+a check of the argument against the input type $\al$ and returns the output type $\al$.
 Checking $()$ against the existential variable $\al$ solves $\al$ to the unit type $1$
 through Rules 18, 24 and 16.
-Immediately after $\al$ is solved, the algorithm substitute every occurrence of $\al$ to $1$.
-Therefore the worklist remains a simple subtyping judgment $1 \le 1$, which finishes off by Rule 4.
+Immediately after $\al$ is solved, the algorithm replaces every occurrence of $\al$ with $1$.
+Therefore the worklist remains a simple subtyping judgment $1 \le 1$, which is finished off by Rule 4.
 Finally we get an empty worklist, indicating the success of the whole derivation.
 
-To summarize, our algorithm accepts the type checking $(\lam x)~() \Lto 1$.
-
-\bruno{Derivation should be explained briefly at least.}
+In summary, our algorithm accepts the type checking $(\lam x)~() \Lto 1$.
 
@@
 
-Such restrictions are not arbitrarily imposed:
-we foramlly proved that the restricted algorithm is still complete.
-In Section~\ref{sec:metatheory:non-overlapping} we discuss the relavant metatheory,
-with the help of a restricted version of declarative system.
+These restrictions have not been imposed arbitrarily:
+we formally prove that the restricted algorithm is still complete.
+In Section~\ref{sec:metatheory:non-overlapping} we discuss the relevant metatheory,
+with the help of a non-overlapping version of the declarative system.

### 2019-02-28 sections/Overview.tex
Appendix and all pointers
Section 3
S2: small fixes

@@
 \text{Type variables}\qquad&a, b
-\\[3mm]
+\\
 \text{Types}\qquad&A, B, C &::=&\quad 1 \mid a \mid \forall a. A \mid A\to B\\
 \text{Monotypes}\qquad&\tau,\sigma &::=&\quad 1 \mid a \mid \tau\to \sigma
-\\[3mm]
+\\
 \text{Expressions}\qquad&e &::=&\quad x \mid () \mid \lam e \mid e_1~e_2 \mid (e:A)
-\\[3mm]
+\\
 \text{Contexts}\qquad&\Psi &::=&\quad \nil \mid \Psi, a \mid \Psi, x:A
@@
 \paragraph{Markers for scoping.}
-The sequential order of entries in the algoritmic context, in combination with
+The sequential order of entries in the algorithmic context, in combination with
 the threading of contexts,  does not perfectly capture the scoping of all
@@
 More seriously, we have also found a lemma that simply does not hold, 
-Lemma 29 (Parallel Admissibility). See Appendix~\ref{} for a detailed explanation
+Lemma 29 (Parallel Admissibility). See Appendix~\ref{appendix:false_lemmas}
+\bruno{reference} for a detailed explanation
 and counterexample. This lemma is a cornerstone of the two metatheoretical results 
@@
 algorithms. For instance, the well-known unification
-algorithm by \citet{unification} decomposes the problem of unifying two terms $s \stackrel{.}{=} t$ into a set
+algorithm by \citet{unification} decomposes the problem of unifying two terms $s \stackrel{.}{=} t$ into a number
 of related unification problems between pairs of terms $s_i \stackrel{.}{=} t_i$. These smaller
@@
 $\mathcal{K}$. The latter consists of constraints like $\Psi \vdash M = N : C$
-that are equipped with their individual context $\Psi$. While acurrately tracking
+that are equipped with their individual context $\Psi$. While accurately tracking
 the scoping of regular and meta-variables, this approach is not ideal because it
@@
 
-\paragraph{Scoping Garbage} Firstly, the single common type ordered context 
+\paragraph{Scoping Garbage} Firstly, the single common ordered context 
 $\Gamma$ does not accurately reflect the type and unification variables
@@
 \emph{checking} whether one type is the subtype of another. While this may not
-so be difficult to scale to the \emph{checking} mode of term typing $\Gamma
+be so difficult to adapt to the \emph{checking} mode of term typing $\Gamma
 \vdash e \Leftarrow A$, it clearly lacks the functionality to support the
@@
 Our worklist can be seen as an all-in-one stack,
-containing variable declarations and subtyping/typing judgments.
+containing variable declarations and subtyping/ typing judgments.
 The stack is an enriched form of ordered context,
@@
 However, the non-trivial forms of chains capture an
-inference task together with the the work that depends on its outcome. For
+inference task together with the work that depends on its outcome. For
 instance, a type inference task takes the form $e \Rightarrow_a \omega$.
@@
 (please refer to Figure~\ref{fig:alg} for detailed derivations).
-Then we substitute every instance of $\al$ and $\bt$ to $1$,
+Then we substitute every instance of $\al$ and $\bt$ with $1$,
 so the variable declarations can be safely removed from the worklist.
@@
 %just as what we have mentioned above about the scope tracking.
-After removing the garbage term variable declaration from the worklist, we process the last judgment chain from the left,
+After removing the garbage term variable declaration from the worklist, we process the last remaining judgment chain,
 namely an inference judgment $(\lam x)~()\To~?$, with the unit type $1$ as its result.
-Finally the last judgment becomes $1 \le 1$, and is satisfied as a base case.
+Finally the last judgment becomes $1 \le 1$, a trivial base case.
 

### 2019-02-28 sections/Overview.tex
blue -> underline

@@
 Take the following worklist as an example
-$$\al \Vdash \blue{(\lam x)~() \To_a a \le \al} , x:\al, \bt \Vdash \blue{\al \le \bt} \Vdash \blue{\bt \le 1}$$
-There are three (blue-colored) judgment chains in the worklist,
+$$\al \Vdash \underline{(\lam x)~() \To_a a \le \al} , x:\al, \bt \Vdash \underline{\al \le \bt} \Vdash \underline{\bt \le 1}$$
+There are three (underlined) judgment chains in the worklist,
 where the first and second judgment chains (from the right) are two subtyping judgments,
@@
 Now we reduce the worklist to the following state
-$$\nil \Vdash \blue{(\lam x)~() \To_a a \le 1}, x:1$$
+$$\nil \Vdash \underline{(\lam x)~() \To_a a \le 1}, x:1$$
 which has a term variable declaration as the top element.

### 2019-02-28 sections/algorithm.tex
Section 3

@@
 reduction steps. If $\Gm\redto\nil$, we say that $\Gm$ reduces successfully;
-this corresponds to a successful type checking.
+this corresponds to successful type checking.
 
 Let us explain the rules in more detail:
-Rules 1-3 pop a variable declaration; this garbage collects variables that are
-out of scope for the remaining judgements.
+Rules 1-3 pop a variable declaration. Essentially this is garbage: variables that are
+out of scope for the remaining judgements in the worklist.
 All other rules concern a judgement at the front of the worklist. Logically we
-can discern 6 group of rules.
+can discern 6 groups of rules.
 
@@
 they are fresh existential variables with the same scope as $\al$.
-We remove replace $\al$ in the worklist with  $\al[1], \al[2]$.
+We replace $\al$ in the worklist with  $\al[1], \al[2]$.
 To propagate the instantiation to the rest of the worklist and maintain well-formedness,
@@
 they \emph{return} a type instead of only accepting or rejecting.
-For algorithmic system, where guessings are involved,
+For the algorithmic system, where guesses are involved,
 it may happen that the output type of an inference judgment refers to new existential variables,
@@
 By making use of the continuation-passing-style judgement chain,
-inner judgments always shares context with its parent judgment,
-therefore we achieve all the desired property in such a neat way.
+inner judgments always share the context with their parent judgment.
 
@@
 the assumption of argument type ($x:\al$) is recycled after checking against the function body;
-the declaration of the function type ($\al,\bt$) is only visible to
-the checking and the remaining chain $[\al\to\bt/a]\jg$.
+the existential variables used by the function type ($\al,\bt$) are only visible in the remaining chain $[\al\to\bt/a]\jg$.
 The recycling process of Rule 25 differs from DK's corresponding rule significantly,
@@
 
-\paragraph{\bf 6. Application inferences judgements}
+\paragraph{\bf 6. Application inference judgements}
 Finally, Rules 27-29 deal with application inference judgments.
@@
 first we infer the type of the function,
-and then the application inference judgment helps determine the return type.
-In the following 5 steps The type of the identity function, $\lam x$, is inferred to be $\al \to \al$:
+and then the application inference judgment helps to determine the return type.
+In the following 5 steps the type of the identity function, $\lam x$, is inferred to be $\al \to \al$:
 checking the body of the lambda function (Rule 25),

### 2019-02-28 sections/Metatheory.tex
Appendix and all pointers
S4.6 Abella and Proof Statistics

@@
 The following inversion lemma for Rule $\mathtt{\forall R}$ resolves the overlap:
-\begin{lemma}[Invertibility of $\mathtt{\forall R}$]
+\begin{lemma}[Invertibility of $\mathtt{\forall R}$]\label{lem:inv_allR}
 If $\Psi \vdash A \le \all B$ then $\Psi, a \vdash A \le B$.
@@
 We have proven two inversion lemmas for these overlaps:
-\begin{lemma}[Invertibility of $\mathtt{Decl\forall I}$]
+\begin{lemma}[Invertibility of $\mathtt{Decl\forall I}$]\label{lem:inv_chkAll}
 If $\Psi \vdash e \Lto \all A$ then $\Psi, a \vdash e \Lto A$.
 \end{lemma}
-\begin{lemma}[Invertibility of $\mathtt{Decl{\to}I}$]
+\begin{lemma}[Invertibility of $\mathtt{Decl{\to}I}$]\label{lem:inv_chkLam}
 If $\Psi \vdash \lam e \Lto A \to B$ then $\Psi, x:A \vdash e \Lto B$.
@@
 
-\begin{lemma}[Subsumption]
+\begin{lemma}[Subsumption]\label{lem:subsumption}
 Given $\Psi' \le \Psi$:
@@
 Fortunately we have found a different way to prove the lemma.
-The details of this issue can be found in Appendix~\ref{}.\bruno{reference}
+The details of this issue can be found in Appendix~\ref{appendix:subsumption}.
 
@@
 
-\begin{lemma}[Prune Transfer for Instantiation]
+\begin{lemma}[Prune Transfer for Instantiation]\label{lem:prune_inst}
 If $(\Gm \Vdash \al \le A \to B) \sto (\Om \Vdash C \le A_1 \to B_1)$ and
@@
 \end{corollary}
-\subsection{Abella}
+
+% ---------------------------------------------
+\subsection{Abella and Proof Statistics}
+We have chosen the Abella (v2.0.7-dev) proof assistant~\cite{AbellaDesc} to develop our formalization.
+Abella is designed to help such formalizations for programming languages,
+due to its built-in support for variable bindings and
+the $\lambda$-tree syntax~\cite{}, i.e. a form of HOAS.
+Nominal variables, or $\nabla$-quantified variables, are used as an unlimited name supply,
+which supports explicit freshness control and substitutions.
+
+Although Abella lacks packages, tactics and support on modules,
+we enjoy its higher-order system and the ease of formalizing substitution-intensive relations,
+which are widely used in our algorithm.
+
+\paragraph{Statistics of the Proof}
+The proof script consists of 7683 lines of code with a total of
+51 definitions and 567 theorems.
+Figure~\ref{fig:proof_statistics} briefly summarizes the contents in each file.
+The files are linealy dependent due to the limitation of Abella.
+
+\begin{figure}
+\centering\begin{tabular}{|c|c|c|l|}\hline
+    File(s) & SLOC & \#Theorems & Description\\\hline
+    olist.thm, nat.thm  &  309 & 56  & Basic data structures\\\hline
+    typing.thm          &  238 & 5   & Declarative \& algorithmic system, debug examples\\\hline
+    decl.thm            &  226 & 33  & Basic declarative properties\\\hline
+    order.thm           &  235 & 27  & The $|\cdot|_\forall$ measure; decl. subtyping strenthening\\\hline
+    alg.thm             &  655 & 72  & Basic algorithmic properties\\\hline
+    trans.thm           &  628 & 53  & \makecell[l]{Transfer and its basic properties \\
+                            Lemmas~\ref{lem:insert}, \ref{lem:extract}}\\\hline
+    declTyping.thm      &  944 & 72  & \makecell[l]{Additional properties for declarative typing \\
+                            Lemmas~\ref{lem:inv_allR}, \ref{lem:inv_chkAll}, \ref{lem:inv_chkLam}}\\\hline
+    soundness.thm       & 1069 & 70  & Soundness theorem; aux. lemmas on transfer\\\hline
+    depth.thm           &  206 & 14  & \makecell[l]{The $|\cdot|_\to$ measure; \\
+                            Lemma~\ref{lem:prune_inst}}\\\hline
+    \makecell{dcl.thm, \\completeness.thm} & 1458 & 73  &
+                            Completeness theorem; aux. lemmas and relations\\\hline
+    inst\_decidable.thm &  830 & 42  & Lemma~\ref{lemma:inst:decidable} Instantiation Decidability\\\hline
+    decidability.thm    &  885 & 50  & Lemma~\ref{lemma:inst:decidable}\\\hline
+    Total               & 7683 & 567 & (51 definitions in total)\\\hline
+\end{tabular}
+\caption{Statistics for the proof scripts}
+\label{fig:proof_statistics}
+\end{figure}

### 2019-02-28 sections/Metatheory.tex
a missing reference

@@
 due to its built-in support for variable bindings and
-the $\lambda$-tree syntax~\cite{}, i.e. a form of HOAS.
+the $\lambda$-tree syntax~\cite{miller2000abstract}, i.e. a form of HOAS.
 Nominal variables, or $\nabla$-quantified variables, are used as an unlimited name supply,

### 2019-02-28 sections/Metatheory.tex
minor typos

@@
 
-\begin{figure}
+\begin{figure}[ht]
 \centering\begin{tabular}{|c|c|c|l|}\hline
@@
                             Completeness theorem; aux. lemmas and relations\\\hline
-    inst\_decidable.thm &  830 & 42  & Lemma~\ref{lemma:inst:decidable} Instantiation Decidability\\\hline
-    decidability.thm    &  885 & 50  & Lemma~\ref{lemma:inst:decidable}\\\hline
+    inst\_decidable.thm &  830 & 42  & Lemma~\ref{lemma:inst:decidable}\\\hline
+    decidability.thm    &  885 & 50  & Decidability theorem\\\hline
     Total               & 7683 & 567 & (51 definitions in total)\\\hline

### 2019-03-01 sections/Metatheory.tex
remove 2 of my comments
Section 4

@@
 The resulting declarative system has no overlapping rules
-and more closely resembles the aglorithmic system, which contains
+and more closely resembles the algorithmic system, which contains
 the same shape constraints.
@@
 \begin{enumerate}
-    \item If $\Psi \vdash e \Lto A$ and $\Psi \vdash A \le A'$ then $\Psi' \vdash e \Lto A$;
+    \item If $\Psi \vdash e \Lto A$ and $\Psi \vdash A \le A'$ then $\Psi' \vdash e \Lto A'$;
     \item If $\Psi \vdash e \To A$ then there exists
-        $A'$ s.t. $\Psi \vdash A' \le A$ and $\Psi' \vdash e \To A$;
+        $A'$ s.t. $\Psi \vdash A' \le A$ and $\Psi' \vdash e \To A'$;
     \item If $\Psi \vdash \appInf{C}{e}{A}$ and $\Psi \vdash C' \le C$,
@@
 We have tried to follow DK's manual proof of this lemma,
-but we discovered several problems in their reasoning that we have been able to address.
+but we discovered several problems in their reasoning that we have been unable to address.
 Fortunately we have found a different way to prove the lemma.
@@
 our non-overlapping declarative system, and our algorithmic system.
-We have already establishes the first relation, that the two declarative
+We have already established the first relation, that the two declarative
 systems are equivalent.
-
 In what follows, we will establish the soundness of our algorithm directly
@@
 
-The proof proceeds by induction on derivation of $\Gm\redto\nil$.
-Interesting cases are the those involving existential variable instantiations,
+The proof proceeds by induction on the derivation of $\Gm\redto\nil$.
+Interesting cases are those involving existential variable instantiations,
 including Rules 10, 11, 21 and 29.
@@
     \item If $\Gm \Vdash e \Lto A \redto \nil$\\
-        then there exist $e', A', \Om$ s.t.
+        then there exist $ A', \Om$ s.t.
         $\Gm \Vdash e \Lto A \sto \Om \Vdash e \Lto A'$ and $\|\Om\| \vdash e \Lto A'$;
@@
 \end{corollary}
+\bruno{if possible double-check against the corresponding Abella ones.}
 
@@
 
-Completeness of our algorithm means that any derivation in the
+The completeness of our algorithm means that any derivation in the
 declarative system has an algorithmic counterpart.
@@
 the induction always analyses the first judgment until it is satisfied.
-
 As the algorithmic system introduces existential variables,
@@
 and thus we analyse each of the possible cases.
-\jimmy{Cite DK's proof as the sketch.}
+% \jimmy{Cite DK's proof as the sketch.}
+% I later think it might not be needed; the induction is not that hard to think of
 
@@
 \end{corollary}
+\bruno{We discussed the form of the lemmas on slack. Please have the first premise
+  be a declarative judgement.
+  If possible also double-check against Abella.}
 
@@
 We have chosen the Abella (v2.0.7-dev) proof assistant~\cite{AbellaDesc} to develop our formalization.
-Abella is designed to help such formalizations for programming languages,
-due to its built-in support for variable bindings and
+Abella is designed to help with such formalizations of programming languages,
+due to its built-in support for variable binding and
 the $\lambda$-tree syntax~\cite{miller2000abstract}, i.e. a form of HOAS.
@@
 which supports explicit freshness control and substitutions.
-
 Although Abella lacks packages, tactics and support on modules,
-we enjoy its higher-order system and the ease of formalizing substitution-intensive relations,
-which are widely used in our algorithm.
+its higher-order unification and the ease of formalizing substitution-intensive relations
+are very helpful for our formalization.
 
@@
 Figure~\ref{fig:proof_statistics} briefly summarizes the contents in each file.
-The files are linealy dependent due to the limitation of Abella.
+The files are linearly dependent due to the limitation of Abella.
 

### 2019-03-01 sections/introduction.tex
Section 4

@@
 Due to those two challenges, our work takes a different approach by refining and
-extending the idea of \emph{worklists judgments}~\cite{itp2018},
+extending the idea of \emph{worklist judgments}~\cite{itp2018},
 proposed recently to mechanically formalize an algorithm for

### 2019-03-01 sections/Metatheory.tex
S4 completeness corollary; statistics table

--- 2019-02-28 sections/Metatheory.tex
@@
 \begin{enumerate}
-    \item If $\Gm \Vdash A \le B \sto \Om \Vdash A' \le B'$ and $\|\Om\| \vdash A' \le B'$
+    \item If $\Om \vdash A' \le B'$ and $\Gm \Vdash A \le B \sto \Om \Vdash A' \le B'$
         then $\Gm \Vdash A \le B \redto \nil$;
-    \item If $\Gm \Vdash e \Lto A \sto \Om \Vdash e \Lto A'$ and $\|\Om\| \vdash e \Lto A'$
+    \item If $\Om \vdash e \Lto A'$ and $\Gm \Vdash e \Lto A \sto \Om \Vdash e \Lto A'$
         then $\Gm \Vdash e \Lto A \redto \nil$;
-    \item If $\Gm \sto \Om$ and exists $A$ s.t. $\|\Om\| \vdash e \To A$
+    \item If $\Om \vdash e \To A$ and $\Gm \sto \Om$
         then $\Gm \Vdash e \To_a 1 \le 1 \redto \nil$;
-    \item If $\Gm \Vdash \appInfAlg{A}{e}[a][1 \to 1] \sto \Om \Vdash \appInfAlg{A'}{e}[a][1 \to 1]$
-        and exists $C$ s.t. $\|\Om\| \vdash \appInf{A'}{e}{C}$
+    \item If $\Om \vdash \appInf{A'}{e}{C}$ and
+        $\Gm \Vdash \appInfAlg{A}{e}[a][1 \to 1] \sto \Om \Vdash \appInfAlg{A'}{e}[a][1 \to 1]$
         \\
@@
 \subsection{Abella and Proof Statistics}
-We have chosen the Abella (v2.0.7-dev) proof assistant~\cite{AbellaDesc} to develop our formalization.
+We have chosen the Abella (v2.0.7-dev
+\footnote{A develop version is used because the dev just fixed a serious bug that
+accepts a simple proof of false, which also affects the completeness of our proof.
+The author compiled the scripts against commit 92829a of the GitHub repository of Abella.}
+) proof assistant~\cite{AbellaDesc} to develop our formalization.
 Abella is designed to help such formalizations for programming languages,
@@
                             Lemmas~\ref{lem:insert}, \ref{lem:extract}}\\\hline
-    declTyping.thm      &  944 & 72  & \makecell[l]{Additional properties for declarative typing \\
+    declTyping.thm      &  944 & 72  & \makecell[l]{Non-overlapping declarative system \\
                             Lemmas~\ref{lem:inv_allR}, \ref{lem:inv_chkAll}, \ref{lem:inv_chkLam}}\\\hline
     soundness.thm       & 1069 & 70  & Soundness theorem; aux. lemmas on transfer\\\hline
-    depth.thm           &  206 & 14  & \makecell[l]{The $|\cdot|_\to$ measure; \\
-                            Lemma~\ref{lem:prune_inst}}\\\hline
-    \makecell{dcl.thm, \\completeness.thm} & 1458 & 73  &
+    depth.thm           &  206 & 14  & The $|\cdot|_\to$ measure; Lemma~\ref{lem:prune_inst}\\\hline
+    dcl.thm             &  383 & 12  & Non-overlapping declarative worklist \\\hline
+    completeness.thm    & 1075 & 61  &
                             Completeness theorem; aux. lemmas and relations\\\hline
-    inst\_decidable.thm &  830 & 42  & Lemma~\ref{lemma:inst:decidable}\\\hline
+    inst\_decidable.thm &  830 & 42  & Other measures; Lemma~\ref{lemma:inst:decidable}\\\hline
     decidability.thm    &  885 & 50  & Decidability theorem\\\hline

### 2019-03-01 sections/Metatheory.tex

@@
 ) proof assistant~\cite{AbellaDesc} to develop our formalization.
-Abella is designed to help such formalizations for programming languages,
-due to its built-in support for variable bindings and
+Abella is designed to help with such formalizations of programming languages,
+due to its built-in support for variable binding and
 the $\lambda$-tree syntax~\cite{miller2000abstract}, i.e. a form of HOAS.

### 2019-03-01 sections/Discussion.tex
Section 5.1

@@
 
+This section discusses some insights that we gained from our work and contrast
+the scoping mechanisms employed by us with those DK's original algorithm.
+We also discuss a way to improve their scoping tracking precision.
+Furthermore we discuss and sketch out an extension of the our algorithm with
+elaboration to a target calculus.
+
+\begin{comment}
 \subsection{Implementation}
 Anything to say about the implementation? Do we have one?
+\end{comment}
 
-\subsection{Making DK's scoping more precise}\label{sec:discussion:scoping}
+\subsection{Contrasting our scoping mechanisms with DK's scoping}\label{sec:discussion:scoping}
 
-DK's algorithm deals with garbage variables in a different approach,
-as discussed in Section~\ref{ssec:DK_Algorithm},
-through type variables or existential variable markers.
-However, there is still a chance that
-unused existential variables leak to output context and accumulate infinitely.
+One nice feature of our worklists is that scoping of variables can be
+made precise, simply by interleaving variable declarations and
+judgement chains.  DK's algorithm deals with garbage variables in a
+different way through type variables or existential variable
+markers (as discussed in Section~\ref{ssec:DK_Algorithm}).  Despite
+the sophistication employed in DK's algorithm to keep scoping precise,
+there is still a chance that unused existential variables leak their
+scope to an output context and accumulate indefinitely.
 For example, the derivation of the judgment $(\lam x)~() \Lto 1$ is as follows
@@
 
-One may blame that the inference rule for lambda functions does not collect garbages correctly.
+One may blame the inference rule for lambda functions for not collecting garbage correctly.
 Here we also show another rule for application inference with a similar problem:
@@
 $$
-But our algorithm collects the existential variables right after the next judgment chain:
-$$\Gm \Vdash \lambda x. e \To_a \jg \rrule{25}
-\Gm,\al,\bt \Vdash [\al\to\bt/a]\jg, x:\al \Vdash e\Lto \bt$$
-$$\Gm \Vdash \appInfAlg{\forall a. A}{e} \rrule{27} \Gm,\al \Vdash \appInfAlg{[\al/a]A}{e}$$
-But it seems impossible to propose a fix by only modifying this rule.
+In contrast our algorithm collects the existential variables right after the next judgment chain:
+\[\Gm \Vdash \lambda x. e \To_a \jg \rrule{25}
+\Gm,\al,\bt \Vdash [\al\to\bt/a]\jg, x:\al \Vdash e\Lto \bt\]
+\[\Gm \Vdash \appInfAlg{\forall a. A}{e} \rrule{27} \Gm,\al \Vdash \appInfAlg{[\al/a]A}{e}\]
+But it seems impossible to achieve a similar outcome by only modifying the two rules in DK's system.
 Taking $\mathtt{DK\_{\to}I{\To}}$ as an example,
-the declaration or solution for $\al$ and $\bt$ may be referred by following judgments,
-therefore leaving $\al$ and $\bt$ in the output context is the only choice,
-when the next judgments are inaccessible to us at the moment.
+the declaration or solution for $\al$ and $\bt$ may be referred to by following judgments.
+Therefore leaving $\al$ and $\bt$ in the output context is the only choice,
+when the next judgments are still inaccessible.
 
-To fix the problem, we suggest a modification on the algorithmic subsumption rule,
+To fix the problem, one possible modification is on the algorithmic subsumption rule,
 as garbage collection for a checking judgment is always safe:
@@
 $$
+Here we employ the markers in a way different from what they were originally intended for.
+We create a dummy fresh $\al$ and push a marker into the inference judgement.
+When the subtyping judgement is used we look for the marker and drop everything afterwards.
 We pick this rule because this is the only one where a checking judgment calls an inference judgment.
@@
 DK's algorithm behaves equivalently to our algorithm in terms of variable scoping.
+However this is exploiting markers in a way that they were not intended for and seems ad-hoc.
 

### 2019-03-01 sections/Discussion.tex
Section 5

@@
 \subsection{Elaboration}
+A common extension to type-inference algorithms is to have an
+associated elaboration.  For example, for languages with implicit
+polymorphism, it is common to have an elaboration to a variant of
+System F~\cite{reynolds1983types}, which recovers type information and explicit type
+applications. Therefore a natural question is whether our algorithm
+can also accomodate an elaboration step.
 While our algorithmic reduction does not elaborate to System F,
-it is not difficult to extend the algorithm with a type-directed translation.
+we believe that it is not difficult to extend the algorithm with a (type-directed) elaboration.
+We explain the rought idea as follows.
+
 Since the judgment form of our algorithmic worklist contains a collection of judgments,
 elaboration expressions are also generated as a list of equal length to
-the number of judgments (NOT judgment chains!) in the worklist.
-Within expectation, subtyping judgments translate to type coercions (denoted by $f$),
+the number of judgments (\emph{not judgment chains}) in the worklist.
+Within expectation, subtyping judgments translate to coercions (denoted by $f$),
 all three other types of judgments translate to terms in System F (denoted by $t$).
+\bruno{Why distinguish coercions and terms? Coercions are just terms, right?}
 
@@
 
-And the form of our algorithmic judgment becomes:
+Then the form of our algorithmic judgment becomes:
 $$\Gamma \hookrightarrow \Phi$$
 
-We take Rule 18 as an example, rewritting small-step reduction into a relational style,
+We take Rule 18 as an example, rewritting small-step reduction in a relational style,
 $$
@@
 If an existential variable is solved, we not only propagate the solution to the rest judgments,
-but also replace ocurrances in the elaboration list.
-If an existential variable is recycled, meaning that it is not constraint,
+but also replace ocurrences in the elaboration list.
+If an existential variable is recycled, meaning that it is not constrained,
 we can pick any well-formed monotype as its solution.

### 2019-03-01 sections/Discussion.tex
slight modifications to S5

@@
 $$
-In contrast our algorithm collects the existential variables right after the next judgment chain:
+In contrast Rule 25 of our algorithm collects the existential variables
+right after the second judgment chain,
+and Rule 27 collects one exsitential variable similarly:
 \[\Gm \Vdash \lambda x. e \To_a \jg \rrule{25}
@@
 We create a dummy fresh $\al$ and push a marker into the inference judgement.
-When the subtyping judgement is used we look for the marker and drop everything afterwards.
+After the subtyping judgement is processed we look for the marker and drop everything afterwards.
 We pick this rule because this is the only one where a checking judgment calls an inference judgment.
@@
 \bruno{Why distinguish coercions and terms? Coercions are just terms, right?}
+\jimmy{please help on the question @Tom.}
 

### 2019-03-01 sections/Discussion.tex
5.1 revision

@@
 
-This section discusses some insights that we gained from our work and contrast
-the scoping mechanisms employed by us with those DK's original algorithm.
-We also discuss a way to improve their scoping tracking precision.
-Furthermore we discuss and sketch out an extension of the our algorithm with
+This section discusses some insights that we gained from our work and contrasts
+the scoping mechanisms we have employed by those in DK's algorithm.
+We also discuss a way to improve the precision of their scope tracking.
+Furthermore we discuss and sketch an extension of our algorithm with
 elaboration to a target calculus.
@@
 
-One nice feature of our worklists is that scoping of variables can be
-made precise, simply by interleaving variable declarations and
-judgement chains.  DK's algorithm deals with garbage variables in a
-different way through type variables or existential variable
+A nice feature of our worklists is that, simply by interleaving variable declarations and
+judgement chains, they make the scope of variables
+precise.  DK's algorithm deals with garbage collecting variables in a
+different way: through type variable or existential variable
 markers (as discussed in Section~\ref{ssec:DK_Algorithm}).  Despite
@@
 or to the very end of a type inference task.
-In that sense, DK's algorithm does not control the scoping of variables precisely enough.
+In that sense, DK's algorithm does not control the scoping of variables precisely.
 %However, it is a minor issue that does not affect the soundness and completeness properties.
 
-One may blame the inference rule for lambda functions for not collecting garbage correctly.
-Here we also show another rule for application inference with a similar problem:
+Two rules we may blame for not garbage collecting correctly are the inference rule for lambda
+functions and an application inference rule:
 $$
@@
 $$
-In contrast Rule 25 of our algorithm collects the existential variables
+In contrast, Rule 25 of our algorithm collects the existential variables
 right after the second judgment chain,
-and Rule 27 collects one exsitential variable similarly:
+and Rule 27 collects one existential variable similarly:
 \[\Gm \Vdash \lambda x. e \To_a \jg \rrule{25}
@@
 \[\Gm \Vdash \appInfAlg{\forall a. A}{e} \rrule{27} \Gm,\al \Vdash \appInfAlg{[\al/a]A}{e}\]
-But it seems impossible to achieve a similar outcome by only modifying the two rules in DK's system.
+It seems impossible to achieve a similar outcome in DK's system by only modifying these two rules.
 Taking $\mathtt{DK\_{\to}I{\To}}$ as an example,
-the declaration or solution for $\al$ and $\bt$ may be referred to by following judgments.
+the declaration or solution for $\al$ and $\bt$ may be referred to by subsequent judgments.
 Therefore leaving $\al$ and $\bt$ in the output context is the only choice,
-when the next judgments are still inaccessible.
+when the subsequent judgments cannot be consulted.
 
@@
 $$
-Here we employ the markers in a way different from what they were originally intended for.
-We create a dummy fresh $\al$ and push a marker into the inference judgement.
+Here we employ the markers in a way they were originally not intended for.
+We create a dummy fresh existential $\al$ and add a marker to the input context of the inference judgement.
 After the subtyping judgement is processed we look for the marker and drop everything afterwards.
-We pick this rule because this is the only one where a checking judgment calls an inference judgment.
-And that's the point where our algorithm recycles variables---right after a judgment chain is satisfied.
-After applying this patch, to our best knowledge,
+We pick this rule because it is the only one where a checking judgment calls an inference judgment.
+That is the point where our algorithm recycles variables---right after a judgment chain is satisfied.
+After applying this patch, to the best of our knowledge,
 DK's algorithm behaves equivalently to our algorithm in terms of variable scoping.
-However this is exploiting markers in a way that they were not intended for and seems ad-hoc.
+However, this exploits markers in a way they were not intended to be used and seems ad-hoc.
 

### 2019-03-01 sections/Appendix.tex
S4 and Appendix lemma sync (minor swap of letters)

@@
 
-\subsection{False Lemmas in DK's papers}
+\subsection{Broken Lemmas in DK's papers}
 \label{appendix:false_lemmas}
@@
 \begin{enumerate}
-    \item If $\Psi \vdash e \Lto A$ and $\Psi \vdash A \le A'$ then $\Psi' \vdash e \Lto A$;
+    \item If $\Psi \vdash e \Lto A$ and $\Psi \vdash A \le A'$ then $\Psi' \vdash e \Lto A'$;
     \item If $\Psi \vdash e \To B$ then there exists
-        $A'$ s.t. $\Psi \vdash B' \le B$ and $\Psi' \vdash e \To B$;
-    \item If $\Psi \vdash \appInf{C}{e}{A}$ and $\Psi \vdash C' \le C$,
-        then there exists $A'$ s.t. $\Psi \vdash A' \le A$ and $\Psi' \vdash \appInf{C'}{e}{A'}$.
+        $B'$ s.t. $\Psi \vdash B' \le B$ and $\Psi' \vdash e \To B'$;
+    \item If $\Psi \vdash \appInf{A}{e}{C}$ and $\Psi \vdash A' \le A$,
+        then there exists $C'$ s.t. $\Psi \vdash C' \le C$ and $\Psi' \vdash \appInf{A'}{e}{C'}$.
 \end{enumerate}
@@
 The proof is by a nested mutual induction on the lexicographical order of the measures
-$$\langle |e|_e, |\cdot|_\Leftrightarrow, |A \to B|_\forall \rangle,$$
+$$\langle |e|_e, |\cdot|_\Leftrightarrow, |A|_\forall + |A'|_\forall \rangle,$$
 where the second measure is 2 for checking (1), 1 for inference (2) and 3 for application inference (3);
@@
 thus the subtyping relation derives from either $\mathtt{{\le}\forall L}$ or $\mathtt{{\le}\forall R}$.
-For each of the case, the third measure $|A \to B|_\forall$ decreases
+For each of the case, the third measure $|A|_\forall + |A'|_\forall$ decreases
 (the $\mathtt{{\le}\forall L}$ case requires a type substitution lemma obtaining
@@
 When the subtyping relation is derived from $\mathtt{{\le}\forall R}$,
-$|B|_\forall$ decreases, and thus the induction hypothesis finishes this case.
+$|A'|_\forall$ decreases, and thus the induction hypothesis finishes this case.

### 2019-03-01 sections/Discussion.tex

@@
 \subsection{Elaboration}
-A common extension to type-inference algorithms is to have an
+
+Type-inference algorithms are often extended with an
 associated elaboration. For example, for languages with implicit
@@
 applications. Therefore a natural question is whether our algorithm
-can also accomodate an elaboration step.
+can also accomodate such elaboration.
 While our algorithmic reduction does not elaborate to System F,
 we believe that it is not difficult to extend the algorithm with a (type-directed) elaboration.
-We explain the rought idea as follows.
+We explain the rough idea as follows.
 
@@
 the number of judgments (\emph{not judgment chains}) in the worklist.
-Within expectation, subtyping judgments translate to coercions (denoted by $f$),
+As usual, subtyping judgments translate to coercions (denoted by $f$ and
+represented by System F functions),
 all three other types of judgments translate to terms in System F (denoted by $t$).
-\bruno{Why distinguish coercions and terms? Coercions are just terms, right?}
-\jimmy{please help on the question @Tom.}
 
@@
 
-We take Rule 18 as an example, rewritting small-step reduction in a relational style,
+We take Rule 18 as an example, rewriting small-step reduction in a relational style,
 $$
@@
 As is shown in the conclusion of the rule,
-a checking judgment on the top of the worklist corresponds to a top element for elaboration.
-The premise indicates that one judgment chain may relate to more than one elaboration elements,
+a checking judgment at the top of the worklist corresponds to a top element for elaboration.
+The premise shows that one judgment chain may relate to more than one elaboration elements,
 and that the outer judgment, being processed before inner ones,
@@
 
-Existential variables need special treatment, since they may be solved at any step,
-or being recycled if not solved in its scope.
-If an existential variable is solved, we not only propagate the solution to the rest judgments,
+Existential variables need special treatment, since they may be solved at any point,
+or be recycled if not solved within its scope.
+If an existential variable is solved, we not only propagate the solution to the other judgments,
 but also replace ocurrences in the elaboration list.

### 2019-03-01 sections/Metatheory.tex
S4 and Appendix lemma sync (minor swap of letters)
S4 double-checked Abella script against corollaries

@@
 The typing contexts of declarative judgements are recovered using an
-auxiliary erasure function $\|\Om\|$. The erasure function
-simply drops all judgment entries from the worklist, keeping only variable
-and type variable declarations.
+auxiliary erasure function $\|\Om\|$
+\footnote{In the proof script we do not use the erasure function,
+for the declarative system and well-formedness judgments automatically fits
+the non-erased declarative worklist just as declarative contexts.}.
+The erasure function simply drops all judgment entries from the worklist,
+keeping only variable and type variable declarations.
 
@@
     \item If $\Psi \vdash e \Lto A$ and $\Psi \vdash A \le A'$ then $\Psi' \vdash e \Lto A'$;
-    \item If $\Psi \vdash e \To A$ then there exists
-        $A'$ s.t. $\Psi \vdash A' \le A$ and $\Psi' \vdash e \To A'$;
-    \item If $\Psi \vdash \appInf{C}{e}{A}$ and $\Psi \vdash C' \le C$,
-        then there exists $A'$ s.t. $\Psi \vdash A' \le A$ and $\Psi' \vdash \appInf{C'}{e}{A'}$.
+    \item If $\Psi \vdash e \To B$ then there exists
+        $B'$ s.t. $\Psi \vdash B' \le B$ and $\Psi' \vdash e \To B'$;
+    \item If $\Psi \vdash \appInf{A}{e}{C}$ and $\Psi \vdash A' \le A$,
+        then there exists $C'$ s.t. $\Psi \vdash C' \le C$ and $\Psi' \vdash \appInf{A'}{e}{C'}$.
 \end{enumerate}
@@
 \end{corollary}
-\bruno{if possible double-check against the corresponding Abella ones.}
 
@@
     \item If $\Om \vdash A' \le B'$ and $\Gm \Vdash A \le B \sto \Om \Vdash A' \le B'$
-        then $\Gm \Vdash A \le B \redto \nil$;
+        \\then $\Gm \Vdash A \le B \redto \nil$;
     \item If $\Om \vdash e \Lto A'$ and $\Gm \Vdash e \Lto A \sto \Om \Vdash e \Lto A'$
-        then $\Gm \Vdash e \Lto A \redto \nil$;
-    \item If $\Om \vdash e \To A$ and $\Gm \sto \Om$
-        then $\Gm \Vdash e \To_a 1 \le 1 \redto \nil$;
+        \\then $\Gm \Vdash e \Lto A \redto \nil$;
+    \item If $\Om \vdash e \To A$ and $\Gm \Vdash e \To_a 1 \le 1 \sto \Om \Vdash e \To_a 1 \le 1$
+        \\then $\Gm \Vdash e \To_a 1 \le 1 \redto \nil$;
     \item If $\Om \vdash \appInf{A'}{e}{C}$ and
         $\Gm \Vdash \appInfAlg{A}{e}[a][1 \to 1] \sto \Om \Vdash \appInfAlg{A'}{e}[a][1 \to 1]$
-        \\
-        then $\Gm \Vdash \appInfAlg{A}{e}[a][1 \le 1] \redto \nil$.
+        \\then $\Gm \Vdash \appInfAlg{A}{e}[a][1 \le 1] \redto \nil$.
 \end{enumerate}
 \end{corollary}
-\bruno{We discussed the form of the lemmas on slack. Please have the first premise
-  be a declarative judgement.
-  If possible also double-check against Abella.}
 
@@
                             Lemmas~\ref{lem:inv_allR}, \ref{lem:inv_chkAll}, \ref{lem:inv_chkLam}}\\\hline
-    soundness.thm       & 1069 & 70  & Soundness theorem; aux. lemmas on transfer\\\hline
+    soundness.thm       & 1086 & 70  & Soundness theorem; aux. lemmas on transfer\\\hline
     depth.thm           &  206 & 14  & The $|\cdot|_\to$ measure; Lemma~\ref{lem:prune_inst}\\\hline
@@
     decidability.thm    &  885 & 50  & Decidability theorem\\\hline
-    Total               & 7683 & 567 & (51 definitions in total)\\\hline
+    Total               & 7700 & 567 & (51 definitions in total)\\\hline
 \end{tabular}

### 2019-03-01 sections/Metatheory.tex
S4 develop, dev -> development, developers

@@
 We have chosen the Abella (v2.0.7-dev
-\footnote{A develop version is used because the dev just fixed a serious bug that
-accepts a simple proof of false, which also affects the completeness of our proof.
+\footnote{A development version is used because the developers
+just fixed a serious bug that accepts a simple proof of false,
+which also affects the completeness of our proof.
 The author compiled the scripts against commit 92829a of the GitHub repository of Abella.}

### 2019-03-01 sections/conclusion.tex
Outline of S7 Conclusion
S7 More on the outline
S7. Future work

@@
 
-\jimmy{Discuss the efficient algorithm and its equivalence}
+% \jimmy{Discuss the efficient algorithm and its equivalence}
+% The algorithm is already efficient
+
+Outline:
+\begin{itemize}
+    \item difficulties following DK's formalization
+    \item a blend of ideas from various sources DK's ordered contexts
+    \item list of judgements + some novel ideas
+    \item lessons learned: problems identified in proofs and scoping
+    \item Future work
+\end{itemize}
+
+\paragraph{Future work}
+\begin{itemize}
+    \item Scoped type variables?
+    \item Go beyond predicative instantiations
+    \item More complicated features, such as dependent type
+\end{itemize}
+
+We have not yet explore lexically-scoped type variables~\cite{jones2003lexically}
+as a practical extension to our system.
+Formalization of a different variable scoping scheme is also
+a challenging task for formal proofs via proof assistants.
+
+Moreover, the garbage collection process for type variables and
+existential variables do worth further attention:
+the inference of $\lam x$ leaves an unsolved $\al$ in the context,
+which we simply remove the variable after it is no longer refered to.
+Is it possible that we infer better types by making better use of that piece of information?
+How do we deal with, at least some ``trivial'' impredicative instantiations?
+
+Last but not least, a further expectation is that our system would
+adapt to various complicated type inference algorithms,
+such as ones for dependent types.

### 2019-03-01 sections/Metatheory.tex
Abelle subsection

--- 2019-03-01 sections/Metatheory.tex
@@
 We have chosen the Abella (v2.0.7-dev
-\footnote{A develop version is used because the dev just fixed a serious bug that
-accepts a simple proof of false, which also affects the completeness of our proof.
-The author compiled the scripts against commit 92829a of the GitHub repository of Abella.}
+\footnote{We use a development version because the developers
+just fixed a serious bug that accepts a simple proof of false,
+which also affects the completeness of one of our proofs.
+Specifically, our scipts compile against commit 92829a of Abella's GitHub repository.}
 ) proof assistant~\cite{AbellaDesc} to develop our formalization.
-Abella is designed to help with such formalizations of programming languages,
+Abella is designed to help with formalizations of programming languages,
 due to its built-in support for variable binding and
-the $\lambda$-tree syntax~\cite{miller2000abstract}, i.e. a form of HOAS.
+the $\lambda$-tree syntax~\cite{miller2000abstract}, which is a form of HOAS.
 Nominal variables, or $\nabla$-quantified variables, are used as an unlimited name supply,
 which supports explicit freshness control and substitutions.
-Although Abella lacks packages, tactics and support on modules,
+Although Abella lacks packages, tactics and support for modules,
 its higher-order unification and the ease of formalizing substitution-intensive relations
@@
 \paragraph{Statistics of the Proof}
-The proof script consists of 7683 lines of code with a total of
+The proof script consists of 7683 lines of Abella code with a total of
 51 definitions and 567 theorems.
-Figure~\ref{fig:proof_statistics} briefly summarizes the contents in each file.
+Figure~\ref{fig:proof_statistics} briefly summarizes the contents of each file.
 The files are linearly dependent due to the limitation of Abella.

### 2019-03-01 sections/conclusion.tex
more on conclusion

@@
     \item difficulties following DK's formalization
-    \item a blend of ideas from various sources DK's ordered contexts
+    \item a blend of ideas from various, including sources DK's ordered contexts
     \item list of judgements + some novel ideas
@@
 \end{itemize}
+
+Our judgment form, the worklist, is shown to be an elegant way to
+formalize higher-order type infernce algorithms.
+By mixing judgment chains with variable declarations in a single sort,
+variable scopings are easily controlled by the ordered context in a natural way.
+
+The formal proof in Abella convinces us of the correctness of our algorithm,
+which further indicates that DK's algorithm, sharing a similar solving procedure as ours,
+should be a correct one.
 

### 2019-03-01 sections/conclusion.tex
conclusion: wip

@@
 
-This is the conclusion.
+In this paper we have provided the first mechanized formalization of type
+inference for higher-ranked polymorphism. This contribution is made possible by
+a new type inference algorithm for DK's declarative type sytsem that
+incorporates several novel mechanisation-friendly ideas. Firstly, we merge the
+traditional type context with the recently proposed concept of judgment chains,
+to accurately track variable scopes and to easily propagate substitutions.
+Secondly, we use continuation-passing style to return types from the
+bidirectional synthesis mode to further tasks.
 

### 2019-03-01 sections/conclusion.tex
finish short conclusion

--- 2019-03-01 sections/conclusion.tex
@@
 
-This is the conclusion.
+In this paper we have provided the first mechanized formalization of type
+inference for higher-ranked polymorphism. This contribution is made possible by
+a new type inference algorithm for DK's declarative type sytsem that
+incorporates two novel mechanisation-friendly ideas. Firstly, we merge the
+traditional type context with the recently proposed concept of judgment chains,
+to accurately track variable scopes and to easily propagate substitutions.
+Secondly, we use continuation-passing style to return types from the
+bidirectional synthesis mode to further tasks.
 
-% \jimmy{Discuss the efficient algorithm and its equivalence}
-% The algorithm is already efficient
+We leave extending our algorithm with elaboration to future work, as well as
+investigating whether the problems we have found in DK's manual proofs for
+their algorithm can be addressed.
 
-Outline:
-\begin{itemize}
-    \item difficulties following DK's formalization
-    \item a blend of ideas from various sources DK's ordered contexts
-    \item list of judgements + some novel ideas
-    \item lessons learned: problems identified in proofs and scoping
-    \item Future work
-\end{itemize}
-
-\paragraph{Future work}
-\begin{itemize}
-    \item Scoped type variables?
-    \item Go beyond predicative instantiations
-    \item More complicated features, such as dependent type
-\end{itemize}
-
-We have not yet explore lexically-scoped type variables~\cite{jones2003lexically}
-as a practical extension to our system.
-Formalization of a different variable scoping scheme is also
-a challenging task for formal proofs via proof assistants.
-
-Moreover, the garbage collection process for type variables and
-existential variables do worth further attention:
-the inference of $\lam x$ leaves an unsolved $\al$ in the context,
-which we simply remove the variable after it is no longer refered to.
-Is it possible that we infer better types by making better use of that piece of information?
-How do we deal with, at least some ``trivial'' impredicative instantiations?
-
-Last but not least, a further expectation is that our system would
-adapt to various complicated type inference algorithms,
-such as ones for dependent types.
+% % \jimmy{Discuss the efficient algorithm and its equivalence}
+% % The algorithm is already efficient
+% 
+% Outline:
+% \begin{itemize}
+%     \item difficulties following DK's formalization
+%     \item a blend of ideas from various, including sources DK's ordered contexts
+%     \item list of judgements + some novel ideas
+%     \item lessons learned: problems identified in proofs and scoping
+%     \item Future work
+% \end{itemize}
+% 
+% Our judgment form, the worklist, is shown to be an elegant way to
+% formalize higher-order type infernce algorithms.
+% By mixing judgment chains with variable declarations in a single sort,
+% variable scopings are easily controlled by the ordered context in a natural way.
+% 
+% The formal proof in Abella convinces us of the correctness of our algorithm,
+% which further indicates that DK's algorithm, sharing a similar solving procedure as ours,
+% should be a correct one.
+% 
+% \paragraph{Future work}
+% \begin{itemize}
+%     \item Scoped type variables?
+%     \item Go beyond predicative instantiations
+%     \item More complicated features, such as dependent type
+% \end{itemize}
+% 
+% We have not yet explore lexically-scoped type variables~\cite{jones2003lexically}
+% as a practical extension to our system.
+% Formalization of a different variable scoping scheme is also
+% a challenging task for formal proofs via proof assistants.
+% 
+% Moreover, the garbage collection process for type variables and
+% existential variables do worth further attention:
+% the inference of $\lam x$ leaves an unsolved $\al$ in the context,
+% which we simply remove the variable after it is no longer refered to.
+% Is it possible that we infer better types by making better use of that piece of information?
+% How do we deal with, at least some ``trivial'' impredicative instantiations?
+% 
+% Last but not least, a further expectation is that our system would
+% adapt to various complicated type inference algorithms,
+% such as ones for dependent types.

### 2019-03-01 sections/Related.tex
S6 RW let arguments go first

@@
 complete with respect to their declarative specification.
+
+\jimmy{Let arguments go first by Ningning}
+In a more recent work, \citet{xie2018letarguments} proposed a variant of
+bidirectional type inference system where type information flows from arguments to
+functions with an additional \emph{application} mode,
+which allows more higher-order typed programs to be inferred without additional annotations.
+Following the new mode, the let-generalization of the Hindley-Milner system
+is well supported as a syntactic sugar.
+Unfortunately, no formal proofs are given for the correctness of its algorithmic system.
 

### 2019-03-01 main.tex
Section 3
S4.6 Abella and Proof Statistics
S7. Future work
Abstract follwoing Jimmy suggestion

@@
 %\documentclass[acmsmall]{acmart}\settopmatter{}
+
+%% Remove almost every footnotes and page numbers, for grammar check use.
+%\documentclass[acmsmall,anonymous]{acmart}\settopmatter{printfolios=true,printccs=false,printacmref=false}
+%\settopmatter{printacmref=false} % Removes citation information below abstract
+%\renewcommand\footnotetextcopyrightpermission[1]{} % removes footnote with conference information in first column
+%\pagestyle{plain} % removes running headers
+%\pagenumbering{gobble}
 
@@
 \usepackage{enumerate}
+%\usepackage{enumitem}
 
 \usepackage{multirow}
+\usepackage{makecell}
 \usepackage{color}
@@
 of DK's declarative type system with a novel algorithmic system. We have
-several reasons to use a new algorithm. Firstly, our new algorithm
+a few reasons to use a new algorithm. Firstly, our new algorithm
 employs \emph{worklist judgements}, which allow precise capture of the
@@
 theorem prover. Secondly, while DK's original formalization comes with very
-well-written manual proofs, there are several details missing,
+well-written manual proofs, there are several details missing and some
+incorrect proofs,
 which complicate the task of writing a formal proof. 
-Moreover, some proofs/lemmas are wrong and it is
-not clear how to replace these by other valid ones. 
 Despite the use of a different algorithm we prove the

### 2019-03-01 sections/Overview.tex
proofreading S3

@@
 If $[\tau/a]A$ is a function type, Rule $\mathtt{Decl{\to}App}$ proceeds;
-if $[\tau/a]A$ has another universal quantifier, Rule $\mathtt{Decl{\to}I{\To}}$ is recursively applied.
+if $[\tau/a]A$ is another universal quantified type,
+Rule $\mathtt{Decl{\to}I{\To}}$ is recursively applied.
 
@@
 both $\mathtt{Decl\forall I}$ and $\mathtt{Decl\to I}$.
-However, we could argue that more specific rules are always the best choices,
+However, we argue that more specific rules are always the best choices,
 i.e. $\mathtt{Decl\forall I}$ and $\mathtt{Decl\to I}$ should have
 higher priority to $\mathtt{DeclSub}$. We will come back to this
-topic in Section~\ref{sec:metatheory}.
+topic in Section~\ref{sec:metatheory:non-overlapping}.
 \begin{comment}
@@
 \text{Algorithmic Contexts}\qquad&\Gamma,\Delta,\Theta &::=&\quad \nil \mid
-\Gamma, a \mid \Gamma, \al \mid \Gamma, \al = \tau \mid
+\Gamma, a \mid \Gamma, x:A \mid \Gamma, \al \mid \Gamma, \al = \tau \mid
 \Gamma, \blacktriangleright_{\al}
@@
 ``solved'', its entry in the context is replaced by the assignment
-$\al = \tau$. With the notation $[\Gamma]$ an algorithmic context
-$\Gamma$ denotes the substitution that replaces all its solved existential type
-variables with their solution.
+$\al = \tau$.
+A context applications on a type, denoted by $[\Gamma]A$,
+substitutes all solved existential type variables in $\Gamma$
+with their solutions on the type $A$.
 
@@
 \[
-\inferrule*
+\inferrule*[right=$\mathtt{DK\_Sub}$]
   {\Gamma \vdash e \Rightarrow A \dashv \Theta \\ 
@@
 \[
-\inferrule*
+\inferrule*[right=$\mathtt{DK\_{\le}\forall L}$]
   {\Gamma,\blacktriangleright_{\al}, \al \vdash [\al/a]A \le B \dashv \Delta,\blacktriangleright_{\al},\Theta}
@@
 
-Due to the complication of its scoping control method,
-it is hard to argue the preciseness of scoping for each variable when designing algorithms.
-Actually we found a scoping issue related to the subsumption rule,
-which might cause existential variables to leak across judgments.
-A detailed discussion on this issue with a possible fix is in Section~\ref{sec:discussion:scoping}.
+%Due to the complication of its scoping control method,
+%it is hard to argue the preciseness of scoping for each variable when designing algorithms.
 
@@
 needed in two places to demarcate the end of an existential variable's scope.
-These are compelling reasons to look for a simpler algorithm that is easier to
+Actually we found a scoping issue related to the subsumption rule $\mathtt{DK\_Sub}$,
+which might cause existential variables to leak across judgments.
+A detailed discussion on this issue with a possible fix is in Section~\ref{sec:discussion:scoping}.
+
+The complications of scoping and propagation are compelling reasons
+to look for a simpler algorithm that is easier to
 reason about in a mechanised setting.

### 2019-03-01 sections/algorithm.tex
remove 2 of my comments
proofreading S3

@@
 only the objects that come after a variable's binding in the worklist can refer to it.
-For example, $[a,x:a,x \Lto a]$ is a valid worklist, but $[\underline{x} \Lto a, x:\underline{a},a]$ is not
+For example, $[\nil, a, x:a \Vdash x \Lto a]$ is a valid worklist,
+but $[\nil \Vdash \underline{x} \Lto \underline{a}, x:\underline{a}, a]$ is not
 (the underlined symbols refer to out-of-scope variables).
@@
 judgment and pushes two back to the worklist.  Rule 8 differs from
-declarative Rule ${\le}{\forall}L$ by introducing an existential
+declarative Rule $\mathtt{{\le}{\forall}L}$ by introducing an existential
 variable $\al$ instead of guessing the monotype $\tau$
@@
 variable $\al$ that is only visible to the judgment $[\al/a]A \le B$.
-Reduction continues until all the subtyping judgments after $\al$
+Reduction continues until all the subtyping judgments in front of $\al$
 are satisfied.  Finally we can safely remove $\al$ since no occurrence
@@
 variable at the perfect time: it leaves the environment immediately
-after being unreferenced.
+after being unreferenced completely for sure.
 
-\jimmy{New example on Feb 27}
 \paragraph{Example} The derivation of the subtyping judgment
@@
 Now the existential variable $\al$, introduced before but still unsolved,
-is at the top of the worklist and Rule 2 garbage collects it.
+is at the top of the worklist and Rule 2 garbage-collects it.
 The process is carefully designed within the algorithmic rules:
-when $\al$ is introduced,
-we foresee the recycling of $\al$ after all the judgments requiring $\al$ have been processed.
+when $\al$ is introduced earlier by Rule 8,
+we foresee the recycling of $\al$ after all the judgments (potentially)
+requiring $\al$ have been processed.
 Finally Rule 4 reduces one of the base cases and finishes the subtyping derivation.
@@
 representing a monotype \emph{function} as in the declarative system
-(it must be a function type, since the expression is a function).
+(it must be a function type, since the expression $\lam e$ is a function).
 % \bruno{Rule 21
@@
 must end up equating $\bt[1] \to \bt[2]$ with $\al$ and thus have the same effect as Rule 21, but in a more roundabout
-fashion. DK's original algorithm also avoids this detour and contains a rule similar to our Rule 21.
+fashion.
+
+In comparison,
+DK's original algorithmic subsumption rule have no restriction on the expression $e$,
+and they do not have a rule that explicitly handles the case $\lam e \Lto \al$.
+Therefore their only way to check a lambda function against an existential variable
+is by combining the inferenced type of the lambda function with the subsumption rule.
 % A combination of Rules $18'$ (Rule 18 without the restriction on $e$) and 25 and further derivations

### 2019-03-01 sections/algorithm.tex
S3 Algorithm 5. missing explainations for Rules 24 and 26

@@
 
-Rules 22-26 deal with type inference judgments.
-Inference judgment chains are written in continuation-passing-style.
+Rules 22-26 deal with type inference judgments, written in continuation-passing-style.
 When an inference judgment succeeds with type $A$,
-the algorithm continues to work on the inner-chain $\jg$ by assigning $A$ to its placeholder variable $a$.
+the algorithm continues to work on the inner-chain $\jg$ by
+assigning $A$ to its placeholder variable $a$.
 Rule 23 infers an annotated expression by changing into checking mode,
 therefore another judgment chain is created.
+Rule 24 is a base case,
+where the unit type $1$ is inferred and thus passed to its child judgment chain.
+Rule 26 infers the type of an application by
+firstly inferring the type of the function $e_1$,
+and then leaving the rest work to an application inference judgment,
+which passes $a$, representing the return type of the application,
+to the reminder of the judgment chain $\jg$.
 

### 2019-03-01 sections/algorithm.tex
have -> has

@@
 In comparison,
-DK's original algorithmic subsumption rule have no restriction on the expression $e$,
+DK's original algorithmic subsumption rule has no restriction on the expression $e$,
 and they do not have a rule that explicitly handles the case $\lam e \Lto \al$.

### 2019-03-01 sections/Metatheory.tex
Several refactors

@@
 \begin{lemma}[Insert]\label{lem:insert}
-If $\Gm_L\vdash \tau$ and $\Gm_L, [\tau/\al]\Gm_R \sto \Om$
+If $\Gm_L, [\tau/\al]\Gm_R \sto \Om$ and $\Gm_L\vdash \tau$
 , then $\Gm_L, \al, \Gm_R \sto \Om$.
@@
 If $\Gm_L, \al, \Gm_R \sto \Om$
-, then $\exists \tau$ s.t. $\Gm_L\vdash\tau$ and $\Gm_L, [\tau/\al]\Gm_R \sto \Om$.
+, then there exists $\tau$ s.t. $\Gm_L\vdash\tau$ and $\Gm_L, [\tau/\al]\Gm_R \sto \Om$.
 \end{lemma}
@@
 \footnote{In the proof script we do not use the erasure function,
-for the declarative system and well-formedness judgments automatically fits
+for the declarative system and well-formedness judgments automatically fit
 the non-erased declarative worklist just as declarative contexts.}.
@@
 The resulting declarative system has no overlapping rules
-and more closely resembles the algorithmic system, which contains
-the same shape constraints.
+and more closely resembles the algorithmic system,
+which contains constraints of the same shape.
 
@@
 \begin{theorem}[Soundness]
-If \emph{wf }$\Gm$ and $\Gm \redto \nil$, then $\exists \Om$ s.t. $\Gm\sto\Om$ and $\Om\redto\nil$.
+If \emph{wf }$\Gm$ and $\Gm \redto \nil$,
+then there exists $\Om$ s.t. $\Gm\sto\Om$ and $\Om\redto\nil$.
 \end{theorem}
@@
 (shown in Figure~\ref{fig:decl:worklist}),
-the induction always analyses the first judgment until it is satisfied.
+the induction always analyses the first judgment by a small step.
 As the algorithmic system introduces existential variables,
@@
 but transferred to a monotype for the declarative system,
-such as Rules 6, 10, 11, 12-17, 21 and 29.
+such as Rules 6, 8, 10, 11, 12-17, 21, 25, 27 and 29.
+
 Algorithmic Rules 10 and 11 require special treatment.
@@
 count for 2, inference judgments for 1 and function inference judgments for 3.
-The next two measures, $|\cdot|_\forall$ and $|\cdot|_\to$, count the total number of
+Other two measures, $|\cdot|_\forall$ and $|\cdot|_\to$, count the total number of
 universal quantifiers and function types, respectively. Finally,
-$|\cdot|_{\al}$ counts the number of existential variables in the worklist.
+$|\cdot|_{\al}$ counts the number of existential variables in the worklist,
+and $|\cdot|$ is simply the length of the worklist.
 
@@
 update judgement $\Gm \jExt \Gm'$ defined in Figure~\ref{fig:worklist_ext}.
-For instance, existential variable instantiation, 
-i.e. $\Gm_L,\al,\Gm_R \jExt \Gm_L, \al[1], \al[2], [\al[1] \to \al[2] / \al]\Gm_R$,
+For instance, an existential variable instantiation, 
+$\Gm_L,\al,\Gm_R \jExt \Gm_L, \al[1], \al[2], [\al[1] \to \al[2] / \al]\Gm_R$,
 can be derived as a combination of the three rules that define the update relation.
@@
 The good news is that worklist update does not affect the three main worklist measures:
-\newcommand{\equivGm}[1]{|\Gm|_{#1} = |\Gm'|_{#1}}\label{lemma:inst:invariant}
-\begin{lemma}[Measure Invariants of Worklist Extension]
+\newcommand{\equivGm}[1]{|\Gm|_{#1} = |\Gm'|_{#1}}
+\begin{lemma}[Measure Invariants of Worklist Extension]\label{lemma:inst:invariant}
 If $\Gm\jExt\Gm'$ then $\equivGm{e}$ and $\equivGm\Leftrightarrow$ and $\equivGm\forall$.
@@
 \begin{lemma}[Instantiation Decidability]\label{lemma:inst:decidable}
-For any well-formed algorithmic worklist $(\Gm, \Gm_i)$,
+For any well-formed algorithmic worklist $(\Gm, \Gm_i)$:
 \begin{enumerate}[1)]
-    \item When $|\Gm_i|_\forall = 0$, $(\Gm, \Gm_i) \redto \Gm'$,
-        where $|\Gm'|_{\al} = |\Gm|_{\al} - |\Gm_i|$ and $\Gm\jExt\Gm'$.
-    \item When $|\Gm_i|_\forall > 0$, $(\Gm, \Gm_i) \redto (\Gm', \Gm_\le)$,
-        where $|\Gm_\le|_\forall = |\Gm_i|_\forall - 1$ and $\Gm\jExt\Gm'$.
+    \item If $|\Gm_i|_\forall = 0$,
+        then there exists $\Gm'$\\
+        s.t. $(\Gm, \Gm_i) \redto \Gm'$ and $|\Gm'|_{\al} = |\Gm|_{\al} - |\Gm_i|$ and $\Gm\jExt\Gm'$.
+    \item If $|\Gm_i|_\forall > 0$,
+        then there exist $\Gm', \Gm_\le$\\
+        s.t. $(\Gm, \Gm_i) \redto (\Gm', \Gm_\le)$ and $|\Gm_\le|_\forall = |\Gm_i|_\forall - 1$ and $\Gm\jExt\Gm'$.
 \end{enumerate}
@@
 make use of Lemma~\ref{lemma:inst:decidable} to show that $|\Gm|_{\al}$ or $|\Gm|_\forall$ decreases
-when that instantiation judgment is consumed. Moreover, Lemma~\ref{lemma:inst:invariant} establishes
-that no higher-priority measure component increases. Hence, in the exceptional case we have an overall measure decrease too.
+when that instantiation judgment is consumed or partially reduced.
+Moreover, Lemma~\ref{lemma:inst:invariant} establishes
+that no higher-priority measure component increases.
+Hence, in the exceptional case we have an overall measure decrease too.
 
@@
     alg.thm             &  655 & 72  & Basic algorithmic properties\\\hline
-    trans.thm           &  628 & 53  & \makecell[l]{Transfer and its basic properties \\
+    trans.thm           &  628 & 53  & \makecell[l]{Worklist instantiation with basic properties \\
                             Lemmas~\ref{lem:insert}, \ref{lem:extract}}\\\hline
@@
     decidability.thm    &  885 & 50  & Decidability theorem\\\hline
-    Total               & 7700 & 567 & (51 definitions in total)\\\hline
+    \hline\emph{Total}               & 7700 & 567 & (51 definitions in total)\\\hline
 \end{tabular}

### 2019-03-01 sections/algorithm.tex
Several refactors

@@
 % \noalign{\jimmy{\text{What do the markers do: $\G \vdash,a \vdash,b\vdash \j_1;\j_2;\j_3$ means $\G,a,b\vdash \j_1 \land \G,a\vdash \j_1,\j_2 \land \G\vdash \j_3$}}} \\[2mm]
-\text{Algorithmic types}\qquad&A, B, C &::=& \quad 1 \mid a \mid \forall a. A \mid A\to B \mid \al % \quad 1 \mid a \mid \al \mid \forall a. A \mid A\to B
+\text{Algorithmic types}\qquad&A, B, C &::=& \quad 1 \mid a \mid \all A \mid A\to B \mid \al % \quad 1 \mid a \mid \al \mid \forall a. A \mid A\to B
 \\
@@
 which passes $a$, representing the return type of the application,
-to the reminder of the judgment chain $\jg$.
+to the remainder of the judgment chain $\jg$.
 

### 2019-03-01 main.tex
Abstract minor fix

@@
 Modern functional programming languages, such as Haskell or OCaml,
-use sophisticated forms of type inference. While an
+use sophisticated forms of type inference. While being an
 important topic in the Programming Languages research, there is little

### 2019-03-01 sections/Discussion.tex
Several refactors
Minor word fixes; forall's and lambda's are in good shape

@@
 and Rule 27 collects one existential variable similarly:
-\[\Gm \Vdash \lambda x. e \To_a \jg \rrule{25}
+\[\Gm \Vdash \lam e \To_a \jg \rrule{25}
 \Gm,\al,\bt \Vdash [\al\to\bt/a]\jg, x:\al \Vdash e\Lto \bt\]
-\[\Gm \Vdash \appInfAlg{\forall a. A}{e} \rrule{27} \Gm,\al \Vdash \appInfAlg{[\al/a]A}{e}\]
+\[\Gm \Vdash \appInfAlg{\all A}{e} \rrule{27} \Gm,\al \Vdash \appInfAlg{[\al/a]A}{e}\]
 It seems impossible to achieve a similar outcome in DK's system by only modifying these two rules.
@@
 
-Existential variables need special treatment, since they may be solved at any point,
+Moreover, existential variables need special treatment, since they may be solved at any point,
 or be recycled if not solved within its scope.

### 2019-03-01 sections/Overview.tex
Minor word fixes; forall's and lambda's are in good shape

@@
 \\
-\text{Types}\qquad&A, B, C &::=&\quad 1 \mid a \mid \forall a. A \mid A\to B\\
+\text{Types}\qquad&A, B, C &::=&\quad 1 \mid a \mid \all A \mid A\to B\\
 \text{Monotypes}\qquad&\tau,\sigma &::=&\quad 1 \mid a \mid \tau\to \sigma
@@
 A declarative type $A$ is either the unit type $1$, a type variable $a$,
-a universal quantification $\forall a. A$ or a function type $A \to B$.
+a universal quantification $\all A$ or a function type $A \to B$.
 Nested universal quantifiers are allowed for types,
@@
     {\Psi\vdash \tau \quad \Psi\vdash [\tau/a] A \le B}
-    {\Psi\vdash \forall a. A \le B}
+    {\Psi\vdash \all A \le B}
 \qquad
@@
     {\Psi, a\vdash A\le B}
-    {\Psi\vdash A \le \forall a. B}
+    {\Psi\vdash A \le \all[b]B}
 \end{gather*}
@@
 \paragraph{Overlapping Rules}
-Some of the declarative rules overlaps with each other.
+Some of the declarative rules overlap with each other.
 Declarative subtyping rules $\mathtt{{\le}\forall L}$ and $\mathtt{{\le}\forall R}$
@@
   {\Gamma,\blacktriangleright_{\al}, \al \vdash [\al/a]A \le B \dashv \Delta,\blacktriangleright_{\al},\Theta}
-  {\Gamma \vdash \forall a. A \le B \dashv \Delta}
+  {\Gamma \vdash \all A \le B \dashv \Delta}
 \]
@@
 
-In the above unification problem, all variables are implicitly in the same
+In the above unification problem, all variables are implicitly bound in the same
 global scope. Recent unification algorithms for dependently-typed languages
@@
 
-Our new algorithmic type system addresses builds on the work above, but
+Our new algorithmic type system builds on the work above, but
 addresses the open problems.

### 2019-03-01 sections/algorithm.tex
Minor word fixes; forall's and lambda's are in good shape

@@
 \Gm \Vdash A_1\to A_2 \le B_1\to B_2 &\algrule \Gm \Vdash A_2 \le B_2 \Vdash B_1\le A_1\\
-\Gm \Vdash \forall a. A\le B &\algrule \Gm,\al \Vdash [\al/a]A\le B \quad\text{when } B \neq \all B'\\
-\Gm \Vdash A\le \forall b. B &\algrule \Gm,b \Vdash A\le B
+\Gm \Vdash \all A\le B &\algrule \Gm,\al \Vdash [\al/a]A\le B \quad\text{when } B \neq \all B'\\
+\Gm \Vdash A\le \all[b]B &\algrule \Gm,b \Vdash A\le B
 \\[2mm]
@@
 % \Gm \Vdash () \Lto 1 &\algrule \Gm\\
-\Gm \Vdash e\Lto \forall a. A &\algrule \Gm,a \Vdash e\Lto A\\
-\Gm \Vdash \lambda x. e \Lto A\to B &\algrule \Gm, x:A  \Vdash e \Lto B\\
-\Gm[\al] \Vdash \lambda x. e \Lto \al &\algrule [\al[1]\to \al[2] / \al](\Gm[\al[1],\al[2]], x:\al[1] \Vdash e \Lto \al[2])
+\Gm \Vdash e\Lto \all A &\algrule \Gm,a \Vdash e\Lto A\\
+\Gm \Vdash \lam e \Lto A\to B &\algrule \Gm, x:A  \Vdash e \Lto B\\
+\Gm[\al] \Vdash \lam e \Lto \al &\algrule [\al[1]\to \al[2] / \al](\Gm[\al[1],\al[2]], x:\al[1] \Vdash e \Lto \al[2])
 % \quad\text{\jimmy{Additional}}
@@
 \Gm \Vdash ()\To_a \jg &\algrule \Gm \Vdash [1/a]\jg\\
-\Gm \Vdash \lambda x. e \To_a \jg &\algrule
+\Gm \Vdash \lam e \To_a \jg &\algrule
     \Gm,\al,\bt \Vdash [\al\to\bt/a]\jg, x:\al \Vdash e\Lto \bt\\
@@
 \\[2mm]
-\Gm \Vdash \appInfAlg{\forall a. A}{e} &\algrule \Gm,\al \Vdash \appInfAlg{[\al/a]A}{e}\\
+\Gm \Vdash \appInfAlg{\all A}{e} &\algrule \Gm,\al \Vdash \appInfAlg{[\al/a]A}{e}\\
 \Gm \Vdash \appInfAlg{A\to C}{e} &\algrule \Gm \Vdash [C/a]\jg \Vdash e \Lto A\\
@@
 
-Let us explain the rules in more detail:
+Let us explain the rules in more details:
 Rules 1-3 pop a variable declaration. Essentially this is garbage: variables that are
@@
 The rule introduces an existential variable $\al$ to the context (even though 
-the type $\forall a. 1 \to 1$ does not actually refer to the quantified type
+the type $\all 1 \to 1$ does not actually refer to the quantified type
 variable $a$).

### 2019-03-01 sections/introduction.tex
hpoly parathesis
Minor word fixes; forall's and lambda's are in good shape

@@
 An important challenge is that full type inference for higher-ranked
-polymorphism is know to be undecidable~\cite{wells1999typability}. Therefore some type
+polymorphism is known to be undecidable~\cite{wells1999typability}. Therefore some type
 annotations are necessary to guide type inference. In response to this 
@@
 \begin{verbatim}
-    hpoly = (\f :: forall a. a -> a) -> (f 1, f 'c')
+    hpoly = \(f :: forall a. a -> a) -> (f 1, f 'c')
 \end{verbatim}
@@
 \emph{compiler correctness}~\cite{leroy2012compcert}, \emph{correctness of
-  optimizations}~\cite{Bertot04}, \emph{program analisys}~\cite{Chang2006}
+  optimizations}~\cite{Bertot04}, \emph{program analysis}~\cite{Chang2006}
 or proofs involving \emph{logical relations}~\cite{abel2018}. The
@@
 \emph{decidable}.
-We present a mechanical formalization in the Abella theorem prover for
+We present a mechanical formalization in the Abella theorem prover~\cite{AbellaDesc} for
 DK's declarative type system using a different algorithm. While our

### 2019-03-01 main.tex
All "judgement" -> "judgment"

@@
 a few reasons to use a new algorithm. Firstly, our new algorithm
-employs \emph{worklist judgements}, which allow precise capture of the
+employs \emph{worklist judgments}, which allow precise capture of the
 scope of variables and simplify the formalization of scoping in a

### 2019-03-01 sections/Discussion.tex
All "judgement" -> "judgment"

@@
 A nice feature of our worklists is that, simply by interleaving variable declarations and
-judgement chains, they make the scope of variables
+judgment chains, they make the scope of variables
 precise.  DK's algorithm deals with garbage collecting variables in a
@@
 Here we employ the markers in a way they were originally not intended for.
-We create a dummy fresh existential $\al$ and add a marker to the input context of the inference judgement.
-After the subtyping judgement is processed we look for the marker and drop everything afterwards.
+We create a dummy fresh existential $\al$ and add a marker to the input context of the inference judgment.
+After the subtyping judgment is processed we look for the marker and drop everything afterwards.
 We pick this rule because it is the only one where a checking judgment calls an inference judgment.

### 2019-03-01 sections/Metatheory.tex
Minor word fixes; forall's and lambda's are in good shape
All "judgement" -> "judgment"

@@
 and thus the algorithmic worklist has become a declarative one.
-In order to maintain well-scopedness
+In order to maintain well-scopedness,
 the substitution is applied to all the judgments and term variable bindings in the scope of $\al$.
@@
 Figure~\ref{fig:decl:worklist} defines a relation $\Om \rto \Om'$,
-which transfers all judgements in the declarative worklists to the
-declarative type system. This relation checks that every judgement entry in the worklist
-holds using a corresponding conventional declarative judgement. 
-The typing contexts of declarative judgements are recovered using an
+which transfers all judgments in the declarative worklists to the
+declarative type system. This relation checks that every judgment entry in the worklist
+holds using a corresponding conventional declarative judgment. 
+The typing contexts of declarative judgments are recovered using an
 auxiliary erasure function $\|\Om\|$
@@
 For any worklist $\Gm$ that reduces successfully,
-there is a valid instantiation $\Om$ that transfers all judgements successfully to the declarative system.
+there is a valid instantiation $\Om$ that transfers all judgments successfully to the declarative system.
 \begin{theorem}[Soundness]
@@
 and finishes the corresponding case. Some immediate corollaries which
-show the soundness for specific judgement forms are:
+show the soundness for specific judgment forms are:
 
@@
 $|\cdot|_e$ and $|\cdot|_\Leftrightarrow$, measure the total size of terms
-and the total difficulty of judgments, respectively. In the latter, check judgements
+and the total difficulty of judgments, respectively. In the latter, check judgments
 count for 2, inference judgments for 1 and function inference judgments for 3.
@@
 smaller problems of the form $\al \le A$ and $A \le \al$ which we call
-\emph{instantiation judgements}.
+\emph{instantiation judgments}.
 
@@
 by solving existing existentials and introducing new ones. This worklist update is captured in the 
-update judgement $\Gm \jExt \Gm'$ defined in Figure~\ref{fig:worklist_ext}.
+update judgment $\Gm \jExt \Gm'$ defined in Figure~\ref{fig:worklist_ext}.
 For instance, an existential variable instantiation, 

### 2019-03-01 sections/Overview.tex
All "judgement" -> "judgment"

@@
 Finally we introduce the novelties of our new algorithm.
-In particular the support for inference judgements in
-judgement worklists, and a new form of judgement worklist
+In particular the support for inference judgments in
+judgment worklists, and a new form of judgment worklist
 that the unifies \emph{ordered contexts} and the worklists themselves.
@@
 % records any new variable instantations; to propagate these instantiations to
-% the remaining judgements, their predecessor's output context---which is their
+% the remaining judgments, their predecessor's output context---which is their
 % input context---is applied to them as a substitution.
@@
 subtyping judgment. This approach overcomes the problem's in DK's formulation
-by using a \emph{worklist}-based judgement of the form $$\Gamma \vdash \Omega$$
+by using a \emph{worklist}-based judgment of the form $$\Gamma \vdash \Omega$$
 where $\Omega$ is a worklist (or sequence) of subtyping problems of the
-form $A \leq B$.  The judgement is defined by case analysis on the first
+form $A \leq B$.  The judgment is defined by case analysis on the first
 element of $\Omega$ and recursively processes the worklist until it is empty.
@@
 \paragraph{No Inference Judgments} 
-Secondly and more importantly, the approach only deals with a judgement for
+Secondly and more importantly, the approach only deals with a judgment for
 \emph{checking} whether one type is the subtype of another. While this may not
@@
 $$\begin{aligned}
-\text{Algorithmic judgement chain}\qquad&\jg &::=&\quad A \le B \mid e\Lto A \mid e\To_{a} \jg \mid \appInfAlg{A}{e}\\
+\text{Algorithmic judgment chain}\qquad&\jg &::=&\quad A \le B \mid e\Lto A \mid e\To_{a} \jg \mid \appInfAlg{A}{e}\\
 \text{Algorithmic worklist}\qquad&\Gm &::=&\quad \nil \mid \Gm, a \mid \Gm, \al \mid \Gm, x: A \mid \Gm \Vdash \jg\\
@@
 
-\paragraph{Inference Judgements}
-To express the DK's inference judgements, we use a novel form of work entries in
-the worklist: our algorithmic judgement chains $\omega$. In its simplest form,
-such a judgement chain is just a check, like a subtyping check $A \leq B$ or a
+\paragraph{Inference Judgments}
+To express the DK's inference judgments, we use a novel form of work entries in
+the worklist: our algorithmic judgment chains $\omega$. In its simplest form,
+such a judgment chain is just a check, like a subtyping check $A \leq B$ or a
 term typecheck $e \Leftarrow A$. 
@@
 
-% Importantly we now deal with inference judgement.
+% Importantly we now deal with inference judgment.

### 2019-03-01 sections/Related.tex
All "judgement" -> "judgment"

@@
 
-\paragraph{Lists of Judgements in Unification}
+\paragraph{Lists of Judgments in Unification}
 Some work~\cite{Reed2009,Abel2011higher} adopts a similar idea to this paper

### 2019-03-01 sections/algorithm.tex
All "judgement" -> "judgment"

@@
 passing style} to enable the transfer of inferred information across
-judgements in a worklist.
+judgments in a worklist.
 
@@
 
-Figure~\ref{fig:alg:syntax} shows the syntax and well-formedness judgements used by the algorithm.
+Figure~\ref{fig:alg:syntax} shows the syntax and well-formedness judgments used by the algorithm.
 
@@
 %\text{Context}\qquad&\Psi &::=&\quad \nil \mid \Psi, a \mid \Psi, x:A' \color{red} \\[2mm]
-% \text{Declarative Judgements}\qquad&\jo &::=&\quad \nil \mid \P\vdash A' \le B' \jc \jo\\
+% \text{Declarative Judgments}\qquad&\jo &::=&\quad \nil \mid \P\vdash A' \le B' \jc \jo\\
 % &&&\quad \mid \P\vdash e\Leftarrow A' \jc \jo \mid \P\vdash e\To A' \jc \jo \mid \P\vdash A' \bullet e\TTo C' \jc \jo\\
@@
 % \\[2mm]
-\text{Algorithmic judgement chain}\qquad&\jg &::=&\quad A \le B \mid e\Lto A \mid e\To_{a} \jg \mid \appInfAlg{A}{e}
+\text{Algorithmic judgment chain}\qquad&\jg &::=&\quad A \le B \mid e\Lto A \mid e\To_{a} \jg \mid \appInfAlg{A}{e}
 \\
@@
 since their result is only success or failure. However both inference and
-application inference return a type that is used in subsequent judgements. We use a
+application inference return a type that is used in subsequent judgments. We use a
 continuation-passing-style encoding to accomplish this. For example, the judgment
@@
 subtype of $B$. The \emph{unknown} type of $e$ is represented by a
-type variable $a$, which is used as a placeholder in the second judgement to denote the 
+type variable $a$, which is used as a placeholder in the second judgment to denote the 
 type of $e$.
@@
 Rules 1-3 pop a variable declaration. Essentially this is garbage: variables that are
-out of scope for the remaining judgements in the worklist.
-All other rules concern a judgement at the front of the worklist. Logically we
+out of scope for the remaining judgments in the worklist.
+All other rules concern a judgment at the front of the worklist. Logically we
 can discern 6 groups of rules.
@@
 
-\paragraph{\bf 4. Checking judgements.}
+\paragraph{\bf 4. Checking judgments.}
 Rules 18-21 deal with checking judgments.
@@
 and the garbage will be collected at the correct time.
-By making use of the continuation-passing-style judgement chain,
+By making use of the continuation-passing-style judgment chain,
 inner judgments always share the context with their parent judgment.
@@
 
-\paragraph{\bf 6. Application inference judgements}
+\paragraph{\bf 6. Application inference judgments}
 Finally, Rules 27-29 deal with application inference judgments.

### 2019-03-01 sections/introduction.tex
All "judgement" -> "judgment"

@@
 typically do not rely simply on local environments but instead
-propagate information across judgements. While local environments are
+propagate information across judgments. While local environments are
 well-studied in mechanical formalizations, there is little work on how
@@
 to adapt the idea of worklist judgments to
-\emph{inference judgements}, which are not needed for polymorphic
+\emph{inference judgments}, which are not needed for polymorphic
 subtyping, but are necessary for type-inference.  The idea is to use a \emph{continuation
 passing style} to enable the transfer of inferred information across
-judgements. A further refinement to the idea of worklist judgements is
+judgments. A further refinement to the idea of worklist judgments is
 the \emph{unification between ordered
@@
 enables precise scope tracking of free variables in
-judgements. Furthermore it avoids the duplication of context
+judgments. Furthermore it avoids the duplication of context
 information across judgments in worklists that occurs in other
@@
   use a continuation passing style to enable the transfer of inferred information across
-  judgements.
+  judgments.
 
@@
   between ordered contexts and worklists. This enables precise scope tracking
-  of variables in judgements, and avoids the duplication of context information across
+  of variables in judgments, and avoids the duplication of context information across
   judgments in worklists.
@@
 \jimmy{Notes @20190211 4 points of novalty:\\
-1) Dealing with inference judgements and CPS-style chains\\
+1) Dealing with inference judgments and CPS-style chains\\
 2) The form of the judgment itself with a single shared context\\

### 2019-03-01 sections/Metatheory.tex
proof script statistics minor update

@@
     alg.thm             &  655 & 72  & Basic algorithmic properties\\\hline
-    trans.thm           &  628 & 53  & \makecell[l]{Worklist instantiation with basic properties \\
+    trans.thm           &  628 & 53  & \makecell[l]{Worklist instantiation and declarative transfer\\
                             Lemmas~\ref{lem:insert}, \ref{lem:extract}}\\\hline
     declTyping.thm      &  944 & 72  & \makecell[l]{Non-overlapping declarative system \\
-                            Lemmas~\ref{lem:inv_allR}, \ref{lem:inv_chkAll}, \ref{lem:inv_chkLam}}\\\hline
+                            Lemmas~\ref{lem:inv_allR}, \ref{lem:inv_chkAll},
+                                \ref{lem:inv_chkLam}, \ref{lem:subsumption}}\\\hline
     soundness.thm       & 1086 & 70  & Soundness theorem; aux. lemmas on transfer\\\hline
@@
                             Completeness theorem; aux. lemmas and relations\\\hline
-    inst\_decidable.thm &  830 & 42  & Other measures; Lemma~\ref{lemma:inst:decidable}\\\hline
-    decidability.thm    &  885 & 50  & Decidability theorem\\\hline
+    inst\_decidable.thm &  830 & 42  & Other worklist measures; Lemma~\ref{lemma:inst:decidable}\\\hline
+    decidability.thm    &  885 & 50  & Termination theorem and decidability corollary\\\hline
     \hline\emph{Total}               & 7700 & 567 & (51 definitions in total)\\\hline

### 2019-03-01 main.tex
undo the change on abstract; many typos fixed

@@
 Modern functional programming languages, such as Haskell or OCaml,
-use sophisticated forms of type inference. While being an
+use sophisticated forms of type inference. While an
 important topic in the Programming Languages research, there is little

### 2019-03-01 sections/Discussion.tex
undo the change on abstract; many typos fixed

@@
 applications. Therefore a natural question is whether our algorithm
-can also accomodate such elaboration.
+can also accommodate such elaboration.
 While our algorithmic reduction does not elaborate to System F,
@@
 If an existential variable is solved, we not only propagate the solution to the other judgments,
-but also replace ocurrences in the elaboration list.
+but also replace occurrences in the elaboration list.
 If an existential variable is recycled, meaning that it is not constrained,

### 2019-03-01 sections/Metatheory.tex
undo the change on abstract; many typos fixed

@@
 Figure~\ref{fig:trans}. A declarative worklist $\Om$ has the same structure as an
-algorithmic worklist $\Gm$, but does not contain any extistential variables $\al$.
+algorithmic worklist $\Gm$, but does not contain any existential variables $\al$.
 
@@
 The relation $\Gm \sto \Om$ expresses that the algorithmic worklist $\Gm$ can 
-be instiantiated to the declarative worklist $\Om$, by appropriately instantiating
+be instantiated to the declarative worklist $\Om$, by appropriately instantiating
 all existential variables $\al$ in $\Gm$ with well-scoped monotypes $\tau$.
@@
 
-The good news is that worklist update does not affect the three main worklist measures:
+The good news is that worklist updates do not affect the three main worklist measures:
 \newcommand{\equivGm}[1]{|\Gm|_{#1} = |\Gm'|_{#1}}
@@
 which also affects the completeness of one of our proofs.
-Specifically, our scipts compile against commit 92829a of Abella's GitHub repository.}
+Specifically, our scripts compile against commit 92829a of Abella's GitHub repository.}
 ) proof assistant~\cite{AbellaDesc} to develop our formalization.
@@
     decl.thm            &  226 & 33  & Basic declarative properties\\\hline
-    order.thm           &  235 & 27  & The $|\cdot|_\forall$ measure; decl. subtyping strenthening\\\hline
+    order.thm           &  235 & 27  & The $|\cdot|_\forall$ measure; decl. subtyping strengthening\\\hline
     alg.thm             &  655 & 72  & Basic algorithmic properties\\\hline

### 2019-03-01 sections/Overview.tex
undo the change on abstract; many typos fixed

@@
 which was adopted from \citet{odersky1996putting}. It compares the
-degree of polymorphism between between $A$ and $B$ in DK's implicit polymorphic type system. 
+degree of polymorphism between $A$ and $B$ in DK's implicit polymorphic type system. 
 Essentially, if $A$ can always be instantiated to match any instantiation of $B$,
@@
 decidable. Allowing an arbitrary type (rather than a monotype) in rule $\mathtt{{\le}\forall L}$
-is know to give rise to an undecidable subtyping relation~\cite{tiuryn1996subtyping}.
+is known to give rise to an undecidable subtyping relation~\cite{tiuryn1996subtyping}.
 \citet{jones2007practical} also impose the restriction of
@@
 $\al = \tau$.
-A context applications on a type, denoted by $[\Gamma]A$,
+A context application on a type, denoted by $[\Gamma]A$,
 substitutes all solved existential type variables in $\Gamma$
-with their solutions on the type $A$.
+with their solutions on type $A$.
 
@@
 Solutions are incrementally propagated by applying the algorithmic output
-context of a previous task as substitution to the next task. This can be seen
+context of a previous task as substitutions to the next task. This can be seen
 in the subsumption rule:
@@
 Preservation) and Lemma 14 (Subsumption)--- wrongly apply induction hypotheses
-in several cases. Fortunately, we have found simple work-arounds that fix these
+in several cases. Fortunately, we have found simple workarounds that fix these
 proofs without affecting the appeals to these lemmas.

### 2019-03-01 sections/Related.tex
undo the change on abstract; many typos fixed

@@
 existentials and indexed types.
-With a more complex declarative system, they developed a proofs of over 150 pages.
+With a more complex declarative system, they developed a proof of over 150 pages.
 It is even more difficult to argue its correctness for every single detail within such a big development.
@@
 \citet{naraschewski1999type} in Isabelle/HOL~\cite{nipkow2002isabelle}.
-The treatment of new variables was tricky at that time, while the overall structure follow the
+The treatment of new variables was tricky at that time, while the overall structure follows the
 structure of Damas's manual proof closely.
@@
 since the information of each local context needs to be synchronized.
-Instead we make use of the natural of ordered context to control the scoping of unification variables.
+Instead we make use of the nature of ordered context to control the scoping of unification variables.
 While their algorithms focus only on unification,
-our algorithm also deals with other type of judgments like type inference.
+our algorithm also deals with other types of judgments like type inference.
 A detailed discussion is in Section~\ref{sec:overview:list}.

### 2019-03-01 sections/algorithm.tex
undo the change on abstract; many typos fixed

@@
 condition of Rule 18 (which avoids overlap). It also streamlines the algorithm
-and the meatatheory as we now treat all cases
+and the metatheory as we now treat all cases
 where we can see that an existential variable should be instantiated to a
@@
 such as Rule 25.
-In comparison to Rule 8 and 9, where new variables are only refered by the sub-derivation,
+In comparison to Rule 8 and 9, where new variables are only referred by the sub-derivation,
 Rule 25 introduces variables $\al, \bt$ that affect the remaining judgment chain.

### 2019-03-01 sections/conclusion.tex
All "judgement" -> "judgment"
undo the change on abstract; many typos fixed

@@
 inference for higher-ranked polymorphism. This contribution is made possible by
-a new type inference algorithm for DK's declarative type sytsem that
+a new type inference algorithm for DK's declarative type system that
 incorporates two novel mechanisation-friendly ideas. Firstly, we merge the
@@
 %     \item a blend of ideas from various, including sources DK's ordered contexts
-%     \item list of judgements + some novel ideas
+%     \item list of judgments + some novel ideas
 %     \item lessons learned: problems identified in proofs and scoping
@@
 % the inference of $\lam x$ leaves an unsolved $\al$ in the context,
-% which we simply remove the variable after it is no longer refered to.
+% which we simply remove the variable after it is no longer referred to.
 % Is it possible that we infer better types by making better use of that piece of information?

### 2019-03-01 sections/introduction.tex
undo the change on abstract; many typos fixed

@@
 Unison. The DK type system has two variants: a declarative
-and an algorithmic one. The two variants have been been
+and an algorithmic one. The two variants have been
 \emph{manually} proved to be \emph{sound}, \emph{complete} and
@@
 keep track of variable scoping, DK's algorithmic version employs input
-and output contexts to track information that is discovered throught
+and output contexts to track information that is discovered through
 type inference. However modelling output contexts in a theorem prover

### 2019-03-02 sections/Related.tex
to match the screenshot
Ningning

@@
 
-\jimmy{Let arguments go first by Ningning}
-In a more recent work, \citet{xie2018letarguments} proposed a variant of
-bidirectional type inference system where type information flows from arguments to
-functions with an additional \emph{application} mode,
-which allows more higher-order typed programs to be inferred without additional annotations.
+In a more recent work, \citet{xie2018letarguments} proposed a variant of a
+bidirectional type inference system for a predicative system with higher-ranked types.
+Type information flows from arguments to
+functions with an additional \emph{application} mode. This variant 
+allows more higher-order typed programs to be inferred without additional annotations.
 Following the new mode, the let-generalization of the Hindley-Milner system
-is well supported as a syntactic sugar.
-Unfortunately, no formal proofs are given for the correctness of its algorithmic system.
+is well supported as a syntactic sugar. The formalization includes some
+mechanized proofs for the declarative type system, but all proofs regarding
+the algorithmic type system are manual.
 
@@
 (predicative and impredicative) discussed above, proofs and
-metatheory are manual. The only partial effort on mechanizing aspects of type inference
+metatheory for the algorithmic aspects are manual. The only partial effort on mechanizing algorithmic
+aspects of type inference
 for higher-ranked types is

### 2019-03-02 sections/Overview.tex
unused comments; bad rule format; bib fix; proof script fix

@@
 Declarative typing rule $\mathtt{DeclSub}$ overlaps with
-both $\mathtt{Decl\forall I}$ and $\mathtt{Decl\to I}$.
+both $\mathtt{Decl\forall I}$ and $\mathtt{Decl{\to}I}$.
 However, we argue that more specific rules are always the best choices,
-i.e. $\mathtt{Decl\forall I}$ and $\mathtt{Decl\to I}$ should have
+i.e. $\mathtt{Decl\forall I}$ and $\mathtt{Decl{\to}I}$ should have
 higher priority to $\mathtt{DeclSub}$. We will come back to this
@@
 More seriously, we have also found a lemma that simply does not hold, 
-Lemma 29 (Parallel Admissibility). See Appendix~\ref{appendix:false_lemmas}
-\bruno{reference} for a detailed explanation
+Lemma 29 (Parallel Admissibility).
+See Appendix~\ref{appendix:false_lemmas} for a detailed explanation
 and counterexample. This lemma is a cornerstone of the two metatheoretical results 

### 2019-03-02 sections/Overview.tex
mechanise -> mechanize

@@
 
-At first sight, the DK algorithm would seem a good basis for mechanisation. After
+At first sight, the DK algorithm would seem a good basis for mechanization. After
 all it comes with a careful description and extensive manual proofs.
 Unfortunately, we ran into several obstacles that have prompted us to formulate
-a different, more mechanisation-friendly algorithm.
+a different, more mechanization-friendly algorithm.
 
@@
 look for an alternative algorithmic formulation of the type system that is more
-amenable to mechanisation. Indeed, two aspects that are particularly
-challenging to mechanise are the scoping of universal and existential type
+amenable to mechanization. Indeed, two aspects that are particularly
+challenging to mechanize are the scoping of universal and existential type
 variables, and the propagation of the instantiation of existential type
@@
 While this works well on paper, this approach is still fairly involved and thus
-hard to reason about in a mechanised setting. Indeed, the instantiations have
+hard to reason about in a mechanized setting. Indeed, the instantiations have
 to be recorded in the context and are applied incrementally to each remaining
@@
 to look for a simpler algorithm that is easier to
-reason about in a mechanised setting.
+reason about in a mechanized setting.
 
@@
 
-In recent work, \citet{itp2018} have shown how to mechanise a variant of DK's
+In recent work, \citet{itp2018} have shown how to mechanize a variant of DK's
 subtyping algorithm and shown it correct with respect to DK's declarative

### 2019-03-02 sections/Related.tex
unused comments; bad rule format; bib fix; proof script fix
mechanise -> mechanize

@@
 the issue in more detail in Appendix~\ref{appendix:false_lemmas}.
-\bruno{I think you will need to justify this better and in the appendix you must show
-the lemma and counter-example.}
 
@@
 More and more type systems are formalized with these tools.
-However, mechanising certain algorithmic aspects, like unification and
+However, mechanizing certain algorithmic aspects, like unification and
 constraint solving, has received very little attention and is still challenging.

### 2019-03-02 sections/conclusion.tex
mechanise -> mechanize

@@
 a new type inference algorithm for DK's declarative type system that
-incorporates two novel mechanisation-friendly ideas. Firstly, we merge the
+incorporates two novel mechanization-friendly ideas. Firstly, we merge the
 traditional type context with the recently proposed concept of judgment chains,

### 2019-03-02 sections/algorithm.tex
unused comments; bad rule format; bib fix; proof script fix
Minor fix for a sentence

@@
 \Gm, x:A \algrule \Gm
-\\[2mm]
+\\[3mm]
 \Gm \Vdash 1\le 1 &\algrule \Gm\\
@@
 \Gm \Vdash A\le \all[b]B &\algrule \Gm,b \Vdash A\le B
-\\[2mm]
+\\[3mm]
 %\\
@@
  &\qquad \text{when }\al\notin FV(A)\cup FV(B)
- \\[2mm]
+ \\[3mm]
 \Gm[\al][\bt] \Vdash \al \le \bt &\algrule [\al/\bt](\Gm[\al][])\\
@@
 \Gm[\bt] \Vdash \bt \le 1 &\algrule [1/\bt](\Gm[])
-\\[2mm]
+\\[3mm]
 \Gm \Vdash e \Lto B &\algrule \Gm \Vdash e\To_a a\le B \quad
@@
 % \quad\text{\jimmy{Additional}}
-\\[2mm]
+\\[3mm]
 \Gm \Vdash x\To_a \jg &\algrule \Gm \Vdash [A/a] \jg \quad \text{when } (x:A)\in \Gm\\
@@
 \Gm \Vdash e_1\ e_2 \To_a \jg &\algrule \Gm \Vdash e_1\To_b (\appInfAlg{b}{e_2})
-\\[2mm]
+\\[3mm]
 \Gm \Vdash \appInfAlg{\all A}{e} &\algrule \Gm,\al \Vdash \appInfAlg{[\al/a]A}{e}\\
@@
 In comparison,
-DK's original algorithmic subsumption rule has no restriction on the expression $e$,
+DK's algorithmic subsumption rule has no restriction on the expression $e$,
 and they do not have a rule that explicitly handles the case $\lam e \Lto \al$.
 Therefore their only way to check a lambda function against an existential variable
-is by combining the inferenced type of the lambda function with the subsumption rule.
+is by applying the subsumption rule, which further breaks into
+a type inference of the lambda function and a subtyping judgment.
 % A combination of Rules $18'$ (Rule 18 without the restriction on $e$) and 25 and further derivations

### 2019-03-02 sections/Appendix.tex
newpage before appendix; move measures to appendix; new ==>> symbol
Better-looking worklist measure figure taking more spaces
typo

@@
 \section{Appendix}
+
+\subsection{Formal Definitions for Worklist Measures}
+
+\begin{figure}[h!]
+\hfill \framebox{$|\Gm|_e$} \hfill  \llap{Size measure.}
+\begin{gather*}
+    \begin{aligned}
+    |\Gm|_e &= \sum_{e\Lto A\in\Gm}|e|_e + \sum_{e\To_a \jg\in\Gm}|e|_e +
+        \sum_{\appInfAlg{A}{e}\in\Gm}|e|_e\\
+    \text{where } |x|_e &= |()|_e = 1\\
+        \qquad |\lam e|_e &= |e|_e + 1\\
+        |e_1~e_2|_e &= |e_1|_e + |e_2|_e + 1\\
+        |e:A|_e &= |e| + 1
+    \end{aligned}
+    \end{gather*}
+    \hfill \framebox{$|\Gm|_\Leftrightarrow$} \hfill  \llap{Judgment measure.}
+    \begin{gather*}
+    \begin{aligned}
+    |\Gm|_\Leftrightarrow &= 2\cdot\#_{e\Lto A\in\Gm} +
+        \#_{e\To_a \jg\in\Gm} + 3\cdot\#_{\appInfAlg{A}{e}\in\Gm}
+    \end{aligned}
+    \end{gather*}
+    \hfill \framebox{$|\Gm|_\forall$} \hfill  \llap{Polymorphism measure.}
+    \begin{gather*}
+    \begin{aligned}
+    |\Gm|_\forall &= \sum_{e\Lto A\in\Gm}|A|_\forall + \sum_{\appInfAlg{A}{e}\in\Gm}|A|_\forall +
+        \sum_{A\le B\in\Gm} |A|_\forall + |B|_\forall\\
+    \text{where } |1|_\forall &= |a|_\forall = |\al|_\forall = 1\\
+        |A\to B|_\forall &= |A|_\forall + |B|_\forall\\
+        |\all A|_\forall &= |A|_\forall + 1
+    \end{aligned}
+    \end{gather*}
+    \hfill \framebox{$|\Gm|_\to$} \hfill  \llap{Function measure.}
+    \begin{gather*}
+    \begin{aligned}
+    |\Gm|_\to &= \sum_{e\Lto A\in\Gm}|A|_\to + \sum_{\appInfAlg{A}{e}\in\Gm}|A|_\to +
+        \sum_{A\le B\in\Gm} |A|_\to + |B|_\to\\
+    \text{where } |1|_\to &= |a|_\to = |\al|_\to = 1\\
+        |A\to B|_\to &= |A|_\to + |B|_\to + 1\\
+        |\all A|_\to &= |A|_\to
+    \end{aligned}
+    \end{gather*}
+    \hfill \framebox{$|\Gm|_{\al}$} \hfill  \llap{Existential measure.}
+    \begin{gather*}
+    \begin{aligned}
+    |\Gm|_{\al} &= \#_{\al\in\Gm}\\
+    \end{aligned}
+\end{gather*}
+\caption{Worklist measures}
+\label{fig:measures}
+\end{figure}
 
@@
 and the proof shares a similar structure to DK's.
-One tricky case related to Rule $\mathtt{Decl\forall I}$ inducates that $A$ has the shape $\all A_0$,
+One tricky case related to Rule $\mathtt{Decl\forall I}$ indicates that $A$ has the shape $\all A_0$,
 thus the subtyping relation derives from either $\mathtt{{\le}\forall L}$ or $\mathtt{{\le}\forall R}$.

### 2019-03-02 main.tex
newpage before appendix; move measures to appendix; new ==>> symbol
Remove Abstract for submission

@@
 %% before \maketitle command
-\begin{abstract}
-Modern functional programming languages, such as Haskell or OCaml,
-use sophisticated forms of type inference. While an
-important topic in the Programming Languages research, there is little
-work on the mechanization of the metatheory of type inference in
-theorem provers. In particular we are unaware of any complete
-formalization of the type inference algorithms that are the backbone of modern 
-functional languages.
-
-This paper presents the first full mechanical formalization of the metatheory for
-higher-ranked polymorphic type inference. The system that we formalize
-is the bidirectional type system by Dunfield and
-Krishnaswami (DK). The DK type system has two variants (a declarative
-and an algorithmic one) that have been \emph{manually} proven
-\emph{sound}, \emph{complete} and \emph{decidable}. We present 
-a mechanical formalization in the Abella theorem prover 
-of DK's declarative type system with a novel algorithmic system. We have
-a few reasons to use a new algorithm. Firstly, our new algorithm
-employs \emph{worklist judgments}, which allow precise capture of the
-scope of variables and simplify the formalization of scoping in a
-theorem prover. Secondly, while DK's original formalization comes with very
-well-written manual proofs, there are several details missing and some
-incorrect proofs,
-which complicate the task of writing a formal proof. 
-Despite the use of a different algorithm we prove the
-same results as DK, although with significantly different proofs and
-proof techniques. Since such type inference algorithms are quite
-subtle and have a complex metatheory, mechanical
-formalizations are an important advance in type-inference research.
-\end{abstract}
+%\begin{abstract}
+%Modern functional programming languages, such as Haskell or OCaml,
+%use sophisticated forms of type inference. While an
+%important topic in the Programming Languages research, there is little
+%work on the mechanization of the metatheory of type inference in
+%theorem provers. In particular we are unaware of any complete
+%formalization of the type inference algorithms that are the backbone of modern 
+%functional languages. 
+%
+%This paper presents the first full mechanical formalization of the metatheory for
+%higher-ranked polymorphic type inference. The system that we formalize
+%is the bidirectional type system by Dunfield and
+%Krishnaswami (DK). The DK type system has two variants (a declarative
+%and an algorithmic one) that have been \emph{manually} proven
+%\emph{sound}, \emph{complete} and \emph{decidable}. We present 
+%a mechanical formalization in the Abella theorem prover 
+%of DK's declarative type system with a novel algorithmic system. We have
+%a few reasons to use a new algorithm. Firstly, our new algorithm
+%employs \emph{worklist judgments}, which allow precise capture of the
+%scope of variables and simplify the formalization of scoping in a
+%theorem prover. Secondly, while DK's original formalization comes with very
+%well-written manual proofs, there are several details missing and some
+%incorrect proofs,
+%which complicate the task of writing a formal proof. 
+%Despite the use of a different algorithm we prove the
+%same results as DK, although with significantly different proofs and
+%proof techniques. Since such type inference algorithms are quite
+%subtle and have a complex metatheory, mechanical
+%formalizations are an important advance in type-inference research.
+%\end{abstract}
 
@@
 %% comma separated list
-\keywords{type inference, higher-rank polymorphism, mechanisation}  %% \keywords are mandatory in final camera-ready submission
+\keywords{type inference, higher-rank polymorphism, mechanization}  %% \keywords are mandatory in final camera-ready submission
 
@@
 %% Appendix
+\newpage
 \appendix

### 2019-03-02 sections/Overview.tex
local changes

--- 2019-03-01 sections/Overview.tex
@@
 More seriously, we have also found a lemma that simply does not hold, 
-Lemma 29 (Parallel Admissibility). See Appendix~\ref{appendix:false_lemmas}
-\bruno{reference} for a detailed explanation
+Lemma 29 (Parallel Admissibility). See Appendix~\ref{appendix:false_lemmas} for a detailed explanation
 and counterexample. This lemma is a cornerstone of the two metatheoretical results 
@@
 subtyping algorithm and shown it correct with respect to DK's declarative
-subtyping judgment. This approach overcomes the problem's in DK's formulation
+subtyping judgment. This approach overcomes the some problems in DK's formulation
 by using a \emph{worklist}-based judgment of the form $$\Gamma \vdash \Omega$$
@@
 \paragraph{No Inference Judgments} 
-Secondly and more importantly, the approach only deals with a judgment for
+Secondly, and more importantly, the approach only deals with a judgment for
 \emph{checking} whether one type is the subtype of another. While this may not

### 2019-03-02 sections/Related.tex

@@
 constraint solving, has received very little attention and is still challenging.
-Moreover, while most tools supports local (input) contexts in a neat way,
+Moreover, while most tools support local (input) contexts in a neat way,
 many practical type-inference algorithms require

### 2019-03-02 sections/Discussion.tex
Typo

@@
 This section discusses some insights that we gained from our work and contrasts
-the scoping mechanisms we have employed by those in DK's algorithm.
+the scoping mechanisms we have employed with those in DK's algorithm.
 We also discuss a way to improve the precision of their scope tracking.
 Furthermore we discuss and sketch an extension of our algorithm with
-elaboration to a target calculus.
+an elaboration to a target calculus.
 

### 2019-04-18 response.md
First draft for the response

Response for Submission 66, ICFP 2019
======

We thank the reviewers for their helpful comments and positive reception of the paper.
We answer main questions below.

Review A
------
<!--
One minor point of criticism I have is that there appear to be some unexplained differences between the presentation in the paper and the Abella mechanization, which made it harder than necessary for me to relate the two and to follow some of the proof sketches in the paper. For example, the (binary) multi-step reduction relations on algorithmic and declarative worklists seem to be represented by unary predicates (judge and dc respectively) in the Abella script, which look more like big-step than small-step reduction relations. This change of representation may even impact the structure of some proofs (e.g. the induction hypothesis used in the proofs of soundness and completeness). I know very little about Abella, so this choice of representation may well be more idiomatic than the one presented in the paper. Still, I would have liked to see a short discussion of the pros/cons of the two representations and why different representations are used in the paper and the proof script. More extensive documentation of the proof script (e.g. more comments) would also be helpful in navigating the mechanization.
-->

> The difference between the presentation in the paper and the Abella mechanization.

In the paper, the main algorithmic rules are written in a simple small step style
with no premise on the reduction relation (-->) itself.
And the rules are not overlapping with each other.
These properties indicates that a trivial translation to
the big step style used in the proofs preserves equivalence.

We believe the presentation of the algorithmic rules in the paper is much clear
and restricted to the form $\Gamma$ --> $\Gamma$,
while the big step unary relation in the proof simplfies the induction scheme.
More importantly, the difference is just a matter of presentation,
but the logic behind remains unchanged.

> The proof scripts are hard to read, and some parts are inconsistent with the paper.

We apologize for the lack of clean-up on the proof scripts.
We are now refactoring the scripts,
after that we will provide a better documentation on the encoding details
that might cover all the concerns by the reviewer,
preferrably to the appendix of the paper.

Review B
------
<!--
On the negative part, I am not sure that one can, from the paper, infer general hints/guidelines about the formalization of the correctness/completeness/... of an inference algorithm for a complex typing discipline. This is mainly due to the fact that the authors concentrate their attention to one single inference algorithm.
-->
<!--
Tom:
- Apparently Reviewer C believes they will be able to try out our approach on new problems.
- I also think there is an obvious generalization to other bidirectional type systems.
- We can say that we will extend the discussion to point this out.
-->

> General hints/guidlines about the formalization of the properties of an inference algorithm for a complex typing descipline.

We believe that the algorithm proposed can adapt to other bidirectional type systems,
and we will further extend the discussion to point this out.

Review C
------
<!--
In the course of their investigation, the authors found several problems with the mathematical proof DK give and provided alternatives. I am not in a position to comment on whether these criticisms of DK are justified, nor do I find this significant for the evaluation of the paper.
-->
<!--
Jimmy:
If the reviewer are not interested in the criticisms,
we can reduce relevant discussion in the paper.
My point was that I am not sure whether DK's algorithm and proofs
could be easily formalized, but it seems no easy.
And I decided not to take time investigate,
given that there are already some issues, though they might be minor.
-->

> Criticisms of DK's proof

...

### 2019-04-18 response.md
Drop Review C response

@@
 
-We thank the reviewers for their helpful comments and positive reception of the paper.
-We answer main questions below.
+We are grateful for the helpful reviews and are elated that you appreciate our work.
+In what follows  we respond to your main comments.
 
@@
 and we will further extend the discussion to point this out.
-
-Review C
-------
-<!--
-In the course of their investigation, the authors found several problems with the mathematical proof DK give and provided alternatives.
-I am not in a position to comment on whether these criticisms of DK are justified, nor do I find this significant for the evaluation of the paper.
--->
-<!--
-Jimmy:
-If the reviewer are not interested in the criticisms,
-we can reduce relevant discussion in the paper.
-My point was that I am not sure whether DK's algorithm and proofs
-could be easily formalized, but it seems no easy.
-And I decided not to take time investigate,
-given that there are already some issues, though they might be minor.
--->
-
-> Criticisms of DK's proof
-
-...

### 2019-04-18 response.md
Update response.md

@@
 
-We believe that the algorithm proposed can adapt to other bidirectional type systems,
-and we will further extend the discussion to point this out.
+That is a good point. We believe that our approach can indeed be adapted to other bidirectional type systems, and we will extend the discussion in the paper to point this out.
+We are also happy to see that Reviewer C already looks forward to trying out our approach in other settings.

### 2019-04-18 response.md
Update response.md

@@
 
-We believe the presentation of the algorithmic rules in the paper is much clear
+We believe the presentation of the algorithmic rules in the paper is much clearer
 and restricted to the form $\Gamma$ --> $\Gamma$,
@@
 
-We apologize for the lack of clean-up on the proof scripts.
-We are now refactoring the scripts,
-after that we will provide a better documentation on the encoding details
-that might cover all the concerns by the reviewer,
-preferrably to the appendix of the paper.
+We apologize for the state of the proof scripts. Currently, we are refactoring them.
+and will document the mapping between the paper and the mechanisation.
 

### 2019-04-18 response.md
reply

@@
 
-We are grateful for the helpful reviews and are elated that you appreciate our work.
-In what follows  we respond to your main comments.
+We are grateful for the helpful reviews and are happy
+that you appreciate our work. In what follows we
+respond to your main comments.
 
@@
 <!--
-One minor point of criticism I have is that there appear to be some unexplained differences between the presentation in the paper and the Abella mechanization, which made it harder than necessary for me to relate the two and to follow some of the proof sketches in the paper.
-For example, the (binary) multi-step reduction relations on algorithmic and declarative worklists seem to be represented by unary predicates (judge and dc respectively) in the Abella script, which look more like big-step than small-step reduction relations.
-This change of representation may even impact the structure of some proofs (e.g.
-the induction hypothesis used in the proofs of soundness and completeness).
-I know very little about Abella, so this choice of representation may well be more idiomatic than the one presented in the paper.
-Still, I would have liked to see a short discussion of the pros/cons of the two representations and why different representations are used in the paper and the proof script.
-More extensive documentation of the proof script (e.g.
-more comments) would also be helpful in navigating the mechanization.
+One minor point of criticism I have is that there appear to be some
+unexplained differences between the presentation in the paper and the
+Abella mechanization, which made it harder than necessary for me to
+relate the two and to follow some of the proof sketches in the
+paper. For example, the (binary) multi-step reduction relations on
+algorithmic and declarative worklists seem to be represented by unary
+predicates (judge and dc respectively) in the Abella script, which
+look more like big-step than small-step reduction relations. This
+change of representation may even impact the structure of some proofs
+(e.g. the induction hypothesis used in the proofs of soundness and
+completeness). I know very little about Abella, so this choice of
+representation may well be more idiomatic than the one presented in
+the paper. Still, I would have liked to see a short discussion of the
+pros/cons of the two representations and why different representations
+are used in the paper and the proof script. More extensive
+documentation of the proof script (e.g. more comments) would also be
+helpful in navigating the mechanization.
 -->
@@
 
-In the paper, the main algorithmic rules are written in a simple small step style
-with no premise on the reduction relation (-->) itself.
-And the rules are not overlapping with each other.
-These properties indicates that a trivial translation to
-the big step style used in the proofs preserves equivalence.
+In the paper, the main algorithmic rules are written in a simple small
+step style with no premise on the reduction relation (-->) itself.
+And the rules are not overlapping with each other.  These properties
+indicates that a trivial translation to the big step style used in the
+proofs preserves equivalence.
 
-We believe the presentation of the algorithmic rules in the paper is much clearer
-and restricted to the form $\Gamma$ --> $\Gamma$,
-while the big step unary relation in the proof simplfies the induction scheme.
-More importantly, the difference is just a matter of presentation,
-but the logic behind remains unchanged.
+We believe the presentation of the algorithmic rules in the paper is
+much clearer and restricted to the form $\Gamma$ --> $\Gamma$, while
+the big step unary relation in the proof simplifies the induction
+scheme.  More importantly, the difference is just a matter of
+presentation, but the logic behind remains unchanged.
+We will mention this difference in presentation clearly in the paper.
+
+> The restriction of subtyping to simple  types is effectively an
+> equivalence (and thus symmetric).  Is this correct?
+
+Yes, this is correct. Note also that DK's algorithm also suggests a
+similar symmetric unification.
 
@@
 We apologize for the state of the proof scripts. Currently, we are refactoring them.
-and will document the mapping between the paper and the mechanisation.
+and will document the mapping between the paper and the mechanisation
+in the final version of the paper.
 
@@
 
-> General hints/guidlines about the formalization of the properties of an inference algorithm for a complex typing descipline.
+> General hints/guidlines about the formalization of the properties of
+> an inference algorithm for a complex typing descipline.
 
-That is a good point. We believe that our approach can indeed be adapted to other bidirectional type systems, and we will extend the discussion in the paper to point this out.
-We are also happy to see that Reviewer C already looks forward to trying out our approach in other settings.
+That is a good point. We believe that our approach can indeed be
+adapted to other bidirectional type systems, and we will extend the
+discussion in the paper to point this out.  We are also happy to see
+that Reviewer C already looks forward to trying out our approach in
+other settings.

### 2019-04-18 response.md
More on Review A

@@
 
-We apologize for the state of the proof scripts. Currently, we are refactoring them.
+We apologize for the state of the proof scripts. Currently, we are refactoring them,
 and will document the mapping between the paper and the mechanisation
 in the final version of the paper.
+
+> How does polynomial complexity follow from the reduction rule being deterministic?
+
+We did not justify the point clearly in the paper.
+A detailed proof will be provided for the final version.
+
+> Why not treat the languages of declarative and algorithmic worklists as subsets?
+
+The proof script allows one to treat them as subsets.
+However, we explicitly treat them as different syntactic sorts in the paper
+to avoid potential confusion.
+
+> I could not find a direct counterpart of this theorem in the Abella script.
+
+The `decidable_thm` Theorem is the one that proves the Termination Theorem (4.12),
+following what is described from L918,
+and the `decidable` Theorem is its pretty version.
+We will make the theorem names consistent for publish.
 

### 2019-04-19 response.md
language
Removing comments from the reply

--- 2019-04-18 response.md
@@
 ------
-<!--
-One minor point of criticism I have is that there appear to be some
-unexplained differences between the presentation in the paper and the
-Abella mechanization, which made it harder than necessary for me to
-relate the two and to follow some of the proof sketches in the
-paper. For example, the (binary) multi-step reduction relations on
-algorithmic and declarative worklists seem to be represented by unary
-predicates (judge and dc respectively) in the Abella script, which
-look more like big-step than small-step reduction relations. This
-change of representation may even impact the structure of some proofs
-(e.g. the induction hypothesis used in the proofs of soundness and
-completeness). I know very little about Abella, so this choice of
-representation may well be more idiomatic than the one presented in
-the paper. Still, I would have liked to see a short discussion of the
-pros/cons of the two representations and why different representations
-are used in the paper and the proof script. More extensive
-documentation of the proof script (e.g. more comments) would also be
-helpful in navigating the mechanization.
--->
 
@@
 We believe the presentation of the algorithmic rules in the paper is
-much clearer and restricted to the form $\Gamma$ --> $\Gamma$, while
+clearer and restricted to the form $\Gamma$ --> $\Gamma$, while
 the big step unary relation in the proof simplifies the induction
@@
 ------
-<!--
-On the negative part, I am not sure that one can, from the paper, infer general hints/guidelines about the formalization of the correctness/completeness/...
-of an inference algorithm for a complex typing discipline.
-This is mainly due to the fact that the authors concentrate their attention to one single inference algorithm.
--->
-<!--
-Tom:
-- Apparently Reviewer C believes they will be able to try out our approach on new problems.
-- I also think there is an obvious generalization to other bidirectional type systems.
-- We can say that we will extend the discussion to point this out.
--->
 

### 2019-04-19 response.md

@@
 
-We apologize for the state of the proof scripts. Currently, we are refactoring them.
+We apologize for the state of the proof scripts. Currently, we are refactoring them,
 and will document the mapping between the paper and the mechanisation
 in the final version of the paper.
+
+> How does polynomial complexity follow from the reduction rule being deterministic?
+
+We did not justify the point clearly in the paper.
+A detailed proof will be provided for the final version.
+
+> Why not treat the languages of declarative and algorithmic worklists as subsets?
+
+The proof script allows one to treat them as subsets.
+However, we explicitly treat them as different syntactic sorts in the paper
+to avoid potential confusion.
+
+> I could not find a direct counterpart of this theorem in the Abella script.
+
+The `decidable_thm` Theorem is the one that proves the Termination Theorem (4.12),
+following what is described from L918, and the `decidable` Theorem is its pretty version.
+We will make the theorem names consistent in the final version.
 

### 2019-04-20 response.md
Removing comments from the reply
minor typo

@@
 We believe the presentation of the algorithmic rules in the paper is
-clearer and restricted to the form $\Gamma$ --> $\Gamma$, while
+clearer and restricted to the form $\Gamma$ --> $\Gamma'$, while
 the big step unary relation in the proof simplifies the induction
@@
 We did not justify the point clearly in the paper.
-A detailed proof will be provided for the final version.
+We will either provide a detailed proof in the final version, or omit
+this informal claim.
 
@@
 
-> General hints/guidlines about the formalization of the properties of
-> an inference algorithm for a complex typing descipline.
+> General hints/guidelines about the formalization of the properties of
+> an inference algorithm for a complex typing discipline.
 

### 2019-05-17 main.tex
Not Abstract, sorry
macros in main.tex that helps generate parts
Modification according to comments and known typos

--- 2019-03-01 main.tex
@@
 \input{sections/macros.tex}
+
+% Appendix only
+%\usepackage[28-]{pagesel}
+% Main article with references
+%\usepackage[-27]{pagesel}
 
@@
 incorrect proofs,
-which complicate the task of writing a formal proof. 
+which complicate the task of writing a machanized proof. 
 Despite the use of a different algorithm we prove the

### 2019-05-17 sections/Metatheory.tex
newpage before appendix; move measures to appendix; new ==>> symbol
Modification according to comments and known typos

@@
     \item If $\Om \vdash \appInf{A'}{e}{C}$ and
-        $\Gm \Vdash \appInfAlg{A}{e}[a][1 \to 1] \sto \Om \Vdash \appInfAlg{A'}{e}[a][1 \to 1]$
+        $\Gm \Vdash \appInfAlg{A}{e}[a][1 \le 1] \sto \Om \Vdash \appInfAlg{A'}{e}[a][1 \le 1]$
         \\then $\Gm \Vdash \appInfAlg{A}{e}[a][1 \le 1] \redto \nil$.
@@
 
-Finally, we show that our algorithm terminates. In other words:
+Finally, we show that our algorithm terminates:
 \begin{theorem}[Termination]
@@
 $\langle |\Gm|_e, |\Gm|_\Leftrightarrow, |\Gm|_\forall, |\Gm|_{\al}, |\Gm|_\to + |\Gm| \rangle$
-on the worklist $\Gm$. The definitions of these measures can be found in Figure~\ref{fig:measures}. The first two, 
+on the worklist $\Gm$. Formal definitions of these measures can be found in
+Figure~\ref{fig:measures} in the Appendix. The first two, 
 $|\cdot|_e$ and $|\cdot|_\Leftrightarrow$, measure the total size of terms
@@
 count for 2, inference judgments for 1 and function inference judgments for 3.
-Other two measures, $|\cdot|_\forall$ and $|\cdot|_\to$, count the total number of
+Another two measures, $|\cdot|_\forall$ and $|\cdot|_\to$, count the total number of
 universal quantifiers and function types, respectively. Finally,
@@
 and $|\cdot|$ is simply the length of the worklist.
-
-\begin{figure}[t]
-\hfill \framebox{$|\Gm|_e$} \hfill  \llap{Size measure.}
-\begin{gather*}
-\begin{aligned}
-|\Gm|_e &= \textstyle\sum_{e\Lto A\in\Gm}|e|_e + \sum_{e\To_a \jg\in\Gm}|e|_e +
-    \sum_{\appInfAlg{A}{e}\in\Gm}|e|_e\\
-\text{where } |x|_e &= |()|_e = 1 \qquad |\lam e|_e = |e|_e + 1\\
-|e_1~e_2|_e &= |e_1|_e + |e_2|_e + 1 \qquad |e:A|_e = |e| + 1
-\end{aligned}
-\end{gather*}
-\hfill \framebox{$|\Gm|_\Leftrightarrow$} \hfill  \llap{Judgment measure.}
-\begin{gather*}
-\begin{aligned}
-|\Gm|_\Leftrightarrow &= 2\cdot\#_{e\Lto A\in\Gm} +
-    \#_{e\To_a \jg\in\Gm} + 3\cdot\#_{\appInfAlg{A}{e}\in\Gm}
-\end{aligned}
-\end{gather*}
-\hfill \framebox{$|\Gm|_\forall$} \hfill  \llap{Polymorphism measure.}
-\begin{gather*}
-\begin{aligned}
-|\Gm|_\forall &= \textstyle\sum_{e\Lto A\in\Gm}|A|_\forall + \sum_{\appInfAlg{A}{e}\in\Gm}|A|_\forall +
-    \sum_{A\le B\in\Gm} |A|_\forall + |B|_\forall\\
-\text{where } |1|_\forall &= |a|_\forall = |\al|_\forall = 1\\
-|A\to B|_\forall &= |A|_\forall + |B|_\forall \qquad |\all A|_\forall = |A|_\forall + 1
-\end{aligned}
-\end{gather*}
-\hfill \framebox{$|\Gm|_\to$} \hfill  \llap{Function measure.}
-\begin{gather*}
-\begin{aligned}
-|\Gm|_\to &= \textstyle\sum_{e\Lto A\in\Gm}|A|_\to + \sum_{\appInfAlg{A}{e}\in\Gm}|A|_\to +
-    \sum_{A\le B\in\Gm} |A|_\to + |B|_\to\\
-\text{where } |1|_\to &= |a|_\to = |\al|_\to = 1\\
-|A\to B|_\to &= |A|_\to + |B|_\to + 1 \qquad |\all A|_\to = |A|_\to
-\end{aligned}
-\end{gather*}
-\hfill \framebox{$|\Gm|_{\al}$} \hfill  \llap{Existential measure.}
-\begin{gather*}
-\begin{aligned}
-|\Gm|_{\al} &= \#_{\al\in\Gm}\\
-\end{aligned}
-\end{gather*}
-\caption{Worklist measures}
-\label{fig:measures}
-\end{figure}
 
@@
 
+\jimmy{Discuss briefly on the difference of encodings in the script and this paper.}
+While the algorithmic rules are in a small-step style,
+the proof script rewrites them into big-step style,
+where inductions are more strightforward to handle.
+In addition, we do prove the equivalence of the two representations.
+
 \paragraph{Statistics of the Proof}

### 2019-05-17 sections/Overview.tex
Modification according to comments and known typos

@@
 Contexts $\Psi$ are sequences of type variable declarations and
-term variables with its type declared $x:A$.
+term variables with their types declared $x:A$.
 
 \begin{figure}[t]
-%\centering \framebox{$\Psi \vdash A$}
-%\begin{gather*}
-%\inferrule*[right=$\mathtt{wf_d unit}$]
-%    {~}{\Psi\vdash 1}
-%\qquad
-%\inferrule*[right=$\mathtt{wf_d var}$]
-%    {a\in\Psi}{\Psi\vdash a}
-%\qquad
-%\inferrule*[right=$\mathtt{wf_d{\to}}$]
-%    {\Psi\vdash A\quad \Psi\vdash B}
-%    {\Psi\vdash A\to B}
-%\qquad
-%\inferrule*[right=$\mathtt{wf_d\forall}$]
-%    {\Psi, a\vdash A}
-%    {\Psi\vdash \forall a. A}
-%\end{gather*}
+\centering \framebox{$\Psi \vdash A$}
+\begin{gather*}
+\inferrule*[right=$\mathtt{wf_d unit}$]
+    {~}{\Psi\vdash 1}
+\qquad
+\inferrule*[right=$\mathtt{wf_d var}$]
+    {a\in\Psi}{\Psi\vdash a}
+\qquad
+\inferrule*[right=$\mathtt{wf_d{\to}}$]
+    {\Psi\vdash A\quad \Psi\vdash B}
+    {\Psi\vdash A\to B}
+\qquad
+\inferrule*[right=$\mathtt{wf_d\forall}$]
+    {\Psi, a\vdash A}
+    {\Psi\vdash \forall a. A}
+\end{gather*}
+
+\centering \framebox{$\Psi \vdash e$}
+\begin{gather*}
+\inferrule*[right=$\mathtt{wf_d tmvar}$]
+    {x:A\in\Psi}{\Psi\vdash x}
+\qquad
+\inferrule*[right=$\mathtt{wf_d tmunit}$]
+    {~}{\Psi\vdash ()}
+\qquad
+\inferrule*[right=$\mathtt{wf_d abs}$]
+    {\Psi,x:A\vdash e}
+    {\Psi\vdash \lam e}
+\\
+\inferrule*[right=$\mathtt{wf_d app}$]
+    {\Psi\vdash e_1 \quad \Psi\vdash e_2}
+    {\Psi\vdash e_1~e_2}
+\qquad
+\inferrule*[right=$\mathtt{wf_d anno}$]
+    {\Psi\vdash A \quad \Psi\vdash e}
+    {\Psi\vdash (e:A)}
+\end{gather*}
 
@@
 \inferrule*[right=$\mathtt{{\le}\forall R}$]
-    {\Psi, a\vdash A\le B}
+    {\Psi, b\vdash A\le B}
     {\Psi\vdash A \le \all[b]B}
@@
 
+\jimmy{TODO: Mention the implicit freshness conditions in the premises throughout the text}
+
 \begin{figure}[t]
 \begin{tabular}{rl}
-    \framebox{$\Psi \vdash A \Lto B$} & $e$ checks against input type $A$.\\[0.5mm]
-    \framebox{$\Psi \vdash A \To B$} & $e$ synthesizes output type $A$.\\[0.5mm]
+    \framebox{$\Psi \vdash e \Lto A$} & $e$ checks against input type $A$.\\[0.5mm]
+    \framebox{$\Psi \vdash e \To A$} & $e$ synthesizes output type $A$.\\[0.5mm]
     \framebox{$\Psi \vdash \appInf{A}{e}{C}$} & Applying a function of type $A$ to $e$ synthesizes type $C$.
@@
 i.e. $\mathtt{Decl\forall I}$ and $\mathtt{Decl{\to}I}$ should have
-higher priority to $\mathtt{DeclSub}$. We will come back to this
+higher priority than $\mathtt{DeclSub}$. We will come back to this
 topic in Section~\ref{sec:metatheory:non-overlapping}.
@@
 
+\jimmy{Comment A is not satisfied with the term "continuation passing style" for our definition.}
+
 $$\begin{aligned}
@@
 in the chain $\omega$ by substituting it for the placeholder type variable $a$.
+Notice that such $a$ binds a fresh type variable for the inner chain $\jg$,
+which behaves similarly to the variable declarations in the context.
 

### 2019-05-17 sections/introduction.tex
mechanise -> mechanize
typos; submitted to the website as a draft
Modification according to comments and known typos

@@
 The function \verb|hpoly| cannot be
-type-checked in Hindley-Milner.  The type of \verb|hpoly| is the rank-2 type:
+type-checked in the Hindley-Milner type system. The type of \verb|hpoly| is the rank-2 type:
 \verb|(forall a. a -> a) -> (Int, Char)|. Notably (and unlike
@@
 solution for these problems. Theorem provers can automatically check and
-validate the proofs, which removes the burden from checking from both
+validate the proofs, which removes the burden of checking from both
 the person doing the proofs as well as readers or reviewers.
@@
 The system
-that we formalize is the bidirectional type system by Dunfield and
-Krishnaswami (DK)~\cite{dunfield2013complete}.
+that we formalize is the bidirectional type system by \citet{dunfield2013complete}.
 We chose DK's type system because it is
@@
 %%Perhaps, with some additional thought, a work-around can be found, but, as our algorithm
-%%differs significantly from DK's, it would not further our mechanisation goal and so we did not pursue this.}
+%%differs significantly from DK's, it would not further our mechanization goal and so we did not pursue this.}
 Despite the problems in DK's manual formalization,
@@
 and output contexts to track information that is discovered through
-type inference. However modelling output contexts in a theorem prover
+type inference. However modeling output contexts in a theorem prover
 is non-trivial.

### 2019-05-17 Proto.hs

@@ -1,4 +1,4 @@
-{-# LANGUAGE FlexibleInstances, OverlappingInstances #-}
+{-# LANGUAGE FlexibleInstances, OverlappingInstances, ScopedTypeVariables #-}
 
 import Prelude hiding (drop)
 import Data.List
@@ -69,6 +69,17 @@ instance Show [Work] where
   show (WBind x t : w) = show w ++ ", " ++ x ++ " : " ++ show t
   show (WJug c : w) = show w ++ " ||- " ++ show c
 
+
+eesubst :: String -> Exp -> Exp -> Exp
+eesubst s e (Lam x b)
+  | s == x    = Lam x b
+  | otherwise = Lam x (eesubst s e b)
+eesubst s e (App e1 e2) = App (eesubst s e e1) (eesubst s e e2)
+eesubst s e (Ann e1 t) = Ann (eesubst s e e1) t
+eesubst s e (Var x)
+  | s == x    = e
+  | otherwise = Var x
+eesubst s e t = t
 
 esubst :: String -> Typ -> Exp -> Exp
 esubst e s (Lam x b) = Lam x (esubst e s b)
@@ -175,13 +186,21 @@ findBind x (WBind y a : w)
 findBind x (_ : w) = findBind x w
 findBind _ [] = Nothing
 
-pickNewVar w = [fromJust $ find (\c -> all (\var -> c `notElem` var) wvars) ['a'..'z']]
+pickNewVar w = [fromJust $ find (\c -> all (\var -> c `notElem` var) wvars) ['a'..'w']]
   where
     wvars = concatMap (\x -> case x of
         WEVar v -> [v]
         WTVar v -> [v]
         _ -> []
       ) w
+
+pickNewBindVar w = fromJust $ find (`notElem` wvars) bvarsupply
+  where
+    wvars = concatMap (\x -> case x of
+        WBind v _ -> [v]
+        _ -> []
+      ) w
+    bvarsupply = "x" : "y" : [ xy : show n | n <- [1..100], xy <- ['x', 'y'] ]
 
 step :: Worklist -> Worklist
 -- First 3 rules
@@ -211,10 +230,14 @@ step (WJug (Sub (EVar b) TInt) : w) = wsubst b [] TInt w
 -- rule 7
[…]

### 2019-05-20 sections/Appendix.tex
Better definition for measures

--- 2019-03-01 sections/Appendix.tex
@@
 \section{Appendix}
+
+\subsection{Formal Definitions for Worklist Measures}
+
+\begin{figure}[h!]
+    \begin{minipage}[t]{.5\textwidth}
+        \centering\framebox{$|\Gm|_e$, $|\jg|_e$, $|e|_e$} {Size measure.}
+        \begin{gather*}
+            \begin{aligned}
+            |\Gm|_e &= \sum_{\jg \in \Gm} |\jg|_e\\[2mm]
+            |A \le B|_e &= 0\\
+            |e \Lto A|_e &= |e|_e\\
+            |e \To \jg|_e &= |e|_e + |\jg|_e\\
+            |\appInfAlg{A}{e}|_e &= |e|_e + |\jg|_e\\[2mm]
+            |x|_e &= |()|_e = 1\\
+            \qquad |\lam e|_e &= |e|_e + 1\\
+            |e_1~e_2|_e &= |e_1|_e + |e_2|_e + 1\\
+            |e:A|_e &= |e|_e + 1
+            \end{aligned}
+        \end{gather*}
+
+        \centering\framebox{$|\Gm|_\Leftrightarrow$, $|\jg|_\Leftrightarrow$} {Judgment measure.}
+        \begin{gather*}
+            \begin{aligned}
+            |\Gm|_\Leftrightarrow &= \sum_{\jg \in \Gm} |\jg|_e\\[2mm]
+            % 2\cdot\#_{e\Lto A\in\Gm} +
+            %    \#_{e\To_a \jg\in\Gm} + 3\cdot\#_{\appInfAlg{A}{e}\in\Gm}
+            |A \le B|_\Leftrightarrow &= 0\\
+            |e \Lto A|_\Leftrightarrow &= 2\\
+            |e \To \jg|_\Leftrightarrow &= |\jg|_\Leftrightarrow + 1\\
+            |\appInfAlg{A}{e}|_\Leftrightarrow &= |\jg|_\Leftrightarrow + 3
+            \end{aligned}
+        \end{gather*}
+        \centering\framebox{$|\Gm|_{\al}$} {Existential measure.}
+        \begin{gather*}
+            \begin{aligned}
+            |\Gm|_{\al} &= \#_{\al\in\Gm}\\
+            \end{aligned}
+        \end{gather*}
+    \end{minipage}%
+    \begin{minipage}[t]{.5\textwidth}
+        \centering\framebox{$|\Gm|_\forall$, $|\jg|_\forall$, $|A|_\forall$} {Polymorphism measure.}
+        \begin{gather*}
+            \begin{aligned}
+            |\Gm|_\Leftrightarrow &= \sum_{\jg \in \Gm} |\jg|_\forall\\[2mm]
+            |A \le B|_\forall &= |A|_\forall + |B|_\forall\\
+            |e \Lto A|_\forall &= |A|_\forall\\
+            |e \To \jg|_\forall &= |\jg|_\forall\\
+            |\appInfAlg{A}{e}|_\forall &= |A|_\forall + |\jg|_\forall\\[2mm]
+            |1|_\forall &= |a|_\forall = |\al|_\forall = 1\\
+            |A\to B|_\forall &= |A|_\forall + |B|_\forall\\
+            |\all A|_\forall &= |A|_\forall + 1
+            \end{aligned}
+        \end{gather*}
+
+        \centering\framebox{$|\Gm|_\to$, $|\jg|_\to$, $|A|_\to$} {Function measure.}
+        \begin{gather*}
+            \begin{aligned}
+            |\Gm|_\to &= \sum_{\jg \in \Gm} |\jg|_\forall\\[2mm]
+            |A \le B|_\forall &= |A|_\forall + |B|_\forall\\
+            |e \Lto A|_\forall &= |A|_\forall\\
+            |e \To \jg|_\forall &= |\jg|_\forall\\
+            |\appInfAlg{A}{e}|_\forall &= |A|_\forall + |\jg|_\forall\\[2mm]
+            |1|_\to &= |a|_\to = |\al|_\to = 1\\
+            |A\to B|_\to &= |A|_\to + |B|_\to + 1\\
+            |\all A|_\to &= |A|_\to
+            % \sum_{e\Lto A\in\Gm}|A|_\to + \sum_{\appInfAlg{A}{e}\in\Gm}|A|_\to +
+            %     \sum_{A\le B\in\Gm} |A|_\to + |B|_\to\\
+            \end{aligned}
+        \end{gather*}
+    \end{minipage}
+\caption{Worklist measures}
+\label{fig:measures}
+\end{figure}
 
@@
 
-\paragraph{Problem in DK's Manual Proof}
+\paragraph{Problems in DK's Manual Proof}
 We tried to follow DK's manual proof for this lemma.

### 2019-05-21 sections/Discussion.tex
Modification according to comments and known typos
Lexically scoped type variable extension

@@
 an elaboration to a target calculus.
+
+\jimmy{TODO: update the summary}
 
@@
 Moreover, existential variables need special treatment, since they may be solved at any point,
-or be recycled if not solved within its scope.
+or be recycled if not solved within their scopes.
 If an existential variable is solved, we not only propagate the solution to the other judgments,
@@
 The unit type $1$, as the simplest type in the system, is a good choice.
+
+\subsection{Lexically-scoped type variables}
+We further extend the type system to support a simple extension, Lexically-scoped type variables.
+From a practical point of view,
+such extension enables the programmer to refer to type variables in a generic function signature
+when implementing of that function.
+For example,
+$$(\lam{\lam[y] (x:a)}) : \all[a~b] a \to b \to a$$
+has an inner annotation $(x:a)$ that refers to the outer generic type variable $a$.
+This is not a surprising feature, since the declarative system already accepts these kind of programs
+$$
+\inferrule*[right=$\mathtt{DeclAnno}$]
+{
+    \Psi \vdash \all A \quad
+    \inferrule*[right=$\mathtt{Decl\forall I}$]
+    {\Psi,a \vdash e \Lto A}
+    {\Psi\vdash e\Lto \all A}
+}
+    {\Psi\vdash (e: \all A)\To \all A}
+$$
+
+But there are issues about the well-formedness condition.
+Normally $\Gm \vdash_{wf} (e : A)$ is defined by $\Gm \vdash_{wf} e$ and $\Gm \vdash_{wf} A$.
+When $A = \all A'$, the type variable $a$ is not introduced to the context of $e$,
+therefore $\Gm \vdash_{wf} e$ is not derivable.
+
+To address the problem, one can either
+1) change the well-formedness judgment by introducing the binder for terms as well;
+or 2) add a syntax that explicitly bind the generic type variable.
+The former approach requires to resolve overlapping cases
+$(e : \all A)$ with $(e : B)$ where $B \neq \all B'$.
+Therefore we choose the latter approach due to its simplicity.
+The syntax of expression is extended with one more constructor:
+$$
+\begin{array}{l@{\qquad}lcl}
+\text{Expressions}\qquad&e &::=&\quad x \mid () \mid \lam e \mid e_1~e_2 \mid (e:A) \mid \Lambda a.~e:A
+\end{array}
+$$
+
+Notice that the type-lambda syntax $\Lambda a.~e:A$ annotates the generic function $e$ with $\all A$,
+while making $a$ visible inside the body of the function.
+Corresponding rules are introduced for both the declarative system and the algorithmic system:
+$$
+\inferrule*[right=$\mathtt{Decl}\Lambda$]
+    {\Psi,a \vdash A \quad \Psi,a \vdash e\Lto A}
+    {\Psi\vdash \Lambda a.~e:A \To \all A}
+$$
+$$\Gm \Vdash \Lambda a.~e:A \To_b \jg \algrule \Gm \Vdash [(\all A)/b]\jg, a \Vdash e \Lto A$$
+
+The new syntax for expressions also affect the well-formedness judgment:
+$$
+\inferrule*[right=$\mathtt{wf\_}\Lambda$]
+    {\Gm,a \vdash_{wf} e \quad \Gm,a \vdash_{wf} A}
+    {\Gm \vdash_{wf} \Lambda a.~e:A}
+$$
+
+In a practical point of view, programmers may not write the syntax $\Lambda a. e : A$ directly.
+The ScopedTypeVariables extension of Haskell is effective only
+when the type signature is explicitly universally quantified
+(the compiler translates into an expression similar to $\Lambda a. e : A$);
+otherwise the program means the normal syntax $e : \all A$,
+and one may not later refer to the type variable $a$.
+
+We also obtained all three desired properties for the extended system,
+namely soundness, completeness and decidability.

### 2019-05-21 sections/Overview.tex
Clarify the freshness condition

@@
 
-\jimmy{TODO: Mention the implicit freshness conditions in the premises throughout the text}
-
 \begin{figure}[t]
@@
 Both systems are adopted by several practical programming languages.
+
+\jimmy{Mention the implicit freshness conditions in the premises throughout the text}
+Please note that when we introduce a new binder in the premise, we implicitly pick a fresh one.
+This applies to rules such as $\mathtt{wf_d\forall}$, $\mathtt{wf_dabs}$, $\mathtt{{\le}\forall R}$,
+throughout the whole text.
+For algorithmic rules in Figure~\ref{fig:alg}, the right-hand side $\Gm'$
+is actually the premise if re-written in bit-step style.
 

### 2019-05-21 sections/Discussion.tex
minor edit

@@
 But there are issues about the well-formedness condition.
-Normally $\Gm \vdash_{wf} (e : A)$ is defined by $\Gm \vdash_{wf} e$ and $\Gm \vdash_{wf} A$.
-When $A = \all A'$, the type variable $a$ is not introduced to the context of $e$,
-therefore $\Gm \vdash_{wf} e$ is not derivable.
+Normally $\Psi \vdash (e : A)$ is defined by $\Psi \vdash e$ and $\Psi \vdash A$.
+However, when $A = \all A'$, the type variable $a$ is not introduced to the context of $e$,
+therefore $\Psi \vdash e$ is not derivable.
 
-To address the problem, one can either
-1) change the well-formedness judgment by introducing the binder for terms as well;
-or 2) add a syntax that explicitly bind the generic type variable.
-The former approach requires to resolve overlapping cases
-$(e : \all A)$ with $(e : B)$ where $B \neq \all B'$.
-Therefore we choose the latter approach due to its simplicity.
+To address the problem,
+we add a syntax that explicitly bind the generic type variable into a function and its annotation.
 The syntax of expression is extended with one more constructor:
@@
 
-Notice that the type-lambda syntax $\Lambda a.~e:A$ annotates the generic function $e$ with $\all A$,
+The new type-lambda syntax $\Lambda a.~e:A$ actually annotates the generic function $e$ with $\all A$,
 while making $a$ visible inside the body of the function.
+The well-formedness judgments are updated naturally:
+$$
+\inferrule*[right=$\mathtt{wf_d}\Lambda$]
+    {\Psi,a \vdash e \quad \Psi,a \vdash A}
+    {\Psi \vdash \Lambda a.~e:A}
+\qquad
+\inferrule*[right=$\mathtt{wf\_}\Lambda$]
+    {\Gm,a \vdash e \quad \Gm,a \vdash A}
+    {\Gm \vdash \Lambda a.~e:A}
+$$
+
 Corresponding rules are introduced for both the declarative system and the algorithmic system:
@@
 $$\Gm \Vdash \Lambda a.~e:A \To_b \jg \algrule \Gm \Vdash [(\all A)/b]\jg, a \Vdash e \Lto A$$
-
-The new syntax for expressions also affect the well-formedness judgment:
-$$
-\inferrule*[right=$\mathtt{wf\_}\Lambda$]
-    {\Gm,a \vdash_{wf} e \quad \Gm,a \vdash_{wf} A}
-    {\Gm \vdash_{wf} \Lambda a.~e:A}
-$$
 

### 2019-05-21 sections/Discussion.tex
minor typos

@@
 
-In a practical point of view, programmers may not write the syntax $\Lambda a. e : A$ directly.
+In a practical point of view, programmers may not write the syntax $\Lambda a.~e : A$ directly.
 The ScopedTypeVariables extension of Haskell is effective only
 when the type signature is explicitly universally quantified
-(the compiler translates into an expression similar to $\Lambda a. e : A$);
+(the compiler translates into an expression similar to $\Lambda a.~e : A$);
 otherwise the program means the normal syntax $e : \all A$,

### 2019-05-22 sections/algorithm.tex
Modification according to comments and known typos
Summarize current changes against reviews

@@
 \end{aligned}
+\end{gather*}
+
+\centering \framebox{$\Gm \vdash A$}
+\begin{gather*}
+\inferrule*[right=$\mathtt{wf\_ unit}$]
+    {~}{\Gm\vdash 1}
+\qquad
+\inferrule*[right=$\mathtt{wf\_ var}$]
+    {a\in\Gm}{\Gm\vdash a}
+\qquad
+\inferrule*[right=$\mathtt{wf\_ exvar}$]
+    {\al\in\Gm}{\Gm\vdash \al}
+\qquad
+\inferrule*[right=$\mathtt{wf\_{\to}}$]
+    {\Gm\vdash A\quad \Gm\vdash B}
+    {\Gm\vdash A\to B}
+\qquad
+\inferrule*[right=$\mathtt{wf\_\forall}$]
+    {\Gm, a\vdash A}
+    {\Gm\vdash \forall a. A}
+\end{gather*}
+
+\centering \framebox{$\Gm \vdash e$}
+\begin{gather*}
+\inferrule*[right=$\mathtt{wf\_ tmvar}$]
+    {x:A\in\Gm}{\Gm\vdash x}
+\qquad
+\inferrule*[right=$\mathtt{wf\_ tmunit}$]
+    {~}{\Gm\vdash ()}
+\qquad
+\inferrule*[right=$\mathtt{wf\_ abs}$]
+    {\Gm,x:A\vdash e}
+    {\Gm\vdash \lam e}
+\\
+\inferrule*[right=$\mathtt{wf\_ app}$]
+    {\Gm\vdash e_1 \quad \Gm\vdash e_2}
+    {\Gm\vdash e_1~e_2}
+\qquad
+\inferrule*[right=$\mathtt{wf\_ anno}$]
+    {\Gm\vdash A \quad \Gm\vdash e}
+    {\Gm\vdash (e:A)}
 \end{gather*}
@@
 \Gm[\al] \Vdash \al \le A\to B &\algrule [\al[1]\to\al[2]/\al] (\Gm[\al[1], \al[2]] \Vdash \al[1]\to \al[2] \le A \to B)\\
- &\qquad \text{when }\al\notin FV(A)\cup FV(B)\\
+ &\qquad\qquad \text{when }\al\notin FV(A)\cup FV(B)\\
 \Gm[\al] \Vdash A\to B \le \al &\algrule [\al[1]\to \al[2]/\al] (\Gm[\al[1], \al[2]] \Vdash A \to B \le \al[1]\to \al[2])\\
- &\qquad \text{when }\al\notin FV(A)\cup FV(B)
+ &\qquad\qquad \text{when }\al\notin FV(A)\cup FV(B)
  \\[3mm]
@@
 
-\paragraph{{\bf 1. Declarative Subtyping}}
+\paragraph{{\bf 1. Algorithmic Subtyping}}
 We have six subtyping rules (Rules 4-9) that are similar to their
@@
 implementable. Indeed, although written in the form of reduction rules, they do
-not overlap and are thus deterministic.  This has the added benefit that our
-algorithm runs in polynomial time, which makes it suitable for practical use.
+not overlap and are thus deterministic.
+%This has the added benefit that our % INCORRECT!
+%algorithm runs in polynomial time, which makes it suitable for practical use.
 

### 2019-05-23 sections/Appendix.tex
Translation table: paper symbol to Abella relation

@@
 $|A'|_\forall$ decreases, and thus the induction hypothesis finishes this case.
+
+\subsection{Translation Table for the Proof Scripts}
+In the proof scripts, we use textual relational definitions
+rather than the symbolic ones used in the paper.
+The mapping should be helpful for anyone who reads the script.
+
+\afterpage{%
+\clearpage% Flush earlier floats (otherwise order might not be correct)
+\begin{landscape}% Landscape page
+
+\begin{figure}[ht]
+\centering
+\begin{tabular}{c|c|c|c}
+Symbol & Abella textual relation & File (.thm) & Description\\\hline
+\hline
+$A$    & {\tt ty, wft, wfta} & typing &
+    The "type" type, decl./alg. well-formedness\\\hline
+$e$    & {\tt tm, wftm} & typing &
+    The "term" type, (alg.) well-formedness\\\hline
+$\Psi, \Gm$ & {\tt olist, wfj} & typing &
+    {\tt olist} is Abella's built-in list type, i.e. {\tt [o]}\\\hline
+$\jg$  & {\tt judgment, wfjg} & typing &
+    Judgment chain and its well-formedness\\\hline
+\hline
+$\Gm\rto\Gm'$ & {\tt reduction} & smallStep &
+    Algorithmic reduction: paper version\\\hline
+$\Gm\redto\nil$ & {\tt judge} & typing &
+    ({\tt judge} $\Gm$): a success reduction on $\Gm$\\\hline
+\hline
+$\Gm \sto \Om$ & {\tt tex} & trans &
+    "transfer existential variables"\\\hline
+$\Om\redto\nil$ & {\tt dc, dcl} & trans, dcl &
+    Declarative chain representation\\\hline
+\hline
+$\Psi'\le\Psi$ & {\tt esub} & declTyping & Declarative context subtyping\\\hline
+\hline
+$\Gm\jExt\Gm'$ & {\tt jExt} & inst\_decidable & "Judgment extension"\\\hline
+$\Gm_i$ & {\tt iexp} & inst\_decidable & instantiation judgments ($\Gm \vdash_{wf} \Gm_i$)\\\hline
+$\Gm_\le$ & {\tt subExp} & inst\_decidable & subtyping judgments\\\hline
+\hline
+$|\cdot|_e$ & {\tt tmSize, tmSizeJ, tmSizel} & declTyping, decidability &
+    Size measure for term, judgment chain or context\\\hline
+$|\cdot|_\Leftrightarrow$ & {\tt m\_judgeJ, m\_judge} & decidability &
+    Judgment measure for judgment chain or context\\\hline
+$|\cdot|_{\al}$ & {\tt nVar} & inst\_decidable &
+    Existential measure for context\\\hline
+$|\cdot|_\forall$ & {\tt order, orderJ, orderl} & order, inst\_decidable &
+    Polymorphism measure for type, judgment chain or context\\\hline
+$|\cdot|_\to$ & {\tt depth, depthJ, depthl} & order, decidability &
+    Function measure for type, judgment chain or context\\\hline
+\end{tabular}
+\caption{Translation Table for the Proof Scripts}
+\label{fig:translation_table}
+\end{figure}
+
+\end{landscape}
+\clearpage% Flush page
+}

### 2019-05-28 main.tex
Translation table: paper symbol to Abella relation
Some revisions

@@
-%% For double-blind review submission, w/o CCS and ACM Reference (max submission space)
-\documentclass[acmsmall,review,anonymous]{acmart}\settopmatter{printfolios=true,printccs=false,printacmref=false}
+%% For double-blind review submission, w/o CCS and ACM Reference (max
+%% submission space)
+\documentclass[acmsmall,review]{acmart}\settopmatter{printfolios=true}
+%%\documentclass[acmsmall,review,anonymous]{acmart}\settopmatter{printfolios=true,printccs=false,printacmref=false}
 %% For double-blind review submission, w/ CCS and ACM Reference
@@
 \usepackage{graphicx}
+
+\usepackage{pdflscape}
+\usepackage{afterpage}
+% \usepackage{rotating}
 
@@
 a few reasons to use a new algorithm. Firstly, our new algorithm
-employs \emph{worklist judgments}, which allow precise capture of the
+employs \emph{worklist judgments}, which precisely capture the
 scope of variables and simplify the formalization of scoping in a
@@
 incorrect proofs,
-which complicate the task of writing a machanized proof. 
+which complicate the task of writing a mechanized proof. 
 Despite the use of a different algorithm we prove the
@@
   %% extraction tools.
-  This material is based upon work supported by the
-  \grantsponsor{GS100000001}{National Science
-    Foundation}{http://dx.doi.org/10.13039/100000001} under Grant
-  No.~\grantnum{GS100000001}{nnnnnnn} and Grant
-  No.~\grantnum{GS100000001}{mmmmmmm}.  Any opinions, findings, and
-  conclusions or recommendations expressed in this material are those
-  of the author and do not necessarily reflect the views of the
-  National Science Foundation.
+
+We sincerely thank the anonymous reviewers for their insightful
+comments. Ningning Xie found the issue with Lemma 29 that we reported 
+on this article. This work has been sponsored by the Hong Kong Research
+Grant Council projects number 17210617 and 17258816, and by the
+Research Foundation - Flanders.
 \end{acks}

### 2019-05-28 sections/Overview.tex
minor typos
Some revisions

@@
 term variables with their types declared $x:A$.
+
+\paragraph{Well-formedness} Well-formedness of types and terms is 
+shown at the top of Figure~\ref{fig:decl:sub}. The rules are standard
+and simply ensure that variables in types and terms are well-scoped.
 
@@
 
+%The DK subtyping relation was adopted from \citet{odersky1996putting}.
+
+\paragraph{Declarative Subtyping}
+The bottom of Figure~\ref{fig:decl:sub} shows DK's declarative subtyping judgment $\Psi \vdash A \le B$,
+which was adopted from \citet{odersky1996putting}. It compares the
+degree of polymorphism between $A$ and $B$ in DK's implicit polymorphic type system. 
+Essentially, if $A$ can always be instantiated to match any instantiation of $B$,
+then A is ``at least as polymorphic as'' $B$. We also 
+say that $A$ is ``more polymorphic than'' $B$ and write $A \le B$.
+
+Subtyping rules $\mathtt{{\le}Var}$, $\mathtt{{\le}Unit}$ and $\mathtt{{\le}{\to}}$
+handle simple cases that do not involve universal quantifiers.
+The subtyping rule for function types $\mathtt{{\le}{\to}}$ is standard,
+being covariant on the return type and contravariant on the argument type.
+Rule $\mathtt{{\le}\forall R}$ states that if $A$ is a subtype of $B$
+in the context $\Psi, a$, where $a$ is fresh in $A$, then $A\le\all B$.
+Intuitively, if $A$ is more general than $\all B$ (where the universal quantifier
+already indicates that $\all B$ is a general type),
+then $A$ must instantiate to $[\tau/a]B$ for every $\tau$.
+
+The most interesting rule is $\mathtt{{\le}\forall L}$.
+If some instantiation of $\all A$, $[\tau/a]A$, is a subtype of $B$,
+then $\all A \le B$.
+The monotype $\tau$ we used to instantiate $a$ is \emph{guessed} in this
+declarative rule, but the algorithmic system does not guess and defers the
+instantiation until it can determine the monotype deterministically.
+The fact that $\tau$ is a monotype rules out the possibility of
+polymorphic (or impredicative) instantiation.
+However this restriction ensures that the subtyping relation remains
+decidable. Allowing an arbitrary type (rather than a monotype) in rule $\mathtt{{\le}\forall L}$
+is known to give rise to an undecidable subtyping relation~\cite{tiuryn1996subtyping}.
+\citet{jones2007practical} also impose the restriction of
+predicative instantiation in their type system.
+Both systems are adopted by several practical programming languages.
+
+\jimmy{Mention the implicit freshness conditions in the premises throughout the text}
+Please note that when we introduce a new binder in the premise, we implicitly pick a fresh one.
+This applies to rules such as $\mathtt{wf_d\forall}$, $\mathtt{wf_dabs}$, $\mathtt{{\le}\forall R}$,
+throughout the whole text.
+For algorithmic rules in Figure~\ref{fig:alg}, the right-hand side $\Gm'$
+is actually the premise if re-written in a big-step style.
+\bruno{Comment about big-step seems out of place. Should come later,
+  when we actually talk about the algorithm.}
+
 \begin{figure}[t]
@@
 
-%The DK subtyping relation was adopted from \citet{odersky1996putting}.
-
-\paragraph{Declarative Subtyping}
-Figure~\ref{fig:decl:sub} shows DK's declarative subtyping judgment $\Psi \vdash A \le B$,
-which was adopted from \citet{odersky1996putting}. It compares the
-degree of polymorphism between $A$ and $B$ in DK's implicit polymorphic type system. 
-Essentially, if $A$ can always be instantiated to match any instantiation of $B$,
-then A is ``at least as polymorphic as'' $B$. We also 
-say that $A$ is ``more polymorphic than'' $B$ and write $A \le B$.
-
-Subtyping rules $\mathtt{{\le}Var}$, $\mathtt{{\le}Unit}$ and $\mathtt{{\le}{\to}}$
-handle simple cases that do not involve universal quantifiers.
-The subtyping rule for function types $\mathtt{{\le}{\to}}$ is standard,
-being covariant on the return type and contravariant on the argument type.
-Rule $\mathtt{{\le}\forall R}$ states that if $A$ is a subtype of $B$
-in the context $\Psi, a$, where $a$ is fresh in $A$, then $A\le\all B$.
-Intuitively, if $A$ is more general than $\all B$ (where the universal quantifier
-already indicates that $\all B$ is a general type),
-then $A$ must instantiate to $[\tau/a]B$ for every $\tau$.
-
-The most interesting rule is $\mathtt{{\le}\forall L}$.
-If some instantiation of $\all A$, $[\tau/a]A$, is a subtype of $B$,
-then $\all A \le B$.
-The monotype $\tau$ we used to instantiate $a$ is \emph{guessed} in this
-declarative rule, but the algorithmic system does not guess and defers the
-instantiation until it can determine the monotype deterministically.
-The fact that $\tau$ is a monotype rules out the possibility of
-polymorphic (or impredicative) instantiation.
-However this restriction ensures that the subtyping relation remains
-decidable. Allowing an arbitrary type (rather than a monotype) in rule $\mathtt{{\le}\forall L}$
-is known to give rise to an undecidable subtyping relation~\cite{tiuryn1996subtyping}.
-\citet{jones2007practical} also impose the restriction of
-predicative instantiation in their type system.
-Both systems are adopted by several practical programming languages.
-
-\jimmy{Mention the implicit freshness conditions in the premises throughout the text}
-Please note that when we introduce a new binder in the premise, we implicitly pick a fresh one.
-This applies to rules such as $\mathtt{wf_d\forall}$, $\mathtt{wf_dabs}$, $\mathtt{{\le}\forall R}$,
-throughout the whole text.
-For algorithmic rules in Figure~\ref{fig:alg}, the right-hand side $\Gm'$
-is actually the premise if re-written in bit-step style.
-
 \paragraph{Declarative Typing}
@@
 we check the body $e$ against $B$ in the context extended with $x:A$.
-However, synthesizing a lambda-function is a lot less easy, and 
+However, synthesizing a lambda-function is a lot harder, and 
 this type system only synthesizes monotypes $\sigma\to\tau$.
@@
 which gives more flexibility on the choice of $\tau$.
-Declarative typing rule $\mathtt{DeclSub}$ overlaps with
+The declarative typing rule $\mathtt{DeclSub}$ overlaps with
 both $\mathtt{Decl\forall I}$ and $\mathtt{Decl{\to}I}$.
@@
 
-More seriously, we have also found a lemma that simply does not hold, 
-Lemma 29 (Parallel Admissibility).
+More seriously, we have also found a lemma that simply does not hold:
+Lemma 29 (Parallel Admissibility)\footnote{Ningning Xie found the issue with Lemma 29 in 2016 while
+  collaborating with the second author on an earlier attempt to mechanically
+  formalize DK's algorithm. The authors acknowledged the problem after we contacted them through email.
+  Although they briefly mentioned that it should be possible to use a weaker lemma instead they did
+  not go into details.}.
 See Appendix~\ref{appendix:false_lemmas} for a detailed explanation
@@
 Unfortunately, we have not yet found a way to fix this problem.
-\footnote{We found the issue with Lemma 29 in 2016 in an earlier attempt to mechanically
-  formalize DK's algorithm. The authors acknowledged the problem after we contacted them through email.
-  Although they briefly mentioned that it should be possible to use a weaker lemma instead they did
-  not go into details.
-}
 
@@
 The complications of scoping and propagation are compelling reasons
-to look for a simpler algorithm that is easier to
+to look for another algorithm that is easier to
 reason about in a mechanized setting.
@@
 \subsection{Judgment Lists}\label{sec:overview:list}
-
+\bruno{Somewhere we should mention the ``Essence of ML type
+  inference'' as pointed out by our reviews/comments?}
 To avoid the problem of incrementally applying a substitution to remaining
@@
 subtyping algorithm and shown it correct with respect to DK's declarative
-subtyping judgment. This approach overcomes the some problems in DK's formulation
+subtyping judgment. This approach overcomes some problems in DK's formulation
 by using a \emph{worklist}-based judgment of the form $$\Gamma \vdash \Omega$$

### 2019-05-28 sections/algorithm.tex
Some revisions

@@
 
-Figure~\ref{fig:alg:syntax} shows the syntax and well-formedness judgments used by the algorithm.
+Figure~\ref{fig:alg:syntax} shows the syntax and well-formedness
+judgments used by the algorithm. Similarly to the declarative system 
+the well-formedness rules are unsurprising and merely ensure 
+well-scoping of variables.
 
@@
 the algorithm garbage-collects the $\al$
-variable at the perfect time: it leaves the environment immediately
+variable at the right time: it leaves the environment immediately
 after being unreferenced completely for sure.

### 2019-05-28 sections/introduction.tex
Some revisions

@@
 
-This paper presents the first full mechanical formalization of the
+This paper presents the first fully mechanized formalization of the
 metatheory for higher-ranked polymorphic type inference.
@@
 
-\item {\bf A full mechanical formalization of type inference with
-  higher-ranked types:} Our work presents the first full mechanical formalization
+\item {\bf A fully mechanized formalization of type inference with
+  higher-ranked types:} Our work presents the first fully mechanized formalization
   for type inference of higher ranked types. The formalization is done in the
   Abella theorem prover~\cite{AbellaDesc} and it is available
-  online.\footnote{{\bf Note to reviewers:} Due to the anonymous submission process
-  it is sent as part of the supplementary materials.}
+  online at \url{http://TODO}.\bruno{fill url please!}
 

### 2019-05-28 sections/Overview.tex
Some revisions

@@
 \subsection{Judgment Lists}\label{sec:overview:list}
-\bruno{Somewhere we should mention the ``Essence of ML type
-  inference'' as pointed out by our reviews/comments?}
 To avoid the problem of incrementally applying a substitution to remaining
@@
 In the above unification problem, all variables are implicitly bound in the same
-global scope. Recent unification algorithms for dependently-typed languages
-are more explicit about scopes. For instance, \citet{Reed2009} represents a unification
+global scope. 
+\bruno{Some constraint solving algorithms for Hindley-Milner type
+  inference use similar ideas, but are more careful tracking the
+  scopes of variables~\cite{}. However they have separate phases for 
+constraint generation and solving.}
+Recent unification algorithms for dependently-typed languages
+are also more explicit about scopes. For instance, \citet{Reed2009} represents a unification
 problem as $\Delta \vdash P$ where $P$ is a set of equations to be solved and $\Delta$ is

### 2019-05-28 sections/Related.tex
Some revisions

@@
 
+\paragraph{The Essence of ML Type-Inference} \bruno{we deal with
+  higher-ranked polymorphism and do not have separate phases for 
+constraint generation and solving. In such multi-phase algorithms 
+there is typically a different constraint language. In contrast in our
+single phase algorithm one could view the judgements as a form 
+of constraints, and no separate constraint language is needed.
+Other more technical differences?}
+
 \paragraph{Lists of Judgments in Unification}

### 2019-05-29 sections/introduction.tex
Update URL & proof scripts; multi-step relation

@@
   Abella theorem prover~\cite{AbellaDesc} and it is available
-  online at \url{http://TODO}.\bruno{fill url please!}
+  online at \url{https://github.com/JimmyZJX/TypingFormalization}.\bruno{fill url please!}
 

### 2019-05-29 sections/Overview.tex
ML Essence

@@
 global scope. 
-\bruno{Some constraint solving algorithms for Hindley-Milner type
+Some constraint solving algorithms for Hindley-Milner type
   inference use similar ideas, but are more careful tracking the
-  scopes of variables~\cite{}. However they have separate phases for 
-constraint generation and solving.}
+  scopes of variables~\cite{remy-attapl}. However they have separate phases for 
+constraint generation and solving.
 Recent unification algorithms for dependently-typed languages

### 2019-05-29 sections/introduction.tex
ML Essence

--- 2019-05-28 sections/introduction.tex
@@
   Abella theorem prover~\cite{AbellaDesc} and it is available
-  online at \url{http://TODO}.\bruno{fill url please!}
+  online at \url{https://github.com/JimmyZJX/TypingFormalization}.
 

### 2019-05-29 sections/Overview.tex
fix several TODOs
Some more changes

@@
 instance, a type inference task takes the form $e \Rightarrow_a \omega$.
-This form models a \emph{continuation passing style}:
-it
-denotes that we need to infer the type, say $A$, for expression $e$ and use it
+This form expresses that we need to infer the type, say $A$, for expression $e$ and use it
 in the chain $\omega$ by substituting it for the placeholder type variable $a$.
@@
 which behaves similarly to the variable declarations in the context.
+\bruno{Removed CPS mention here. But if we remove it here should we remove it
+everywhere? Tom, your input would be good regarding this.}
 

### 2019-05-29 sections/algorithm.tex
Update URL & proof scripts; multi-step relation
fix several TODOs
Some more changes

@@
 judgments onto the worklist.  The syntax $\Gm\redto \Gm'$ denotes multiple
-reduction steps. If $\Gm\redto\nil$, we say that $\Gm$ reduces successfully;
-this corresponds to successful type checking.
+reduction steps.
 
-Let us explain the rules in more details:
-Rules 1-3 pop a variable declaration. Essentially this is garbage: variables that are
+$$\inferrule*[right=$\mathtt{{\redto} id}$]
+{~}{\Gm\redto\Gm}
+\qquad
+\inferrule*[right=$\mathtt{{\redto} step}$]
+{\Gm \rto \Gm_1 \quad \Gm_1 \redto \Gm'}{\Gm \redto \Gm'}$$
+
+\noindent In the case that $\Gm\redto\nil$ this corresponds to successful type checking.
+
+Rules 1-3 pop variable declarations that are essentially garbage.
+That is variables that are
 out of scope for the remaining judgments in the worklist.
@@
 and they do not have a rule that explicitly handles the case $\lam e \Lto \al$.
-Therefore their only way to check a lambda function against an existential variable
+Therefore the only way to check a lambda function against an existential variable
 is by applying the subsumption rule, which further breaks into
-a type inference of the lambda function and a subtyping judgment.
+type inference of a lambda function and a subtyping judgment.
 % A combination of Rules $18'$ (Rule 18 without the restriction on $e$) and 25 and further derivations
@@
 \paragraph{Example}
+Figure~\ref{fig:alg:sample} shows a sample derivation of the algorithm.
+It checks the application $(\lam x)~()$ against the unit type.
+According to the algorithm, we apply Rule 18 (subsumption), changing to inference mode.
+Type inference of the application breaks into two steps by Rule 26:
+first we infer the type of the function,
+and then the application inference judgment helps to determine the return type.
+In the following 5 steps the type of the identity function, $\lam x$, is inferred to be $\al \to \al$:
+checking the body of the lambda function (Rule 25),
+switching from check mode to inference mode (Rule 18),
+inferring the type of a term variable (Rule 22),
+solving a subtyping between existential variables (Rule 12) and
+garbage collecting the term variable $x$ (Rule 3).
+
+After that, Rule 28 changes the application inference judgment to
+a check of the argument against the input type $\al$ and returns the output type $\al$.
+Checking $()$ against the existential variable $\al$ solves $\al$ to the unit type $1$
+through Rules 18, 24 and 16.
+Immediately after $\al$ is solved, the algorithm replaces every occurrence of $\al$ with $1$.
+Therefore the worklist remains a simple subtyping judgment $1 \le 1$, which is finished off by Rule 4.
+Finally we get an empty worklist, indicating the success of the whole derivation.
 
@@
 
-Figure~\ref{fig:alg:sample} shows a sample derivation of the algorithm.
-It checks the application $(\lam x)~()$ against the unit type.
-According to the algorithm, we apply Rule 18 (subsumption), changing to inference mode.
-Type inference of the application breaks into two steps by Rule 26:
-first we infer the type of the function,
-and then the application inference judgment helps to determine the return type.
-In the following 5 steps the type of the identity function, $\lam x$, is inferred to be $\al \to \al$:
-checking the body of the lambda function (Rule 25),
-switching from check mode to inference mode (Rule 18),
-inferring the type of a term variable (Rule 22),
-solving a subtyping between existential variables (Rule 12) and
-garbage collecting the term variable $x$ (Rule 3).
-
-After that, Rule 28 changes the application inference judgment to
-a check of the argument against the input type $\al$ and returns the output type $\al$.
-Checking $()$ against the existential variable $\al$ solves $\al$ to the unit type $1$
-through Rules 18, 24 and 16.
-Immediately after $\al$ is solved, the algorithm replaces every occurrence of $\al$ with $1$.
-Therefore the worklist remains a simple subtyping judgment $1 \le 1$, which is finished off by Rule 4.
-Finally we get an empty worklist, indicating the success of the whole derivation.
-
-In summary, our algorithm accepts the type checking $(\lam x)~() \Lto 1$.
+In summary, our type checking algorithm accepts $(\lam x)~() \Lto 1$.
 

### 2019-05-29 sections/Metatheory.tex
Some revisions

@@
 %-------------------------------------------------------------------------------
-\subsection{Termination and Decidability}
+\subsection{Decidability}
 
-Finally, we show that our algorithm terminates:
-\begin{theorem}[Termination]
-Given a well-formed input $\Gm$, the algorithm terminates.
+Finally, we show that our algorithm is decidable:
+\begin{theorem}[Decidability]
+Given a well-formed input $\Gm$, either $\Gm\redto\nil$ or not.
 \end{theorem}
-Our termination proof is based on a lexicographic group of induction measures\\
+Our decidability proof is based on a lexicographic group of induction measures\\
 $\langle |\Gm|_e, |\Gm|_\Leftrightarrow, |\Gm|_\forall, |\Gm|_{\al}, |\Gm|_\to + |\Gm| \rangle$
@@
 Let us go back to the whole algorithm and summarize our findings.
-The termination theorem is shown through a lexicographic group of induction measures
+The decidability theorem is shown through a lexicographic group of induction measures
 $\langle |\Gm|_e, |\Gm|_\Leftrightarrow, |\Gm|_\forall, |\Gm|_{\al}, |\Gm|_\to + |\Gm| \rangle$
@@
 
-Combining all three main results (soundness, completeness and termination), we conclude that the declarative system is decidable
+Combining all three main results (soundness, completeness and decidability), we conclude that the declarative system is decidable
 by means of our algorithm.
@@
 \subsection{Abella and Proof Statistics}
+\bruno{Comment about correspondence between small and big step should
+  probably come here.}
+
+\jimmy{Discuss briefly on the difference of encodings in the script and this paper.}
+While the algorithmic rules are in a small-step style,
+the proof script rewrites them into a big-step style,
+where inductions are more strightforward to handle.
+In addition, we do prove the equivalence of the two representations.
+
 We have chosen the Abella (v2.0.7-dev
@@
 are very helpful for our formalization.
-
-\jimmy{Discuss briefly on the difference of encodings in the script and this paper.}
-While the algorithmic rules are in a small-step style,
-the proof script rewrites them into big-step style,
-where inductions are more strightforward to handle.
-In addition, we do prove the equivalence of the two representations.
 
@@
     inst\_decidable.thm &  830 & 42  & Other worklist measures; Lemma~\ref{lemma:inst:decidable}\\\hline
-    decidability.thm    &  885 & 50  & Termination theorem and decidability corollary\\\hline
+    decidability.thm    &  885 & 50  & Decidability theorem and corollary\\\hline
+    smallStep.thm       &  120 &  2  & The equivalence between big-step and small-step\\\hline
     \hline\emph{Total}               & 7700 & 567 & (51 definitions in total)\\\hline

### 2019-05-29 sections/Metatheory.tex
Abella

@@
 \subsection{Abella and Proof Statistics}
-\bruno{Comment about correspondence between small and big step should
-  probably come here.}
-
-\jimmy{Discuss briefly on the difference of encodings in the script and this paper.}
-While the algorithmic rules are in a small-step style,
-the proof script rewrites them into a big-step style,
-where inductions are more strightforward to handle.
-In addition, we do prove the equivalence of the two representations.
 
@@
 
+While the algorithmic rules are in a small-step style,
+the proof script rewrites them into a big-step style,
+where inductions are easier to handle.
+In addition, we do prove the equivalence of the two representations.
+
 \paragraph{Statistics of the Proof}
@@
 Figure~\ref{fig:proof_statistics} briefly summarizes the contents of each file.
-The files are linearly dependent due to the limitation of Abella.
+The files are linearly dependent due to limitations of Abella.
 

### 2019-05-29 sections/Related.tex
RW: the Essence of ML TypeInference

@@
 
-\paragraph{The Essence of ML Type-Inference} \bruno{we deal with
-  higher-ranked polymorphism and do not have separate phases for 
+\paragraph{The Essence of ML Type-Inference}
+Constraint-based type inference is adopted by \citep{remy-attapl} for ML type systems,
+which does not employ higher-ranked polymorphism.
+Constraint generating and solving are typical phases for a constraint-based algorithm,
+and an intermediate language is used to describe constaints.
+Furthermore, their constraint solver utilizes a stack to track the state of
+the solving process.
+In contrast, our algorithm has a single phase,
+where the judgment chains themselves act as constraints,
+thus no separate constaint language is needed.
+However, the constraint language provides a general and reuseable interface
+that interprets the desired properties for type inference.
+
+\bruno{we deal with higher-ranked polymorphism and do not have separate phases for 
 constraint generation and solving. In such multi-phase algorithms 

### 2019-05-29 sections/Related.tex
"Type Inference"

@@
 
-\paragraph{The Essence of ML Type-Inference}
+\paragraph{The Essence of ML Type Inference}
 Constraint-based type inference is adopted by \citep{remy-attapl} for ML type systems,

### 2019-05-29 sections/Discussion.tex
Discussion

@@
 Furthermore we discuss and sketch an extension of our algorithm with
-an elaboration to a target calculus.
+an elaboration to a target calculus, and discuss an extension of our algorithm
+with scoped type variables~\cite{scoped-type-variables}.
 
@@
 \subsection{Lexically-scoped type variables}
-We further extend the type system to support a simple extension, Lexically-scoped type variables.
-From a practical point of view,
-such extension enables the programmer to refer to type variables in a generic function signature
-when implementing of that function.
-For example,
+We further extend the type system to support a simple extension,
+lexically-scoped type variables~\cite{scoped-type-variables}.
+We have a complete Abella formalization for this extension proving
+all the metatheory discussed in the paper.
+
+From a
+practical point of view, such extension enables the programmer to
+refer to type variables in a generic function signature when
+implementing that function.  For example,
 $$(\lam{\lam[y] (x:a)}) : \all[a~b] a \to b \to a$$
@@
 
-In a practical point of view, programmers may not write the syntax $\Lambda a.~e : A$ directly.
+In practice, programmers would not write the syntax $\Lambda a.~e : A$ directly.
 The ScopedTypeVariables extension of Haskell is effective only

### 2019-05-29 sections/Metatheory.tex
Script statistics update; appendix table style

@@
 \paragraph{Statistics of the Proof}
-The proof script consists of 7683 lines of Abella code with a total of
-51 definitions and 567 theorems.
+The proof script consists of 7977 lines of Abella code with a total of
+51 definitions and 596 theorems.
 Figure~\ref{fig:proof_statistics} briefly summarizes the contents of each file.
@@
     File(s) & SLOC & \#Theorems & Description\\\hline
-    olist.thm, nat.thm  &  309 & 56  & Basic data structures\\\hline
-    typing.thm          &  238 & 5   & Declarative \& algorithmic system, debug examples\\\hline
+    olist.thm, nat.thm  &  311 & 57  & Basic data structures\\\hline
+    typing.thm          &  245 & 7   & Declarative \& algorithmic system, debug examples\\\hline
     decl.thm            &  226 & 33  & Basic declarative properties\\\hline
     order.thm           &  235 & 27  & The $|\cdot|_\forall$ measure; decl. subtyping strengthening\\\hline
-    alg.thm             &  655 & 72  & Basic algorithmic properties\\\hline
-    trans.thm           &  628 & 53  & \makecell[l]{Worklist instantiation and declarative transfer\\
+    alg.thm             &  679 & 80  & Basic algorithmic properties\\\hline
+    trans.thm           &  616 & 53  & \makecell[l]{Worklist instantiation and declarative transfer\\
                             Lemmas~\ref{lem:insert}, \ref{lem:extract}}\\\hline
-    declTyping.thm      &  944 & 72  & \makecell[l]{Non-overlapping declarative system \\
+    declTyping.thm      &  909 & 70  & \makecell[l]{Non-overlapping declarative system \\
                             Lemmas~\ref{lem:inv_allR}, \ref{lem:inv_chkAll},
                                 \ref{lem:inv_chkLam}, \ref{lem:subsumption}}\\\hline
-    soundness.thm       & 1086 & 70  & Soundness theorem; aux. lemmas on transfer\\\hline
+    soundness.thm       & 1107 & 78  & Soundness theorem; aux. lemmas on transfer\\\hline
     depth.thm           &  206 & 14  & The $|\cdot|_\to$ measure; Lemma~\ref{lem:prune_inst}\\\hline
-    dcl.thm             &  383 & 12  & Non-overlapping declarative worklist \\\hline
-    completeness.thm    & 1075 & 61  &
+    dcl.thm             &  380 & 12  & Non-overlapping declarative worklist \\\hline
+    completeness.thm    & 1124 & 61  &
                             Completeness theorem; aux. lemmas and relations\\\hline
-    inst\_decidable.thm &  830 & 42  & Other worklist measures; Lemma~\ref{lemma:inst:decidable}\\\hline
-    decidability.thm    &  885 & 50  & Decidability theorem and corollary\\\hline
-    smallStep.thm       &  120 &  2  & The equivalence between big-step and small-step\\\hline
-    \hline\emph{Total}               & 7700 & 567 & (51 definitions in total)\\\hline
+    inst\_decidable.thm &  837 & 45  & Other worklist measures; Lemma~\ref{lemma:inst:decidable}\\\hline
+    decidability.thm    &  983 & 57  & Decidability theorem and corollary\\\hline
+    smallStep.thm       &  119 &  2  & The equivalence between big-step and small-step\\\hline
+    \hline\emph{Total}               & 7977 & 596 & (51 definitions in total)\\\hline
 \end{tabular}

### 2019-05-29 sections/Overview.tex
further updates

@@
 
-\jimmy{Mention the implicit freshness conditions in the premises throughout the text}
+\jimmy{Mention the implicit freshness conditions in the premises}
 Please note that when we introduce a new binder in the premise, we implicitly pick a fresh one.
@@
 throughout the whole text.
-For algorithmic rules in Figure~\ref{fig:alg}, the right-hand side $\Gm'$
-is actually the premise if re-written in a big-step style.
-\bruno{Comment about big-step seems out of place. Should come later,
-  when we actually talk about the algorithm.}
 

### 2019-05-29 sections/algorithm.tex
Fig 4 spacing and title
further updates

@@
 
-\centering \framebox{$\Gm \vdash A$}
+\centering \framebox{$\Gm \vdash A$} Well-formed algorithmic type
 \begin{gather*}
@@
     {~}{\Gm\vdash 1}
-\qquad
+\quad
 \inferrule*[right=$\mathtt{wf\_ var}$]
     {a\in\Gm}{\Gm\vdash a}
-\qquad
+\quad
 \inferrule*[right=$\mathtt{wf\_ exvar}$]
     {\al\in\Gm}{\Gm\vdash \al}
-\qquad
+\quad
 \inferrule*[right=$\mathtt{wf\_{\to}}$]
@@
     {\Gm\vdash A\to B}
-\qquad
+\quad
 \inferrule*[right=$\mathtt{wf\_\forall}$]
@@
 
-\centering \framebox{$\Gm \vdash e$}
+\centering \framebox{$\Gm \vdash e$} Well-formed algorithmic expression
 \begin{gather*}
@@
 
+\jimmy{The implicit freshness conditions for algorithmic system}
+
+Please note that when a new binder is introduced in the right-hand side worklist $\Gm'$,
+we implicitly pick a fresh one,
+since the right-hand side can be seen as the premise of the reduction.
+
 Rules 1-3 pop variable declarations that are essentially garbage.
@@
 existential variable $\bt$ only if $\bt$ occurs after $\al$.
+
+One may notice that the subtyping relation for simple types is just equivalence,
+which is true according to the declarative system.
+The DK's system works in a similar way.
 

### 2019-05-29 sections/algorithm.tex
typos

@@
 
-Please note that when a new binder is introduced in the right-hand side worklist $\Gm'$,
+Please note that when a new variable is introduced in the right-hand side worklist $\Gm'$,
 we implicitly pick a fresh one,

### 2019-05-29 sections/Metatheory.tex
further updates
English and typos

@@
 in nature, but has no overlap. This intermediate system differs only in a few 
-rules from DK's declarative system; these are the changes:
+rules from DK's declarative system. The changes are:
 \begin{enumerate}
@@
 declarative system has an algorithmic counterpart.
+\jimmy{In response to "treat declarative and algorithmic worklists as a subsets":}
+We explicitly relate between an algorithmic context $\Gm$ and
+a declarative context $\Omega$ to avoid potential confusion.
 

### 2019-05-29 sections/Related.tex
English and typos

--- 2019-05-28 sections/Related.tex
@@
 
-\paragraph{The Essence of ML Type-Inference} \bruno{we deal with
-  higher-ranked polymorphism and do not have separate phases for 
-constraint generation and solving. In such multi-phase algorithms 
-there is typically a different constraint language. In contrast in our
-single phase algorithm one could view the judgements as a form 
-of constraints, and no separate constraint language is needed.
-Other more technical differences?}
+\paragraph{The Essence of ML Type Inference}
+Constraint-based type inference is adopted by \citet{remy-attapl} for
+ML type systems, which do not employ higher-ranked polymorphism. An
+interesting feature of their algorithm is that it keeps precise
+scoping of variables, similarly to our approach.  Their algorithm is
+divided into constraint generation and solving phases (which are
+typical of constraint-based algorithms). Furthermore an intermediate
+language is used to describe constraints and their constraint solver
+utilizes a stack to track the state of the solving process.  In
+contrast, our algorithm has a single phase, where the judgment chains
+themselves act as constraints, thus no separate constraint language is
+needed.
+
+%A possible advantage of a separate constraint language is to provide
+%a general and reuseable interface
+%that interprets the desired properties for type inference.
 

### 2019-05-29 sections/Discussion.tex
further updates
edits in 5.3

@@
 with scoped type variables~\cite{scoped-type-variables}.
-
-\jimmy{TODO: update the summary}
 
@@
 \subsection{Lexically-scoped type variables}
-We further extend the type system to support a simple extension,
+We fhave urther extended the type system with support for
 lexically-scoped type variables~\cite{scoped-type-variables}.
-We have a complete Abella formalization for this extension proving
-all the metatheory discussed in the paper.
+Our Abella formalization for this extension proves
+all the metatheory we discuss here.
 
 From a
-practical point of view, such extension enables the programmer to
-refer to type variables in a generic function signature when
-implementing that function.  For example,
+practical point of view, this extension allows the implementation of a function
+to refer to type variables from its type signature. For example,
 $$(\lam{\lam[y] (x:a)}) : \all[a~b] a \to b \to a$$
-has an inner annotation $(x:a)$ that refers to the outer generic type variable $a$.
-This is not a surprising feature, since the declarative system already accepts these kind of programs
+has an annotation $(x:a)$ that refers to the type variable $a$ in the type signature.
+This is not a surprising feature, since the declarative system already accepts similar programs
 $$
@@
 
-But there are issues about the well-formedness condition.
-Normally $\Psi \vdash (e : A)$ is defined by $\Psi \vdash e$ and $\Psi \vdash A$.
-However, when $A = \all A'$, the type variable $a$ is not introduced to the context of $e$,
+The main issue is the well-formedness condition.
+Normally $\Psi \vdash (e : A)$ follows from $\Psi \vdash e$ and $\Psi \vdash A$.
+However, when $A = \all A'$, the type variable $a$ is not in scope at $e$,
 therefore $\Psi \vdash e$ is not derivable.
 
-To address the problem,
-we add a syntax that explicitly bind the generic type variable into a function and its annotation.
-The syntax of expression is extended with one more constructor:
+To address the problem, we add a new syntactic form that explicitly binds a
+type variable simulatenously in a function and its annotation.
 $$
 \begin{array}{l@{\qquad}lcl}
-\text{Expressions}\qquad&e &::=&\quad x \mid () \mid \lam e \mid e_1~e_2 \mid (e:A) \mid \Lambda a.~e:A
+\text{Expressions}\qquad&e &::=&\quad \ldots \mid \Lambda a.~e:A
 \end{array}
@@
 
-The new type-lambda syntax $\Lambda a.~e:A$ actually annotates the generic function $e$ with $\all A$,
+This new type-lambda syntax $\Lambda a.~e:A$ actually annotates its body $e$ with $\all A$,
 while making $a$ visible inside the body of the function.
-The well-formedness judgments are updated naturally:
+The well-formedness judgments are extended accordingly:
 $$
@@
 when the type signature is explicitly universally quantified
-(the compiler translates into an expression similar to $\Lambda a.~e : A$);
-otherwise the program means the normal syntax $e : \all A$,
-and one may not later refer to the type variable $a$.
+(which the compiler translates into an expression similar to $\Lambda a.~e : A$);
+otherwise the program means the normal syntax $e : \all A$
+and may not later refer to the type variable $a$.
 
-We also obtained all three desired properties for the extended system,
+We have proven all three desired properties for the extended system,
 namely soundness, completeness and decidability.

### 2019-05-30 sections/Discussion.tex
typo

@@
 \subsection{Lexically-scoped type variables}
-We fhave urther extended the type system with support for
+We have further extended the type system with support for
 lexically-scoped type variables~\cite{scoped-type-variables}.
 Our Abella formalization for this extension proves
-all the metatheory we discuss here.
+all the metatheory we discuss in Section~\ref{sec:metatheory}.
 

### 2019-05-30 sections/Overview.tex
"worklist judgments"

@@
 In particular the support for inference judgments in
-judgment worklists, and a new form of judgment worklist
+worklist judgments, and a new form of worklist judgment
 that the unifies \emph{ordered contexts} and the worklists themselves.
@@
 % before analysing its sub-judgment(s).
-% This enables us to write the judgment worklist in such a form:
+% This enables us to write the worklist judgment in such a form:
 % $$\Gm, \{\text{variable declarations}\}, \{\text{judgment}\},

### 2019-05-30 sections/Metatheory.tex
shrinking text

--- 2019-05-29 sections/Metatheory.tex
@@
 in nature, but has no overlap. This intermediate system differs only in a few 
-rules from DK's declarative system; these are the changes:
+rules from DK's declarative system:
 \begin{enumerate}
@@
 For any worklist $\Gm$ that reduces successfully,
-there is a valid instantiation $\Om$ that transfers all judgments successfully to the declarative system.
+there is a valid instantiation $\Om$ that transfers all judgments
+to the declarative system.
 \begin{theorem}[Soundness]
@@
 declarative system has an algorithmic counterpart.
+% \jimmy{In response to "treat declarative and algorithmic worklists as a subsets":}
+We explicitly relate between an algorithmic context $\Gm$ and
+a declarative context $\Omega$ to avoid potential confusion.
 
@@
 \begin{corollary}[Completeness, single judgment form]
-Given \emph{wf }$\Gm$ containing no judgment chains:
+Given \emph{wf }$\Gm$ containing no judgments:
 \begin{enumerate}
@@
 \begin{theorem}[Decidability]
-Given a well-formed input $\Gm$, either $\Gm\redto\nil$ or not.
+Given \emph{wf} $\Gm$, it is decidable whether $\Gm\redto\nil$ or not.
 \end{theorem}
@@
 in a state with an overall smaller measure. Our starting point is a 
-worklist of the form $\Gm,\Gm_i$ where $\Gm_i$ are instantiation judgments.
+worklist $\Gm,\Gm_i$ where $\Gm_i$ are instantiation judgments.
 \begin{gather*}
@@
 just fixed a serious bug that accepts a simple proof of false,
-which also affects the completeness of one of our proofs.
+which also affects our proof.
 Specifically, our scripts compile against commit 92829a of Abella's GitHub repository.}
@@
 its higher-order unification and the ease of formalizing substitution-intensive relations
-are very helpful for our formalization.
+are very helpful.
 
 While the algorithmic rules are in a small-step style,
-the proof script rewrites them into a big-step style,
-where inductions are easier to handle.
+the proof script rewrites them into a big-step style
+for easier inductions.
 In addition, we do prove the equivalence of the two representations.
@@
 The proof script consists of 7977 lines of Abella code with a total of
-51 definitions and 596 theorems.
+60 definitions and 596 theorems.
 Figure~\ref{fig:proof_statistics} briefly summarizes the contents of each file.
@@
 
-\begin{figure}[ht]
+\begin{figure}[t]
 \centering\begin{tabular}{|c|c|c|l|}\hline
@@
     smallStep.thm       &  119 &  2  & The equivalence between big-step and small-step\\\hline
-    \hline\emph{Total}               & 7977 & 596 & (51 definitions in total)\\\hline
+    \hline\emph{Total}               & 7977 & 596 & (60 definitions in total)\\\hline
 \end{tabular}

### 2019-05-30 sections/Overview.tex
shrinking text

@@
 \caption{%Well-formedness of Declarative Types and 
-Declarative Subtyping}\label{fig:decl:sub}
+Declarative Well-formedness and Subtyping}\label{fig:decl:sub}
 \end{figure}
@@
 
-\jimmy{Mention the implicit freshness conditions in the premises}
+% \jimmy{Mention the implicit freshness conditions in the premises}
 Please note that when we introduce a new binder in the premise, we implicitly pick a fresh one.
@@
 Because the whole set is explicitly available, the variable $x$ can be
-simultaneously substituted everywhere.
+simultaneously substituted.
 
@@
 %just as what we have mentioned above about the scope tracking.
-After removing the garbage term variable declaration from the worklist, we process the last remaining judgment chain,
-namely an inference judgment $(\lam x)~()\To~?$, with the unit type $1$ as its result.
+After removing the garbage term variable declaration from the worklist, we process the last remaining
+inference judgment $(\lam x)~()\To~?$, with the unit type $1$ as its result.
 Finally the last judgment becomes $1 \le 1$, a trivial base case.

### 2019-05-30 sections/Related.tex
shrinking text

@@
 mechanized a type inference algorithm,
-with the help of locally nameless cofinite quantification~\cite{LocallyNameless},
+with the help of locally nameless~\cite{LocallyNameless},
 for Core ML extended with structural polymorphism and recursion.
@@
 Our algorithm, in contrast,
-eliminates the use of output contexts and their variable solution records ($\al = \tau$),
+eliminates output contexts and solution records ($\al = \tau$),
 simplifying the information propagation process through immediate substitution
@@
 resulting in a simpler judgment form,
-while processing them in a similar way to those algorithms.
+while processing them in a similar way.
 One important difference is that contexts are
@@
 since the information of each local context needs to be synchronized.
-Instead we make use of the nature of ordered context to control the scoping of unification variables.
+Instead we make use of the nature of ordered context to control the scopes of unification variables.
 While their algorithms focus only on unification,
-our algorithm also deals with other types of judgments like type inference.
+our algorithm also deals with other types of judgments like synthesis.
 A detailed discussion is in Section~\ref{sec:overview:list}.

### 2019-05-30 sections/algorithm.tex
shrinking text

@@
 the well-formedness rules are unsurprising and merely ensure 
-well-scoping of variables.
+well-scopedness.
 
@@
 
-\jimmy{The implicit freshness conditions for algorithmic system}
-
+% \jimmy{The implicit freshness conditions for algorithmic system}
 Please note that when a new variable is introduced in the right-hand side worklist $\Gm'$,
@@
 \paragraph{Example} The derivation of the subtyping judgment
-$(1\to 1)\to 1 \le (\all 1\to 1) \to 1$ proceeds as follows:
+$(1\to 1)\to 1 \le (\all 1\to 1) \to 1$:
 \begin{gather*}
@@
 by Rule 25 to $\Gm, \bt[1], \bt[2] \Vdash \bt[1] \to \bt[2] \le \al, x:\bt[1] \Vdash e \Lto \bt[2]$.
-As a subtyping constraint between two monotypes only succeeds when the types are equal, $\bt[1] \to \bt[2] \le \al$
+As a subtyping constraint between monotypes is simply equality, $\bt[1] \to \bt[2] \le \al$
 must end up equating $\bt[1] \to \bt[2]$ with $\al$ and thus have the same effect as Rule 21, but in a more roundabout

### 2019-06-20 sections/Overview.tex
remove comments about CPS
typos

@@
 % \jimmy{Mention the implicit freshness conditions in the premises}
-Please note that when we introduce a new binder in the premise, we implicitly pick a fresh one.
+Note that when we introduce a new binder in the premise, we implicitly pick a fresh one.
 This applies to rules such as $\mathtt{wf_d\forall}$, $\mathtt{wf_dabs}$, $\mathtt{{\le}\forall R}$,
@@
 
-\jimmy{Comment A is not satisfied with the term "continuation passing style" for our definition.}
+%\jimmy{Comment A is not satisfied with the term "continuation passing style" for our definition.}
 
@@
 which behaves similarly to the variable declarations in the context.
-\bruno{Removed CPS mention here. But if we remove it here should we remove it
-everywhere? Tom, your input would be good regarding this.}
+%\bruno{Removed CPS mention here. But if we remove it here should we remove it
+%everywhere? Tom, your input would be good regarding this.}
 

### 2019-06-24 sections/algorithm.tex
typos
shrink to fit in 27 pages

@@
 
-\paragraph{Example} The derivation of the subtyping judgment
+\paragraph{Example} Consider the derivation of the subtyping judgment
 $(1\to 1)\to 1 \le (\all 1\to 1) \to 1$:
@@
 Immediately after $\al$ is solved, the algorithm replaces every occurrence of $\al$ with $1$.
-Therefore the worklist remains a simple subtyping judgment $1 \le 1$, which is finished off by Rule 4.
-Finally we get an empty worklist, indicating the success of the whole derivation.
+Therefore the worklist remains $1 \le 1$, which is finished off by Rule 4.
+Finally, the empty worklist indicates the success of the whole derivation.
 

### 2019-06-25 main.tex
minor typo and email
shrink to fit in 27 pages
minor format: capitalizations, etc

@@
 %% submission space)
-\documentclass[acmsmall,review]{acmart}\settopmatter{printfolios=true}
+%\documentclass[acmsmall,review]{acmart}\settopmatter{printfolios=true}
 %%\documentclass[acmsmall,review,anonymous]{acmart}\settopmatter{printfolios=true,printccs=false,printacmref=false}
@@
 %% For final camera-ready submission, w/ required CCS and ACM Reference
-%\documentclass[acmsmall]{acmart}\settopmatter{}
+\documentclass[acmsmall]{acmart}\settopmatter{}
 
@@
 %% use defaults for review submission.
-\acmJournal{PACMPL}
-\acmVolume{1}
-\acmNumber{ICFP} % CONF = POPL or ICFP or OOPSLA
-\acmArticle{1}
-\acmYear{2018}
-\acmMonth{1}
-\acmDOI{} % \acmDOI{10.1145/nnnnnnn.nnnnnnn}
-\startPage{1}
+%\acmJournal{PACMPL}
+%\acmVolume{1}
+%\acmNumber{ICFP} % CONF = POPL or ICFP or OOPSLA
+%\acmArticle{1}
+%\acmYear{2018}
+%\acmMonth{1}
+%\acmDOI{} % \acmDOI{10.1145/nnnnnnn.nnnnnnn}
+%\startPage{1}
 
@@
 %% use 'none' for review submission.
-\setcopyright{none}
+%\setcopyright{none}
 %\setcopyright{acmcopyright}
@@
 %\usepackage[-27]{pagesel}
+
+%%% The following is specific to ICFP '19 and the paper
+%%% 'A Mechanical Formalization of Higher-Ranked Polymorphic Type Inference'
+%%% by Jinxu Zhao, Bruno C. d. S. Oliveira, and Tom Schrijvers.
+%%%
+\setcopyright{acmcopyright}
+\acmPrice{}
+\acmDOI{10.1145/3341716}
+\acmYear{2019}
+\copyrightyear{2019}
+\acmJournal{PACMPL}
+\acmVolume{3}
+\acmNumber{ICFP}
+\acmArticle{112}
+\acmMonth{8}
 
@@
 }
+\email{bruno@cs.hku.hk}          %% \email is recommended
 
@@
 <concept>
-<concept_id>10011007.10011006.10011008</concept_id>
-<concept_desc>Software and its engineering~General programming languages</concept_desc>
+<concept_id>10011007.10011006.10011008.10011009.10011012</concept_id>
+<concept_desc>Software and its engineering~Functional languages</concept_desc>
 <concept_significance>500</concept_significance>
@@
 <concept>
-<concept_id>10003456.10003457.10003521.10003525</concept_id>
-<concept_desc>Social and professional topics~History of programming languages</concept_desc>
-<concept_significance>300</concept_significance>
+<concept_id>10011007.10011006.10011008.10011024.10011025</concept_id>
+<concept_desc>Software and its engineering~Polymorphism</concept_desc>
+<concept_significance>500</concept_significance>
+</concept>
+<concept>
+<concept_id>10011007.10011006.10011039</concept_id>
+<concept_desc>Software and its engineering~Formal language definitions</concept_desc>
+<concept_significance>500</concept_significance>
 </concept>
@@
 
-\ccsdesc[500]{Software and its engineering~General programming languages}
-\ccsdesc[300]{Social and professional topics~History of programming languages}
+\ccsdesc[500]{Software and its engineering~Functional languages}
+\ccsdesc[500]{Software and its engineering~Polymorphism}
+\ccsdesc[500]{Software and its engineering~Formal language definitions}
 %% End of generated code
@@
 
+%% Appendix
+\newpage
+\appendix
+\input{sections/Appendix.tex}
+
+\newpage
 %% Acknowledgments
@@
 We sincerely thank the anonymous reviewers for their insightful
-comments. Ningning Xie found the issue with Lemma 29 that we reported 
+comments. Ningning Xie found the issue with Lemma 29 in DK's formalization that we reported 
 on this article. This work has been sponsored by the Hong Kong Research
@@
 
-%% Appendix
-\newpage
-\appendix
-\input{sections/Appendix.tex}
-
 \end{document}

### 2019-06-25 sections/Discussion.tex
shrink to fit in 27 pages
minor format: capitalizations, etc

@@
 
-\subsection{Contrasting our scoping mechanisms with DK's scoping}\label{sec:discussion:scoping}
+\subsection{Contrasting Our Scoping Mechanisms with DK's}\label{sec:discussion:scoping}
 
@@
 
-\subsection{Lexically-scoped type variables}
+\subsection{Lexically-Scoped Type Variables}
 We have further extended the type system with support for
@@
 therefore $\Psi \vdash e$ is not derivable.
-
 To address the problem, we add a new syntactic form that explicitly binds a

### 2019-06-25 sections/Metatheory.tex
minor typo and email
minor format: capitalizations, etc

@@
 
-Moreover, we can characterise the reduction of the instantiation judgments as follows.
+Moreover, we can characterize the reduction of the instantiation judgments as follows.
 \begin{lemma}[Instantiation Decidability]\label{lemma:inst:decidable}
@@
 60 definitions and 596 theorems.
-Figure~\ref{fig:proof_statistics} briefly summarizes the contents of each file.
+Figure~\ref{table:proof_statistics} briefly summarizes the contents of each file.
 The files are linearly dependent due to limitations of Abella.
 
-\begin{figure}[t]
+\begin{table}[t]
+\caption{Statistics for the proof scripts}
 \centering\begin{tabular}{|c|c|c|l|}\hline
@@
 \end{tabular}
-\caption{Statistics for the proof scripts}
-\label{fig:proof_statistics}
-\end{figure}
+\label{table:proof_statistics}
+\end{table}

### 2019-06-25 sections/algorithm.tex
minor format: capitalizations, etc
camera-ready: acmart ver., fig. Description

@@
 \end{gather*}
+\Description{Extended Syntax and Well-Formedness for the Algorithmic System}
 \caption{Extended Syntax and Well-Formedness for the Algorithmic System}\label{fig:alg:syntax}
@@
 \end{gather*}
+\Description{Algorithmic Typing}
 \caption{Algorithmic Typing}\label{fig:alg}
@@
 
-\paragraph{{\bf 1. Algorithmic Subtyping}}
+\paragraph{{\bf 1. Algorithmic subtyping}}
 We have six subtyping rules (Rules 4-9) that are similar to their
@@
 \end{gather*}
+\Description{A Sample Derivation for Algorithmic Typing}
 \caption{A Sample Derivation for Algorithmic Typing}

### 2019-06-25 sections/Appendix.tex
Script statistics update; appendix table style
minor format: capitalizations, etc
camera-ready: acmart ver., fig. Description
pretty tables

--- 2019-05-20 sections/Appendix.tex
@@
     \end{minipage}
+\Description{Worklist measures}
 \caption{Worklist measures}
@@
 
-\subsection{Broken Lemmas in DK's papers}
+\subsection{Broken Lemmas in DK's Papers}
 \label{appendix:false_lemmas}
@@
 $|A'|_\forall$ decreases, and thus the induction hypothesis finishes this case.
+
+\subsection{Translation Table for the Proof Scripts}
+In the proof scripts, we use textual relational definitions
+rather than the symbolic ones used in the paper.
+The mapping should be helpful for anyone who reads the script.
+
+\afterpage{%
+\clearpage% Flush earlier floats (otherwise order might not be correct)
+\begin{landscape}% Landscape page
+
+\begin{table}[ht]
+\renewcommand{\arraystretch}{1.3} 
+\caption{Translation Table for the Proof Scripts}
+\centering
+\begin{tabular}{@{}llll@{}}
+\toprule
+Symbol & Abella textual relation & File (.thm) & Description\\
+\midrule
+$A$    & {\tt ty, wft, wfta} & typing &
+    The "type" type, decl./alg. well-formedness\\
+$e$    & {\tt tm, wftm} & typing &
+    The "term" type, (alg.) well-formedness\\
+$\Psi, \Gm$ & {\tt olist, wfj} & typing &
+    {\tt olist} is Abella's built-in list type, i.e. {\tt [o]}\\
+$\jg$  & {\tt judgment, wfjg} & typing &
+    Judgment chain and its well-formedness\\
+
+$\Gm\rto\Gm'$ & {\tt reduction} & smallStep &
+    Algorithmic reduction: paper version\\
+$\Gm\redto\nil$ & {\tt judge} & typing &
+    ({\tt judge} $\Gm$): a success reduction on $\Gm$\\
+
+$\Gm \sto \Om$ & {\tt tex} & trans &
+    "transfer existential variables"\\
+$\Om\redto\nil$ & {\tt dc, dcl} & trans, dcl &
+    Declarative chain representation\\
+
+$\Psi'\le\Psi$ & {\tt esub} & declTyping & Declarative context subtyping\\
+
+$\Gm\jExt\Gm'$ & {\tt jExt} & inst\_decidable & "Judgment extension"\\
+$\Gm_i$ & {\tt iexp} & inst\_decidable & instantiation judgments ($\Gm \vdash_{wf} \Gm_i$)\\
+$\Gm_\le$ & {\tt subExp} & inst\_decidable & subtyping judgments\\
+
+$|\cdot|_e$ & {\tt tmSize, tmSizeJ, tmSizel} & declTyping, decidability &
+    Size measure for term, judgment chain or context\\
+$|\cdot|_\Leftrightarrow$ & {\tt m\_judgeJ, m\_judge} & decidability &
+    Judgment measure for judgment chain or context\\
+$|\cdot|_{\al}$ & {\tt nVar} & inst\_decidable &
+    Existential measure for context\\
+$|\cdot|_\forall$ & {\tt order, orderJ, orderl} & order, inst\_decidable &
+    Polymorphism measure for type, judgment chain or context\\
+$|\cdot|_\to$ & {\tt depth, depthJ, depthl} & order, decidability &
+    Function measure for type, judgment chain or context\\
+\bottomrule
+\end{tabular}
+\label{table:translation_table}
+\end{table}
+
+\end{landscape}
+\clearpage% Flush page
+}

### 2019-06-25 sections/Metatheory.tex
camera-ready: acmart ver., fig. Description
pretty tables

@@
 \end{gather*}
+\Description{Declarative Worklists and Instantiation}
 \caption{Declarative Worklists and Instantiation}
@@
 \end{gather*}
-\caption{Declarative Transfer.}
+\Description{Declarative Transfer}
+\caption{Declarative Transfer}
 \label{fig:decl:worklist}
@@
 \end{aligned}$$
+\Description{Context Subtyping}
 \caption{Context Subtyping}
@@
 \end{gather*}
+\Description{Worklist Update}
 \caption{Worklist Update}\label{fig:worklist_ext}
@@
 \paragraph{Statistics of the Proof}
-The proof script consists of 7977 lines of Abella code with a total of
+The proof script consists of 7,977 lines of Abella code with a total of
 60 definitions and 596 theorems.
@@
 \begin{table}[t]
+\renewcommand{\arraystretch}{1.2}
 \caption{Statistics for the proof scripts}
-\centering\begin{tabular}{|c|c|c|l|}\hline
-    File(s) & SLOC & \#Theorems & Description\\\hline
-    olist.thm, nat.thm  &  311 & 57  & Basic data structures\\\hline
-    typing.thm          &  245 & 7   & Declarative \& algorithmic system, debug examples\\\hline
-    decl.thm            &  226 & 33  & Basic declarative properties\\\hline
-    order.thm           &  235 & 27  & The $|\cdot|_\forall$ measure; decl. subtyping strengthening\\\hline
-    alg.thm             &  679 & 80  & Basic algorithmic properties\\\hline
-    trans.thm           &  616 & 53  & \makecell[l]{Worklist instantiation and declarative transfer\\
-                            Lemmas~\ref{lem:insert}, \ref{lem:extract}}\\\hline
-    declTyping.thm      &  909 & 70  & \makecell[l]{Non-overlapping declarative system \\
+\centering\begin{tabular}{@{}lrrl@{}}
+\toprule
+    File(s) & SLOC & \#Theorems & Description\\
+\midrule
+    olist.thm, nat.thm  &   311 & 57  & Basic data structures\\
+    typing.thm          &   245 & 7   & Declarative \& algorithmic system, debug examples\\
+    decl.thm            &   226 & 33  & Basic declarative properties\\
+    order.thm           &   235 & 27  & The $|\cdot|_\forall$ measure; decl. subtyping strengthening\\
+    alg.thm             &   679 & 80  & Basic algorithmic properties\\
+    trans.thm           &   616 & 53  & \makecell[l]{Worklist instantiation and declarative transfer;\\
+                            Lemmas~\ref{lem:insert}, \ref{lem:extract}}\\
+    declTyping.thm      &   909 & 70  & \makecell[l]{Non-overlapping declarative system; \\
                             Lemmas~\ref{lem:inv_allR}, \ref{lem:inv_chkAll},
-                                \ref{lem:inv_chkLam}, \ref{lem:subsumption}}\\\hline
-    soundness.thm       & 1107 & 78  & Soundness theorem; aux. lemmas on transfer\\\hline
-    depth.thm           &  206 & 14  & The $|\cdot|_\to$ measure; Lemma~\ref{lem:prune_inst}\\\hline
-    dcl.thm             &  380 & 12  & Non-overlapping declarative worklist \\\hline
-    completeness.thm    & 1124 & 61  &
-                            Completeness theorem; aux. lemmas and relations\\\hline
-    inst\_decidable.thm &  837 & 45  & Other worklist measures; Lemma~\ref{lemma:inst:decidable}\\\hline
-    decidability.thm    &  983 & 57  & Decidability theorem and corollary\\\hline
-    smallStep.thm       &  119 &  2  & The equivalence between big-step and small-step\\\hline
-    \hline\emph{Total}               & 7977 & 596 & (60 definitions in total)\\\hline
+                                \ref{lem:inv_chkLam}, \ref{lem:subsumption}}\\
+    soundness.thm       & 1,107 & 78  & Soundness theorem; aux. lemmas on transfer\\
+    depth.thm           &   206 & 14  & The $|\cdot|_\to$ measure; Lemma~\ref{lem:prune_inst}\\
+    dcl.thm             &   380 & 12  & Non-overlapping declarative worklist \\
+    completeness.thm    & 1,124 & 61  &
+                            Completeness theorem; aux. lemmas and relations\\
+    inst\_decidable.thm &   837 & 45  & Other worklist measures; Lemma~\ref{lemma:inst:decidable}\\
+    decidability.thm    &   983 & 57  & Decidability theorem and corollary\\
+    smallStep.thm       &   119 &  2  & The equivalence between big-step and small-step\\
+\midrule
+    \emph{Total}        & 7,977 & 596 & (60 definitions in total)\\
+\bottomrule
 \end{tabular}

### 2019-06-26 sections/Overview.tex
shrink to fit in 27 pages
camera-ready: acmart ver., fig. Description
split appendix from main & fix reference issues

@@
 \]
+\Description{Syntax of Declarative System}
 \caption{Syntax of Declarative System}\label{fig:decl:syntax}
@@
 \begin{figure}[t]
-\centering \framebox{$\Psi \vdash A$}
+\centering \framebox{$\Psi \vdash A$} Well-formed declarative type
 \begin{gather*}
@@
 
-\centering \framebox{$\Psi \vdash e$}
+\centering \framebox{$\Psi \vdash e$} Well-formed declarative expression
 \begin{gather*}
@@
 
-\centering \framebox{$\Psi \vdash A \le B$}
+\centering \framebox{$\Psi \vdash A \le B$} Declarative subtyping
 \begin{gather*}
@@
 \end{gather*}
+\Description{Declarative Well-formedness and Subtyping}
 \caption{%Well-formedness of Declarative Types and 
@@
 \end{gather*}
+\Description{Declarative Typing}
 \caption{Declarative Typing}\label{fig:decl:typing}
@@
   not go into details.}.
-See Appendix~\ref{appendix:false_lemmas} for a detailed explanation
+See Appendix %~\ref{appendix:false_lemmas}
+for a detailed explanation
 and counterexample. This lemma is a cornerstone of the two metatheoretical results 
@@
 which might cause existential variables to leak across judgments.
-A detailed discussion on this issue with a possible fix is in Section~\ref{sec:discussion:scoping}.
+In Section~\ref{sec:discussion:scoping} we give a detailed discussion.
+%A detailed discussion on this issue with a possible fix is .
 

### 2019-06-26 sections/Related.tex
split appendix from main & fix reference issues

@@
 Though carefully written, some of the proofs are incorrect
-(see discussion in Appendix~\ref{appendix:false_lemmas} and \ref{appendix:subsumption}),
+(see discussion in Appendix
+%~\ref{appendix:false_lemmas} and \ref{appendix:subsumption}
+),
 which creates difficulties when formalizing them in a proof assistant.
@@
 as lemma 29 in \citet{dunfield2013complete}: the conclusion is false. We also discuss
-the issue in more detail in Appendix~\ref{appendix:false_lemmas}.
+the issue in more detail in Appendix. %~\ref{appendix:false_lemmas}.
 
